### 2023-08-03 article.tex

\documentclass[review,12pt,authoryear]{elsarticle}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

% for adjusting table width automatically
\usepackage{adjustbox}
\usepackage{tabulary, ragged2e}
\usepackage{booktabs}

% Below is Elseviers requirements - they are similar to most articles and a good point of reference when writting scientific articles or analyses in general.
%1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
% The main text should be approximately 6,000 words in length, but it should not exceed 8,000 words (excluding abstract, references, tables, figures, and appendices).
%A maximum of 250 words abstract and up to 10 displayed items (figures and tables) is allowed. A full-length article should include an Introduction,
%Materials and methods, Results, Discussion, Conclusions, and References, which can be accompanied by Supplementary material.

\begin{document}
\begin{linenumbers}
\begin{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%              Start Matter            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:

%% \title{Title\tnoteref{label1}}
\title{Grape Quality and its Link to Regional Differences in the Australian Winegrowing Industry}
% Regional differences in Australian winegrowing (Quality)?

%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%\affiliation[label1]{organization={QUT},
%  addressline={},
%  city={},
%  postcode={},
%  state={QLD},
%  country={}}
%\affiliation[label2]{organization={AWRI},
%  addressline={},
%  city={},
%  postcode={},
%  state={SA},
%  country={}}
%\affiliation[label3]{organization={Food Agility CRC},
%  addressline={},
%  city={},
%  postcode={},
%  state={Vic},
%  country={}}
\author[label1,label2,label3]{Author}
\date{02/08/2023}

\begin{abstract}
\end{abstract}
%%Graphical abstract`
%\begin{graphicalabstract}
 % \includegraphics{graphical_abstract.jpeg}
%\end{graphicalabstract}'

%\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
%Keyword one \sep{} keyword two
%% PACS codes here, in the form: \PACS code \sep code
%\PACS{} 0000 \sep{} 1111
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC{} 0000 \sep{} 1111
%\end{keyword}
%%Research highlights
\begin{highlights}
  \item Comparative analysis of resource use, quality and quantity in Australian winegrowing.
  \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
  \item Baseline models for comparing wine crops.
  \item Analysis of national, decade long data source.
\end{highlights}
\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                main text             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple regions that help to describe the unique varietals and qualities that are produced there. While a great deal has been done regarding regional properties and traits, there has been little statistical insight into how regions differ due to a lack of cross-regional and in-depth data sources (Keith Jones, 2002; Knight et al., 2019). In this study we use Classification Trees to compare regional differences and how they relate to sustainable practices and grape quality.
The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}. This analysis addresses the knowledge gap regarding the effectiveness of regional level strategies employed in the wine industry and their relation to grape quality. Through the use of classification trees this study aims to highlight the key differences in sustainable practices at a regional level and how these practices relate to the different grades of grape quality.

\subsubsection{A figure sub-subsection}

\if{false}
\begin{figure}
    \includegraphics{violinplot.pdf}
    \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
    \end{figure}
  %
\fi{}

\section{Methods}

\subsection{Data}
The Australian wine industry is divided into 65 regions, known as a Geographical Indicator Regions (GI Region). Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties (Halliday, 2009; Oliver et al., 2013; SOAR et al., 2008). Each region is explicitly defined under the Wine Australia Corporation Act of 1980 (Attorney-General's Department, 2010). The climatic properties of a GI Region are summarised by Sustainable Winegrowing Australia (2021), where regions of similar climates are amalgamated together into superset regions. The climatic regions were utilised to illustrate similar trends and explain differences between sets of regions.
The data used in this analysis comes from Sustainable Winegrowing Australia and covers the period 2015 to 2022. The dataset contained 3342 samples across 52 GI Regions and 1072 individual vineyards.

\subsection{Classification Trees}

Classification Trees were developed to discern the different practices within regions and climates,  comparing these relationships to those linked to grape quality. This was done using the rparts and caret packages (Kuhn, 2008; Terry Therneau and Beth Atkinson, 2022) in the R statistical programming language (R Core Team, 2021). Three classifications were undertaken for region, climate and grape quality. Climate was further classified into two subcategories of rainfall and temperature, resulting in a total of 5 classification trees being created.
Classification trees were validated using K-fold cross validation. Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the  classification trees. A summary confusion matrix was then constructed to show the class bias and overall accuracy of each tree.

\section{Results}

\subsection{Model 1 GI Regions}
The first Model was used to classify GI regions and resulted in an accuracy of 36.48\% across 52 classes. The most prominent features used to classify regions were the types of water resources available (see Figure 1). Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively. These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region. However, they are both similar in operational scales, with vineyards being relatively large compared with other regions. The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source. Many of the regions had significantly lower reporting rates, resulting much poorer classification performance. The regions with the most samples performed the best (see Table 1). Notably bordering regions were routinely grouped together and misclassified as the same region, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively. The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland. These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another. Other misclassifications were most likely due to lower reporting rates with many regions being under represented.

\begin{table}[]
    \caption{Classification accuracy of the most prominent GI Regions.}
    \label{tab:accuracy}
    \resizebox{\textwidth}{!}{
        \begin{tabular}{@{}llll@{}}
            \toprule
            \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
            \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
            \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
            \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
            \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
            \textbf{Margaret River} & 78.82\% & 201 & 255 \\
            \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
            \textbf{Riverland} & 92.74\% & 345 &
        \end{tabular}
    }
\end{table}

\subsection{Climate}
Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly. These categories were divided into 12 climatic classifications with 3 and 4 separate subsets for rainfall and temperature respectively. The decision tree behaved similarly and over classified climates with higher response rates. The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area. The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure (Reynolds, 2010). This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.

\subsubsection{Rainfall}
The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).

\subsubsection{Temperature}
The classification of GI Regions by their temperatures (see Figure 5) showed similarities to the other trees, with a heavy reliance on the types of water resources used as dominant predictors. The use of contractors was again used to differentiate between warm and cool regions, likely being due to disease pressure. The temperature classification tree was only a minor improvement over the regional classification tree, with an accuracy of 49.26\% as shown in the confusion matrix (see Table 4).

\subsection{Model 3 Grape Quality}
The classification of grape quality through its grade had an accuracy of 55.72\% across 5 separate grades. There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2). The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner. Although not surprising the number of contractor tractor passes is new deciding factor due disease and pests reducing the potential quality of a crop. The prevalence of contractor use is greater in regions such as the Barossa Valley and the McLaren Vale, this could be due to the difference in operational scales, with larger sites being more likely to have ownership of their own equipment for weeding and spraying due to the cost benefit.

\section{Discussion}
The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively. Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012).
The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities. The most prominent being what readily available resources of each region were, particular the types of water available. Heavy water consumption is often linked to the mass production of grapes, where lower quality grapes are targeted in a quantity over quality strategy. These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans. Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield ((Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I. Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018)). There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales. This results in unique issues, where on-the-ground choices are influenced by other wine industry’s decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019). It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them. It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season (Leilei He et al., 2022).

\section{Conclusion}

The type and availability of water resources were a major contributing factor when classifying grape quality and region. This was seen in the two most accurately classified regions, Coonawarra and the Riverland, with the Riverland predominantly utilising river water. Furthermore, the study highlighted the influence of water use, fungicide application, and contractor use in differentiating grape quality, climate and region respectively. These models provide insight into the complex dynamics between regional characteristics, sustainable practices, and grape quality in the Australian winegrowing industry. It is important to acknowledge that grape quality is subject to external influences such as market demands and prior established business arrangements. Further in-depth data and understanding are necessary to fully grasp the nuances of decision-making and the interplay of factors impacting grape quality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%              End Matter              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{references} % This points to the references.bib file - the file extention is automatically added.
\bibliographystyle{elsarticle-harv}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
 \appendix

\end{linenumbers}
  \end{document}

\endinput

### 2023-08-07 article.tex
added draft qustions
initial commit toget references to work.

@@
 
-The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple regions that help to describe the unique varietals and qualities that are produced there.
-While a great deal has been done regarding regional properties and traits, there has been little statistical insight into how regions differ due to a lack of cross-regional and in-depth data sources (Keith Jones, 2002; Knight et al., 2019).
+% You obviously have no leading question for this paper, and although you do a good job at creating accurate models; it is more important that the research is asking a good question.
+A cool question as Kate puts it.
+%
+% The questions I put forward to you are:
+%
+%   How do winegrowing regions differ?
+%   Why is a region different?
+%   How do regional differences affect winegrowing?
+%   What differences are important between winegrowing regions?
+%   
+%
+
+The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple regions describing the unique varietals and qualities of wine produced there.
+While a great deal has been done regarding regional properties and traits, there has been little statistical insight into how regions differ; this has been due to a lack of cross-regional and in-depth data sources   (Keith Jones, 2002; Knight et al., 2019).
 In this study we use Classification Trees to compare regional differences and how they relate to sustainable practices and grape quality.

### 2023-08-08 article.tex

@@
 
-The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple regions describing the unique varietals and qualities of wine produced there.
-While a great deal has been done regarding regional properties and traits, there has been little statistical insight into how regions differ; this has been due to a lack of cross-regional and in-depth data sources   (Keith Jones, 2002; Knight et al., 2019).
-In this study we use Classification Trees to compare regional differences and how they relate to sustainable practices and grape quality.
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-This analysis addresses the knowledge gap regarding the effectiveness of regional level strategies employed in the wine industry and their relation to grape quality.
+The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
+Each region describing unique reputations, qualities and varietals of wine produced there.
+While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \autocite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
+\newline
+A vineyard's region predetermines several physical parameters, such as: climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+%
+%This analysis addresses the knowledge gap regarding the effectiveness of regional level strategies employed in the wine industry and their relation to grape quality.
+% What the hell is a regional strategy
+% And how do you define, or know what trategy each region is using?
+%
 Through the use of classification trees this study aims to highlight the key differences in sustainable practices at a regional level and how these practices relate to the different grades of grape quality.
+
+% This does not really address what the methods are - aside from classisifcation trees in the most vague sense.
+% What are the results briefly? why should someone read this article?
+% What is interesting - what will be discussed in the article with regard to these results? What makes me want to read the article?
+% And what is the take away point of the article? If I cant be bothered to read this article, what piece of knowledge can you leave mewiwth that I could recommend the article to someone else with?
 

### 2023-08-08 article.tex
methods comments

@@
 
+% As with the previous paper, put the models up front it is very unclear what you have created - and less clear to understand it if there is no way to reference to it.
+
 \subsection{Data}
 The Australian wine industry is divided into 65 regions, known as a Geographical Indicator Regions (GI Region).
-Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties (Halliday, 2009; Oliver et al., 2013; SOAR et al., 2008).
-Each region is explicitly defined under the Wine Australia Corporation Act of 1980 (Attorney-General's Department, 2010).
+Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties \autocite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \autocite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 The climatic properties of a GI Region are summarised by Sustainable Winegrowing Australia (2021), where regions of similar climates are amalgamated together into superset regions.
@@
 
+% how was it collected
+
 \subsection{Classification Trees}
 
-Classification Trees were developed to discern the different practices within regions and climates,  comparing these relationships to those linked to grape quality.
-This was done using the rparts and caret packages (Kuhn, 2008; Terry Therneau and Beth Atkinson, 2022) in the R statistical programming language (R Core Team, 2021).
+Classification Trees were developed to discern the different practices within regions and climates, 
+% Were they? Or was it to discern regional differences and then compare them?
+comparing these relationships to those linked to grape quality.
+This was done using the rparts and caret packages \autocite{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language (R Core Team, 2021).
+\newline
 Three classifications were undertaken for region, climate and grape quality.
-Climate was further classified into two subcategories of rainfall and temperature, resulting in a total of 5 classification trees being created.
-Classification trees were validated using K-fold cross validation.
+% Is this three classifications for each of these, being 9 in total? Or three overall, one for each?
+Climate was further classified into two subcategories of rainfall and temperature,
+% Why was this done?
+resulting in a total of 5 classification trees being created. % You say this too late!
+Classification trees were validated using K-fold cross validation.% This could be its own section!
  Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the  classification trees.
 A summary confusion matrix was then constructed to show the class bias and overall accuracy of each tree.
+ % To what end? Why do we want them to be accurate?? What does this analysis achieve?
+
+% Was the analysis validated?
+% How where the carts compared??
+% I still do not know what the models really where
 

### 2023-08-08 article.tex
Comments for edits throughout the paper

@@
 \subsection{Model 1 GI Regions}
-The first Model was used to classify GI regions and resulted in an accuracy of 36.48\% across 52 classes.
+The first Model was used to classify GI regions % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
+and resulted in an accuracy of 36.48\% across 52 classes.
 The most prominent features used to classify regions were the types of water resources available (see Figure 1).
 Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively.
+% it might be worth only talking about the regions that were well classifified - and then talk about their determining traits.
 These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
 However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
+%Summary statistics regarding these two regions would be a better point of comparison, for example:
+%
+% How much in size do these operational scales differ?
+% How much yield do they differ by?
+% What are the differences in rainfall etc
+% What makes them unique, different and similar?
+%
 The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
+% What are the specific differences? This is a really critical part of this paper
 Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
+% This is almost its own paragraph.
+Why was this case?
+is it important?
+what are the similarities and differences of regions that were poorly classified?
  The regions with the most samples performed the best (see Table 1).
@@
 The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
+ % THis speaks highly to why regions were poorly classified.
 These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
@@
 The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area.
-The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure (Reynolds, 2010).
+The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \autocite{reynoldsManagingWineQuality2010}.
+% interesting although not really appropriate for the results section.
  This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
+
+% You really need to show every decision tree.
 
@@
 The classification of GI Regions by their temperatures (see Figure 5) showed similarities to the other trees, with a heavy reliance on the types of water resources used as dominant predictors.
-The use of contractors was again used to differentiate between warm and cool regions, likely being due to disease pressure.
+The use of contractors % This is super important.
+Talk about this at lengthi nt the discussion.
+Find more literature on agriculture and the use of contractors as oppose to self managing issues such as weeds.
+was again used to differentiate between warm and cool regions, likely being due to disease pressure.
 The temperature classification tree was only a minor improvement over the regional classification tree, with an accuracy of 49.26\% as shown in the confusion matrix (see Table 4).
@@
 \subsection{Model 3 Grape Quality}
-The classification of grape quality through its grade had an accuracy of 55.72\% across 5 separate grades.
-There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2).
+The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
+had an accuracy of 55.72\% across 5 separate grades. There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2). % This is actually really interesting.
  The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
@@
 The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
-Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012).
+Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
+It is a bit unclear what you are really trying to cite and its relevance.
+I think it important to chop this up.
 The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
 The most prominent being what readily available resources of each region were, particular the types of water available.
-Heavy water consumption is often linked to the mass production of grapes, where lower quality grapes are targeted in a quantity over quality strategy.
+% A citation or reference to a study looking at the disparity in wine regions and their resources - or use the data more to describe these regional differences.
+Is water availability apparent?
+Do some regions have access to more sohpisticated water sources such as pressurised water etc.
+ Heavy water consumption is often linked to the mass production of grapes, % Really?
+proce it?
+Where in the results is this really found??
+This sounds like a link to your previous paper but you cannot really show this and you do not really produce evidence to support this claim.
+ where lower quality grapes are targeted in a quantity over quality strategy.
 These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
-Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield ((Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
-Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018)).
+ % This point is incredibly important but it, itself, is obfuscated by you odd claims that are not supported by facts.
+Better to talk about how this could improve and the weaknesses than to start spouting unsupported claims.
+ Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \autocite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
+ % Woah woah woah
+%
+ % what are these decisions??? Slow down speed gonzales!! This is really interseteing and important. Discuss it, its the discussion! God damn I want to know and I wrote this!.
+ %
+%
+% This is not all one paragraph surely! There are so many ideas here that should be discussed!
+%
+% This leads heavily into the next paper and shows that natural progression.
+Especially in the sense of: if regions have so much influences, what determines ones use of water, fertiliser, etc.
+And, what determines the success financially, in quality or quantity?
+How are thes evariables interacting!
+The next paper is less about predicting and much more about what interacts with what?
+Where as this paper is about how things change over landscapes and what localised events helpe to determine decisions and their outcomes.
+%
+
+ .
 There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues, where on-the-ground choices are influenced by other wine industry’s decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+This results in unique issues, where on-the-ground choices are influenced by other wine industry's decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+% This is an interesting point.
+However it would be better to include other industries, such as the ones you touch on in your previous chapter.
+If industries such as Sake are similar - how do their decisions effect the outcomes of rice quality and Sake?
+A point of comparison and difference would be great here!
  It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
-It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season (Leilei He et al., 2022).
+% What are these business decisions? Can you give some examples?
+ It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \autocite{leileiheFruitYieldPrediction2022}.
+% It is difficult yes, but please at least try!
+Highlight some of these affects, some of these different contributing factors!
+What did we really learn by doing this?
+What can someone take away from this and say thank you for?
+(like Josh said)
 
@@
 Further in-depth data and understanding are necessary to fully grasp the nuances of decision-making and the interplay of factors impacting grape quality.
+% Interesting - the conclusion is actually reasonable, especially considering the disconnection of ideas and models in this paper.
 

### 2023-08-10 article.tex
Added commentry about the structure of the article and analytics.

@@
 
+\if{false}
+
+We classify regions to find the most prominent and well-defined relationships to regions.
+We examine the variables that are the most decisive in determining region.
+We compare the similarities and differences of these regions
+- within the analysis and
+- within the literature (also in their definition and history)
+
+We examine the similarity between these classifications and the variables that are significant in classifying profitable vs not.
+How do regional changes affect that. What is similar and and what is different.
+
+We can link these contributing factors back to the baselines - at least anaecdotally, or in an unpublished section. THis will be good for the thesis.
+
+\fi
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.

### 2023-08-10 analysis.r
removed unnecessary data transform

@@ -8,7 +8,7 @@ trControl <- trainControl(
     , repeats=10
     , verboseIter=TRUE
 )
-    
+
 df <- read.csv("no_trans.csv")
 df$profitable <- factor(df$profitable)
 
@@ -100,22 +100,3 @@ rf <- train(
     , na.action = na.omit
     , tuneLength=10
 )
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2023-08-10 data_prep.py
added req file

@@ -84,7 +84,9 @@ df_o = df.select_dtypes("O")
 df_int = df.select_dtypes(int)
 df_floats = df_floats.apply(np.log)
 df_floats = df_floats.replace({np.inf: np.nan, -np.inf: np.nan})
-df_floats = centre_scale(df_floats)
+
+# We do not want to scale and centre things as we are using classification techniques. In particular these are partition techniques, and so the raw data is important.
+#df_floats = centre_scale(df_floats)
 
 # original df
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("df.csv")
@@ -114,7 +116,7 @@ df_floats["herbicide_spraying_number_of_times_passes_per_year"] = df["herbicide_
 df_floats["insecticide_spraying_number_of_times_passes_per_year"] = df["insecticide_spraying_number_of_times_passes_per_year"]
 
 for col in cols_to_transform:
-    df_floats[col] = df[col].div(df["area_harvested"], axis=0)*1000
+    df_floats[col] = df[col].div(df["area_harvested"], axis=0)*10000
 
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfp.csv")
 
@@ -164,7 +166,7 @@ prop_cols = [
     , 'vineyard_area_red_grapes'
 ]
 for col in prop_cols:
-    df_floats[col] = df[col].div(df["vineyard_area"], axis=0)*1000
+    df_floats[col] = df[col].div(df["vineyard_area"], axis=0)*10000
 
 # These are a kind of proportion 2 passes is 200%
 df_floats["total_tractor_passes"] = df["total_tractor_passes"]

### 2023-08-10 data_prep.py
removed log transform that was not needed.

@@ -82,7 +82,7 @@ dfcsv.to_csv("dft.csv")
 df_floats = df.select_dtypes(float)
 df_o = df.select_dtypes("O")
 df_int = df.select_dtypes(int)
-df_floats = df_floats.apply(np.log)
+#df_floats = df_floats.apply(np.log)
 df_floats = df_floats.replace({np.inf: np.nan, -np.inf: np.nan})
 
 # We do not want to scale and centre things as we are using classification techniques. In particular these are partition techniques, and so the raw data is important.

### 2023-08-10 analysis.r
edits sanity checking analysis

@@ -9,80 +9,36 @@ trControl <- trainControl(
     , verboseIter=TRUE
 )
 
-df <- read.csv("no_trans.csv")
+df <- read.csv("df.csv")
 df$profitable <- factor(df$profitable)
 
-# We select the columns we are going to use!
-cols <- c('tonnes_grapes_harvested'
-#, 'area_harvested'
-#, 'water_used'
-, 'total_tractor_passes'
-#, 'total_vineyard_fuel'
-#, 'total_vineyard_electricity'
-#, 'total_irrigation_area'
-, 'synthetic_nitrogen_applied'
-, 'organic_nitrogen_applied'
-, 'synthetic_fertiliser_applied'
-, 'organic_fertiliser_applied'
-, 'area_not_harvested'
-#, 'total_irrigation_electricity'
-#, 'total_irrigation_fuel'
-, 'giregion'
-, 'vineyard_area_white_grapes'
-, 'vineyard_area_red_grapes'
-, 'river_water'
-, 'groundwater'
-, 'surface_water_dam'
-, 'recycled_water_from_other_source'
-, 'mains_water'
-, 'other_water'
-, 'water_applied_for_frost_control'
-, 'bare_soil'
-, 'annual_cover_crop'
-, 'permanent_cover_crop_native'
-, 'permanent_cover_crop_non_native'
-, 'permanent_cover_crop_volunteer_sward'
-, 'irrigation_energy_diesel'
-, 'irrigation_energy_electricity'
-, 'irrigation_energy_pressure'
-, 'irrigation_energy_solar'
-, 'irrigation_type_dripper'
-, 'irrigation_type_flood'
-, 'irrigation_type_non_irrigated'
-, 'irrigation_type_overhead_sprinkler'
-, 'irrigation_type_undervine_sprinkler'
-, 'diesel_vineyard'
-, 'electricity_vineyard'
-, 'petrol_vineyard'
-, 'vineyard_solar'
-, 'vineyard_wind'
-, 'lpg_vineyard'
-, 'biodiesel_vineyard'
-, 'slashing_number_of_times_passes_per_year'
-, 'fungicide_spraying_number_of_times_passes_per_year'
-, 'herbicide_spraying_number_of_times_passes_per_year'
-, 'insecticide_spraying_number_of_times_passes_per_year'
-, 'nh_disease'
-, 'nh_frost'
-, 'nh_new_development'
-, 'nh_non_sale'
-, 'off_farm_income'
[…]

### 2023-08-10 data_prep.py

@@ -3,6 +3,7 @@ import pandas as pd
 import numpy as np
 
 df = pd.read_feather("df.feather")
+
 # we get the indices of those that made profit
 indices = df[
     (df["total_operating_costs"]>0) & (df["total_operating_costs"]>0)].index
@@ -11,7 +12,7 @@ df["profit"] = np.nan
 # calculate profit
 df.loc[indices, "profit"] = df.loc[
     indices, "total_grape_revenue"] - df.loc[indices, "total_operating_costs"]
-df["area_harvested"] = df["area_harvested"]*1000
+df["area_harvested"] = df["area_harvested"]*10000
 # We create a binary classification of those that made a profit
 # as oppose to a loss
 df["profitable"] = np.nan
@@ -53,10 +54,10 @@ for i, row in df.iterrows():
         df.at[i, "prev_avg"] = np.nan
 df["prev_avg"] = df["prev_avg"].replace({0: np.nan})
 
-dfcsv = pd.read_csv("no_trans.csv")
-dfcsv["profit"] = df["profit"].copy()
-dfcsv["profitable"] = df["profitable"].copy()
-dfcsv["prev_avg"] = df["prev_avg"]
+#dfcsv = pd.read_csv("no_trans.csv")
+#dfcsv["profit"] = df["profit"].copy()
+#dfcsv["profitable"] = df["profitable"].copy()
+#dfcsv["prev_avg"] = df["prev_avg"]
 
 # We add the wine australia survey data:
 df["average_per_tonne"] = df["average_per_tonne"].replace({0: np.nan})
@@ -77,7 +78,7 @@ df.loc[(1500<=df["average_per_tonne"])&(df["average_per_tonne"]<2000), "grade"] 
 df.loc[(2000<=df["average_per_tonne"]), "grade"] = "A"
 
 # no transformation
-dfcsv.to_csv("dft.csv")
+#dfcsv.to_csv("dft.csv")
 
 df_floats = df.select_dtypes(float)
 df_o = df.select_dtypes("O")

### 2023-08-21 analysis.r
changed to rpart + caret. with simple example.

@@ -2,19 +2,30 @@ library(randomForest)
 library(mlbench)
 library(caret)
 
+#######################
+#
+# We want:
+#
+# 
+# giregion
+# profitability
+# profit
+# operating costs
+#
+########################
+
 trControl <- trainControl(
-    method='repeatedcv'
-    , number=10, 
-    , repeats=10
-    , verboseIter=TRUE
+    method = "repeatedcv"
+    , number = 2
+    , repeats = 2
+    , verboseIter = TRUE
 )
 
 df <- read.csv("df.csv")
 df$profitable <- factor(df$profitable)
 
 profit_tree_col <- c(
-        "profitable"
-        , "profit"
+         "profit"
         , "operating_cost_per_ha"
         , "operating_cost_per_t"
         , "total_operating_cost"
@@ -25,34 +36,4 @@ profit_tree_col <- c(
         , "gross_margin"
         , "cost_of_debt_servicing")
 
-profit_tree <- train(
-    df[complete.cases(df[ , "profitable"]),
-     !(names(df) %in% profit_tree_col)]
-    , df[complete.cases(df[ , "profitable"]), "profitable"]
-    , method = "rf"
-    , metric = "Accuracy"
-    , trControl = trControl
-    , na.action = na.omit
-    , tuneLength = 10
-)
-plot(profit_tree)
-
-varImp(profit_tree)
-plot(varImp(profit_tree))
-
-###################################################
-
-trControl <- trainControl(
-    , number=10, 
-    , repeats=1
-    , verboseIter=TRUE
-)
-
-operating_tree <- train(                           
-    df[complete.cases(df[ , "total_operating_costs"]), (names(df) %in% cols)]
-    , df[complete.cases(df[ , "total_operating_costs"]), "total_operating_costs"]
-    , method="xgbTree"
-    , trControl=trControl
-    , na.action = na.omit
-    , tuneLength=10
-)
+rpart.plot(rf$finalModel, extra=102, legend.x=-100)

### 2023-08-21 analysis.py
working xgboost

@@ -1,0 +1,228 @@
+import pandas as pd
+import numpy as np
+import xgboost as xgb
+from xgboost import cv
+from sklearn.model_selection import train_test_split, KFold, cross_val_score
+from sklearn.metrics import mean_squared_error
+import matplotlib.pyplot as plt
+from data import bin_col
+from sklearn.inspection import permutation_importance
+from sklearn.metrics import accuracy_score
+
+import random
+
+random.seed(0)
+
+################################################################################
+# Setup
+df = pd.read_csv("dfb.csv")
+
+df["nh_disease"] = bin_col(df["nh_disease"].replace({np.nan: 0})**2)
+
+df["fertiliser"] =  (df['synthetic_nitrogen_applied']
+    + df['organic_nitrogen_applied']
+    + df['synthetic_fertiliser_applied']
+    + df['organic_fertiliser_applied']
+)
+
+cols = ['tonnes_grapes_harvested'
+    , 'area_harvested'
+    , 'water_used'
+    , 'total_tractor_passes'
+    , "fertiliser"
+    #, 'total_vineyard_fuel'
+    #, 'total_vineyard_electricity'
+    #, 'total_irrigation_area'
+    , 'synthetic_nitrogen_applied'
+    , 'organic_nitrogen_applied'
+    , 'synthetic_fertiliser_applied'
+    , 'organic_fertiliser_applied'
+    #, 'area_not_harvested'
+    #, 'total_irrigation_electricity'
+    #, 'total_irrigation_fuel'
+    , 'giregion'
+    #, 'vineyard_area_white_grapes'
+    #, 'vineyard_area_red_grapes'
+    , 'river_water'
+    , 'groundwater'
+    , 'surface_water_dam'
+    , 'recycled_water_from_other_source'
+    , 'mains_water'
+    , 'other_water'
+    , 'water_applied_for_frost_control'
+    , 'bare_soil'
+    , 'annual_cover_crop'
+    , 'permanent_cover_crop_native'
+    , 'permanent_cover_crop_non_native'
+    , 'permanent_cover_crop_volunteer_sward'
+    #, 'irrigation_energy_diesel'
+    #, 'irrigation_energy_electricity'
+    #, 'irrigation_energy_pressure'
+    #, 'irrigation_energy_solar'
+    #, 'irrigation_type_dripper'
+    #, 'irrigation_type_flood'
+    #, 'irrigation_type_non_irrigated'
+    #, 'irrigation_type_overhead_sprinkler'
[…]

### 2023-08-21 article.tex
minor changes

@@
 %% \title{Title\tnoteref{label1}}
-\title{Grape Quality and its Link to Regional Differences in the Australian Winegrowing Industry}
+\title{???Grape Quality and its Link to Regional Differences in the Australian Winegrowing Industry}
 % Regional differences in Australian winegrowing (Quality)?
@@
 \begin{highlights}
-  \item Comparative analysis of resource use, quality and quantity in Australian winegrowing.
-  \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
-  \item Baseline models for comparing wine crops.
-  \item Analysis of national, decade long data source.
+  \item ???
+  \item ???
+  \item ????
+  \item ????
 \end{highlights}
@@
 We examine the similarity between these classifications and the variables that are significant in classifying profitable vs not.
-How do regional changes affect that. What is similar and and what is different.
+How do regional changes affect that. What is similar and what is different.
 
-We can link these contributing factors back to the baselines - at least anaecdotally, or in an unpublished section. THis will be good for the thesis.
+We can link these contributing factors back to the baselines - at least anecdotally, or in an unpublished section. This will be good for the thesis.
 

### 2023-08-22 analysis.r
finalised ctrees.

@@ -1,18 +1,50 @@
 library(randomForest)
 library(mlbench)
 library(caret)
+library(rpart.plot)
 
-#######################
-#
-# We want:
-#
-# 
-# giregion
-# profitability
-# profit
-# operating costs
-#
-########################
+df <- read.csv("dfb.csv")
+df$profitable <- factor(df$profitable)
+
+selected_cols <- c("giregion"
+    , "profitable"
+    , "profit"
+    , "total_operating_costs"
+    , "tonnes_grapes_harvested"
+    , "area_harvested"
+    , "water_used"
+    , "total_tractor_passes"
+    , "total_fertiliser"
+    , "synthetic_nitrogen_applied"
+    , "organic_nitrogen_applied"
+    , "synthetic_fertiliser_applied"
+    , "organic_fertiliser_applied"
+    , "river_water"
+    , "groundwater"
+    , "surface_water_dam"
+    , "recycled_water_from_other_source"
+    , "mains_water"
+    , "other_water"
+    , "water_applied_for_frost_control"
+    , "bare_soil"
+    , "annual_cover_crop"
+    , "permanent_cover_crop_native"
+    , "permanent_cover_crop_non_native"
+    , "permanent_cover_crop_volunteer_sward"
+    , "diesel_vineyard"
+    , "electricity_vineyard"
+    , "petrol_vineyard"
+    , "vineyard_solar"
+    , "vineyard_wind"
+    , "lpg_vineyard"
+    , "biodiesel_vineyard"
+    , "slashing_number_of_times_passes_per_year"
+    , "fungicide_spraying_number_of_times_passes_per_year"
+    , "herbicide_spraying_number_of_times_passes_per_year"
+    , "insecticide_spraying_number_of_times_passes_per_year"
+)
+
+# Training function
 
 trControl <- trainControl(
     method = "repeatedcv"
@@ -21,19 +53,85 @@ trControl <- trainControl(
     , verboseIter = TRUE
 )
 
-df <- read.csv("df.csv")
-df$profitable <- factor(df$profitable)
+###########################
+#       GI region         #
+###########################
 
-profit_tree_col <- c(
-         "profit"
-        , "operating_cost_per_ha"
-        , "operating_cost_per_t"
-        , "total_operating_cost"
-        , "total_grape_revenue"
-        , "highest_per_tonne"
-        , "lowest_per_tonne"
-        , "average_per_tonne"
-        , "gross_margin"
[…]

### 2023-08-22 data_prep.py
changed some variables to not be binary such as tractor passes.

@@ -12,7 +12,7 @@ df["profit"] = np.nan
 # calculate profit
 df.loc[indices, "profit"] = df.loc[
     indices, "total_grape_revenue"] - df.loc[indices, "total_operating_costs"]
-df["area_harvested"] = df["area_harvested"]*10000
+# df["area_harvested"] = df["area_harvested"]*10000
 # We create a binary classification of those that made a profit
 # as oppose to a loss
 df["profitable"] = np.nan
@@ -77,6 +77,12 @@ df.loc[(600<=df["average_per_tonne"])&(df["average_per_tonne"]<1500), "grade"] =
 df.loc[(1500<=df["average_per_tonne"])&(df["average_per_tonne"]<2000), "grade"] = "B"
 df.loc[(2000<=df["average_per_tonne"]), "grade"] = "A"
 
+df["total_fertiliser"] =  (df['synthetic_nitrogen_applied']
+    + df['organic_nitrogen_applied']
+    + df['synthetic_fertiliser_applied']
+    + df['organic_fertiliser_applied']
+)
+
 # no transformation
 #dfcsv.to_csv("dft.csv")
 
@@ -116,8 +122,8 @@ df_floats["fungicide_spraying_number_of_times_passes_per_year"] = df["fungicide_
 df_floats["herbicide_spraying_number_of_times_passes_per_year"] = df["herbicide_spraying_number_of_times_passes_per_year"]
 df_floats["insecticide_spraying_number_of_times_passes_per_year"] = df["insecticide_spraying_number_of_times_passes_per_year"]
 
-for col in cols_to_transform:
-    df_floats[col] = df[col].div(df["area_harvested"], axis=0)*10000
+# for col in cols_to_transform:
+#     df_floats[col] = df[col].div(df["area_harvested"], axis=0)*10000
 
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfp.csv")
 
@@ -125,11 +131,11 @@ pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfp.csv")
 for col in cols_to_transform:
     df_floats[col] = bin_col(df[col])
 
-df_floats["total_tractor_passes"] = bin_col(df["total_tractor_passes"])
-df_floats["slashing_number_of_times_passes_per_year"] = bin_col(df["slashing_number_of_times_passes_per_year"])
-df_floats["fungicide_spraying_number_of_times_passes_per_year"] = bin_col(df["fungicide_spraying_number_of_times_passes_per_year"])
[…]

### 2023-08-22 analysis.py
xgboost now runs!

@@ -19,17 +19,11 @@ df = pd.read_csv("dfb.csv")
 
 df["nh_disease"] = bin_col(df["nh_disease"].replace({np.nan: 0})**2)
 
-df["fertiliser"] =  (df['synthetic_nitrogen_applied']
-    + df['organic_nitrogen_applied']
-    + df['synthetic_fertiliser_applied']
-    + df['organic_fertiliser_applied']
-)
-
 cols = ['tonnes_grapes_harvested'
     , 'area_harvested'
     , 'water_used'
     , 'total_tractor_passes'
-    , "fertiliser"
+    , "total_fertiliser"
     #, 'total_vineyard_fuel'
     #, 'total_vineyard_electricity'
     #, 'total_irrigation_area'

### 2023-08-22 data_prep.py
fixed binary list, proportions was changed.

@@ -114,6 +114,15 @@ cols_to_transform = [
     , "irrigation_type_non_irrigated"
     , "irrigation_type_overhead_sprinkler"
     , "irrigation_type_undervine_sprinkler"
+    , 'river_water'
+    , 'groundwater'
+    , 'surface_water_dam'
+    , 'recycled_water_from_other_source'
+    , 'mains_water'
+    , 'other_water'
+    , 'water_applied_for_frost_control'
+    , "nh_frost"
+    , "nh_disease"
 ]
 
 df_floats["total_tractor_passes"] = df["total_tractor_passes"]
@@ -162,6 +171,8 @@ bin_cols = [
     , 'mains_water'
     , 'other_water'
     , 'water_applied_for_frost_control'
+    , "nh_frost"
+    , "nh_disease"
 ]
 
 for col in bin_cols:

### 2023-08-22 analysis.py
fixed importance and categoric variables

@@ -11,32 +11,64 @@ from sklearn.metrics import accuracy_score
 
 import random
 
-random.seed(0)
+random.seed(100)
 
 ################################################################################
 # Setup
 df = pd.read_csv("dfb.csv")
 
-df["nh_disease"] = bin_col(df["nh_disease"].replace({np.nan: 0})**2)
-
-cols = ['tonnes_grapes_harvested'
-    , 'area_harvested'
-    , 'water_used'
-    , 'total_tractor_passes'
+cols = ["tonnes_grapes_harvested"
+    , "area_harvested"
+    , "water_used"
+    , "total_tractor_passes"
     , "total_fertiliser"
-    #, 'total_vineyard_fuel'
-    #, 'total_vineyard_electricity'
-    #, 'total_irrigation_area'
-    , 'synthetic_nitrogen_applied'
-    , 'organic_nitrogen_applied'
-    , 'synthetic_fertiliser_applied'
-    , 'organic_fertiliser_applied'
-    #, 'area_not_harvested'
-    #, 'total_irrigation_electricity'
-    #, 'total_irrigation_fuel'
-    , 'giregion'
-    #, 'vineyard_area_white_grapes'
-    #, 'vineyard_area_red_grapes'
+    , "synthetic_nitrogen_applied"
+    , "organic_nitrogen_applied"
+    , "synthetic_fertiliser_applied"
+    , "organic_fertiliser_applied"
+    , "giregion"
+    , "data_year_id"
+    , "river_water"
+    , "groundwater"
+    , "surface_water_dam"
+    , "recycled_water_from_other_source"
+    , "mains_water"
+    , "other_water"
+    , "water_applied_for_frost_control"
+    , "bare_soil"
+    , "annual_cover_crop"
+    , "permanent_cover_crop_native"
+    , "permanent_cover_crop_non_native"
+    , "permanent_cover_crop_volunteer_sward"
+    , "diesel_vineyard"
+    , "electricity_vineyard"
+    , "petrol_vineyard"
+    , "vineyard_solar"
+    , "vineyard_wind"
+    , "lpg_vineyard"
+    , "biodiesel_vineyard"
+    , "slashing_number_of_times_passes_per_year"
+    , "fungicide_spraying_number_of_times_passes_per_year"
+    , "herbicide_spraying_number_of_times_passes_per_year"
+    , "insecticide_spraying_number_of_times_passes_per_year"
+]
+
+
+cat_cols = [ # These are binary
+    "bare_soil"
+    , "annual_cover_crop"
[…]

### 2023-08-23 analysis.py
Added all xgboost models!!!

@@ -82,76 +82,28 @@ cat_cols = [ # These are binary
     , "giregion"
 ]
 
-
+# We need to declare each column that is categorical
+# as a categorical column!
+#
+# later these need to be one hot encoded.
 for column in cat_cols:
     df[column] = pd.Categorical(df[column])
 
-y_name = "bare_soil"
-categoric = True
+for y_name in cols:
+    model, scores = train_model(
+        df[cols]
+        , y_name
+        , df[cols].dtypes.loc[y_name] == "category"
+        , val_size=.1/.9)
+    
 
-model = train_model(df[cols], y_name, categoric)
 
-def train_model(df: pd.DataFrame, y_name: str, categoric: bool, test_size=0.1, val_size=0):
-    """Trains an XGBoosted Tree given a y variable and dataframe. """
-    # X, y setup
-    # Currently giregion is in both the predicted and predictors! You gotta fix that
-    X = df.drop([y_name], axis=1)
-    X = pd.get_dummies(X)
-    # Make sure that the y variable is not in the X variable or you will have a lot of trouble!!
 
-    # we add a random number for baselining importance
-
-    #X["random"] = np.random.normal(0, 1, len(X))
-    # i = 0
-    # giregion_lookup = {}
-    # for giregion in df["giregion"].unique():
-    #     giregion_lookup[giregion] = i
-    #     i += 1
-    # df["giregion"] = df["giregion"].replace(giregion_lookup)
-
-    # split into train test val
-    y = df[y_name]
-    X_train, X_test, y_train, y_test \
-        = train_test_split(X, y, test_size=test_size, random_state=1)
-
-    if val_size != 0:
-        X_train, X_val, y_train, y_val \
-            = train_test_split(X_train, y_train, test_size=val_size, random_state=1) 
-    else:
-        y_val = 0
-        X_val = 0
-
-    dtrain = xgb.DMatrix(X_train, label=y_train.apply(int), enable_categorical=categoric)
-    dtest = xgb.DMatrix(X_test, label=y_test, enable_categorical=categoric)
-    if val_size != 0:
-        dval = xgb.DMatrix(X_val, label=y_val, enable_categorical=categoric)
-
-    # setup model
-
-    param = {"colsample_bytree": 0.3, "learning_rate": 0.00025, "objective": "binary:logistic"}
-
-    param["eval_metric"] = "auc"
-
-    evallist = [(dtrain, "train"), (dtest, "eval")]
-
-    num_round = 100
-
[…]

### 2023-08-23 analysis.py
all models run!

@@ -1,104 +1,13 @@
 import pandas as pd
-import numpy as np
 import xgboost as xgb
-from xgboost import cv
-from sklearn.model_selection import train_test_split, KFold, cross_val_score
-from sklearn.metrics import mean_squared_error
-import matplotlib.pyplot as plt
-from data import bin_col
+from sklearn.model_selection import train_test_split, cross_val_score, RepeatedKFold
+from sklearn.metrics import r2_score
 from sklearn.inspection import permutation_importance
 from sklearn.metrics import accuracy_score
 
 import random
 
 random.seed(100)
-
-################################################################################
-# Setup
-df = pd.read_csv("dfb.csv")
-
-cols = ["tonnes_grapes_harvested"
-    , "area_harvested"
-    , "water_used"
-    , "total_tractor_passes"
-    , "total_fertiliser"
-    , "synthetic_nitrogen_applied"
-    , "organic_nitrogen_applied"
-    , "synthetic_fertiliser_applied"
-    , "organic_fertiliser_applied"
-    , "giregion"
-    , "data_year_id"
-    , "river_water"
-    , "groundwater"
-    , "surface_water_dam"
-    , "recycled_water_from_other_source"
-    , "mains_water"
-    , "other_water"
-    , "water_applied_for_frost_control"
-    , "bare_soil"
-    , "annual_cover_crop"
-    , "permanent_cover_crop_native"
-    , "permanent_cover_crop_non_native"
-    , "permanent_cover_crop_volunteer_sward"
-    , "diesel_vineyard"
-    , "electricity_vineyard"
-    , "petrol_vineyard"
-    , "vineyard_solar"
-    , "vineyard_wind"
-    , "lpg_vineyard"
-    , "biodiesel_vineyard"
-    , "slashing_number_of_times_passes_per_year"
-    , "fungicide_spraying_number_of_times_passes_per_year"
-    , "herbicide_spraying_number_of_times_passes_per_year"
-    , "insecticide_spraying_number_of_times_passes_per_year"
-]
-
-
-cat_cols = [ # These are binary
-    "bare_soil"
-    , "annual_cover_crop"
-    , "permanent_cover_crop_native"
-    , "permanent_cover_crop_non_native"
-    , "permanent_cover_crop_volunteer_sward"
-    , "irrigation_energy_diesel"
-    , "irrigation_energy_electricity"
-    , "irrigation_energy_pressure"
[…]

### 2023-08-23 analysis.py
added saving

@@ -34,8 +34,8 @@ def kfold_val(model: xgb.sklearn.XGBClassifier
 
     accuracy = accuracy_score(y_val, predictions)
 
-    print("validation data prediction:")
-    print("Accuracy: {0:.2%}".format(accuracy))
+    # print("validation data prediction:")
+    # print("Accuracy: {0:.2%}".format(accuracy))
 
     # conf with val data
     predictions = [round(value) for value in y_pred]
@@ -43,8 +43,9 @@ def kfold_val(model: xgb.sklearn.XGBClassifier
     conf = pd.DataFrame({"actual": y_val.values, "predicted": predictions})
     conf["actual"] = conf["actual"].apply(int)
 
-    print("confusion matrix for validation set")
-    print(conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0))
+    # print("confusion matrix for validation set")
+    # print(conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0))
+    val = conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0)
 
     # confusion matrix
     y_pred = model.predict(X)
@@ -53,8 +54,11 @@ def kfold_val(model: xgb.sklearn.XGBClassifier
     conf = pd.DataFrame({"actual": y.values, "predicted": predictions})
     conf["actual"] = conf["actual"].apply(int)
 
-    print("confusion matrix for all data")
-    print(conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0))
+    # print("confusion matrix for all data")
+    # print(conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0))
+    all = conf.groupby(["actual", "predicted"]).size().unstack(fill_value=0)
+
+    return {"accuracy": accuracy, "all": all, "val": val}
 
 # Feature importance
 # feature_importances = rf_gridsearch.best_estimator_.feature_importances_
@@ -154,7 +158,10 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     feature_loss["loss"] = model.feature_importances_
     feature_loss.index = model.feature_names_in_
 
-    return model, scores
+    model.save_model("{}.json".format(y_name))
+    pd.DataFrame(scores).to_csv("{}.csv".format(y_name))
+
+    return
 
 # binary class
 def train_model_b(df: pd.DataFrame, y_name: str):
[…]

### 2023-08-23 analysis.py
added loss saVING FUNCTION

@@ -118,7 +118,10 @@ def kfold_val(model: xgb.sklearn.XGBClassifier
 # Training scripts #
 ###################
 
-# Regressor
+#############
+# Regressor #
+#############
+
 def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     """Trains an XGBoosted Regression Tree given a y variable and dataframe. """
     # X, y setup
@@ -157,9 +160,10 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     feature_loss = pd.DataFrame()
     feature_loss["loss"] = model.feature_importances_
     feature_loss.index = model.feature_names_in_
+    feature_loss.tocsv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
-    pd.DataFrame(scores).to_csv("{}.csv".format(y_name))
+    pd.DataFrame(scores).to_csv("{}_scores.csv".format(y_name))
 
     return
 
@@ -194,13 +198,15 @@ def train_model_b(df: pd.DataFrame, y_name: str):
         , verbose = True
     )
 
-    validation = kfold_val(model, X, y, X_val, y_val)
-
+    feature_loss = pd.DataFrame()
+    feature_loss["loss"] = model.feature_importances_
+    feature_loss.index = model.feature_names_in_
+    feature_loss.tocsv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
 
-    with open(y_name, 'w') as f:
-        f.writelines(validation["accuracy"])
+    validation = kfold_val(model, X, y, X_val, y_val)
+    validation["accuracy"].tofile("{}_accuracy".format(y_name))
     validation["val"].to_csv("{}_val.csv".format(y_name))
     validation["all"].to_csv("{}_all.csv".format(y_name))
 
@@ -250,12 +256,15 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
         , verbose = True
     )
 
-    validation = kfold_val(model, X, y, X_val, y_val)
+    feature_loss = pd.DataFrame()
+    feature_loss["loss"] = model.feature_importances_
+    feature_loss.index = model.feature_names_in_
+    feature_loss.tocsv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
 
-    with open(y_name, 'w') as f:
-        f.writelines(validation["accuracy"])
[…]

### 2023-08-23 analysis.py
changed to rmse metric, fixed loss save

@@ -140,7 +140,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
         , importance_type="gain"
         , learning_rate=0.00025
         , objective="reg:squarederror"
-        , eval_metric=r2_score
+        , eval_metric="rmse"
     )
 
     model.fit(X_train
@@ -160,7 +160,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     feature_loss = pd.DataFrame()
     feature_loss["loss"] = model.feature_importances_
     feature_loss.index = model.feature_names_in_
-    feature_loss.tocsv("{}_loss.csv".format(y_name))
+    feature_loss.to_csv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
     pd.DataFrame(scores).to_csv("{}_scores.csv".format(y_name))
@@ -201,7 +201,7 @@ def train_model_b(df: pd.DataFrame, y_name: str):
     feature_loss = pd.DataFrame()
     feature_loss["loss"] = model.feature_importances_
     feature_loss.index = model.feature_names_in_
-    feature_loss.tocsv("{}_loss.csv".format(y_name))
+    feature_loss.to_csv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
 
@@ -259,7 +259,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     feature_loss = pd.DataFrame()
     feature_loss["loss"] = model.feature_importances_
     feature_loss.index = model.feature_names_in_
-    feature_loss.tocsv("{}_loss.csv".format(y_name))
+    feature_loss.to_csv("{}_loss.csv".format(y_name))
 
     model.save_model("{}.json".format(y_name))
 

### 2023-08-23 analysis.py
fixed saving!

@@ -1,3 +1,4 @@
+from numpy import savetxt
 import pandas as pd
 import xgboost as xgb
 from sklearn.model_selection import train_test_split, cross_val_score, RepeatedKFold
@@ -206,7 +207,9 @@ def train_model_b(df: pd.DataFrame, y_name: str):
     model.save_model("{}.json".format(y_name))
 
     validation = kfold_val(model, X, y, X_val, y_val)
-    validation["accuracy"].tofile("{}_accuracy".format(y_name))
+    f = open("{}_accuracy.csv".format(y_name), "w")
+    f.write("{}".format(validation["accuracy"]))
+    f.close()
     validation["val"].to_csv("{}_val.csv".format(y_name))
     validation["all"].to_csv("{}_all.csv".format(y_name))
 
@@ -264,7 +267,9 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     model.save_model("{}.json".format(y_name))
 
     validation = kfold_val(model, X, y, X_val, y_val)
-    validation["accuracy"].tofile("{}_accuracy".format(y_name))
+    f = open("{}_accuracy.csv".format(y_name), "w")
+    f.write("{}".format(validation["accuracy"]))
+    f.close()
     validation["val"].to_csv("{}_val.csv".format(y_name))
     validation["all"].to_csv("{}_all.csv".format(y_name))
 

### 2023-08-24 analysis.py
added chord diagram

@@ -317,6 +317,17 @@ cols = ["tonnes_grapes_harvested"
     , "fungicide_spraying_number_of_times_passes_per_year"
     , "herbicide_spraying_number_of_times_passes_per_year"
     , "insecticide_spraying_number_of_times_passes_per_year"
+    , "irrigation_energy_diesel"
+    , "irrigation_energy_electricity"
+    , "irrigation_energy_pressure"
+    , "irrigation_energy_solar"
+    , "irrigation_type_dripper"
+    , "irrigation_type_flood"
+    , "irrigation_type_non_irrigated"
+    , "irrigation_type_overhead_sprinkler"
+    , "irrigation_type_undervine_sprinkler"
+    , "nh_disease"
+    , "nh_frost"
 ]
 
 # These are the columns that will be classes!

### 2023-08-24 chord_diagram.py

@@ -1,0 +1,57 @@
+from os import listdir
+from os.path import isfile, join
+
+import pandas as pd
+import re
+
+cat_cols = [ # These are binary
+    "bare_soil"
+    , "annual_cover_crop"
+    , "permanent_cover_crop_native"
+    , "permanent_cover_crop_non_native"
+    , "permanent_cover_crop_volunteer_sward"
+    , "irrigation_energy_diesel"
+    , "irrigation_energy_electricity"
+    , "irrigation_energy_pressure"
+    , "irrigation_energy_solar"
+    , "irrigation_type_dripper"
+    , "irrigation_type_flood"
+    , "irrigation_type_non_irrigated"
+    , "irrigation_type_overhead_sprinkler"
+    , "irrigation_type_undervine_sprinkler"
+    , 'river_water'
+    , 'groundwater'
+    , 'surface_water_dam'
+    , 'recycled_water_from_other_source'
+    , 'mains_water'
+    , 'other_water'
+    , 'water_applied_for_frost_control'
+    , "nh_frost"
+    , "nh_disease"
+    , "data_year_id" # These are one hot encoded
+    , "giregion"
+]
+
+files = [f for f in listdir("./") if isfile(join("./", f))]
+
+r = re.compile(".*loss")
+files = list(filter(r.match, files))
+
+df = pd.read_csv(files[0], index_col=0, names=[files[0][:-9]], header=0)
+
+for filename in files[1:]:
+    df[filename[:-9]] = pd.read_csv(files[0], index_col=0, header=0)
+
+df = df[df.index[:3]].iloc[:3]
+
+for col in list(set(cat_cols) - set(df.index)):
+    r = re.compile("{}.*".format(col))
+    rows = list(filter(r.match, list(df.index)))
+    df.loc[col] = df.loc[rows].mean()
+    df = df.drop(rows).copy()
+
+df = df.stack().reset_index()
+df.columns = ["source", "links", "value"]
+
+df["value"] = round(df["value"]*1000000000).apply(int)
+

### 2023-08-24 analysis.py
set first training parameters

@@ -136,7 +136,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
         = train_test_split(X, y, test_size=test_size, random_state=1)
 
     model = xgb.XGBRegressor(
-         n_estimators=10
+         n_estimators=10000
         , scale_pos_weight=45
         , importance_type="gain"
         , learning_rate=0.00025
@@ -185,7 +185,7 @@ def train_model_b(df: pd.DataFrame, y_name: str):
     # We define the measure of objective for the different types of variables.
 
     model = xgb.XGBClassifier(
-         n_estimators=10
+         n_estimators=10000
         , scale_pos_weight=45
         , importance_type="gain"
         , objective="binary:logistic"
@@ -245,7 +245,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
         = train_test_split(X_train, y_train, test_size=0.1/0.9, stratify=y_train) 
 
     model = xgb.XGBClassifier(
-            n_estimators=10
+            n_estimators=10000
         , importance_type="gain"
         , objective="multi:softmax"
         , learning_rate=0.00025

### 2023-08-24 analysis.py
set kfold values

@@ -152,7 +152,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
         , verbose = True
     )
 
-    cv = RepeatedKFold(n_splits=10, n_repeats=3, random_state=1)
+    cv = RepeatedKFold(n_splits=10, n_repeats=10, random_state=1)
     scores = cross_val_score(model, X, y, scoring='r2', cv=cv, n_jobs=-1)
     # scores = cross_val_score(model, X.drop("tonnes_grapes_harvested", axis=1), y, scoring='r2', cv=cv, n_jobs=-1)
     print(scores)

### 2023-08-24 analysis.r
set folds to 10

@@ -48,8 +48,8 @@ selected_cols <- c("giregion"
 
 trControl <- trainControl(
     method = "repeatedcv"
-    , number = 2
-    , repeats = 2
+    , number = 10
+    , repeats = 10
     , verboseIter = TRUE
 )
 

### 2023-08-24 analysis.py
Looks for missing models then trains them.

@@ -1,4 +1,6 @@
 from numpy import savetxt
+from os import listdir
+from os.path import isfile, join
 import pandas as pd
 import xgboost as xgb
 from sklearn.model_selection import train_test_split, cross_val_score, RepeatedKFold
@@ -6,6 +8,7 @@ from sklearn.metrics import r2_score
 from sklearn.inspection import permutation_importance
 from sklearn.metrics import accuracy_score
 
+import re
 import random
 
 random.seed(100)
@@ -367,7 +370,13 @@ cat_cols = [ # These are binary
 for column in cat_cols:
     df[column] = pd.Categorical(df[column])
 
-for y_name in cols:
+files = [f for f in listdir("./") if isfile(join("./", f))]
+r = re.compile(".*json")
+files = list(filter(r.match, files))
+files = [file[:-5] for file in files]
+files = list(set(cols) - set(files))
+
+for y_name in files:
     if y_name in cat_cols:
         if y_name in ["data_year_id", "giregion"]:
             train_model_multi(df[cols]

### 2023-08-24 article.tex
methods save

@@
 Each region describing unique reputations, qualities and varietals of wine produced there.
-While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \autocite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
@@
 
-\subsubsection{A figure sub-subsection}
-
 \if{false}
@@
 The Australian wine industry is divided into 65 regions, known as a Geographical Indicator Regions (GI Region).
-Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties \autocite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \autocite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 The climatic properties of a GI Region are summarised by Sustainable Winegrowing Australia (2021), where regions of similar climates are amalgamated together into superset regions.
@@
 
+\subsection{XGBoosted Trees}
+XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+They were chosen for this analysis as they provide a both high predictive performance and ability to effectively capture complex relationships.
+A separate XGBoosted tree was used to predict each variable.
+As variables were both continuous, binary and multiclass, separate functions were created to handle the three types of variables.
+
+%
+With judicious choices for , we may express a variety of tasks, such as regression, classification, and ranking.
+The task of training the model amounts to finding the best parameters that best fit the training data and labels
+
+. In order to train the model, we need to define the objective function to measure how well the model fit the training data.
+
+A salient characteristic of objective functions is that they consist of two parts: training loss and regularization term:
+$obj(\theta)=L(\theta)+\Omega(\theta)$
+
+where $L$ is the training loss function, and $\Omega$ is the regularization term.
+The training loss measures how predictive our model is with respect to the training data.
+A common choice of $L$ is the mean squared error, which is given by
+
+$L(\theta)=\sum_i (y_i - \hat{y_i})^2$
+
+Another commonly used loss function is logistic loss, to be used for logistic regression:
+
+The regularization term is what people usually forget to add.
+The regularization term controls the complexity of the model, which helps us to avoid overfitting.
+This sounds a bit abstract, so let us consider the following problem in the following picture.
+You are asked to fit visually a step function given the input data points on the upper left corner of the image.
+Which solution among the three do you think is the best fit?
+
+% For a given data set with n examples and m features
+% D = {(xi, yi)} (|D| = n, xi ∈ Rm, yi ∈ R), a tree ensem-
+% ble model (shown in Fig. 1) uses K additive functions to
+% predict the output.
+
+\[ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F} \]
+
+%
+
+XGBoosted Regression trees were used to predict continuous variables. With data being split into 80\% training data and 20\% testing data.
+
+XGBoosted classification trees were used to classify the binary and multiclass variables. Data was split into 80\% training, 10\% testing and 10\% validation data.
+
+The modelled relationships are able to be scrutinised by using techniques such as feature importance analysis.
+The use of the XGBoost library also incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
+The further use of cross validated grid search functions allowed for the selection of better performing hyper-parameters when selecting the final model.
+
 \subsection{Classification Trees}
@@
 comparing these relationships to those linked to grape quality.
-This was done using the rparts and caret packages \autocite{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language (R Core Team, 2021).
+This was done using the rparts and caret packages \citep{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language (R Core Team, 2021).
 \newline
@@
 The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area.
-The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \autocite{reynoldsManagingWineQuality2010}.
+The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \citep{reynoldsManagingWineQuality2010}.
 % interesting although not really appropriate for the results section.
@@
 The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
-Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
+Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
 It is a bit unclear what you are really trying to cite and its relevance.
@@
 Better to talk about how this could improve and the weaknesses than to start spouting unsupported claims.
- Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \autocite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
+ Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
  % Woah woah woah
@@
 % What are these business decisions? Can you give some examples?
- It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \autocite{leileiheFruitYieldPrediction2022}.
+ It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \citep{leileiheFruitYieldPrediction2022}.
 % It is difficult yes, but please at least try!

### 2023-08-24 analysis.py
changed to check for loss.csv instead. commented out saving models

@@ -166,7 +166,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     feature_loss.index = model.feature_names_in_
     feature_loss.to_csv("{}_loss.csv".format(y_name))
 
-    model.save_model("{}.json".format(y_name))
+    # model.save_model("{}.json".format(y_name))
     pd.DataFrame(scores).to_csv("{}_scores.csv".format(y_name))
 
     return
@@ -207,7 +207,7 @@ def train_model_b(df: pd.DataFrame, y_name: str):
     feature_loss.index = model.feature_names_in_
     feature_loss.to_csv("{}_loss.csv".format(y_name))
 
-    model.save_model("{}.json".format(y_name))
+    # model.save_model("{}.json".format(y_name))
 
     validation = kfold_val(model, X, y, X_val, y_val)
     f = open("{}_accuracy.csv".format(y_name), "w")
@@ -267,7 +267,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     feature_loss.index = model.feature_names_in_
     feature_loss.to_csv("{}_loss.csv".format(y_name))
 
-    model.save_model("{}.json".format(y_name))
+    # model.save_model("{}.json".format(y_name))
 
     validation = kfold_val(model, X, y, X_val, y_val)
     f = open("{}_accuracy.csv".format(y_name), "w")
@@ -371,9 +371,9 @@ for column in cat_cols:
     df[column] = pd.Categorical(df[column])
 
 files = [f for f in listdir("./") if isfile(join("./", f))]
-r = re.compile(".*json")
+r = re.compile(".*_loss.csv")
 files = list(filter(r.match, files))
-files = [file[:-5] for file in files]
+files = [file[:-9] for file in files]
 files = list(set(cols) - set(files))
 
 for y_name in files:

### 2023-08-25 analysis.py
Grid search and f-score importance

@@ -3,7 +3,7 @@ from os import listdir
 from os.path import isfile, join
 import pandas as pd
 import xgboost as xgb
-from sklearn.model_selection import train_test_split, cross_val_score, RepeatedKFold
+from sklearn.model_selection import train_test_split, cross_val_score, RepeatedKFold, GridSearchCV
 from sklearn.metrics import r2_score
 from sklearn.inspection import permutation_importance
 from sklearn.metrics import accuracy_score
@@ -132,22 +132,35 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     # Currently giregion is in both the predicted and predictors! You gotta fix that
     X = df.drop([y_name], axis=1)
     X = pd.get_dummies(X)
-    
+
     y = df[y_name]
 
     X_train, X_test, y_train, y_test \
         = train_test_split(X, y, test_size=test_size, random_state=1)
 
     model = xgb.XGBRegressor(
-         n_estimators=10000
-        , scale_pos_weight=45
-        , importance_type="gain"
-        , learning_rate=0.00025
+        scale_pos_weight=45
+        , importance_type="weight"
         , objective="reg:squarederror"
         , eval_metric="rmse"
     )
 
-    model.fit(X_train
+    parameters = {
+        'max_depth': range (2, 10, 1),
+        'n_estimators': range(60, 220, 40),
+        'learning_rate': [0.1, 0.01, 0.05]
+    }
+
+    grid_search = GridSearchCV(
+        estimator=model,
+        param_grid=parameters,
+        scoring = 'roc_auc',
+        n_jobs = 10,
+        cv = 10,
+        verbose=True
+    )
+
+    grid_search.fit(X_train
         , y_train
         #, num_boost_round = num_round
         #, nfold = 5
@@ -156,15 +169,16 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     )
 
     cv = RepeatedKFold(n_splits=10, n_repeats=10, random_state=1)
-    scores = cross_val_score(model, X, y, scoring='r2', cv=cv, n_jobs=-1)
+    scores = cross_val_score(grid_search.best_estimator_, X, y, scoring='r2', cv=cv, n_jobs=-1)
     # scores = cross_val_score(model, X.drop("tonnes_grapes_harvested", axis=1), y, scoring='r2', cv=cv, n_jobs=-1)
     print(scores)
     # we want the validation
     # we want the loss
-    feature_loss = pd.DataFrame()
[…]

### 2023-08-27 analysis.py
added predictor variables and multiclassified binary variables

@@ -162,8 +162,6 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
 
     grid_search.fit(X_train
         , y_train
-        #, num_boost_round = num_round
-        #, nfold = 5
         , eval_set=[(X_test, y_test)]
         , verbose = True
     )
@@ -174,7 +172,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     print(scores)
     # we want the validation
     # we want the loss
-    grid_search.best_estimator_.get_booster().get_score(importance_type="weight").values()
+
     importance = pd.DataFrame()
     importance["values"] = grid_search.best_estimator_.get_booster().get_score(importance_type="weight").values()
     importance.index = grid_search.best_estimator_.get_booster().get_score(importance_type="weight").keys()
@@ -202,28 +200,41 @@ def train_model_b(df: pd.DataFrame, y_name: str):
     # We define the measure of objective for the different types of variables.
 
     model = xgb.XGBClassifier(
-         n_estimators=4
-        , scale_pos_weight=45
-        , importance_type="gain"
+         scale_pos_weight=45
+        , importance_type="weight"
         , objective="binary:logistic"
-        , learning_rate=0.00025
         , eval_metric="auc"
         )
 
-    model.fit(X_train
+    parameters = {
+        'max_depth': range (2, 10, 1),
+        'n_estimators': range(60, 220, 40),
+        'learning_rate': [0.1, 0.01, 0.05]
+    }
+
+    grid_search = GridSearchCV(
+        estimator=model,
+        param_grid=parameters,
+        scoring = 'roc_auc',
+        n_jobs = 10,
+        cv = 10,
+        verbose=True
+    )
+
+    grid_search.fit(X_train
         , y_train
         , eval_set=[(X_test, y_test)]
         , verbose = True
     )
 
-    feature_loss = pd.DataFrame()
-    feature_loss["loss"] = model.feature_importances_
-    feature_loss.index = model.feature_names_in_
-    feature_loss.to_csv("{}_loss.csv".format(y_name))
+    importance = pd.DataFrame()
+    importance["values"] = grid_search.best_estimator_.get_booster().get_score(importance_type="weight").values()
+    importance.index = grid_search.best_estimator_.get_booster().get_score(importance_type="weight").keys()
[…]

### 2023-08-27 article.tex
chord diagram and method edits

@@
 They were chosen for this analysis as they provide a both high predictive performance and ability to effectively capture complex relationships.
-A separate XGBoosted tree was used to predict each variable.
-As variables were both continuous, binary and multiclass, separate functions were created to handle the three types of variables.
+Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees use a given set of data, to predict $y_i$ from the input $x_i$.
+The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
 
-%
-With judicious choices for , we may express a variety of tasks, such as regression, classification, and ranking.
+\begin{equation} \label{eqn:objective}
+  \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F}
+\end{equation}
+
+Each function $f_K$ is a classification or regression tree, such that all functions are in defined in the set $\mathcal{F}$ of trees given by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
+Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights.
+Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf.
+The final prediction is determined by the sum of the score of the corresponding leaves, given  by $\omega$.
+The set of functions used by the tree is determined by minimising the regularised objective function, given by:
+
+\begin{equation}
+ \mathcal{L}(\phi)=\sum_i l(\hat{y_i}, y_i^{t-1} + f_t(x_i)) + \sum_k \Omega (f_K)
+\end{equation}
+
+The difference between the prediction and actual variable is a convex loss function $l$.
+To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
+The function $f_t$ is selected according to which value minimises \ref{eqn:objective}.
+The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
+As predictions are made using additive tree functions, it can be used for both classification and regression.
+For this analysis both classification and regression trees were used.
+The major difference between the types of trees created was the objective function.
+As variables were both continuous, binary and multiclass, three different objective functions were used, root mean squared error, binary:logistic and the soft max functions respectively.
+
+\par
+Chen and Guestrin \citep{chenXGBoostScalableTree2016} further illustrate, using Taylor expansions how, for a fixed structure $q(x)$ the the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
+Furthermore they show how to successfully enumarate tree structures using left $I_L$ and right $I_R$ instance sets of nodes and letting $I = I_L \union I_R$.
+The loss reduction after the split is given by the function:
+
+\begin{equation}
+\mathcal{L}_{split} = {{1}\over{2}} \left[ 
+  {{(\sum_{i\in I_L}g_i)^{2}}\over{\sum_{i\in I_L} h_i + \lambda}} +
+  {{(\sum_{i\in I_R}g_i)^{2}}\over{\sum_{i\in I_R} h_i + \lambda}} -
+  {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma
+\end{equation}
+
+#
+# Connect variable frequency in trees to loss to the objective function.
+#
+
+This means we greedily add the ft that most improves our
+model according to Eq. (2). Second-order approximation
+can be used to quickly optimize the objective in the general
+setting [12].
+
+A common example is a linear model, where the prediction is given as $y_i = \Sigma_j \theta_j x_{ij}$, a linear combination of weighted input features.
+The prediction value can have different interpretations, depending on the task, i.e., regression or classification.
+For example, it can be logistic transformed to get the probability of positive class in logistic regression, and it can also be used as a ranking score when we want to rank the outputs.
+
+The parameters are the undetermined part that we need to learn from data.
+In linear regression problems, the parameters are the coefficients $\theta$.
+Usually we will use $\theta$ to denote the parameters (there are many parameters in a model, our definition here is sloppy).
+
+With judicious choices for $y_i$, we may express a variety of tasks, such as regression, classification, and ranking.
 The task of training the model amounts to finding the best parameters that best fit the training data and labels
@@
 % predict the output.
-
-\[ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F} \]
 

### 2023-08-27 chord_diagram.py

@@ -34,7 +34,7 @@ cat_cols = [ # These are binary
 
 files = [f for f in listdir("./") if isfile(join("./", f))]
 
-r = re.compile(".*loss")
+r = re.compile(".*imp")
 files = list(filter(r.match, files))
 
 df = pd.read_csv(files[0], index_col=0, names=[files[0][:-9]], header=0)

### 2023-08-27 analysis.py
removed dummy functions

@@ -481,15 +481,6 @@ cat_cols = [ # These are binary
     , "giregion"
 ]
 
-def train_model_b(df, y_name):
-    print(y_name)
-
-def train_model_multi(df, y_name):
-    print(y_name)
-
-def train_model_reg(df, y_name):
-    print(y_name)
-
 for y_name in files:
     if y_name in cat_cols:
         if y_name in ["nh_disease"]:

### 2023-08-27 analysis.py
Changed the population sample to those who reported profit

@@ -495,11 +495,11 @@ for y_name in files:
 
 # We also do the predicted variables!
 
-train_model_b(df[cols + ["profitable"]]
+train_model_b(df[df["profitable"].notnull()][cols + ["profitable"]]
                 , "profitable")
 
-train_model_reg(df[cols+["profit"]]
+train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
         , "profit")
 
-train_model_reg(df[cols+["total_operating_costs"]]
+train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
                 , "total_operating_costs")

### 2023-08-27 analysis.py
changed gridsearch metrics to be compatible

@@ -154,7 +154,7 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
     grid_search = GridSearchCV(
         estimator=model,
         param_grid=parameters,
-        scoring = 'roc_auc',
+        scoring = 'neg_root_mean_squared_error',
         n_jobs = 10,
         cv = 10,
         verbose=True
@@ -286,7 +286,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     grid_search = GridSearchCV(
         estimator=model,
         param_grid=parameters,
-        scoring = 'roc_auc',
+        scoring = 'roc_auc_ovr',
         n_jobs = 10,
         cv = 10,
         verbose=True

### 2023-08-27 analysis.py
fixed model fit error

@@ -294,7 +294,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
 
     # Currently there is no eval metric for multi_class classifiers
     # So it trains off the soft max.
-    grid_search.model.fit(X_train
+    grid_search.fit(X_train
         , y_train
         , eval_set=[(X_test, y_test)]
         , verbose = True

### 2023-08-27 analysis.py
only using total fertiliser

@@ -325,10 +325,10 @@ cols = ["tonnes_grapes_harvested"
     , "water_used"
     , "total_tractor_passes"
     , "total_fertiliser"
-    , "synthetic_nitrogen_applied"
-    , "organic_nitrogen_applied"
-    , "synthetic_fertiliser_applied"
-    , "organic_fertiliser_applied"
+    # , "synthetic_nitrogen_applied"
+    # , "organic_nitrogen_applied"
+    # , "synthetic_fertiliser_applied"
+    # , "organic_fertiliser_applied"
     , "giregion"
     , "data_year_id"
     , "river_water"

### 2023-08-27 article.tex
xgboost edits

@@
 XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-They were chosen for this analysis as they provide a both high predictive performance and ability to effectively capture complex relationships.
-Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees use a given set of data, to predict $y_i$ from the input $x_i$.
+They were chosen for this analysis as they provide both a high predictive performance and ability to effectively capture complex relationships.
+Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees predict a value $y_i$ from the input $x_i$.
 The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
@@
 
-Each function $f_K$ is a classification or regression tree, such that all functions are in defined in the set $\mathcal{F}$ of trees given by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
+Each function $f_K$ is a classification or regression tree, such that all functions are in the set $\mathcal{F}$ of trees given by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
 Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights.
@@
 Chen and Guestrin \citep{chenXGBoostScalableTree2016} further illustrate, using Taylor expansions how, for a fixed structure $q(x)$ the the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
-Furthermore they show how to successfully enumarate tree structures using left $I_L$ and right $I_R$ instance sets of nodes and letting $I = I_L \union I_R$.
+Furthermore they show how to successfully enumarate tree structures using left $I_L$ and right $I_R$ instance sets of nodes and letting $I = I_L \cup I_R$.
 The loss reduction after the split is given by the function:
@@
 
-#
-# Connect variable frequency in trees to loss to the objective function.
-#
+%
+% Connect variable frequency in trees to loss to the objective function.
+%
 

### 2023-08-28 analysis.py
fixed comments, refactored class validation. func

@@ -27,12 +27,12 @@ random.seed(100)
 
 # validation results
 
-def kfold_val(model: xgb.sklearn.XGBClassifier
+def class_val(model: xgb.sklearn.XGBClassifier
               , X
               , y
               , X_val: pd.DataFrame
               , y_val: pd.DataFrame):
-    """Kfold validation of xgboost model given its data."""
+    """Validation of classification model."""
     y_pred = model.predict(X_val)
     predictions = [round(value) for value in y_pred]
 
@@ -183,7 +183,10 @@ def train_model_reg(df: pd.DataFrame, y_name: str, test_size=0.2):
 
     return
 
-# binary class
+################
+# binary class #
+################
+
 def train_model_b(df: pd.DataFrame, y_name: str):
     """Trains an XGBoosted Tree given a y variable and dataframe. """
     X = df.drop([y_name], axis=1)
@@ -234,7 +237,7 @@ def train_model_b(df: pd.DataFrame, y_name: str):
 
     # model.save_model("{}.json".format(y_name))
 
-    validation = kfold_val(grid_search.best_estimator_, X, y, X_val, y_val)
+    validation = class_val(grid_search.best_estimator_, X, y, X_val, y_val)
     f = open("{}_accuracy.csv".format(y_name), "w")
     f.write("{}".format(validation["accuracy"]))
     f.close()
@@ -307,7 +310,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
 
     # model.save_model("{}.json".format(y_name))
 
-    validation = kfold_val(grid_search.best_estimator_, X, y, X_val, y_val)
+    validation = class_val(grid_search.best_estimator_, X, y, X_val, y_val)
     f = open("{}_accuracy.csv".format(y_name), "w")
     f.write("{}".format(validation["accuracy"]))
     f.close()

### 2023-08-28 article.tex
methods XGBoost section

--- 2023-08-24 article.tex
@@
 XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-They were chosen for this analysis as they provide a both high predictive performance and ability to effectively capture complex relationships.
-A separate XGBoosted tree was used to predict each variable.
-As variables were both continuous, binary and multiclass, separate functions were created to handle the three types of variables.
+They were chosen for this analysis as they provide both a high predictive performance and ability to effectively capture complex relationships.
+An XGBoosted tree was created for each variable to show how they interacted.
+Each tree included all but the economic variables (profit and operating cost), which were only included once as predicted variables.
+\par
+Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees predict a value $y_i$ from the input $x_i$.
+The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
 
-%
-With judicious choices for , we may express a variety of tasks, such as regression, classification, and ranking.
-The task of training the model amounts to finding the best parameters that best fit the training data and labels
+\begin{equation}
+  \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F},
+\end{equation}
 
-. In order to train the model, we need to define the objective function to measure how well the model fit the training data.
+where each function $f_K$ is a classification or regression tree, such that all functions are in the set of all decision trees $\mathcal{F}$, defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
+Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights.
+Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf.
+The final prediction is determined by the sum of the score of the corresponding leaves, given  by $\omega$.
+The set of functions used by the tree is determined by minimising the regularised objective function, given by:
 
-A salient characteristic of objective functions is that they consist of two parts: training loss and regularization term:
-$obj(\theta)=L(\theta)+\Omega(\theta)$
+\begin{equation}\label{eqn:objective}
+ \mathcal{L}(\phi)=\sum_i l(\hat{y_i}, y_i^{t-1} + f_t(x_i)) + \sum_k \Omega (f_K).
+\end{equation}
 
-where $L$ is the training loss function, and $\Omega$ is the regularization term.
-The training loss measures how predictive our model is with respect to the training data.
-A common choice of $L$ is the mean squared error, which is given by
+The difference between the prediction and actual variable is a convex loss function $l$.
+To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
+The function $f_t$ is selected according to which value minimises (\ref{eqn:objective}).
+The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
+\par
+As predictions are made using additive tree functions, XGboosted trees can be used for classification and regression.
+Due to the mixture of continuous, binary and multiclass variables in this analysis, both classification and regression trees were created.
+The difference between the trees created for this analysis was the objective function used.
+XGBoosted regression trees were created for continuous variables, using the root-mean-square as the objective function.
+Binary class variables utilised the logistic loss function as the objective.
+And, Multiclass variable used the soft max function.
+All objective functions are defined within the SKlearn library \citep{sklearn_api}, linked via an API to the XGBoost library \citep{chenXGBoostScalableTree2016}.
+\par
+Chen and Guestrin \citep{chenXGBoostScalableTree2016} further illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
+Furthermore, they show the loss reduction after the split is given by the function:
 
-$L(\theta)=\sum_i (y_i - \hat{y_i})^2$
+\begin{equation} \label{eqn:enumerate}
+\mathcal{L}_{split} = {{1}\over{2}} \left[ 
+  {{(\sum_{i\in I_L}g_i)^{2}}\over{\sum_{i\in I_L} h_i + \lambda}} +
+  {{(\sum_{i\in I_R}g_i)^{2}}\over{\sum_{i\in I_R} h_i + \lambda}} -
+  {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma,
+\end{equation}
 
-Another commonly used loss function is logistic loss, to be used for logistic regression:
-
-The regularization term is what people usually forget to add.
-The regularization term controls the complexity of the model, which helps us to avoid overfitting.
-This sounds a bit abstract, so let us consider the following problem in the following picture.
-You are asked to fit visually a step function given the input data points on the upper left corner of the image.
-Which solution among the three do you think is the best fit?
-
-% For a given data set with n examples and m features
-% D = {(xi, yi)} (|D| = n, xi ∈ Rm, yi ∈ R), a tree ensem-
-% ble model (shown in Fig. 1) uses K additive functions to
-% predict the output.
-
-\[ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F} \]
-
-%
-
-XGBoosted Regression trees were used to predict continuous variables. With data being split into 80\% training data and 20\% testing data.
-
-XGBoosted classification trees were used to classify the binary and multiclass variables. Data was split into 80\% training, 10\% testing and 10\% validation data.
-
-The modelled relationships are able to be scrutinised by using techniques such as feature importance analysis.
-The use of the XGBoost library also incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
-The further use of cross validated grid search functions allowed for the selection of better performing hyper-parameters when selecting the final model.
+with the tree structure is defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
+Instead of enumerating all possible tree structures, a greedy algorithm iteratively adds branches to the tree minimising $ \mathcal{L}_{split} $ in (\ref{eqn:enumerate}).
+The frequency of a variable's occurrence within a tree is directly attributed to the minimisation of the objective function (or loss) through the minimisation of  $ \mathcal{L}_{split} $.
+\par
+The frequency of a variable appearing as a node within the ensemble was used as a measure of importance.
+This measure was chosen as it connected a variable to the minimisation of its associated objective function, translating the value into a simple count metric.
+Creating XGBoosted trees for each variable allowed the use of importance to show how strongly variables were associated with each other.
+The importance of predictor variables to economic variables was illustrated through the use of Sankey diagrams constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+Other variable's interconnectedness was demonstrated through the use of a chord diagram also created using Holoviews.
+\par
+Each variable utilised 80\% of the data to train the XGBoost ensemble, with 20\% reserved for testing and validation.
+Testing was done through the iterative minimisation of the respective objective function for the variables type.
+For continuous variables 20\% was used as testing data, minimising the root-mean-square function.
+The final model was validated using repeated k-fold cross validation for 10 folds, repeated 10 times.
+ For  binary and multiclass variables data was split into 80\% training, 10\% testing and 10\% validation data.
+Due to class disparity in multiclass variables (most prominently in region) data was stratified into each subset at the same ratio of class occurrence.
+Validation was summarised through confusion matrices and their associated accuracy
+\par
+The use of the XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
+The further use of cross validated grid search functions allowed for the selection of better performing hyperparameters when selecting the final model.
+The performance measure for model selection was root-mean-square error for continuous variables.
+The receiver operator characteristic's area under the curve was used for category variables; with multiclass variables utilising the one verse the rest approach \citep{hanley1982meaning}.
 

### 2023-08-28 analysis.py
changed to ovo grid search

@@ -289,7 +289,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     grid_search = GridSearchCV(
         estimator=model,
         param_grid=parameters,
-        scoring = 'roc_auc_ovr',
+        scoring = 'roc_auc_ovo',
         n_jobs = 10,
         cv = 10,
         verbose=True

### 2023-08-28 analysis.py
set for fixing multiclass grid search.

@@ -479,12 +479,12 @@ cat_cols = [ # These are binary
     , "cover_crop"
     , "irrigation_type"
     , "irrigation_energy"
-    , "nh_disease"
+    # , "nh_disease"
     , "data_year_id" # These are one hot encoded
     , "giregion"
 ]
 
-for y_name in files:
+for y_name in cat_cols:
     if y_name in cat_cols:
         if y_name in ["nh_disease"]:
             train_model_b(df[cols]

### 2023-08-28 article.tex
add ovo to methods.

@@
 The performance measure for model selection was root-mean-square error for continuous variables.
-The receiver operator characteristic's area under the curve was used for category variables; with multiclass variables utilising the one verse the rest approach \citep{hanley1982meaning}.
+The receiver operator characteristic's area under the curve was used for category variables.
+Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{hanley1982meaning}.
 

### 2023-08-28 article.tex
ovo references

--- 2023-08-28 article.tex
@@
 The performance measure for model selection was root-mean-square error for continuous variables.
-The receiver operator characteristic's area under the curve was used for category variables; with multiclass variables utilising the one verse the rest approach \citep{hanley1982meaning}.
+The receiver operator characteristic's area under the curve was used for category variables \citep{hanley1982meaning}.
+Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
 

### 2023-08-28 article.tex
CARTS

@@
 
-\subsection{Classification Trees}
+\subsection{Classification and Regression Trees}
 
-Classification Trees were developed to discern the different practices within regions and climates, 
-% Were they? Or was it to discern regional differences and then compare them?
-comparing these relationships to those linked to grape quality.
-This was done using the rparts and caret packages \citep{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language (R Core Team, 2021).
-\newline
-Three classifications were undertaken for region, climate and grape quality.
-% Is this three classifications for each of these, being 9 in total? Or three overall, one for each?
-Climate was further classified into two subcategories of rainfall and temperature,
-% Why was this done?
-resulting in a total of 5 classification trees being created. % You say this too late!
-Classification trees were validated using K-fold cross validation.% This could be its own section!
+Classification and Regression Trees were created for region, year, profit and operating cost.
+These models describe the partitions that are useful in predicting these variables; giving insight into the trees that make up the ensembles created by XGBoost.
+These trees were created using the rparts and caret packages \citep{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
+\par
+Classification trees were validated using K-fold cross validation.
 Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the  classification trees.
 A summary confusion matrix was then constructed to show the class bias and overall accuracy of each tree.
- % To what end? Why do we want them to be accurate?? What does this analysis achieve?
-
-% Was the analysis validated?
-% How where the carts compared??
-% I still do not know what the models really where
 

### 2023-08-28 analysis.py
fixed error with ovo and features being missed.

@@ -255,7 +255,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     
     # We need at least 3 entries into a region for it to be able to be classified.
 
-    X = df[df.groupby(y_name)[y_name].transform('count')>6].copy() 
+    X = df[df.groupby(y_name)[y_name].transform('count')>10].copy() 
 
     y = X[y_name]
     y = y.cat.remove_categories(list(set(y.unique().categories) - set(y.unique())))
@@ -326,7 +326,7 @@ df = pd.read_csv("dfb.csv")
 cols = ["tonnes_grapes_harvested"
     , "area_harvested"
     , "water_used"
-    , "total_tractor_passes"
+    # , "total_tractor_passes"
     , "total_fertiliser"
     # , "synthetic_nitrogen_applied"
     # , "organic_nitrogen_applied"
@@ -419,8 +419,9 @@ irrigation_type = [
 
 for col in irrigation_type:
     df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col[16:])}).copy()
-df["irrigation_type"] = df[irrigation_type].astype(str).sum(axis=1)
+df["irrigation_type"] = pd.Categorical(df[irrigation_type].astype(str).sum(axis=1))
 cols = list(set(cols) - set(irrigation_type))
+cols.append("irrigation_type")
 
 ####################
 
@@ -433,8 +434,9 @@ irrigation_energy = [
 
 for col in irrigation_energy:
     df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col[18:])}).copy()
-df["irrigation_energy"] = df[irrigation_energy].astype(str).sum(axis=1)
+df["irrigation_energy"] = pd.Categorical(df[irrigation_energy].astype(str).sum(axis=1))
 cols = list(set(cols) - set(irrigation_energy))
+cols.append("irrigation_energy")
 
 ####################
 
@@ -447,8 +449,9 @@ cover_crops = [
 
 for col in cover_crops:
     df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col[12:])}).copy()
-df["cover_crops"] = df[cover_crops].astype(str).sum(axis=1)
+df["cover_crops"] = pd.Categorical(df[cover_crops].astype(str).sum(axis=1))
 cols = list(set(cols) - set(cover_crops))
+cols.append("cover_crops")
 
 ####################
 
@@ -463,8 +466,9 @@ water_type = [
 
[…]

### 2023-08-28 analysis.py
changed overwrite of files to skip for training.

@@ -488,8 +488,8 @@ cat_cols = [ # These are binary
 ]
 
 # cols += cat_cols
-print(cols)
-for y_name in cols:
+print(files)
+for y_name in files:
     print(y_name)
     if y_name in cat_cols:
         if y_name in ["nh_disease", "nh_frost"]:

### 2023-08-29 analysis.py
removed variables that were not used.

@@ -350,7 +350,7 @@ cols = ["tonnes_grapes_harvested"
     , "electricity_vineyard"
     , "petrol_vineyard"
     , "vineyard_solar"
-    , "vineyard_wind"
+    # , "vineyard_wind"
     , "lpg_vineyard"
     , "biodiesel_vineyard"
     , "slashing_number_of_times_passes_per_year"
@@ -367,6 +367,7 @@ cols = ["tonnes_grapes_harvested"
     , "irrigation_type_overhead_sprinkler"
     , "irrigation_type_undervine_sprinkler"
     , "nh_disease"
+    # , "nh_frost"
 ]
 
 # These are the columns that will be classes!
@@ -393,7 +394,6 @@ cat_cols = [ # These are binary
     , 'mains_water'
     , 'other_water'
     , 'water_applied_for_frost_control'
-    , "nh_frost"
     , "nh_disease"
     , "data_year_id" # These are one hot encoded
     , "giregion"

### 2023-08-29 analysis.r
changed script to include years and appropriate variables

@@ -3,53 +3,25 @@ library(mlbench)
 library(caret)
 library(rpart.plot)
 
-df <- read.csv("dfb.csv")
+# The csv file used is the state of the dataframe used in
+#  the python analysis. After variables are made multiclass and the limit of 10 classes is used.
+df <- read.csv("dftree.csv")
 df$profitable <- factor(df$profitable)
-
-selected_cols <- c("giregion"
-    , "profitable"
-    , "profit"
-    , "total_operating_costs"
-    , "tonnes_grapes_harvested"
-    , "area_harvested"
-    , "water_used"
-    , "total_tractor_passes"
-    , "total_fertiliser"
-    , "synthetic_nitrogen_applied"
-    , "organic_nitrogen_applied"
-    , "synthetic_fertiliser_applied"
-    , "organic_fertiliser_applied"
-    , "river_water"
-    , "groundwater"
-    , "surface_water_dam"
-    , "recycled_water_from_other_source"
-    , "mains_water"
-    , "other_water"
-    , "water_applied_for_frost_control"
-    , "bare_soil"
-    , "annual_cover_crop"
-    , "permanent_cover_crop_native"
-    , "permanent_cover_crop_non_native"
-    , "permanent_cover_crop_volunteer_sward"
-    , "diesel_vineyard"
-    , "electricity_vineyard"
-    , "petrol_vineyard"
-    , "vineyard_solar"
-    , "vineyard_wind"
-    , "lpg_vineyard"
-    , "biodiesel_vineyard"
-    , "slashing_number_of_times_passes_per_year"
-    , "fungicide_spraying_number_of_times_passes_per_year"
-    , "herbicide_spraying_number_of_times_passes_per_year"
-    , "insecticide_spraying_number_of_times_passes_per_year"
-)
+df$water_type <- factor(df$water_type)
+df$cover_crops <- factor(df$cover_crops)
+df$irrigation_type <- factor(df$irrigation_type)
+df$irrigation_energy <- factor(df$irrigation_energy)
+df$data_year_id <- factor(df$data_year_id)
+df$giregion <- factor(df$giregion)
+df$nh_disease <- factor(df$nh_disease)
+# df$nh_frost <- factor(df$nh_frost)
 
 # Training function
 
 trControl <- trainControl(
     method = "repeatedcv"
-    , number = 10
-    , repeats = 10
+    , number = 2
+    , repeats = 2
     , verboseIter = TRUE
 )
 
[…]

### 2023-08-29 article.tex
added variable summary table

@@
 \subsection{Data}
+
+% 6091 samples - total
+% 58 regions
+% 10 years
+
+% 853 with reported profits
+% 46 regions
+% reported since 2018
+Data used in this analysis were obtained from Sustainable Winegrowing Australia.
+Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data recorded by SWA is entered manually by winegrowers using a web based interface.
+A total of 6091 observations has been collected over 10 years.
+Each observation contained 23 variables see Table \ref{tab:vars}.
+
+\begin{table}[]
+  \caption{Summary of variables used in the analysis. The recorded column indicate values that were either greater than zero or that were not missing.}\label{tab:summary}
+\begin{tabular}{@{}cccc@{}}
+  \toprule
+  \textbf{} & \textbf{Units} & \textbf{Recorded} & \textbf{\begin{tabular}[c]{@{}c@{}}Number of\\ Classes\end{tabular}} \\ \midrule
+  Water Used & Mega Litres & 5846 &  \\
+  Diesel & Litres & 5585 &  \\
+  Biodiesel & Litres & 25 &  \\
+  LPG & Litres & 958 &  \\
+  Herbicide Spray & Times per year & 2026 &  \\
+  Year & Class & 6091 & 10 \\
+  Disease & Class & 6091 & 2 \\
+  Region & Class & 6091 & 58 \\
+  Solar & Kilowatt Hours & 622 &  \\
+  Wind & Kilowatt Hours & 2 &  \\
+  Irrigation Type & Class & 6091 & 20 \\
+  Petrol & Litres & 4309 &  \\
+  Slashing & Times per year & 2290 &  \\
+  Yield & Tonnes & 5935 &  \\
+  Irrigation Energy & Class & 6091 & 16 \\
+  Area Harvested & Hectares & 6091 &  \\
+  Electricity & Kilowatt Hours & 1015 &  \\
+  Insecticide Spray & Times per year & 1092 &  \\
+  Fertiliser & \begin{tabular}[c]{@{}c@{}}Kilograms \\ of Nitrogen\end{tabular} & 795 &  \\
+  Fungicide Spray & Times per year & 2260 &  \\
+  Cover Crop & Class & 6091 & 32 \\
+  Water Type & Class & 6091 & 39 \\
+  Profit & AUD & 853 &  \\
+  Operating Costs & AUD & 853 &  \\ \bottomrule
+  \end{tabular}
+
+ were collected
+
+The data collected pertains to both the physical and
+
 The Australian wine industry is divided into 65 regions, known as a Geographical Indicator Regions (GI Region).

### 2023-08-30 analysis.r
fixed plots script (prettified)

@@ -6,22 +6,21 @@ library(rpart.plot)
 # The csv file used is the state of the dataframe used in
 #  the python analysis. After variables are made multiclass and the limit of 10 classes is used.
 df <- read.csv("dftree.csv")
-df$profitable <- factor(df$profitable)
-df$water_type <- factor(df$water_type)
-df$cover_crops <- factor(df$cover_crops)
-df$irrigation_type <- factor(df$irrigation_type)
-df$irrigation_energy <- factor(df$irrigation_energy)
-df$data_year_id <- factor(df$data_year_id)
-df$giregion <- factor(df$giregion)
-df$nh_disease <- factor(df$nh_disease)
-# df$nh_frost <- factor(df$nh_frost)
+df[, "Profitable"] <- factor(df[, "Profitable"])
+df[, "Water.Source"] <- factor(df[, "Water.Source"])
+df[, "Cover.Crop.Type"] <- factor(df[, "Cover.Crop.Type"])
+df[, "Irrigation.System.Type"] <- factor(df[, "Irrigation.System.Type"])
+df[, "Irrigation.Energy.Type"] <- factor(df[, "Irrigation.Energy.Type"])
+df[, "Year"] <- factor(df[, "Year"])
+df[, "Region"] <- factor(df[, "Region"])
+df[, "Disease.Presence"] <- factor(df[, "Disease.Presence"])
 
 # Training function
 
 trControl <- trainControl(
     method = "repeatedcv"
-    , number = 2
-    , repeats = 2
+    , number = 10
+    , repeats = 10
     , verboseIter = TRUE
 )
 
@@ -29,12 +28,12 @@ trControl <- trainControl(
 #       GI region         #
 ###########################
 
-giregion_tree <- train(giregion ~ .
+model_tree <- train(Region ~ .
     , data = df[, colnames(df)[
         !(colnames(df) %in% c(
-            "profit",
-            "profitable",
-            "total_operating_costs"))]]
+            "Profit",
+            "Profitable",
+            "Operating Costs"))]]
     , method = "rpart"
     , metric = "Accuracy"
     , trControl = trControl
@@ -42,22 +41,29 @@ giregion_tree <- train(giregion ~ .
     , tuneLength = 10
 )
 
-rpart.plot(giregion_tree$finalModel
+tree_plot <- rpart.plot(model_tree$finalModel
     , extra = 102
     , legend.x = -100
[…]

### 2023-08-30 analysis.r
fixed figures

@@ -41,14 +41,13 @@ model_tree <- train(Region ~ .
     , tuneLength = 10
 )
 
-tree_plot <- rpart.plot(model_tree$finalModel
+pdf("region.pdf")
+rpart.plot(model_tree$finalModel
     , extra = 102
     , legend.x = -100
     , box.palette = "auto"
 )
 
-pdf("region.pdf")
-rpart.plot(tree_plot)
 dev.off()
 
 sink("region.txt")
@@ -71,14 +70,12 @@ model_tree <- train(Year ~ .
     , tuneLength = 10
 )
 
-tree_plot <- rpart.plot(model_tree$finalModel
+pdf("year.pdf")
+rpart.plot(model_tree$finalModel
     , extra = 102
     , legend.x = -100
     , box.palette = "auto"
 )
-
-pdf("year.pdf")
-rpart.plot(tree_plot)
 dev.off()
 
 
@@ -101,14 +98,12 @@ model_tree <- train(Profitable ~ .
     , tuneLength = 10
 )
 
-tree_plot <- rpart.plot(model_tree$finalModel
+pdf("profitable.pdf")
+rpart.plot(model_tree$finalModel
     , extra = 102
     , legend.x = -100
     , box.palette = "auto"
 )
-
-pdf("profitable.pdf")
-rpart.plot(tree_plot)
 dev.off()
 
 sink("profitable.txt")
@@ -134,13 +129,11 @@ model_tree <- train(Profit ~ .
 # Note that each terminating node is a regression model
 # - the percentage of observations in the node
 
-tree_plot <- rpart.plot(model_tree$finalModel
+pdf("profit.pdf")
+rpart.plot(model_tree$finalModel
     , legend.x = -100
     , box.palette = "auto"
 )
-
-pdf("profit.pdf")
-rpart.plot(tree_plot)
 dev.off()
 
 sink("profit.txt")
@@ -166,13 +159,11 @@ model_tree <- train(Operating.Costs ~ .
 # Note that each terminating node is a regression model
 # - the percentage of observations in the node
 
-tree_plot <- rpart.plot(model_tree$finalModel
+pdf("operating_costs.pdf")
+rpart.plot(model_tree$finalModel
     , legend.x = -100
     , box.palette = "auto"
 )
-
-pdf("operating_costs.pdf")
-rpart.plot(tree_plot)
 dev.off()
 
 sink("operating_costs.txt")

### 2023-08-30 article.tex
Results include figures (looking good!)

--- 2023-08-28 article.tex
@@
 \subsection{Data}
-The Australian wine industry is divided into 65 regions, known as a Geographical Indicator Regions (GI Region).
-Each GI Region is used to describe different unique localised traits of vineyards across Australia; with each having its own mixture of climatic and geophysical properties \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+
+% 6091 samples - total
+% 58 regions
+% 10 years
+
+% 853 with reported profits
+% 46 regions
+% reported since 2018
+Data used in this analysis were obtained from Sustainable Winegrowing Australia.
+Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data recorded by the SWA is entered manually by winegrowers using a web based interface tool.
+A total of 6091 observations were collected from 2012 to 2022.
+Each observation contained 23 variables reflecting a vineyards account for the given year (see Table \ref{tab:vars}).
+\par
+\begin{table}[] 
+  \label{tab:vars}
+  \caption{Summary of variables used in the analysis. The recorded column indicate values that were either greater than zero or that were not missing.}\label{tab:summary}
+  \begin{tabular}{@{}ccccl@{}}
+    \toprule
+    \textbf{Variable} & \textbf{Units} & \textbf{Recorded} & \textbf{\begin{tabular}[c]{@{}c@{}}Number of\\ Classes\end{tabular}} &  \\ \midrule
+    Water Used & Mega Litres & 5846 &  &  \\
+    Diesel & Litres & 5585 &  &  \\
+    Biodiesel & Litres & 25 &  &  \\
+    LPG & Litres & 958 &  &  \\
+    Herbicide Spray & Times per year & 2026 &  &  \\
+    Year & Class & 6091 & 10 &  \\
+    Disease & Class & 6091 & 2 &  \\
+    Region & Class & 6091 & 58 &  \\
+    Solar & Kilowatt Hours & 622 &  &  \\
+    Irrigation Type & Class & 6091 & 20 &  \\
+    Petrol & Litres & 4309 &  &  \\
+    Slashing & Times per year & 2290 &  &  \\
+    Yield & Tonnes & 5935 &  &  \\
+    Irrigation Energy & Class & 6091 & 16 &  \\
+    Area Harvested & Hectares & 6091 &  &  \\
+    Electricity & Kilowatt Hours & 1015 &  &  \\
+    Insecticide Spray & Times per year & 1092 &  &  \\
+    Fertiliser & \begin{tabular}[c]{@{}c@{}}Kilograms \\ of Nitrogen\end{tabular} & 795 &  &  \\
+    Fungicide Spray & Times per year & 2260 &  &  \\
+    Cover Crop & Class & 6091 & 32 &  \\
+    Water Type & Class & 6091 & 39 &  \\
+    Profit & AUD & 853 &  &  \\
+    Operating Costs & AUD & 853 &  &  \\ \bottomrule
+    \end{tabular}
+\end{table}
+
+The data originally contained only two multiclass variables: year and region.
+Variables that measured the same metric from different sources (such as water collected from rivers versus water from dams) were converted into multiclass variables representing the source.
+The total amount used from these variables was retained as a separate variable.
+Occurrences of multiple sources were defined as separate classes.
+As harvest does not run by calendar year, years are in financial years.
+Region represents one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-The climatic properties of a GI Region are summarised by Sustainable Winegrowing Australia (2021), where regions of similar climates are amalgamated together into superset regions.
-The climatic regions were utilised to illustrate similar trends and explain differences between sets of regions.
-The data used in this analysis comes from Sustainable Winegrowing Australia and covers the period 2015 to 2022. The dataset contained 3342 samples across 52 GI Regions and 1072 individual vineyards.
+Profit was also used as a binary variable, depicting whether a vineyard was profitable or not.
+\par
+%
+% Mention one-hot-encoding.
+%
 
-% how was it collected
-
+%
+% How deep into the classes do we go?
+%
 \subsection{XGBoosted Trees}
@@
 
-with the tree structure is defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
+with the tree structure defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
 Instead of enumerating all possible tree structures, a greedy algorithm iteratively adds branches to the tree minimising $ \mathcal{L}_{split} $ in (\ref{eqn:enumerate}).
@@
 
+\subsection{Region}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{region.pdf}}
+  \caption{Decision tree predicting Region.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:region_tree}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{region.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:region_sankey}
+\end{figure}
+
+\subsection{Year}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{year.pdf}}
+  \caption{Decision tree predicting Year.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:year_tree}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{year.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:year_sankey}
+\end{figure}
+
+\subsection{Operating Costs}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{operating_costs.pdf}}
+  \caption{Decision tree predicting Operating Costs.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_costs_tree}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{operating_costs.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:operating_costs_sankey}
+\end{figure}
+
+\subsection{Profit}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.pdf}}
+  \caption{Decision tree predicting Profit.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:profit_tree}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:profit_sankey}
+\end{figure}
+
+\subsection{Validation}
+
+\begin{table}[]
+  \caption{Validation and training accuracies of each multiclass variable.}
+  \begin{tabular}{@{}ccc@{}}
+  \toprule
+  \textbf{Variable} & \textbf{Validation} & \textbf{Training} \\ \midrule
+  \textbf{cover crops} & 0.364086 & 0.396418 \\
+  \textbf{water type} & 0.742097 & 0.928905 \\
+  \textbf{profitable} & 0.705882 & 0.719737 \\
+  \textbf{irrigation type} & 0.841845 & 0.847554 \\
+  \textbf{giregion} & 0.505824 & 0.568242 \\
+  \textbf{irrigation energy} & 0.746293 & 0.836405 \\
+  \textbf{data year id} & 0.422003 & 0.518059 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
 \subsection{Model 1 GI Regions}
+
 The first Model was used to classify GI regions % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
@@
 
-\begin{table}[]
-    \caption{Classification accuracy of the most prominent GI Regions.}
-    \label{tab:accuracy}
-    \resizebox{\textwidth}{!}{
-        \begin{tabular}{@{}llll@{}}
-            \toprule
-            \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
-            \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
-            \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
-            \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
-            \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
-            \textbf{Margaret River} & 78.82\% & 201 & 255 \\
-            \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
-            \textbf{Riverland} & 92.74\% & 345 &
-        \end{tabular}
-    }
-\end{table}
+% \begin{table}[]
+%     \caption{Classification accuracy of the most prominent GI Regions.}
+%     \label{tab:accuracy}
+%     \resizebox{\textwidth}{!}{
+%         \begin{tabular}{@{}llll@{}}
+%             \toprule
+%             \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
+%             \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
+%             \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
+%             \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
+%             \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
+%             \textbf{Margaret River} & 78.82\% & 201 & 255 \\
+%             \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
+%             \textbf{Riverland} & 92.74\% & 345 &
+%         \end{tabular}
+%     }
+% \end{table}
 

### 2023-08-30 article.tex
results additions

@@
 
+Region classification performed at 36.10\% and 56.82\% accuracy, for the classification tree and XGBoosted ensemble respectively.
+The most prominent features used to classify regions with the classification tree was water sources (see Figure \ref{fig:region_tree}).
+This differed from the variables that illustrated the greatest importance for the XGBoosted ensemble (see Figure (\ref{fig:region_sankey}), with predictor variables being highly interrelated in importance.
+Area, water, fuel and yield were more determining factors when predicting region using XGBoost.
+Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
+\par
+It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
+The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
+Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of other activities within the vineyard, such as pruning and harvesting.
+With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
+\par
+Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
+The regions with the most samples performed the best.
+Notably bordering regions were routinely grouped together and misclassified as the same region.
+Two areas taht suffered the most from this, specifically with the classification tree were the lime coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
+%, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
+The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
+These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
+Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
+
 \begin{figure}
@@
 \subsection{Year}
+
+The classification tree and XGBoosted ensemble performed similarly for classifying year with 41.15\% and 51.81\% respectively.
+Electricity and the type of irrigation were highly influential within the classification tree.
+Similarly, electricity was the most frequently occurring node in the XGBoost ensemble.
+However, other variables such as slashing passes, and fungicide and herbicide spraying were more prevalenct than in the classification tree.
+Weed and disease outbreaks are likely an influential factor when classifying different years, making the decisions to spray and slash unique factors that differ year to year.
+Climatic differences between years are likely tied to the influence of yield and water use.
+\par
+Over half of the interrelated importance of the predictor variables is dominated by area harvested, yield and slashing passes.
+Although all the predictor variables are highly connected, their relative importance is not as prominent as the three major variables.
+It is of particular note of the relative importance of slashing to area, fuel and yield; as these are not directly related activities.
+The connection between slashing and spraying is that those who do a set number of spraying or slashing passes tended to do that many passes for all slashing and spraying activities.
 
@@
 \section{Discussion}
+
+% %Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74% and 96.97% respectively. 
+These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
+However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
+The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
+
 The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.

### 2023-08-31 analysis.r
Operating costs was missing the "." instead of space

@@ -33,7 +33,7 @@ model_tree <- train(Region ~ .
         !(colnames(df) %in% c(
             "Profit",
             "Profitable",
-            "Operating Costs"))]]
+            "Operating.Costs"))]]
     , method = "rpart"
     , metric = "Accuracy"
     , trControl = trControl
@@ -62,7 +62,7 @@ model_tree <- train(Year ~ .
         !(colnames(df) %in% c(
             "Profit",
             "Profitable",
-            "Operating Costs"))]]
+            "Operating.Costs"))]]
     , method = "rpart"
     , metric = "Accuracy"
     , trControl = trControl
@@ -90,7 +90,7 @@ model_tree <- train(Profitable ~ .
     , data = df[, colnames(df)[
         !(colnames(df) %in% c(
             "Profit",
-            "Operating Costs"))]]
+            "Operating.Costs"))]]
     , method = "rpart"
     , metric = "Accuracy"
     , trControl = trControl
@@ -117,7 +117,7 @@ model_tree <- train(Profit ~ .
     , data = df[, colnames(df)[
         !(colnames(df) %in% c(
             "Profitable",
-            "Operating Costs"))]]
+            "Operating.Costs"))]]
     , method = "rpart"
     , trControl = trControl
     , na.action = na.omit

### 2023-08-31 analysis.r
removed white space

@@ -47,7 +47,6 @@ rpart.plot(model_tree$finalModel
     , legend.x = -100
     , box.palette = "auto"
 )
-
 dev.off()
 
 sink("region.txt")

### 2023-08-31 analysis.py
fixed categoryu naming convention for future sanity

@@ -448,7 +448,7 @@ cover_crops = [
     , "permanent_cover_crop_volunteer_sward"]
 
 for col in cover_crops:
-    df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col[12:])}).copy()
+    df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col)}).copy()
 df["cover_crops"] = pd.Categorical(df[cover_crops].astype(str).sum(axis=1))
 cols = list(set(cols) - set(cover_crops))
 cols.append("cover_crops")
@@ -465,7 +465,7 @@ water_type = [
     , 'water_applied_for_frost_control']
 
 for col in water_type:
-    df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col[11:])}).copy()
+    df[col] = df[col].cat.rename_categories({0: "", 1: "{} ".format(col)}).copy()
 df["water_type"] = pd.Categorical(df[water_type].astype(str).sum(axis=1))
 cols = list(set(cols) - set(water_type))
 cols.append("water_type")

### 2023-08-31 article.tex
discussion backup

@@
 \newline
-A vineyard's region predetermines several physical parameters, such as: climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+
 %
@@
 
-Region classification performed at 36.10\% and 56.82\% accuracy, for the classification tree and XGBoosted ensemble respectively.
+Region classification performed at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
 The most prominent features used to classify regions with the classification tree was water sources (see Figure \ref{fig:region_tree}).
@@
 \subsection{Year}
+0.35196134
+0.0628483907157039
 
-The classification tree and XGBoosted ensemble performed similarly for classifying year with 41.15\% and 51.81\% respectively.
+The classification tree and XGBoosted ensemble performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
 Electricity and the type of irrigation were highly influential within the classification tree.
@@
 
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{operating_costs.pdf}}
-  \caption{Decision tree predicting Operating Costs.
+There was a pronounced difference in accuracy between the regression tree and the XGBoost model when predicting Operating costs.
+With the regression tree achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) in its cross validation.
+The XGBoosted regression ensemble achieved an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+\par
+Within the XGBoost ensemble's nodes for operating costs (see figure \ref{fig:operating_costs_sankey}) fuel, water, area and yield occurred the most, similarly to region.
+Both diesel and petrol were of more relative importance (being ranked higher) in operating costs than water was compared with region.
+It is surprising that electricity, slashing and spraying was not more prominent in operating costs.
+However, Figure \ref{fig:year_sankey} shows that electricity, slashing and spraying are important variables in determining area and yield.
+Electricity in particular is used predominantly for irrigation and so is related largely to the size of vineyard.
+However, slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, as they are not scaled to any measure of size.
+This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
+
+%
+% Is there any point in including this if its so bad?
+%
+% \begin{figure}
+%   \resizebox{\textwidth}{!}{
+%   \includegraphics{operating_costs.pdf}}
+%   \caption{Decision tree predicting Operating Costs.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_costs_tree}
-\end{figure}
+% \end{figure}
 
@@
 \subsection{Profit}
+
+Predictions of profit performed poorly compared to operating costs with the regression tree having an $R^2$ of 0.1873 (with a standard deviation of 0.0522) and the XGBoosted ensemble achieving an $R^2$ of 0.2535 (with a standard deviation of 0.3126).
+The high standard deviation in the XGBoosted tree was a bias in more accurately predicting vineyards that made profit compared to those that lost money.With much higher $R^2$ values being achieved in k-folds containing only those that made profits (recording a maximum of 0.7634).
+% profit summary stats
+% max 3345000
+% min -3937878
+% std 396027.633119089
+% range 7282878
+\par
+There was a disparity of 66.63\% of vineyards recording a profit than those that did not.
+When predicting if a vineyard would be profitable or not the classification tree and XGBoosted ensemble did not perform considerably differently from this proportion.
+With the regression tree achieving an accuracy of 68.66\% (and a standard deviation of 0.01\%) and the XGBoost ensemble achieving 70.59\% accuracy (with a validation accuracy of 71.97\%).
+\par
+It was surprising that operating costs performed substantially better in $R^2$ compared to profit.
+Interestingly the important variables when attempting to determine profit were similar to those used to classify region (see Figure \ref{fig:profit_sankey}), with the exception water used.
+Both the regression tree and the XGBoosted ensemble used region, specifically the Hunter Valley.
+The regression tree also used Tasmania when determining profit.
+Both the Hunter valley and Tasmania are known for the production of high quality grapes used in export wines
+% ref
+. A major difference between region and profit was the importance given to water use, with water use being a more important variable in predicting region than profit.
 
@@
 \end{figure}
-
-\subsection{Validation}
 
@@
 
-\subsection{Model 1 GI Regions}
+%%%%%%%%%%%%%%%%%%%%%
+% Points to address
+% 
+\section{Discussion}
 
-The first Model was used to classify GI regions % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
-and resulted in an accuracy of 36.48\% across 52 classes.
-The most prominent features used to classify regions were the types of water resources available (see Figure 1).
-Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively.
-% it might be worth only talking about the regions that were well classifified - and then talk about their determining traits.
-These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
-However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
-%Summary statistics regarding these two regions would be a better point of comparison, for example:
-%
-% How much in size do these operational scales differ?
-% How much yield do they differ by?
-% What are the differences in rainfall etc
-% What makes them unique, different and similar?
-%
-The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
-% What are the specific differences? This is a really critical part of this paper
-Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
-% This is almost its own paragraph.
-Why was this case?
-is it important?
-what are the similarities and differences of regions that were poorly classified?
- The regions with the most samples performed the best (see Table 1).
-Notably bordering regions were routinely grouped together and misclassified as the same region, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
-The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
- % THis speaks highly to why regions were poorly classified.
-These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
-Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
+\subsection{Region}
 
-% \begin{table}[]
-%     \caption{Classification accuracy of the most prominent GI Regions.}
-%     \label{tab:accuracy}
-%     \resizebox{\textwidth}{!}{
-%         \begin{tabular}{@{}llll@{}}
-%             \toprule
-%             \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
-%             \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
-%             \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
-%             \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
-%             \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
-%             \textbf{Margaret River} & 78.82\% & 201 & 255 \\
-%             \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
-%             \textbf{Riverland} & 92.74\% & 345 &
-%         \end{tabular}
-%     }
-% \end{table}
+A vineyard's region predetermines several physical parameters, such as: climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The association between yield and region is demonstrated by its position as fourth most occurring variable within the nodes of the XGBoosted ensemble which determined region (see Figure \ref{fig:region_sankey}).
+The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019}.
+\par
+Regions with lower land costs are also warmer \citep{willchancellorMeasuringAustralianBroadacre2019}, which is known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
+Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
+In general warmer regions have been associated with lower yields due to their generally lower rainfall, which can be mitigated through applying excess water \citep{campsGrapeHarvestYield2012}.
+It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of lower quality grapes; reflected through the variables' importance of water use in the XGBoost ensemble.
+The practice of utilising larger quantities of water for inland Australian wine crops is partly reflected in the prior use of flood style irrigation to saturate soil \citep{bgcoombeGrapeBerryDevelopment2004}.
+This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
+This is possibly the connection between the presence of the Hunter Valley within the XGBoost ensemble that determined profit (see Figure \ref{fig:profit_sankey}).
+With this connection reflecting the restriction of possible strategies employable by winegrowers between different regions.
+\par
+In part some winegrowing strategies are restricted simply through access to water resources, being reflected through the region classification tree (see Figure \ref{fig:region_tree}).
+Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops compared with coastal regions.
+Similarly, the connection between region and fuel is likely an indicator of the level of infrastructure within the region.
+Where, the need to pressurise irrigation systems from river water or to generate power would require larger amount of diesel and petrol.
+\par
+Although less important, the variables cover crops, fungicide spraying and slashing are likely linked to broad environmental properties of regions.
+Rainfall being related to fungal growth and disease, as well as weeds.
+With cover crops being an effective and sustainable method to alleviate these issues \autocite{delpuechAdaptingCoverCrop2018}.
+It is difficult to extrapolate findings to these methods and the reason for their use due to the broad and varying definition of the regions.
+Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation, as it is too broad to fully capture a vineyards location and its influence on more granular variables.
+The reasoning for using approaches such as cover crops can be widely varying.
 
-\subsection{Climate}
-Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly.
-These categories were divided into 12 climatic classifications with 3 and 4 separate subsets for rainfall and temperature respectively.
-The decision tree behaved similarly and over classified climates with higher response rates.
-The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area.
-The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \citep{reynoldsManagingWineQuality2010}.
-% interesting although not really appropriate for the results section.
- This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
+A cover crop can help to increase soil water retention, reduce erosion, increase biodiversity and reduce weeds \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020,delpuechAdaptingCoverCrop2018}.
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
+A more granular definition of region may help to better discern the differences in practices, and the reason for employing them.
 
-% You really need to show every decision tree.
+\subsection{Year}
 
-\subsubsection{Rainfall}
-The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).
+This may be particularly important in rainfed areas, like in the study case, due to the lack of irrigation possibilities.
+The result would be a shortening of the ripening period, with harvest occurring during the period with high temperatures, which could have a negative impact on wine quality (Salazar Parra et al.
+2010; Duchne and Schneider 2005; Jones and Davis 2000) and yield (Mira de Ordua 2010; Iglesias et al.
+2010).
+Climate change in the future might move the north and south latitude boundaries of areas suitable for good quality wines (Schultz and Jones 2010), and could even lead to improvements in fruit production and quality in some areas (Olesen and Bindi 2002).
+However, other areas may be negatively affected by high temperatures and water stress due to a reduction in the amount of water available.
 
-\subsubsection{Temperature}
-The classification of GI Regions by their temperatures (see Figure 5) showed similarities to the other trees, with a heavy reliance on the types of water resources used as dominant predictors.
-The use of contractors % This is super important.
-Talk about this at lengthi nt the discussion.
-Find more literature on agriculture and the use of contractors as oppose to self managing issues such as weeds.
-was again used to differentiate between warm and cool regions, likely being due to disease pressure.
-The temperature classification tree was only a minor improvement over the regional classification tree, with an accuracy of 49.26\% as shown in the confusion matrix (see Table 4).
+% %% Climate change
 
-\subsection{Model 3 Grape Quality}
-The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
-had an accuracy of 55.72\% across 5 separate grades. There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2). % This is actually really interesting.
- The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
-Although not surprising the number of contractor tractor passes is new deciding factor due disease and pests reducing the potential quality of a crop.
-The prevalence of contractor use is greater in regions such as the Barossa Valley and the McLaren Vale, this could be due to the difference in operational scales, with larger sites being more likely to have ownership of their own equipment for weeding and spraying due to the cost benefit.
+There are several environmental concerns that affect viticulture, including loss of soil quality, lack of rain, hail, disease, fire, and frost; with climate change exacerbating these issues.
+In 2020, 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase (Canadell et al., 2021).
+In comparison to countrywide pressures such as drought, this damage made up only 3% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought (Wine Australia, 2020).
+Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
+For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
+Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
+The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
+When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
+Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
+Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
+The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
+Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
 
-\section{Discussion}
+% %% Pests Insects and weeds
+% %% Disease
+The winegrowing industry holds significant importance for Australia and its economy.
+There exists many challenges that the industry has to contend against, with disease from sources such as Mildew and Botrytis being a considerable one (Cole, 2010; Magarey et al., 1994).
+This analysis looks at the prediction of disease in crops across Australia, linking disease pressure and its potential mitigation through the use of sustainable practices.
+The dataset for this analysis includes multiple vineyard attributes such as water source types, cover crop extents, renewable energy use, and fuel and electricity use.
+A major consideration within these sustainable practices for this analysis was the use of cover cropping.
+Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
+The primary reason for employing cover crops is to increase water retention and reduce the presence of disease and weeds (Delpuech and Metay, 2018).
+Prior studies have placed an emphasis on optimisations of fungicide sprays through the development of Bayesian models to forecast disease risk (Lu et al., 2020).
+This analysis investigates at the synergy between spray strategies as a proxy of fuel use to different types of water sources and sustainable strategies with an emphasis on the use of cover crops; allowing for the modelling of interactions between the use of multiple strategies.
+With the need to balance yield, quality, and combat adversities such as disease; the creation of tools to inform decisions and assist growers with warnings of disease risks is becoming crucial in ensuring sustainable and profitable wine production  (Abbal et al., 2016).
+An interesting observation within the relative importance of variables computed through SHAP values and variable permutation importance, is the similarity between cover crops and the use of contractors to spray herbicide.
+The presence of a cover crop is known to help in reducing disease and weeds (Capello et al., 2019).
+Cover crops can further help to increase soil water retention, reducing erosion and water runoff in shallow soils (Capello et al., 2020), which could be the interaction between water and cover crop use in the SHAP variable interactions; where the interaction might be more indicative of how well established a cover crop is, as it would require greater water resources to maintain but potentially offer more protection in return (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
+A further important consideration with this comparison is that diesel vineyard encompasses all vineyard operation aside from irrigation, being a proxy for actions such as weeding and spraying.
+The disparity between the importance of vineyards being sprayed by owners compared to the use of contractors warrants further investigation; as it would be reasonable to presume similar importance between diesel as a proxy for a vineyard’s management of its own disease prevention and the hired preventative strategies of a third party.
+There is a possibility that the use of contractors may be a vector for disease spread through lax bio-security practices, but would require a rigorous study to inform the matter.
+There is the potential that the types of herbicides used also have a long term effect on crops, reducing the presence of microorganisms and soil health, making the area more prone to disease in the long run, becoming dependent on these sprays (Coll et al., 2011; Gosling and Shepherd, 2005).
+
+\subsection{Operating Costs}
+
+% Reasons for operational costs (connect to disease and pests)
+% %% Electricity use and reasons for (connect to year)
+% %% Fuel use and what it is used for
+The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
+An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
+Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
+It is important that the interaction of events such as erosion have on other considerations such as soil health.
+
+\subsection{Profit}
+
+% %% Address yield and how predictable it is
+% %% different sales markets and performance?
+% %% Reasons for profit?
+
+% %% Grape quality
+The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
+Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012).
+The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
+The most prominent being what readily available resources of each region were, particular the types of water available.
+Heavy water consumption is often linked to the mass production of grapes, where lower quality grapes are targeted in a quantity over quality strategy.
+These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
+Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield ((Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
+Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018)).
+There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues, where on-the-ground choices are influenced by other wine industry’s decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
+It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season (Leilei He et al., 2022).
+
+% %% connection to watrer resource availability
+% What are the water resources in the hunter valley and tasmania?
+
+% economic pressures
+Historically strong demands for Australian wine have helped to create a thriving industry, however recently sharp reductions in exports to mainland China due to significant deposit tariffs have caused  a decline of 19% in Australian export value during the 2021-2022 financial year (see Figure 1) (Wine Australia, 2022).
+The pressure brought on by the drop in export value has been exacerbated by loss of tourism and labour due to the COVID-19 pandemic, global freight crisis, war in Europe and rising inflation (Wine Australia, 2021a, 2020).
+These pressures within the wine industry have trickled down to vineyards, where winemakers retain unwanted wine, creating an oversupply of grapes.
+Currently the proposed strategy to soften pressures proposed by Wine Australia includes initiatives focused on market diversification, and sustainability to build resilience in the coming years.
+Figure 1:  The exports of Australian wine over time in Australian Dollars Free On Board, comparing exports between China and the rest of the world.
+This graphic is taken from the Wine Australia Annual Report of 2020-21(Wine Australia, 2022).
+
+%%%%%%%%%%%%%%%
 
@@
 
+\subsection{Model 1 GI Regions}
+
+The first Model was used to classify GI regions % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
+and resulted in an accuracy of 36.48\% across 52 classes.
+The most prominent features used to classify regions were the types of water resources available (see Figure 1).
+Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively.
+% it might be worth only talking about the regions that were well classifified - and then talk about their determining traits.
+These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
+However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
+%Summary statistics regarding these two regions would be a better point of comparison, for example:
+%
+% How much in size do these operational scales differ?
+% How much yield do they differ by?
+% What are the differences in rainfall etc
+% What makes them unique, different and similar?
+%
+The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
+% What are the specific differences? This is a really critical part of this paper
+Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
+% This is almost its own paragraph.
+Why was this case?
+is it important?
+what are the similarities and differences of regions that were poorly classified?
+ The regions with the most samples performed the best (see Table 1).
+Notably bordering regions were routinely grouped together and misclassified as the same region, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
+The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
+ % This speaks highly to why regions were poorly classified.
+These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
+Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
+
+% \begin{table}[]
+%     \caption{Classification accuracy of the most prominent GI Regions.}
+%     \label{tab:accuracy}
+%     \resizebox{\textwidth}{!}{
+%         \begin{tabular}{@{}llll@{}}
+%             \toprule
+%             \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
+%             \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
+%             \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
+%             \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
+%             \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
+%             \textbf{Margaret River} & 78.82\% & 201 & 255 \\
+%             \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
+%             \textbf{Riverland} & 92.74\% & 345 &
+%         \end{tabular}
+%     }
+% \end{table}
+
+\subsection{Climate}
+Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly.
+These categories were divided into 12 climatic classifications with 3 and 4 separate subsets for rainfall and temperature respectively.
+The decision tree behaved similarly and over classified climates with higher response rates.
+The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area.
+The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \citep{reynoldsManagingWineQuality2010}.
+% interesting although not really appropriate for the results section.
+ This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
+
+% You really need to show every decision tree.
+
+\subsubsection{Rainfall}
+The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).
+
+was again used to differentiate between warm and cool regions, likely being due to disease pressure. The temperature classification tree
+
+\subsection{Model 3 Grape Quality}
+The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
+had an accuracy of 55.72\% across 5 separate grades. There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2). % This is actually really interesting.
+ The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
+Although not surprising the number of contractor tractor passes is new deciding factor due disease and pests reducing the potential quality of a crop.
+The prevalence of contractor use is greater in regions such as the Barossa Valley and the McLaren Vale, this could be due to the difference in operational scales, with larger sites being more likely to have ownership of their own equipment for weeding and spraying due to the cost benefit.
+
 \section{Conclusion}

### 2023-08-31 article.tex
discussion and comments

--- 2023-08-30 article.tex
@@
 Data recorded by the SWA is entered manually by winegrowers using a web based interface tool.
-A total of 6091 observations were collected from 2012 to 2022.
-Each observation contained 23 variables reflecting a vineyards account for the given year (see Table \ref{tab:vars}).
+A total of 6091 observations were collected from 2012/2013 to 2021/2022 financial years.
+23 variables were used for each observation reflecting a vineyards account for the given year (see Table \ref{tab:vars}).
+The profit variable was additionally transformed and included as another separate variable, profitable; depicting whether a vineyard was profitable or not.
 \par
+
 \begin{table}[] 
@@
 Occurrences of multiple sources were defined as separate classes.
-As harvest does not run by calendar year, years are in financial years.
-Region represents one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+\par
+The variable region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-Profit was also used as a binary variable, depicting whether a vineyard was profitable or not.
 \par
@@
 An XGBoosted tree was created for each variable to show how they interacted.
-Each tree included all but the economic variables (profit and operating cost), which were only included once as predicted variables.
+Each tree included all but the economic variables (profit and operating cost), which were only included once as response variables.
 \par
@@
 The final model was validated using repeated k-fold cross validation for 10 folds, repeated 10 times.
+$R^2$ scores were used to determine the best regression models during validation.
 For binary and multiclass variables data was split into 80\% training, 10\% testing and 10\% validation data.
 Due to class disparity in multiclass variables (most prominently in region) data was stratified into each subset at the same ratio of class occurrence.
-Validation was summarised through confusion matrices and their associated accuracy
+Validation was summarised through the accuracy, the proportion of true negatives and positives.
 \par
@@
 \par
-Classification trees were validated using K-fold cross validation.
-Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the  classification trees.
-A summary confusion matrix was then constructed to show the class bias and overall accuracy of each tree.
+Decision trees were validated using K-fold cross validation.
+Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the decision trees.
+The same measure of accuracy as the XGBoosted trees was used for comparison.
 
@@
 
-Region classification performed at 36.10\% and 56.82\% accuracy, for the classification tree and XGBoosted ensemble respectively.
-The most prominent features used to classify regions with the classification tree was water sources (see Figure \ref{fig:region_tree}).
+Region classification performed at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
+The most prominent feature used to classify regions with the classification tree was water source (see Figure \ref{fig:region_tree}).
 This differed from the variables that illustrated the greatest importance for the XGBoosted ensemble (see Figure (\ref{fig:region_sankey}), with predictor variables being highly interrelated in importance.
@@
 \par
-Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
+Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
 The regions with the most samples performed the best.
 Notably bordering regions were routinely grouped together and misclassified as the same region.
-Two areas taht suffered the most from this, specifically with the classification tree were the lime coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
 The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
@@
 
-The classification tree and XGBoosted ensemble performed similarly for classifying year with 41.15\% and 51.81\% respectively.
+The classification tree and XGBoosted ensemble performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
 Electricity and the type of irrigation were highly influential within the classification tree.
@@
 
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{operating_costs.pdf}}
-  \caption{Decision tree predicting Operating Costs.
+There was a pronounced difference in accuracy between the regression tree and the XGBoost model when predicting Operating costs.
+With the regression tree achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) in its cross validation.
+The XGBoosted regression ensemble achieved an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+\par
+Within the XGBoost ensemble's nodes for operating costs (see figure \ref{fig:operating_costs_sankey}) fuel, water, area and yield occurred the most, similarly to region.
+Both diesel and petrol were of more relative importance (being ranked higher) in operating costs than water was compared with region.
+It is surprising that electricity, slashing and spraying was not more prominent in operating costs.
+However, Figure \ref{fig:year_sankey} shows that electricity, slashing and spraying are important variables in determining area and yield.
+Electricity in particular is used predominantly for irrigation and so is related largely to the size of a vineyard.
+However, slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
+This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
+
+%
+% Is there any point in including this if its so bad?
+%
+% \begin{figure}
+%  \resizebox{\textwidth}{!}{
+%  \includegraphics{operating_costs.pdf}}
+%  \caption{Decision tree predicting Operating Costs.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_costs_tree}
-\end{figure}
+% \end{figure}
 
@@
 \subsection{Profit}
+
+Predictions of profit performed poorly compared to operating costs with the regression tree having an $R^2$ of 0.1873 (with a standard deviation of 0.0522) and the XGBoosted ensemble achieving an $R^2$ of 0.2535 (with a standard deviation of 0.3126).
+The high standard deviation in the XGBoosted tree was a bias in more accurately predicting vineyards that made profit compared to those that lost money.With much higher $R^2$ values being achieved in k-folds containing only those that made profits (recording a maximum of 0.7634).
+% profit summary stats
+% max 3345000
+% min -3937878
+% std 396027.633119089
+% range 7282878
+\par
+There was a disparity of 66.63\% of vineyards recording a profit than those that did not.
+When predicting if a vineyard would be profitable or not the classification tree and XGBoosted ensemble did not perform considerably differently from this proportion.
+With the regression tree achieving an accuracy of 68.66\% (and a standard deviation of 0.01\%) and the XGBoost ensemble achieving 71.97\% accuracy (with a validation accuracy of 70.59\%).
+\par
+It was surprising that operating costs performed substantially better in $R^2$ compared to profit.
+Interestingly the important variables when attempting to determine profit were similar to those used to classify region (see Figure \ref{fig:profit_sankey}), with the exception of water used.
+Both the regression tree and the XGBoosted ensemble used region, specifically the Hunter Valley.
+The regression tree also used Tasmania when determining profit.
+Both the Hunter valley and Tasmania are known for the production of high quality grapes used in export wines \citep{wineaustraliaNationalVintageReport2022}.
+A major difference between region and profit was the importance given to water use, with water use being a more important variable in predicting region than profit.
 
@@
 
-\subsection{Validation}
+% \begin{table}[]
+%  \caption{Validation and training accuracies of each multiclass variable.}
+%  \begin{tabular}{@{}ccc@{}}
+%  \toprule
+%  \textbf{Variable} & \textbf{Validation} & \textbf{Training} \\ \midrule
+%  \textbf{cover crops} & 0.364086 & 0.396418 \\
+%  \textbf{water type} & 0.742097 & 0.928905 \\
+%  \textbf{profitable} & 0.705882 & 0.719737 \\
+%  \textbf{irrigation type} & 0.841845 & 0.847554 \\
+%  \textbf{giregion} & 0.505824 & 0.568242 \\
+%  \textbf{irrigation energy} & 0.746293 & 0.836405 \\
+%  \textbf{data year id} & 0.422003 & 0.518059 \\ \bottomrule
+%  \end{tabular}
+%  \end{table}
 
-\begin{table}[]
-  \caption{Validation and training accuracies of each multiclass variable.}
-  \begin{tabular}{@{}ccc@{}}
-  \toprule
-  \textbf{Variable} & \textbf{Validation} & \textbf{Training} \\ \midrule
-  \textbf{cover crops} & 0.364086 & 0.396418 \\
-  \textbf{water type} & 0.742097 & 0.928905 \\
-  \textbf{profitable} & 0.705882 & 0.719737 \\
-  \textbf{irrigation type} & 0.841845 & 0.847554 \\
-  \textbf{giregion} & 0.505824 & 0.568242 \\
-  \textbf{irrigation energy} & 0.746293 & 0.836405 \\
-  \textbf{data year id} & 0.422003 & 0.518059 \\ \bottomrule
-  \end{tabular}
-  \end{table}
+%%%%%%%%%%%%%%%%%%%%%
+% Points to address
+% 
+\section{Discussion}
 
-\subsection{Model 1 GI Regions}
+Several physical parameters such as climate, geology and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The association between yield and region is demonstrated by its position as fourth most occurring variable within the nodes of the XGBoosted ensemble which determined region (see Figure \ref{fig:region_sankey}).
+The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019}.
+\par
+Regions with lower land costs are also warmer \citep{willchancellorMeasuringAustralianBroadacre2019}, which is known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
+Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
+In general warmer regions have been associated with lower yields due to their generally lower rainfall, which can be mitigated through applying excess water \citep{campsGrapeHarvestYield2012}.
+It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of lower quality grapes; reflected through the variables' importance of water use in the XGBoost ensemble.
+The practice of utilising larger quantities of water for inland Australian wine crops is partly reflected in the prior use of flood style irrigation to saturate soil \citep{bgcoombeGrapeBerryDevelopment2004}.
+This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
+This is possibly the connection between the presence of the Hunter Valley within the XGBoost ensemble that determined profit (see Figure \ref{fig:profit_sankey}).
+With this connection reflecting the restriction of possible strategies employable by winegrowers between different regions.
+\par
+In part some winegrowing strategies are restricted simply through access to water resources, being reflected through the region classification tree (see Figure \ref{fig:region_tree}).
+Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops unlike most coastal regions.
+Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region.
+Where, the need to pressurise irrigation systems from river water or to generate power would require larger amounts of diesel and petrol.
+\par
+Operational costs showed similar importance across fuel, water and tractor use.
+The dominating factor of area likely played a large part in determining how costly a tractor pass would be, or in defining the ratio of water applied to the amount of vines.
+The node frequency was high for area but much lower in general across the other variables, which could indicate the need to be more circumstantial in determining operational costs.
+Although it was attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), it is likely yet more complicated still.
+An example of how interrelated operational costs can be, is the optimisation of tractor passes being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
+\par
+Decisions made on the ground have far-reaching effects and are difficult to completely capture.
+Greater tractor use as a preventative measure for occurrences such as disease, may incur higher operational costs but could be critical in preventing long term losses.
+With factors such as erosion and soil health being difficult to capture but also influenced by tractor use \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
+Although, performing well in $R^2$, the ability to predict operational costs is limited by the variables incorporated.
+Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions.
+Without fully capturing more granular activities such as the specifics of what fuel was used for, it is hard to determine what decisions specifically influence the operational costs.
+\par
+Although less important in the XGBoost ensembles for profit, the variables: cover crops, fungicide spraying and slashing are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
+Rainfall being related to fungal growth and disease, as well as weeds.
+With cover crops being an effective and sustainable method to alleviate these issues \citep{delpuechAdaptingCoverCrop2018}.
+It is difficult to extrapolate findings to these methods and the reason for their use due to the broad and varying definition of the regions.
+Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation, as it is too broad to fully capture a vineyards location and its influence on more granular variables.
+The reasoning for using approaches such as cover crops can be widely varying.
+Where, a cover crop may be employed to help increase soil water retention, reduce erosion, increase biodiversity and reduce weeds \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020,delpuechAdaptingCoverCrop2018}.
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
+A more granular definition of region may help to better discern the differences in practices, and the reason for employing them.
+More sophisticated models, specifically those that utilise expert opinion, may also help to capture and address the decision making process.
+An example is the optimisation of fungicide sprays using Bayesian models that forecast disease risk \citep{luDiseaseRiskForecasting2020}.
 
-The first Model was used to classify GI regions % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
-and resulted in an accuracy of 36.48\% across 52 classes.
+% In particular There is the potential that the types of herbicides used also have a long term effect on crops, reducing the presence of microorganisms and soil health, making the area more prone to disease in the long run, becoming dependent on these sprays (Coll et al., 2011; Gosling and Shepherd, 2005).
+
+% The classification tree shows a likely change in avaialble infrastructure across the years, with mroe current years being associated with pressurised water and more slashing and herbicide passes.
+
+% The availability of solar panels for electrical irrigation and electrical irrigation in general
+
+% Electricity is predominantly used for pumping and irrigation in vineyards \citep{longbottomRoleVineyardPractices2015}
+
+% Fungicides are likely associated with higher preventive sprays when disease breaks out.
+
+% Slashing passes are likely weed outbreaks or yearly cycles?
+
+% %% Climate change
+
+% %% Pests Insects and weeds
+% %% Disease
+
+% %% Electricity use and reasons for (connect to year)
+% %% Fuel use and what it is used for
+
+% Operational costs saw the only occurrence of renewable resources being included in the ten most important variables for its XGBoost ensemble.
+
+% %% Address yield and how predictable it is
+% %% different sales markets and performance?
+% %% Reasons for profit?
+
+% The winegrowing industry holds significant importance for Australia and its economy. There exists many challenges that the industry has to contend against, 
+% with disease from sources such as Mildew and Botrytis being a considerable one (Cole, 2010; Magarey et al., 1994).
+
+The disparity in accuracy between profit and operational costs is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
+The difference between turning a profit or loss is dependent on decisions made and chance.
+The difference between vineyards that make profit and those that do not could be a multitude of factors including differences in farming practices not captured within this study.
+Some decisions leading to latent effects such as large scale soil deposition in extreme rain events can be caused by soil compaction due to overworking a vineyard \citep{capelloPermanentCoverSoil2020}.
+
+\iffalse
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  Prior studies have shown that decisions can also have long-lasting positive effects at the cost of short term negatives.
+In particular the use of cover crops and remedial soil techniques offer such as trade.
+
+% with well established a cover crops requiring greater water resources to maintain but offering more protection in return (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
+
+% Prior studies have placed an emphasis on optimisations of fungicide sprays through the development of Bayesian models to forecast disease risk (Lu et al., 2020).
+This analysis investigates at the synergy between spray strategies as a proxy of fuel use to different types of water sources and sustainable strategies with an emphasis on the use of cover crops; allowing for the modelling of interactions between the use of multiple strategies.
+With the need to balance yield, quality, and combat adversities such as disease; the creation of tools to inform decisions and assist growers with warnings of disease risks is becoming crucial in ensuring sustainable and profitable wine production (Abbal et al., 2016).
+
+% Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
+For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
+Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
+The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
+
+% When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
+Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
+Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
+The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
+Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
+
+% Reasons for operational costs (connect to disease and pests)
+
+% There are several environmental concerns that affect viticulture, including loss of soil quality, lack of rain, hail, disease, fire, and frost; with climate change exacerbating these issues.
+In 2020, 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase (Canadell et al., 2021).
+In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought (Wine Australia, 2020).
+
+% In rainfed areas due to the lack of irrigation possibilities The result would be a shortening of the ripening period, with harvest occurring during the period with high temperatures, which could have a negative impact on wine quality (Salazar Parra et al.
+2010; Duchne and Schneider 2005; Jones and Davis 2000) and yield (Mira de Ordua 2010; Iglesias et al.
+2010).
+
+% Climate change in the future might move the north and south latitude boundaries of areas suitable for good quality wines (Schultz and Jones 2010), and could even lead to improvements in fruit production and quality in some areas (Olesen and Bindi 2002).
+However, other areas may be negatively affected by high temperatures and water stress due to a reduction in the amount of water available.
+
+% %% Grape quality
+
+% The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
+Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to complex forces such as international market pressure, fire, pests and disease (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012).
+
+% The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
+The most prominent being what readily available resources of each region were, particular the types of water available.
+Heavy water consumption is often linked to the mass production of grapes, where lower quality grapes are targeted in a quantity over quality strategy.
+These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
+Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield ((Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
+Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018)).
+There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues, where on-the-ground choices are influenced by other wine industry’s decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
+It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season (Leilei He et al., 2022).
+
+% %% connection to watrer resource availability
+% What are the water resources in the hunter valley and tasmania?
+
+% economic pressures
+
+% Historically strong demands for Australian wine have helped to create a thriving industry, however recently sharp reductions in exports to mainland China due to significant deposit tariffs have caused a decline of 19% in Australian export value during the 2021-2022 financial year (see Figure 1) (Wine Australia, 2022).
+The pressure brought on by the drop in export value has been exacerbated by loss of tourism and labour due to the COVID-19 pandemic, global freight crisis, war in Europe and rising inflation (Wine Australia, 2021a, 2020).
+These pressures within the wine industry have trickled down to vineyards, where winemakers retain unwanted wine, creating an oversupply of grapes.
+Currently the proposed strategy to soften pressures proposed by Wine Australia includes initiatives focused on market diversification, and sustainability to build resilience in the coming years.
+
+% Figure 1: The exports of Australian wine over time in Australian Dollars Free On Board, comparing exports between China and the rest of the world.
+This graphic is taken from the Wine Australia Annual Report of 2020-21(Wine Australia, 2022).
+
+%%%%%%%%%%%%%%%
+
+% %Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74% and 96.97% respectively.
+
+% These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
+However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
+The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
+
+% The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
+Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to complex forces such as international market pressure, fire, pests and disease \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
+It is a bit unclear what you are really trying to cite and its relevance.
+I think it important to chop this up.
+
+% The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
+The most prominent being what readily available resources of each region were, particular the types of water available.
+% A citation or reference to a study looking at the disparity in wine regions and their resources - or use the data more to describe these regional differences.
+Is water availability apparent?
+Do some regions have access to more sohpisticated water sources such as pressurised water etc.
+
+%  Heavy water consumption is often linked to the mass production of grapes,
+
+ % Really?
+proce it?
+Where in the results is this really found??
+This sounds like a link to your previous paper but you cannot really show this and you do not really produce evidence to support this claim.
+
+%  where lower quality grapes are targeted in a quantity over quality strategy.
+These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
+
+ % This point is incredibly important but it, itself, is obfuscated by you odd claims that are not supported by facts.
+Better to talk about how this could improve and the weaknesses than to start spouting unsupported claims.
+
+%  Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
+ % Woah woah woah
+%
+ % what are these decisions??? Slow down speed gonzales!! This is really interseteing and important. Discuss it, its the discussion! God damn I want to know and I wrote this!.
+ %
+%
+% This is not all one paragraph surely! There are so many ideas here that should be discussed!
+%
+% This leads heavily into the next paper and shows that natural progression.
+Especially in the sense of: if regions have so much influences, what determines ones use of water, fertiliser, etc.
+And, what determines the success financially, in quality or quantity?
+How are thes evariables interacting!
+The next paper is less about predicting and much more about what interacts with what?
+Where as this paper is about how things change over landscapes and what localised events helpe to determine decisions and their outcomes.
+%
+%  .
+There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues, where on-the-ground choices are influenced by other wine industry's decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+
+% This is an interesting point.
+However it would be better to include other industries, such as the ones you touch on in your previous chapter.
+If industries such as Sake are similar - how do their decisions effect the outcomes of rice quality and Sake?
+A point of comparison and difference would be great here!
+
+%  It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
+
+% What are these business decisions? Can you give some examples?
+
+%  It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \citep{leileiheFruitYieldPrediction2022}.
+
+% It is difficult yes, but please at least try!
+Highlight some of these affects, some of these different contributing factors!
+What did we really learn by doing this?
+What can someone take away from this and say thank you for?
+(like Josh said)
+
+% \subsection{Model 1 GI Regions}
+
+% The first Model was used to classify GI regions
+ % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
+
+% and resulted in an accuracy of 36.48\% across 52 classes.
 The most prominent features used to classify regions were the types of water resources available (see Figure 1).
 Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively.
+
 % it might be worth only talking about the regions that were well classifified - and then talk about their determining traits.
-These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
+
+% These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
 However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
+
 %Summary statistics regarding these two regions would be a better point of comparison, for example:
@@
 %
-The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
+
+% The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
 % What are the specific differences? This is a really critical part of this paper
-Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
-% This is almost its own paragraph.
-Why was this case?
-is it important?
-what are the similarities and differences of regions that were poorly classified?
- The regions with the most samples performed the best (see Table 1).
+
+% Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
+ % This is almost its own paragraph. Why was this case? is it important? what are the similarities and differences of regions that were poorly classified?
+
+%  The regions with the most samples performed the best (see Table 1).
 Notably bordering regions were routinely grouped together and misclassified as the same region, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
 The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
- % THis speaks highly to why regions were poorly classified.
-These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
+ % This speaks highly to why regions were poorly classified.
+
+% These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
 Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
@@
 
-\subsection{Climate}
-Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly.
+% \subsection{Climate}
+% Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly.
 These categories were divided into 12 climatic classifications with 3 and 4 separate subsets for rainfall and temperature respectively.
@@
 % interesting although not really appropriate for the results section.
- This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
+%  This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
 
@@
 
-\subsubsection{Rainfall}
-The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).
+% \subsubsection{Rainfall}
+% The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).
 
-\subsubsection{Temperature}
-The classification of GI Regions by their temperatures (see Figure 5) showed similarities to the other trees, with a heavy reliance on the types of water resources used as dominant predictors.
-The use of contractors % This is super important.
-Talk about this at lengthi nt the discussion.
-Find more literature on agriculture and the use of contractors as oppose to self managing issues such as weeds.
-was again used to differentiate between warm and cool regions, likely being due to disease pressure.
-The temperature classification tree was only a minor improvement over the regional classification tree, with an accuracy of 49.26\% as shown in the confusion matrix (see Table 4).
+% was again used to differentiate between warm and cool regions, likely being due to disease pressure. The temperature classification tree
 
-\subsection{Model 3 Grape Quality}
-The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
-had an accuracy of 55.72\% across 5 separate grades. There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2). % This is actually really interesting.
- The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
+% \subsection{Model 3 Grape Quality}
+% The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
+% had an accuracy of 55.72\% across 5 separate grades.
+There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2).
+% This is actually really interesting.
+%  The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
 Although not surprising the number of contractor tractor passes is new deciding factor due disease and pests reducing the potential quality of a crop.
@@
 
-\section{Discussion}
-
-% %Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74% and 96.97% respectively. 
-These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
-However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
-The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
-
-The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
-Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to  complex forces such as international market pressure, fire, pests and disease \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
-It is a bit unclear what you are really trying to cite and its relevance.
-I think it important to chop this up.
-The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
-The most prominent being what readily available resources of each region were, particular the types of water available.
-% A citation or reference to a study looking at the disparity in wine regions and their resources - or use the data more to describe these regional differences.
-Is water availability apparent?
-Do some regions have access to more sohpisticated water sources such as pressurised water etc.
- Heavy water consumption is often linked to the mass production of grapes, % Really?
-proce it?
-Where in the results is this really found??
-This sounds like a link to your previous paper but you cannot really show this and you do not really produce evidence to support this claim.
- where lower quality grapes are targeted in a quantity over quality strategy.
-These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
- % This point is incredibly important but it, itself, is obfuscated by you odd claims that are not supported by facts.
-Better to talk about how this could improve and the weaknesses than to start spouting unsupported claims.
- Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
- % Woah woah woah
-%
- % what are these decisions??? Slow down speed gonzales!! This is really interseteing and important. Discuss it, its the discussion! God damn I want to know and I wrote this!.
- %
-%
-% This is not all one paragraph surely! There are so many ideas here that should be discussed!
-%
-% This leads heavily into the next paper and shows that natural progression.
-Especially in the sense of: if regions have so much influences, what determines ones use of water, fertiliser, etc.
-And, what determines the success financially, in quality or quantity?
-How are thes evariables interacting!
-The next paper is less about predicting and much more about what interacts with what?
-Where as this paper is about how things change over landscapes and what localised events helpe to determine decisions and their outcomes.
-%
-
- .
-There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues, where on-the-ground choices are influenced by other wine industry's decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
-% This is an interesting point.
-However it would be better to include other industries, such as the ones you touch on in your previous chapter.
-If industries such as Sake are similar - how do their decisions effect the outcomes of rice quality and Sake?
-A point of comparison and difference would be great here!
- It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
-% What are these business decisions? Can you give some examples?
- It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \citep{leileiheFruitYieldPrediction2022}.
-% It is difficult yes, but please at least try!
-Highlight some of these affects, some of these different contributing factors!
-What did we really learn by doing this?
-What can someone take away from this and say thank you for?
-(like Josh said)
+\fi
 
@@
 
-The type and availability of water resources were a major contributing factor when classifying grape quality and region.
+% The type and availability of water resources were a major contributing factor when classifying grape quality and region.
 This was seen in the two most accurately classified regions, Coonawarra and the Riverland, with the Riverland predominantly utilising river water.
@@
 Further in-depth data and understanding are necessary to fully grasp the nuances of decision-making and the interplay of factors impacting grape quality.
+
 % Interesting - the conclusion is actually reasonable, especially considering the disconnection of ideas and models in this paper.

### 2023-08-31 article.tex
conclusion

@@
 
+This study has provided valuable insights into the multifaceted dynamics governing operational costs, different yearly effects and vineyard regions.
+Highlighting the complex interrelatedness of variables within a vineyard.
+The paper underscores how factors such as water and fuel use intersect to impact operational costs.
+How different yearly events affect these operations and the significance of context-specific decision-making.
+While this investigation utilised a broad regional classification, the potential benefits of adopting a more nuanced approach and incorporating expert knowledge have been highlighted.
+By delving deeper into the complex interplay of variables, further advancements can be made in optimising vineyard management strategies for lowering operational costs and enhancing sustainability.
+
+% The intricate relationships between water resources, and management practices have been examined through advanced techniques like XGBoost ensembles, shedding light on their roles in shaping production outcomes.
+
 % The type and availability of water resources were a major contributing factor when classifying grape quality and region.

### 2023-08-31 article.tex
its a title

@@
 %% \title{Title\tnoteref{label1}}
-\title{???Grape Quality and its Link to Regional Differences in the Australian Winegrowing Industry}
+\title{An analysis of interrelations between economic and environmental variables in Australian Winegrowing.}
+
+% 
+% Variable importance in determining vineyard operational costs and region using statistical machine learning.
+% How are region and operational costs related?
+
 % Regional differences in Australian winegrowing (Quality)?
@@
 
-\begin{abstract}
-\end{abstract}
+% \begin{abstract}
+% \end{abstract}
 %%Graphical abstract`
@@
 
-\if{false}
+In the past decade, the Australian winegrowing industry has undergone a variety of pressures, such as changing market demands, disease and drought \citep{wineaustraliaAustralianWineProduction2021}.
+Furthermore, natural resources are likely to decrease, as pressures from climate change increase, making it more important than ever to improve the efficiency and sustainability of crops \citep{agdeeNationalGreenhouseAccounts2021}.
+It has become crucial for those in the wine industry to address issues relating to environmental sustainability and economic viability, with a growing need for the industry to close the research gaps between sustainable practices and their real and perceived environmental and economic advantages \citep{montalvo-falconSustainabilityResearchWine2023,ouvrardDoesSustainabilityPush2020}.
+This paper aims to provide a comprehensive analysis of the intricate relationships between economic and environmental variables within the winegrowing sector.
+By employing statistical machine learning we show how interconnected vineyard variables are.
+Utilising a ten year data set that spans Australia we show the predominant elements in classifying region and year, as well as determining a vineyards potential operational costs and profit.
+\par
+This analysis utilises XGBoosted trees to classify region and year; and to determine operating cost and profit of Australian vineyards.
+Classification and regression trees are utilised as surrogate models to shed insight into the key partitions used by the XGBoost ensembles.
+Variable importance is further used to illustrate the interconnectedness of the different predictor variables, and to show the similarity in variable importance between different response variables (particularly between region and operational cost).
+This study aims to assist in uncovering the complexity of variables that are affected by a variety of vineyard management decisions to illustrate complex interplay of variables.
+This study endeavours to gain insight into the similarity in predictor variable importance between year, region, operational costs and profit.
 
-We classify regions to find the most prominent and well-defined relationships to regions.
-We examine the variables that are the most decisive in determining region.
-We compare the similarities and differences of these regions
-- within the analysis and
-- within the literature (also in their definition and history)
+% Initiatives to improve the environmental impacts of the Australian Wine industry have been ongoing since the 1990s; with recommendations to implement strategies for national programs that would help educate industry members in sustainable practices (Keith Jones, 2002).
+The first legislative initiative to support national programs was the Environmental Management Guidelines for Vineyards in Western Australia during 2002 (Nind, 2002).
+The first national program was the Australian Wine Industry Stewardship, beginning shortly after, during 2004 (WFA, 2009).
+This program was replaced multiple times during the 2000s, first by the Australian Wine Environmental Stewardship, and then by EnviroWine, finally becoming EntWine in 2009.
+The period in the 2000s was further influenced by several key researchers and program leaders, where concurrently multiple regions developed approaches to sustainable management tied to sustainability programs (SWA, 2023).
+EntWine was eventually acquired by the Australian Wine Research Institute and Australian Grape & Wine during 2015; which would lead to its redevelopment into Sustainable Winegrowing Australia (SWA) during 2019/2020 (SWA, 2023).
+%             2.1.2. Previous Research and Motivation
+% During the period of 2010 to 2023 data was collected voluntarily from wine industry members through the national sustainability programs, ultimately being in SWA possession as of 2023.
+The data set was used in 2020 to analyse sustainable practices, resulting in the development of a Resource Intensity Score (RIS); where the RIS was a summary indicator value of a vineyard’s energy use, emissions and water use (Nayak et al., 2020).
+The analysis then fit vineyards’ RIS values to 55 variables sourced from 23 separate wine regions during the period of 2015 to 2018 through the use of Stepwise Regression modelling.
+It was found that the top 10% gross margin vineyards had a lower operating cost and a lower RIS score than vineyards in the bottom 10% gross margin.
+% The motivation for this thesis is to investigate the differences between the top 10% gross margin vineyards and those of the bottom 10%, hoping to uncover what the key differences were and to relate them to sustainable practices; as there is a known gap in the literature due to the difficulty to evaluate relationships using empirical evidence of both the environmental and economic impacts of sustainable practices in viticulture.
+This difficulty stems from data scarcity, industry complexity and transparency.
+While there have been attempts to fill these gaps, methodologies and indicators have lacked in ability to compare different regions and scales (Baiano, 2021).
+%             2.1.3. Economic Pressures
+% Historically strong demands for Australian wine have helped to create a thriving industry, however recently sharp reductions in exports to mainland China due to significant deposit tariffs have caused  a decline of 19% in Australian export value during the 2021-2022 financial year (see Figure 1) (Wine Australia, 2022).
+The pressure brought on by the drop in export value has been exacerbated by loss of tourism and labour due to the COVID-19 pandemic, global freight crisis, war in Europe and rising inflation (Wine Australia, 2021a, 2020).
+These pressures within the wine industry have trickled down to vineyards, where winemakers retain unwanted wine, creating an oversupply of grapes.
+Currently the proposed strategy to soften pressures proposed by Wine Australia includes initiatives focused on market diversification, and sustainability to build resilience in the coming years.
+% Figure 1:  The exports of Australian wine over time in Australian Dollars Free On Board, comparing exports between China and the rest of the world.
+This graphic is taken from the Wine Australia Annual Report of 2020-21(Wine Australia, 2022).
+%             2.1.4. Environmental Pressures
+% There are several environmental concerns that affect viticulture, including loss of soil quality, lack of rain, hail, disease, fire, and frost; with climate change exacerbating these issues.
+In 2020, 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase (Canadell et al., 2021).
+In comparison to countrywide pressures such as drought, this damage made up only 3% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought (Wine Australia, 2020).
 
-We examine the similarity between these classifications and the variables that are significant in classifying profitable vs not.
-How do regional changes affect that. What is similar and what is different.
+% \if{false}
 
-We can link these contributing factors back to the baselines - at least anecdotally, or in an unpublished section. This will be good for the thesis.
+% We classify regions to find the most prominent and well-defined relationships to regions.
+% We examine the variables that are the most decisive in determining region.
+% We compare the similarities and differences of these regions
+% - within the analysis and
+% - within the literature (also in their definition and history)
 
-\fi
+% We examine the similarity between these classifications and the variables that are significant in classifying profitable vs not.
+% How do regional changes affect that. What is similar and what is different.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
+% We can link these contributing factors back to the baselines - at least anecdotally, or in an unpublished section. This will be good for the thesis.
+
+% \fi
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
 Each region describing unique reputations, qualities and varietals of wine produced there.
@@
 In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
-\newline
+% \newline
 
@@
 %
-Through the use of classification trees this study aims to highlight the key differences in sustainable practices at a regional level and how these practices relate to the different grades of grape quality.
+% Through the use of classification trees this study aims to highlight the key differences in sustainable practices at a regional level and how these practices relate to the different grades of grape quality.
 
@@
 % And what is the take away point of the article? If I cant be bothered to read this article, what piece of knowledge can you leave mewiwth that I could recommend the article to someone else with?
-
-\if{false}
-\begin{figure}
-  \includegraphics{violinplot.pdf}
-  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
-  \end{figure}
- %
-\fi{}
 

### 2023-09-27 article.tex
half-way save on Kates edits

@@
 
-% 
 % Variable importance in determining vineyard operational costs and region using statistical machine learning.
@@
 
+% TODO:
+% 
+% Below is the original comment. BUT you have since come up with more of a scheme that answers these questions (in the bayesian paper). Use this to extrapolate on the introduction.
+% 
 % You obviously have no leading question for this paper, and although you do a good job at creating accurate models; it is more important that the research is asking a good question.
@@
 %  
-%
+%%
+
+% FIXME:
+% Whats been done? Whats the persistant gap? Why is a data-driven approach the answer?
 
@@
 By employing statistical machine learning we show how interconnected vineyard variables are.
-Utilising a ten year data set that spans Australia we show the predominant elements in classifying region and year, as well as determining a vineyards potential operational costs and profit.
+Utilising a ten year data set that spans Australia we show the predominant
+elements in classifying region and year, % FIXME: A reader will ask: why would i want to predict region and year?
+ as well as determining a vineyards potential operational costs and profit.
 \par
-This analysis utilises XGBoosted trees to classify region and year; and to determine operating cost and profit of Australian vineyards.
-Classification and regression trees are utilised as surrogate models to shed insight into the key partitions used by the XGBoost ensembles.
-Variable importance is further used to illustrate the interconnectedness of the different predictor variables, and to show the similarity in variable importance between different response variables (particularly between region and operational cost).
-This study aims to assist in uncovering the complexity of variables that are affected by a variety of vineyard management decisions to illustrate complex interplay of variables.
-This study endeavours to gain insight into the similarity in predictor variable importance between year, region, operational costs and profit.
+This analysis utilises 
+XGBoosted trees % FIXME: Why XGBoosted trees?
+ to classify region and year; and to determine operating cost and profit of Australian vineyards. Classification and regression trees are 
+ utilised as surrogate models %FIXME: Why are we utilising surrogate models?
+to shed insight into the key partitions used by the XGBoost ensembles. Variable importance is further used to illustrate the 
+interconnectedness % FIXME: What kind of interconnectedness?
+of the different predictor variables, and to show the similarity in
+variable importance % FIXME: What does variable importance mean in this context?
+between different response variables (particularly between region and operational cost). This study aims to assist in uncovering 
+the complexity of variables %FIXME: This is too vague, what makes them complex? why are the complex? etc
+that are affected by a variety of vineyard management decisions to illustrate 
+complex interplay of variables. % FIXME: What do you mean by complex interplay? Like what? An example?
+This study endeavours to gain insight into the 
+similarity % FIXME: Why do we want to know the similarity?
+ in predictor variable importance between year, region, operational costs and profit.
 
@@
 
+% FIXME:
+% While overall this methods section seems technically correct, more work is needed on justification, motivation, and context.
+% Through the methods, motivate each decision youve made with context from the particular problem/data/question.
+Why is the basic standard procedure not adequate?
+This helps to slow it down and help the reader understand complex methods.
+%
+% As kate put it in person you are currently writting like ||||| fact fact fact. You need to be more |...||...|| with context etc inbetween each fact so that people can digest it.
+
 % As with the previous paper, put the models up front it is very unclear what you have created - and less clear to understand it if there is no way to reference to it.
+
+% FIXME:
+% In this methods section, what is novel/innovative? Not currently clear, In particular - if I gave this data & these packages to a 3rd year class, which part here couldn't they do?
 
@@
 23 variables were used for each observation reflecting a vineyards account for the given year (see Table \ref{tab:vars}).
-The profit variable was additionally transformed and included as another separate variable, profitable; depicting whether a vineyard was profitable or not.
+The profit variable was transformed % FIXME: What was it before?
+and included as another separate variable, profitable; depicting whether a vineyard was profitable or not.
 \par
@@
 \subsection{XGBoosted Trees}
+% FIXME: What do boosted trees do?
 XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-They were chosen for this analysis as they provide both a high predictive performance and ability to effectively capture complex relationships.
+They were chosen for this analysis as they provide both a
+high predictive performance % FIXME: For what?
+and ability to effectively 
+capture complex relationships % FIXME: How?
+.
 An XGBoosted tree was created for each variable to show how they interacted.
 Each tree included all but the economic variables (profit and operating cost), which were only included once as response variables.
+% FIXME: what does this mean in terms of the method and outputs?
 \par
@@
 The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
-
+% FIXME: All equations must be in the middle or end of sentances, not at the beginning.
 \begin{equation}
- \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, f_{K}\in \mathcal{F},
+ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, \space f_{K}\in \mathcal{F},
 \end{equation}
 
+% FIXME: The methods would benefit from some subsections (signposts)
+% FIXME: What does the equation do? Give intuition.
+
 where each function $f_K$ is a classification or regression tree, such that all functions are in the set of all decision trees $\mathcal{F}$, defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
-Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights.
+Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights. % FIXME: Sentance fragment.
 Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf. 
-The final prediction is determined by the sum of the score of the corresponding leaves, given by $\omega$.
-The set of functions used by the tree is determined by minimising the regularised objective function, given by:
+The final prediction % FIXME: What is this?
+is determined by the sum of the score of the corresponding leaves, given by $\omega$. 
+The set of functions % FIXME: You haven't described the tree as having a set of functions?
+%  I believe when you refer to the set of functions F you might benefit from putting the F in brackets here or being more explicit where the functions are.
+used by the tree is determined by minimising the regularised objective function, given by:
 

### 2023-09-28 article.tex
Kate edits

@@
 
+% FIXME: Stop writing chronologically. Write by significance.
+
 \section{Introduction}
@@
 
+% FIXME: The methods needs to be restructured. Try the following order:
+%     - Why do it?
+%     - What does it do? (at a high level)
+%     - How did you do it?
+% 
+% You should also include subsections. I think subsections are a good crutch because they can always be removed if the flow is done well.
+% 
+% Avoid following a narrative. Do not do the who/what/when/where/how/why and the big reveal.
+% An aside comment might be that you can attempt to follow the narrative writing paradigm across the paper as a whole but limit your methods section to the why, what then how.
+And use the results as your reveal.
+% Or maybe dont do that - it might just be that you need to write it like a narrative as a first draft then butcher it into the necessary sections.
+That writing style is incredibly ingrained at this point.
+
+% FIXME: Adde a section for CV.
+
+% FIXME: Another section that needs to be added is one that describes what the sankey and chord diagrams are and what they represent.
+The reader needs to be prepped for this before the results.
+They need to be able to properly interpret them.
+
+% FIXME: Explain how some variables are one-hot-encoded and how this relates to variable importance. This is a significant resulti n the xgboost tree for the huntervalley
+
+% FIXME: the motivation is missing from methods.
+% FIXME: What is interrelated importance?
+
 \subsection{Data}
@@
 %  I believe when you refer to the set of functions F you might benefit from putting the F in brackets here or being more explicit where the functions are.
-used by the tree is determined by minimising the regularised objective function, given by:
+used by the tree is determined by minimising the 
+regularised % What is the regularised objective?
+objective function, given by:
 
@@
 
-The difference between the prediction and actual variable is a convex loss function $l$.
-To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
-The function $f_t$ is selected according to which value minimises (\ref{eqn:objective}).
-The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
+The difference between the 
+prediction and actual variable % FIXME: what are the predition and actual variables?
+% When you are talking broadly about elements of xgboost I believe it will help to refer to the equation, an example or something more specific.
+This would help to address the other comment regarding giving an intuition to the method.
+is a convex loss function $l$. To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration. The function 
+$f_t$ % FIXME: Kate asked if this was $f_K$.
+This is not the case as $f_K$ is part of the penalty function, and I know this at first glance due to working with the equation regularly for this paper.
+If Kate is confused about this then your explanation sucks.
+It is to obtuse, like Kate has said you are hitting people with too many facts and not padding it enough to the context of this paper and its application or even its purpose.
+The function $f_t$ is incredibly important, especially because for the purposes of this study it changes depending on the variable and its nature.
+You are clearly not
+is selected according to which value minimises (\ref{eqn:objective}). The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
 \par
 As predictions are made using additive tree functions, XGboosted trees can be used for classification and regression.
-Due to the mixture of continuous, binary and multiclass variables in this analysis, both classification and regression trees were created.
+Due to the mixture of continuous, binary and multiclass variables in this analysis,
+both classification and regression trees were created. % FIXME: What were these created for?
 The difference between the trees created for this analysis was the objective function used.
-XGBoosted regression trees were created for continuous variables, using the root-mean-square as the objective function.
+% FIXME: Kate marked this with a "?" clearly highlighting the issue with the previous comment regarding f_t (at some point was on line 321)
+XGBoosted regression trees were created for continuous variables, using the
+
+% FIXME: The below section talks about the objective functions.
+Kate askes what they are and what they related to.
+You will need to go into the objective functions, how they are different and why different functions are required to be used.
+% Perhaps a reference that talks about the importance of objective functions.
+root-mean-square as the objective function.
 Binary class variables utilised the logistic loss function as the objective.
@@
 \par
-Chen and Guestrin \citep{chenXGBoostScalableTree2016} further illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
-Furthermore, they show the loss reduction after the split is given by the function:
+% FIXME: Why are you doing this? give me reasons related to this work to keep reading.
+Chen and Guestrin \citep{chenXGBoostScalableTree2016} illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
+Furthermore, they show the loss reduction after
+the split % FIXME: What is the split?
+is given by the function:
 
@@
 \par
+% FIXME: The below section needs to come sooner
 The frequency of a variable appearing as a node within the ensemble was used as a measure of importance.
-This measure was chosen as it connected a variable to the minimisation of its associated objective function, translating the value into a simple count metric.
+This measure was chosen as it connected a variable to the minimisation of its associated objective function, translating the value into a
+simple count metric. % FIXME: Variable importance might need to be in its own section to fully explain it and its motivation.
+% FIXME: The creation of xgboost trees for each variable needs to be earlier and more clear as well contextualised by its motivation.
 Creating XGBoosted trees for each variable allowed the use of importance to show how strongly variables were associated with each other.
-The importance of predictor variables to economic variables was illustrated through the use of Sankey diagrams constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+The importance of predictor variables to economic variables was illustrated through the use of
+Sankey diagrams constructed % FIXME: More needs to be said on this.
+More needs to be talked about regarding the chord diagrams as well.
+There is no thorough explanation of how the diagrams work and what you have done with them to fully explain them!
+using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
 Other variable's interconnectedness was demonstrated through the use of a chord diagram also created using Holoviews.
+% FIXME: The above comment about sankey diagrams applies to chord diagrams.
+These need fuller descriptions.
 \par
+% FIXME: More info is needed regarding the validation and how it was performed.
+This section can be expaned on as different validation and scoring is used due to the differencein objective functions.
+This is a rather critical section and very little is said about it.
+% 
 Each variable utilised 80\% of the data to train the XGBoost ensemble, with 20\% reserved for testing and validation.
@@
 For binary and multiclass variables data was split into 80\% training, 10\% testing and 10\% validation data. 
-Due to class disparity in multiclass variables (most prominently in region) data was stratified into each subset at the same ratio of class occurrence.
-Validation was summarised through the accuracy, the proportion of true negatives and positives.
+Due to class disparity in multiclass variables (most prominently in region) data was stratified % FIXME: What does it means to stratify the data and why is this done?
+into each subset at the same ratio of class occurrence. Validation was summarised through the accuracy, the proportion of true negatives and positives.
 \par
+% FIXME: rewrite the below paragraph in the active voice.
 The use of the XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
@@
 Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
-
-\subsection{Classification and Regression Trees}
-
+% FIXME: What is the output?
+\subsection{Classification and Regression Trees} % FIXME: perhaps this should be renamed surrogate models
+% FIXME: It is not clear what you are trying to achieve with these models. No where do you give a definition of what a surrogate model is and why you are using it!
+% Be specific about the purpose of what these models are being used for!
+Justify why you have not simply shown one of the ensembles trees as well!
+The use of these algorithms means that the tree is more general than any one of the ensembles manifestations.
 Classification and Regression Trees were created for region, year, profit and operating cost.
@@
 
+% FIXME: The first result should be in the langauge of the question.
+In the original document the first result presented was region, implying that the question is: 'are regions different?' which is not the forefocus of the article.
+The results need to be rearranged to show grossmargin and operating costs as more significant variables as they are the curious factors that we do not understand.
+
+% FIXME: The irritating placement of the figures is due to the way latex uses the /ref function to determine where a figure should be.
+Have a google of this to overcome it as the figure placement is all over the shop.
+
 \subsection{Region}
-
-Region classification performed at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
+% FIXME: there are artifacts from the previous paper.
+Although the previous paper is pertinant to some of this (quite critically so) it is no longer the primary focus.
+On top of this some of these terms such as 'Region Classification' are not well defined.
+This is because previously a great deal of effort was put into defining them and describing the potential regions.
+However this has been removed.
+% FIXME: This section is much more of a: 'isn't it weird?' That region is not as significant in some areas but not others.
+That is can be used to predict things like yield but no profit etc.
+This phrasing shows why it goes after other sections.
+if you are writting by significance then it is not a significant find but it is something that is worth noting - unlike year which was just oh dear we dont know what is going on.
+Region classification performed % FIXME: what does this mean - see the above comment.
+% Note that it is not about the implications but getting in the definition of these terms.
+These things should be prepared for the reader well before the results.
+There should be no new terms here, we should just be refering to the outcomes that were established in the methods, and they should already have their motivations and contexts estabished (there is more that Kate has written if you are confused this is my translation after our meeting).
+at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
 The most prominent feature used to classify regions with the classification tree was water source (see Figure \ref{fig:region_tree}).
@@
 Area, water, fuel and yield were more determining factors when predicting region using XGBoost.
+% FIXME: Is this perhaps your primary test?
+The fact that Kate was asking this means that the tools you are using do not have a well defined purpose in this analysis.
+I believe this stems from the lack of contextual motivation behind these tools.
+What is it that you are testing?
+Why are these important?
+This problem is throughout the paper and cannot simply be fixed here.
 Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
 \par
-It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
-The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
-Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of other activities within the vineyard, such as pruning and harvesting.
-With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
-\par
+
 Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
 The regions with the most samples performed the best.
-Notably bordering regions were routinely grouped together and misclassified as the same region.
+Bordering regions were routinely grouped together and misclassified as the same region.
 Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
@@
  \includegraphics{region.png}}
+% FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
+% FIXME: Mark one Panel A and the other B
  \caption{The left-hand side depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
@@
 
-\subsection{Year}
+\subsection{Operating Costs}
 
-The classification tree and XGBoosted ensemble performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
-Electricity and the type of irrigation were highly influential within the classification tree.
-Similarly, electricity was the most frequently occurring node in the XGBoost ensemble.
-However, other variables such as slashing passes, and fungicide and herbicide spraying were more prevalenct than in the classification tree.
-Weed and disease outbreaks are likely an influential factor when classifying different years, making the decisions to spray and slash unique factors that differ year to year.
-Climatic differences between years are likely tied to the influence of yield and water use.
-\par
-Over half of the interrelated importance of the predictor variables is dominated by area harvested, yield and slashing passes.
-Although all the predictor variables are highly connected, their relative importance is not as prominent as the three major variables.
-It is of particular note of the relative importance of slashing to area, fuel and yield; as these are not directly related activities.
-The connection between slashing and spraying is that those who do a set number of spraying or slashing passes tended to do that many passes for all slashing and spraying activities.
-
-\begin{figure}
- \resizebox{\textwidth}{!}{
- \includegraphics{year.pdf}}
- \caption{Decision tree predicting Year.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:year_tree}
-\end{figure}
-
-\begin{figure}
- \resizebox{\textwidth}{!}{
- \includegraphics{year.png}}
- \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:year_sankey}
-\end{figure}
-
-\subsection{Operating Costs}
+% FIXME: Try to follow the same format for each of the section in the results. It will help to include a leading sentance about why the metrics are different everytime.
 
@@
 It is surprising that electricity, slashing and spraying was not more prominent in operating costs.
+% FIXME: Me and Kate talked about this.
+The diagrams, their context and justification go into the methods.
+How to interpret them needs to be better outlined here.
+You need to affirm peoples intition of them and the significance of more ribbons going out than in as well as why we are only using 10 variables.
 However, Figure \ref{fig:year_sankey} shows that electricity, slashing and spraying are important variables in determining area and yield.
+
 Electricity in particular is used predominantly for irrigation and so is related largely to the size of a vineyard.
@@
  \includegraphics{operating_costs.png}}
+ % FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
+ % FIXME: Mark one Panel A and the other B
  \caption{The left-hand side depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
@@
 Predictions of profit performed poorly compared to operating costs with the regression tree having an $R^2$ of 0.1873 (with a standard deviation of 0.0522) and the XGBoosted ensemble achieving an $R^2$ of 0.2535 (with a standard deviation of 0.3126).
-The high standard deviation in the XGBoosted tree was a bias in more accurately predicting vineyards that made profit compared to those that lost money.With much higher $R^2$ values being achieved in k-folds containing only those that made profits (recording a maximum of 0.7634).
+The high standard deviation in the XGBoosted tree was a bias in more accurately predicting vineyards that made profit compared to those that lost money.
+% FIXME: Kate pointed out that it is unclear what this implies (the below sentance).
+% The following paragraph goes into it though, it would be better to meld these together. Talking more about the fact that it was easier to predict those who made a profit than those who did not.
+% FIXME: It is wise to run a model purely on gross margin as well. It might have better insights than just profit! And it might over come this disparity.
+With much higher $R^2$ values being achieved in k-folds containing only those that made profits (recording a maximum of 0.7634).
 % profit summary stats
@@
 It was surprising that operating costs performed substantially better in $R^2$ compared to profit.
+% Is there a better way to phrase this?
+Look at the papers that use the same analysis techniques for phrases to use when soething is good, bad predictor.
+
 Interestingly the important variables when attempting to determine profit were similar to those used to classify region (see Figure \ref{fig:profit_sankey}), with the exception of water used.
@@
  \includegraphics{profit.png}}
+ % FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
+ % FIXME: Mark one Panel A and the other B
  \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
@@
 
+% FIXME: Moved here from results. Find a place for this lost puppy!
+% It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
+The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
+Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of other activities within the vineyard, such as pruning and harvesting.
+With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
+
 Several physical parameters such as climate, geology and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 The association between yield and region is demonstrated by its position as fourth most occurring variable within the nodes of the XGBoosted ensemble which determined region (see Figure \ref{fig:region_sankey}).
+% FIXME: Clunky phrasing
+
 The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019}.
@@
 
+ \subsection{Year}
+
+ The classification tree and XGBoosted ensemble performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
+Electricity and the type of irrigation were highly influential within the classification tree.
+Similarly, electricity was the most frequently occurring node in the XGBoost ensemble.
+Other variables such as slashing passes, and fungicide and herbicide spraying were more prevalent than in the classification tree.
+Weed and disease outbreaks are likely an influential factor when classifying different years, making the decisions to spray and slash unique factors that differ year to year.
+Climatic differences between years are likely tied to the influence of yield and water use.
+ \par
+ Over half of the interrelated importance of the predictor variables is dominated by area harvested, yield and slashing passes.
+Although all the predictor variables are highly connected, their relative importance is not as prominent as the three major variables.
+It is of particular note of the relative importance of slashing to area, fuel and yield; as these are not directly related activities.
+The connection between slashing and spraying is that those who do a set number of spraying or slashing passes tended to do that many passes for all slashing and spraying activities.
+
+ \begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{year.pdf}}
+  \caption{Decision tree predicting Year.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:year_tree}
+ \end{figure}
+
+ \begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{year.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:year_sankey}
+ \end{figure}
+
 \end{linenumbers}

### 2023-10-03 article.tex
Mardi edits

@@
 % FIXME: Stop writing chronologically. Write by significance.
+% TODO: Should the models be redone as a factor of area? instead of including area? The ensemble should effectively pick this up I do not think it needs to be done.
 
@@
 
-In the past decade, the Australian winegrowing industry has undergone a variety of pressures, such as changing market demands, disease and drought \citep{wineaustraliaAustralianWineProduction2021}.
+In the past decade, the Australian grape and wine industry has undergone a variety of pressures, including changing market demands, disease and drought \citep{wineaustraliaAustralianWineProduction2021}.
 Furthermore, natural resources are likely to decrease, as pressures from climate change increase, making it more important than ever to improve the efficiency and sustainability of crops \citep{agdeeNationalGreenhouseAccounts2021}.
@@
 Data used in this analysis were obtained from Sustainable Winegrowing Australia.
-Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Australia's national wine industry sustainability program.
+The program aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
 Data recorded by the SWA is entered manually by winegrowers using a web based interface tool.
@@
  \label{tab:vars}
- \caption{Summary of variables used in the analysis. The recorded column indicate values that were either greater than zero or that were not missing.}\label{tab:summary}
+ \caption{Summary of variables used in the analysis. The recorded column indicate the number of values that were either greater than zero or that were not missing.}\label{tab:summary}
  \begin{tabular}{@{}ccccl@{}}
@@
 Both diesel and petrol were of more relative importance (being ranked higher) in operating costs than water was compared with region.
-It is surprising that electricity, slashing and spraying was not more prominent in operating costs.
+It is surprising that electricity, slashing and spraying passes was not more prominent in operating costs.
 % FIXME: Me and Kate talked about this.
@@
 This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
+% FIXME: May need to maintain bigger vineyards are more likely to have more staff verse small owner operators.
+Those extra staff may be trying to keep employers busy be sending them slashing (this comment was from Mardi and very hard to validate in the literature.)
 
@@
 
-\begin{figure}
+\begin{figure} % FIXME: Interesting these 2 regions probably have the highest price of grapes/tonne.
+They are also 2 extreme climates (Hunter = lots of grazing, season rainfall = of tractor passes.
+Tasmania = V.Cool weather = lots of tractor passes.)
  \resizebox{\textwidth}{!}{
@@
 
+% FIXME: Mardi again asked about whether yield was per hectare.
+This is areally interesting consideration - we do not look at predicting yield specifically in this though.
+perhaps we should be very wary about these analysis with this respect.
+
 % FIXME: Moved here from results. Find a place for this lost puppy!
+%
 % It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
 The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
-Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of other activities within the vineyard, such as pruning and harvesting.
-With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
-
-Several physical parameters such as climate, geology and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of
+% other activities within the vineyard, such as pruning and harvesting. FIXME: This is new infor tht doesnt have any context (e.g no discussion yet about what fuel may be used for)
+% With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
+
+Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 The association between yield and region is demonstrated by its position as fourth most occurring variable within the nodes of the XGBoosted ensemble which determined region (see Figure \ref{fig:region_sankey}).
@@
 
-The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019}.
+The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % FIXME: Maybe but not sure this assumption holds.
+It may be an artefact of the members being corporate in larger holdings.
+The larger percentage of blocks in the riverland are small.
 \par
@@
 Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
-In general warmer regions have been associated with lower yields due to their generally lower rainfall, which can be mitigated through applying excess water \citep{campsGrapeHarvestYield2012}.
+In general if irrigated warmer regions have been associated with lower yields due to their generally lower rainfall, which can be mitigated through applying more water \citep{campsGrapeHarvestYield2012}.
+% FIXME: Rework this sentance
 It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of lower quality grapes; reflected through the variables' importance of water use in the XGBoost ensemble.
+% FIXME: Need to link lower quality to profitability.
+% FIXME: Warm inland regions are close to a major source of irrigation water which means that they have always taken advantage of this to grow higher crops.
+Which are also extremely important to our sector.)
+% FIXME: May need to explain that growers are paid @ a $ value / tonne and that the $ value varies in region.
 The practice of utilising larger quantities of water for inland Australian wine crops is partly reflected in the prior use of flood style irrigation to saturate soil \citep{bgcoombeGrapeBerryDevelopment2004}.
+% FIXME: More likely is that they have access to large water licences (Prev used for flood irrigation) That ythey now use on vineyards.
+% FIXME: Examples of regions??
 This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
@@
 With this connection reflecting the restriction of possible strategies employable by winegrowers between different regions.
+% FIXME:  No this is because the hunter valley/tasmania get paid more / tonne.
 \par
 In part some winegrowing strategies are restricted simply through access to water resources, being reflected through the region classification tree (see Figure \ref{fig:region_tree}).
-Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops unlike most coastal regions.
+Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surgace or underground water.
 Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region.
 Where, the need to pressurise irrigation systems from river water or to generate power would require larger amounts of diesel and petrol.
+% FIXME: Yes but this requires some explanation
 \par
@@
 The node frequency was high for area but much lower in general across the other variables, which could indicate the need to be more circumstantial in determining operational costs.
-Although it was attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), it is likely yet more complicated still.
-An example of how interrelated operational costs can be, is the optimisation of tractor passes being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
+Although it was attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), it is likely yet more complicated.
+An example of how interrelated operational costs can be, is
+the optimisation of tractor passes % FIXME: What does this mean? Multiple activities with a single pass?
+being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}. 
 \par
 Decisions made on the ground have far-reaching effects and are difficult to completely capture.
-Greater tractor use as a preventative measure for occurrences such as disease, may incur higher operational costs but could be critical in preventing long term losses.
+A higher number of tractor passes used as a preventative measure for occurrences such as disease, may incur higher operational costs but could be critical in preventing long term losses.
 With factors such as erosion and soil health being difficult to capture but also influenced by tractor use \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
@@
 \par
-Although less important in the XGBoost ensembles for profit, the variables: cover crops, fungicide spraying and slashing are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
-Rainfall being related to fungal growth and disease, as well as weeds.
+Although less important in the XGBoost ensembles for profit, the variables: cover crops, and number of fungicide spraying and slashing passes are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
+Rainfall being related to vine fungal growth, disease and weed pressure.
 With cover crops being an effective and sustainable method to alleviate these issues \citep{delpuechAdaptingCoverCrop2018}.
+% FIXME: Re-structure this sentance.
 It is difficult to extrapolate findings to these methods and the reason for their use due to the broad and varying definition of the regions.
@@
 
-This study has provided valuable insights into the multifaceted dynamics governing operational costs, different yearly effects and vineyard regions.
-Highlighting the complex interrelatedness of variables within a vineyard.
-The paper underscores how factors such as water and fuel use intersect to impact operational costs.
-How different yearly events affect these operations and the significance of context-specific decision-making.
+This study has provided valuable insights into the multifaceted dynamics governing operational costs, different yearly effects and vineyard regions highlighting the complex interrelatedness of variables within a vineyard.
+The paper underscores how factors such as water and fuel intersect to impact operational costs and How different seasonal events affect these operations and the significance of context-specific decision-making.
 While this investigation utilised a broad regional classification, the potential benefits of adopting a more nuanced approach and incorporating expert knowledge have been highlighted.
@@
 Although all the predictor variables are highly connected, their relative importance is not as prominent as the three major variables.
-It is of particular note of the relative importance of slashing to area, fuel and yield; as these are not directly related activities.
-The connection between slashing and spraying is that those who do a set number of spraying or slashing passes tended to do that many passes for all slashing and spraying activities.
+It is of particular note of the relative importance of slashing passes to area, fuel and yield; as these are not directly related activities.
+The connection between the number of slashing and spraying passes is that those who do a set number of spraying or slashing passes tended to do that many passes for all slashing and spraying activities.
 

### 2023-10-04 article.tex
answered basic questions re intro

@@
 % TODO:
-% 
-% Below is the original comment. BUT you have since come up with more of a scheme that answers these questions (in the bayesian paper). Use this to extrapolate on the introduction.
-% 
-% You obviously have no leading question for this paper, and although you do a good job at creating accurate models; it is more important that the research is asking a good question.
-A cool question as Kate puts it.
-%
-% The questions I put forward to you are:
-%
-%  How do winegrowing regions differ?
-%  Why is a region different?
-%  How do regional differences affect winegrowing?
-%  What differences are important between winegrowing regions?
-%  
-%%
-
+% %%%%%%%%%%%%%%%
+% Paper setup %
+%%%%%%%%%%%%%%%
+
+% What is your topic?
+Economic sustainability of vineyards.
+
+% What is the context for your reserach?
+An unprecedented amount of data regarding the Australian wine and vine industry has been collected through Sustainable Winegrowing Australia offering new insights into the driving economic forces of the Australian wine industry.
+
+% What do you want to achieve?
+We hope to uncover the driving relationships and factors of economic sustainability in Australian Winegrowing.
+
+% Are there any relationships you want to explore?
+We explore the relationship between various operational and resource use variables to operational costs and vineyard gross margin.
+
+%%%%%%%%%%%%%%%%
+% Working title
+%
+
+% Description
+% NB Avoid titles along the lines of: “Effects of ...”, “The role of ...”, etc.  Be specific about the effect and its significance so that your reader knows what is on offer.
+%  TODO:
+
+% For example, rather than write a title like “The effect of factor X on astrophysical properties of green cheese.”,
+
+% be specific about the effect and write something more like ...
+
+% “Factor X halves the lunar thermal diffusivity of  green cheese”.
+
+% You will usually find it easier to write an effective title if you make your title a sentence.
+
+% Notes
+% N/a
+
+%%%%%%%%%%%%%%%%%%%%
+% Intended readers
+%
+% Agricultural economists
+% Growers
+% Property managers (beyond winegrowing)
+% Mardi seemed super interseted - I know a lot of growers would be interested, however I am not sure of any names in particular beyond the usual cases such as Hans, David Klassen etc.
+% Perhaps those seeking to purchase a property.
+
+% Description:
+% Name 4 or 5 potential readers - give their names and why they would be interested  (e.g.
+“Ichabod Crane, paleo-fudgologist interested in polygalactic fudginomiality”, not “assorted paleo-fudgologists”).
+
+% Your readers should be outside your institution.
+
+%%%%%%%%%%%%%%%%%%%%%
+% Anticipated Journal
+%
+% Wiley Journal of Agricultural Economics
+% https://onlinelibrary.wiley.com/journal/14779552
+
+% Descrption
+% Ensure that all readers are likely to read the nominated journal  (e.g.  few non-researchers read refereed academic journals; politicians simply don’t read).
+
+% Notes
+% Contributors to the JAE should note that, in choosing subjects for articles, the objects of the Society are ‘to promote the study and teaching of agricultural economics and relevant disciplines, and their application to issues in the agricultural, food and related industries, rural communities and the environment.
+The relevant disciplines include economics, statistics, marketing, business management, politics, history and sociology'.
+Preference will be given to articles which deal either with new developments in research and methods of analysis, or apply existing methods and techniques to new problems and situations which are likely to be of general interest to the Journal’s international readership.
+The Journal is not a suitable publication for articles of purely local interest.
+Articles should be clearly and concisely written and should normally not exceed 7,000 words.
+Articles greatly in excess of this limit cannot be considered for publication.
+
+%%%%%%%%%%%%
+% Question %
+%%%%%%%%%%%%
+
+% Things to consider
+%
+% Why is the knowledge important?
+A better understanding of driving economic factors within agricultural industries helps to drive sustainability through understanding relationships between factors better.
+%  FIXME: work on this^
+
+% What is the significance?
+This broad dataset has not yet been analysed, showing new insights into the driving factors of the Australian winegrowing industries economy
+
+% How will the findings be utilised?
+These findings can be used by growers and property managers to help understand the driving factors that affect their operational costs and gross margins.
+
+% What improvements may be derived from this result?
+These insights help to optimise the gross margin and minimise the operational cost of a vineyard.
+
+% Are the terms well defined?
+% Some terms need more definition.
+
+% Is it doable? (cost & ethics)
+%  ITs done!
+%       Can you finish it in time?
+% Do experts think your question is important/relevant/doable?
+%  Mardi was pretty excited about it!
+
+% What is the nature of your question?
+%         who, what, where, when, why, how?
+The primary nature of this question is what, and then how: What are the driving factors for operational cost? And, how do these factors relate to each other?
+
+% You can have more than one question.
+
+% What is the most important question your paper will pose?
+What are the driving factors for operational cost and gross margin in Australian Winegrowing?
+
+%
+% NB It is essential that your answer is framed as a direct question.  Your response must end with a question mark.
+
+% Why is the question/issue/problem worth examining?
+There is a lot of economic pressure on Australian winegrowers and the wine and vine industry in general due to international market changes, tighter requirements for organic and sustainability credentials and a shrinking market.
+% FIXME: Get some references in here!
+
+% Briefly outline the problem you are tackling and explain why the problem is important to knowledge in general.
+“Nothing much is known” is not sufficient justification by itself.
+You have to show why the gap in knowledge is important.
+Expect to draw heavily on your reading of the literature in framing your answer but do not get into detail of  author and year.
+
+%%%%%%%%%%
+% Answer %
+%%%%%%%%%%
+
+% What is the answer to question
+Region is a driving force in determining profits.
+Fuel and water are the primary driving operational costs
+
+%
+% 
+% NB  You must give a direct answer to the question posed.
+% Notes
+
+% Will the findings be considered significant?
+The finding themselves are not that significant, it is the evidence of a complex system governing these relationships that is the significant finding.
+
+% How did you gather the evidence? 
+%
+% Briefly outline the methods you used to gather your evidence.
+They were gathered by Sustainable Winegrowing Australia
+
+% What is the main evidence?
+%
+% Briefly outline the key results.  Focus on outcomes.
+
+% What can you add to theory?
+%  
+% A research paper has to add to broader understanding. What will yours contribute?  Think about how your results and conclusions will change how people see the world.
+What we add is that agricultural economics maybe complex systems that are obfuscated by the expertise of growers to navigate problems such as drought, pests and disease.
+These systems, although predictable through some variables are not causal.
+We see this in the evidence of some years not affecting outcomes even though they are known to contain the aforementioned issues.
+% TODO: use the evidence of draught to be an example.
+Show why year should be important but is not!
+
+%
+% Many people have trouble with this section.
+Do not recycle the results.
+Focus on the conceptual models that explain why your results are as they are, or why they are different from what might have been expected.
+Your contribution may be something new or it may be confirmation of something already known but in a slightly different context.
+%
+% Sometimes the contribution to theory is not a simple answer but a better understanding of the questions that ought to be asked in future.
+%
+% Again, expect to draw heavily on the literature in framing your answer, but cite the literature only sparingly here (you can go into full detail when you prepare your discussion).
+
+% What can you add to practice?
+%
+% Superior research also has practical consequences.
+What are the consequences of your work?
+Think about how your results and conclusions might change what people do.
+Do not merely restate your results.
+It maybe better to pursue more causal models and create decision support systems than to model factors directly.
+It is difficult to untangle the predictive and correlative nature of the variable compared to the causal reasons.
+
+%%%%%%%%%%%%%%%
+% Future work %
+%%%%%%%%%%%%%%%
+
+% What remains unresolved?
+% 
+% You may or may not have a lot to say here.  Some of it may be useful in your discussion.
+The causal reasoning behind these variables behaviours. The untangling the multicollinearity of these variables.
 % FIXME:
 % Whats been done? Whats the persistant gap? Why is a data-driven approach the answer?
+A data driven approach can highlight the significance between different factors. And, highlight how heavily these factors relate through their relative importance.
+% Predictive Modeling: When the research goal involves making predictions or forecasts, a data-driven approach, particularly using machine learning algorithms, can analyze historical data to build predictive models.
+These models can be used to make future predictions or test hypothetical scenarios.
+% Exploratory Research: When the research area is relatively unexplored, and there is limited theoretical framework or prior research to build upon, a data-driven approach can help in exploring patterns, trends, and relationships in the available data.
+This can lead to the formulation of new hypotheses and research directions.
+% Large and Complex Datasets: When dealing with large volumes of data that are too complex to be analyzed using traditional methods, data-driven techniques such as machine learning algorithms, data mining, or big data analytics can help in uncovering meaningful insights that might not be apparent through conventional approaches.
 

### 2023-10-05 analysis.py
adding graph of op vs tgr

@@ -504,11 +504,22 @@ for y_name in files:
 
 # We also do the predicted variables!
 
-train_model_b(df[df["profitable"].notnull()][cols + ["profitable"]]
-                , "profitable")
+# These are artefacts. As profit can be negative it is harder to predict! We know after some further mapping that both operational costs and gross margin have a cubic root relationships to the logarithm of other variables. An interesting relationship to have!
+# 
+# Importantly this cubic root relationship exists between operational costs and gross margin, which is worth noting. However we want to know what tips it either side of them being equal as that shows that a vineyard is profitable!!
+# 
+# train_model_b(df[df["profitable"].notnull()][cols + ["profitable"]]
+#                 , "profitable")
 
-train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
-        , "profit")
+# train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
+#         , "profit")
 
 train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
                 , "total_operating_costs")
+train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
+                , "total_operating_costs")
+
+train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
+                , "total_grape_revenue")
+train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
+                , "total_grape_revenue")

### 2023-10-05 opc_vs_tgr.r

@@ -1,0 +1,61 @@
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+library(sf)
+library(cowplot)
+library(stringi)
+
+# this comes from some interesting relationships.
+#
+# In [55]: plt.scatter(df_floats["total_grape_revenue"].apply(lambda x: math.sqrt(math.sqrt(x))), df_floats["total_operating_costs"])
+#     ...: plt.show()
+
+# In [56]: plt.scatter(df_floats["total_grape_revenue"].apply(lambda x: math.sqrt(math.sqrt(x))), df_floats["total_operating_costs"].apply(lambda x: math.sqrt(math.sqrt(x))))
+#     ...: plt.show()
+
+# In [57]: plt.scatter(df_floats["total_grape_revenue"].apply(lambda x: math.sqrt(math.sqrt(x))), df_floats["water_used"].apply(lambda x: np.log(x+1)))
+#     ...: plt.show()
+;z;l
+df <- read.csv("dfb.csv")
+df <- df[df$total_operating_costs != 0, ]
+df <- df[df$total_grape_revenue != 0, ]
+
+total <- ggplot(
+    df
+    , aes(
+        x = total_operating_costs^(1 / 3)
+        , y = total_grape_revenue^(1 / 3)
+        , col = giregion)
+        ) +
+        xlim(-2.5, 2.5) +
+        ylim(-2.5, 2.5) +
+        geom_point(size = .5) +
+        xlab("Total Operating Costs") +
+        ylab("Total Grape Revenue") +
+        ggtitle("Operating cost vs Revenue") +
+        theme_light() +
+        theme(panel.grid.minor = element_blank()
+            , legend.title = element_blank()
+            , legend.position = "top", legend.direction = "horizontal")
+
+by_area <- ggplot(
+    df
+    , aes(
+        x = total_operating_costs^(1 / 3)
+        , y = total_grape_revenue^(1 / 3)
+        , col = giregion)
+        ) +
+        xlim(-2.5, 2.5) +
+        ylim(-2.5, 2.5) +
+        geom_point(size = .5) +
+        xlab("Total Operating Costs") +
+        ylab("Total Grape Revenue") +
+        ggtitle("Operating cost vs Revenue by area") +
+        theme_light() +
+        theme(panel.grid.minor = element_blank()
+            , legend.title = element_blank()
+            , legend.position = "top", legend.direction = "horizontal")
+
[…]

### 2023-10-09 data_prep.py
added data as a ratio area.

@@ -7,8 +7,10 @@ df = pd.read_feather("df.feather")
 # we get the indices of those that made profit
 indices = df[
     (df["total_operating_costs"]>0) & (df["total_operating_costs"]>0)].index
+
 # we set them to nan as oppose to 0 because profit will take negative values
 df["profit"] = np.nan
+
 # calculate profit
 df.loc[indices, "profit"] = df.loc[
     indices, "total_grape_revenue"] - df.loc[indices, "total_operating_costs"]
@@ -148,59 +150,9 @@ for col in cols_to_transform:
 
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfb.csv")
 
-# a mixture!
-bin_cols = [
-    "bare_soil"
-    , "annual_cover_crop"
-    , "permanent_cover_crop_native"
-    , "permanent_cover_crop_non_native"
-    , "permanent_cover_crop_volunteer_sward"
-    , "irrigation_energy_diesel"
-    , "irrigation_energy_electricity"
-    , "irrigation_energy_pressure"
-    , "irrigation_energy_solar"
-    , "irrigation_type_dripper"
-    , "irrigation_type_flood"
-    , "irrigation_type_non_irrigated"
-    , "irrigation_type_overhead_sprinkler"
-    , "irrigation_type_undervine_sprinkler"
-    , 'river_water'
-    , 'groundwater'
-    , 'surface_water_dam'
-    , 'recycled_water_from_other_source'
-    , 'mains_water'
-    , 'other_water'
-    , 'water_applied_for_frost_control'
-    , "nh_frost"
-    , "nh_disease"
-]
+# We also create a set of data that is a ratio to area:
 
-for col in bin_cols:
-    df_floats[col] = bin_col(df[col])
-
-prop_cols = [
-    'vineyard_area_white_grapes'
-    , "area_not_harvested"
-    , 'vineyard_area_red_grapes'
-]
-for col in prop_cols:
-    df_floats[col] = df[col].div(df["vineyard_area"], axis=0)*10000
-
-# These are a kind of proportion 2 passes is 200%
-df_floats["total_tractor_passes"] = df["total_tractor_passes"]
-df_floats["slashing_number_of_times_passes_per_year"] = df["slashing_number_of_times_passes_per_year"]
-df_floats["fungicide_spraying_number_of_times_passes_per_year"] = df["fungicide_spraying_number_of_times_passes_per_year"]
[…]

### 2023-10-09 opc_vs_tgr.r

@@ -15,7 +15,7 @@ library(stringi)
 
 # In [57]: plt.scatter(df_floats["total_grape_revenue"].apply(lambda x: math.sqrt(math.sqrt(x))), df_floats["water_used"].apply(lambda x: np.log(x+1)))
 #     ...: plt.show()
-;z;l
+
 df <- read.csv("dfb.csv")
 df <- df[df$total_operating_costs != 0, ]
 df <- df[df$total_grape_revenue != 0, ]

### 2023-10-09 analysis.py
removed duplicate lines at the end.

@@ -518,8 +518,3 @@ train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"
                 , "total_operating_costs")
 train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
                 , "total_operating_costs")
-
-train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
-                , "total_grape_revenue")
-train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
-                , "total_grape_revenue")

### 2023-10-09 analysis.py
commenting out loop of all variables.

@@ -487,20 +487,22 @@ cat_cols = [ # These are binary
     , "giregion"
 ]
 
+# We comment out these unless we really need to redo every single variable. As they are not compared to the target variables at the end!
+# 
 # cols += cat_cols
-print(files)
-for y_name in files:
-    print(y_name)
-    if y_name in cat_cols:
-        if y_name in ["nh_disease", "nh_frost"]:
-            train_model_b(df[cols]
-                , y_name)
-        else:
-            train_model_multi(df[cols]
-                , y_name)
-    else:
-        train_model_reg(df[cols]
-                , y_name)
+# print(files)
+# for y_name in files:
+#     print(y_name)
+#     if y_name in cat_cols:
+#         if y_name in ["nh_disease", "nh_frost"]:
+#             train_model_b(df[cols]
+#                 , y_name)
+#         else:
+#             train_model_multi(df[cols]
+#                 , y_name)
+#     else:
+#         train_model_reg(df[cols]
+#                 , y_name)
 
 # We also do the predicted variables!
 

### 2023-10-09 analysis.py
temp minor change to fix duplicate line

@@ -516,7 +516,8 @@ cat_cols = [ # These are binary
 # train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
 #         , "profit")
 
-train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
-                , "total_operating_costs")
-train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
-                , "total_operating_costs")
+# train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
+#                 , "total_operating_costs")
+
+train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
+                , "total_grape_revenue")

### 2023-10-09 analysis.py
removed the temp comment markers

@@ -516,8 +516,8 @@ cat_cols = [ # These are binary
 # train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
 #         , "profit")
 
-# train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
-#                 , "total_operating_costs")
+train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
+                , "total_operating_costs")
 
 train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
                 , "total_grape_revenue")

### 2023-10-09 data_prep.py
fixed missing axis=0

@@ -153,6 +153,6 @@ pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfb.csv")
 # We also create a set of data that is a ratio to area:
 
 area = df_floats["area_harvested"]
-df_floats = df.div(df_floats["area_harvested"]*10000)/10000
+df_floats = df.div(df_floats["area_harvested"]*10000, axis=0)/10000
 df_floats["area_harvested"] = area
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfa.csv")

### 2023-10-09 data_prep.py
WOW

@@ -153,6 +153,6 @@ pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfb.csv")
 # We also create a set of data that is a ratio to area:
 
 area = df_floats["area_harvested"]
-df_floats = df.div(df_floats["area_harvested"]*10000, axis=0)/10000
+df_floats = df_floats.div(df_floats["area_harvested"]*10000, axis=0)/10000
 df_floats["area_harvested"] = area
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfa.csv")

### 2023-10-09 data_prep.py
removed infinity error

@@ -155,4 +155,5 @@ pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfb.csv")
 area = df_floats["area_harvested"]
 df_floats = df_floats.div(df_floats["area_harvested"]*10000, axis=0)/10000
 df_floats["area_harvested"] = area
+df_floats = df_floats.replace({np.inf: np.nan, -np.inf: np.nan})
 pd.concat([df_floats, df_o, df_int], axis=1).to_csv("dfa.csv")

### 2023-10-10 article.tex
notes on edits
additions to indtroduction

@@
 \begin{highlights}
- \item ???
- \item ???
- \item ????
- \item ????
+ \item Highlight 1
+ \item Highlight 2
+ \item Highlight 3
+ \item Highlight 4
 \end{highlights}
@@
 
-% FIXME: Stop writing chronologically. Write by significance.
-% TODO: Should the models be redone as a factor of area? instead of including area? The ensemble should effectively pick this up I do not think it needs to be done.
-
 \section{Introduction}
 
-% TODO:
-% %%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%
 % Paper setup %
@@
 
-% What is your topic?
-Economic sustainability of vineyards.
+% Why is the question/issue/problem worth examining?
+There is a lot of economic pressure on Australian winegrowers and the wine and vine industry in general due to international market changes, tighter requirements for organic and sustainability credentials and a shrinking market.
+% FIXME: Get some references in here!
+
+ There is a need for increased understanding of different underlying conditions leading to improved performance in agricultural productivity and sustainability \citep{oecdInnovationProductivitySustainability2019}.
+
+ Specifically within the Australian Vine and Wine industry there is a need to further understand the driving relationships between resource use and economic output .
+ Environmentally sustainable
+ UC Davis in California has created a new green-tech winery, a sustainable entity and blueprint
+ on how to differentiate a brand. Growers need to determine better and efficient ways to make
+ wine, develop benchmarks with local growers and market this approach. \citep{lukemanciniUnderstandingAustralianWine2020}
 
@@
 Articles greatly in excess of this limit cannot be considered for publication.
+% 
+% A definition of agricultural economics from the encyclopedia britanica:
+% 
+% agricultural economics, study of the allocation, distribution, and utilization of the resources used, along with the commodities produced, by farming.
+Agricultural economics plays a role in the economics of development, for a continuous level of farm surplus is one of the wellsprings of technological and commercial growth.
+
+We study the economic outcomes of Australian vineyards and their statistical relationships to the utilisation of the resources used.
+This is done through analysing a new comprehensive nationwide data set using XGBoosted models.
+We further compare the relationships between different resources to address the extensive collinearity found within the data.
+XGBoosted models were used because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have; with importance being able to be statistically defined through different measures.
+
+applies existing techniques to a new data set
+
+agricultural economics and relevant disciplines, and their application to issues in the agricultural, food and related industries, rural communities and the environment.
+The relevant disciplines include economics, statistics, marketing, business management, politics, history and sociology'.
+Preference will be given to articles which apply existing methods and techniques to new problems and situations which are likely to be of general interest to the Journal’s international readership.
+The Journal is not a suitable publication for articles of purely local interest.
+Articles should be clearly and concisely written and should normally not exceed 7,000 words.
+Articles greatly in excess of this limit cannot be considered for publication.
 
@@
 
-% Why is the question/issue/problem worth examining?
-There is a lot of economic pressure on Australian winegrowers and the wine and vine industry in general due to international market changes, tighter requirements for organic and sustainability credentials and a shrinking market.
-% FIXME: Get some references in here!
-
 % Briefly outline the problem you are tackling and explain why the problem is important to knowledge in general.
@@
 Expect to draw heavily on your reading of the literature in framing your answer but do not get into detail of  author and year.
+% 
+% ^ is this almost boiler plate writing at this point.
+Considering that not much is known but we are applying it in a specific field, in this case to economics.
+We can use a general, nothing is known but further understanding this for economic reasons will help to improve the economic sustainability of these practices.
+and that these authors have postulated about these connections etc.
+We can start generic and funnel it into the reasoning of importance using the journals general frame work and objectives.
 
@@
 % Large and Complex Datasets: When dealing with large volumes of data that are too complex to be analyzed using traditional methods, data-driven techniques such as machine learning algorithms, data mining, or big data analytics can help in uncovering meaningful insights that might not be apparent through conventional approaches.
+
+% TODO:
+%  From the code comments~ We might want to show the relationship between revenue and operational cost! then we have an idea of how these two are related.
+% # These are artefacts of.
+As profit can be negative it is harder to predict!
+We know after some further mapping that both operational costs and gross margin have a cubic root relationships to the logarithm of other variables.
+An interesting relationship to have!
+% # 
+% # Importantly this cubic root relationship exists between operational costs and gross margin, which is worth noting.
+However we want to know what tips it either side of them being equal as that shows that a vineyard is profitable!!
 

### 2023-10-10 article.tex
Introduction!

--- 2023-10-03 article.tex
@@
 %%        main text       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% FIXME: Stop writing chronologically. Write by significance.
-% TODO: Should the models be redone as a factor of area? instead of including area? The ensemble should effectively pick this up I do not think it needs to be done.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%
+% Working title
+%
+
+% Description
+% NB Avoid titles along the lines of: “Effects of ...”, “The role of ...”, etc.  Be specific about the effect and its significance so that your reader knows what is on offer.
+%  TODO:
+
+% For example, rather than write a title like “The effect of factor X on astrophysical properties of green cheese.”,
+
+% be specific about the effect and write something more like ...
+
+% “Factor X halves the lunar thermal diffusivity of  green cheese”.
+
+% You will usually find it easier to write an effective title if you make your title a sentence.
+
+% Notes
+% N/a
+
+\section{Abstract}
+
+Region is a driving force in determining profits.
+Fuel and water are the primary driving operational costs
+
+The finding themselves are not that significant, it is the evidence of a complex system governing these relationships that is the significant finding.
+
+They were gathered by Sustainable Winegrowing Australia
+
+What we add is that agricultural economics maybe complex systems that are obfuscated by the expertise of growers to navigate problems such as drought, pests and disease.
+These systems, although predictable through some variables are not causal.
+We see this in the evidence of some years notaffecting outcomes even though they are known to contain the aforementioned issues.
 
@@
 
+Historically strong demands for Australian wine have helped to create a thriving industry, however recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation has negatively effected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
+The 2021-2022 financial year alone saw a decline of 19\% in exports solely due to tariffs \citep{wineaustraliaNationalVintageReport2022}.
+A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale are key to  introducing stronger policy and information to aid in increasing a nations agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
+Specifically within the Australian Wine and vine industry there is a need to further understand the driving relationships between resource use and economic output.
+Where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers.
+\citep{lukemanciniUnderstandingAustralianWine2020}
+
+An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
+We use this data to study the economic outcomes of Australian vineyards and their statistical relationships to vineyards' utilisation of the resources.
+Answering what the driving factors are behind vineyard economic outcomes, and linking these outcomes to predictor importance.
+This is done through analysing a new comprehensive nationwide data set using XGBoosted models.
+We further compare the relationships between different resources to address the extensive collinearity found within the data FIXME: This needs references||.
+XGBoosted models were used because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables; with importance being able to be statistically defined through multiple methods.
+
+%%%%%%%%%%%%%%%
+% Future work %
+%%%%%%%%%%%%%%%
+
+% What remains unresolved?
+% 
+% You may or may not have a lot to say here.  Some of it may be useful in your discussion.
+The causal reasoning behind these variables behaviours. The untangling the multicollinearity of these variables.
+
 % TODO:
-% 
-% Below is the original comment. BUT you have since come up with more of a scheme that answers these questions (in the bayesian paper). Use this to extrapolate on the introduction.
-% 
-% You obviously have no leading question for this paper, and although you do a good job at creating accurate models; it is more important that the research is asking a good question.
-A cool question as Kate puts it.
-%
-% The questions I put forward to you are:
-%
-%  How do winegrowing regions differ?
-%  Why is a region different?
-%  How do regional differences affect winegrowing?
-%  What differences are important between winegrowing regions?
-%  
-%%
-
-% FIXME:
-% Whats been done? Whats the persistant gap? Why is a data-driven approach the answer?
-
-In the past decade, the Australian grape and wine industry has undergone a variety of pressures, including changing market demands, disease and drought \citep{wineaustraliaAustralianWineProduction2021}.
-Furthermore, natural resources are likely to decrease, as pressures from climate change increase, making it more important than ever to improve the efficiency and sustainability of crops \citep{agdeeNationalGreenhouseAccounts2021}.
-It has become crucial for those in the wine industry to address issues relating to environmental sustainability and economic viability, with a growing need for the industry to close the research gaps between sustainable practices and their real and perceived environmental and economic advantages \citep{montalvo-falconSustainabilityResearchWine2023,ouvrardDoesSustainabilityPush2020}.
-This paper aims to provide a comprehensive analysis of the intricate relationships between economic and environmental variables within the winegrowing sector.
-By employing statistical machine learning we show how interconnected vineyard variables are.
-Utilising a ten year data set that spans Australia we show the predominant
-elements in classifying region and year, % FIXME: A reader will ask: why would i want to predict region and year?
- as well as determining a vineyards potential operational costs and profit.
-\par
-This analysis utilises 
-XGBoosted trees % FIXME: Why XGBoosted trees?
- to classify region and year; and to determine operating cost and profit of Australian vineyards. Classification and regression trees are 
- utilised as surrogate models %FIXME: Why are we utilising surrogate models?
-to shed insight into the key partitions used by the XGBoost ensembles. Variable importance is further used to illustrate the 
-interconnectedness % FIXME: What kind of interconnectedness?
-of the different predictor variables, and to show the similarity in
-variable importance % FIXME: What does variable importance mean in this context?
-between different response variables (particularly between region and operational cost). This study aims to assist in uncovering 
-the complexity of variables %FIXME: This is too vague, what makes them complex? why are the complex? etc
-that are affected by a variety of vineyard management decisions to illustrate 
-complex interplay of variables. % FIXME: What do you mean by complex interplay? Like what? An example?
-This study endeavours to gain insight into the 
-similarity % FIXME: Why do we want to know the similarity?
- in predictor variable importance between year, region, operational costs and profit.
-
-% Initiatives to improve the environmental impacts of the Australian Wine industry have been ongoing since the 1990s; with recommendations to implement strategies for national programs that would help educate industry members in sustainable practices (Keith Jones, 2002).
-The first legislative initiative to support national programs was the Environmental Management Guidelines for Vineyards in Western Australia during 2002 (Nind, 2002).
-The first national program was the Australian Wine Industry Stewardship, beginning shortly after, during 2004 (WFA, 2009).
-This program was replaced multiple times during the 2000s, first by the Australian Wine Environmental Stewardship, and then by EnviroWine, finally becoming EntWine in 2009.
-The period in the 2000s was further influenced by several key researchers and program leaders, where concurrently multiple regions developed approaches to sustainable management tied to sustainability programs (SWA, 2023).
-EntWine was eventually acquired by the Australian Wine Research Institute and Australian Grape & Wine during 2015; which would lead to its redevelopment into Sustainable Winegrowing Australia (SWA) during 2019/2020 (SWA, 2023).
-%             2.1.2. Previous Research and Motivation
-% During the period of 2010 to 2023 data was collected voluntarily from wine industry members through the national sustainability programs, ultimately being in SWA possession as of 2023.
-The data set was used in 2020 to analyse sustainable practices, resulting in the development of a Resource Intensity Score (RIS); where the RIS was a summary indicator value of a vineyard’s energy use, emissions and water use (Nayak et al., 2020).
-The analysis then fit vineyards’ RIS values to 55 variables sourced from 23 separate wine regions during the period of 2015 to 2018 through the use of Stepwise Regression modelling.
-It was found that the top 10% gross margin vineyards had a lower operating cost and a lower RIS score than vineyards in the bottom 10% gross margin.
-% The motivation for this thesis is to investigate the differences between the top 10% gross margin vineyards and those of the bottom 10%, hoping to uncover what the key differences were and to relate them to sustainable practices; as there is a known gap in the literature due to the difficulty to evaluate relationships using empirical evidence of both the environmental and economic impacts of sustainable practices in viticulture.
-This difficulty stems from data scarcity, industry complexity and transparency.
-While there have been attempts to fill these gaps, methodologies and indicators have lacked in ability to compare different regions and scales (Baiano, 2021).
-%             2.1.3. Economic Pressures
-% Historically strong demands for Australian wine have helped to create a thriving industry, however recently sharp reductions in exports to mainland China due to significant deposit tariffs have caused  a decline of 19% in Australian export value during the 2021-2022 financial year (see Figure 1) (Wine Australia, 2022).
-The pressure brought on by the drop in export value has been exacerbated by loss of tourism and labour due to the COVID-19 pandemic, global freight crisis, war in Europe and rising inflation (Wine Australia, 2021a, 2020).
-These pressures within the wine industry have trickled down to vineyards, where winemakers retain unwanted wine, creating an oversupply of grapes.
-Currently the proposed strategy to soften pressures proposed by Wine Australia includes initiatives focused on market diversification, and sustainability to build resilience in the coming years.
-% Figure 1:  The exports of Australian wine over time in Australian Dollars Free On Board, comparing exports between China and the rest of the world.
-This graphic is taken from the Wine Australia Annual Report of 2020-21(Wine Australia, 2022).
-%             2.1.4. Environmental Pressures
-% There are several environmental concerns that affect viticulture, including loss of soil quality, lack of rain, hail, disease, fire, and frost; with climate change exacerbating these issues.
-In 2020, 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase (Canadell et al., 2021).
-In comparison to countrywide pressures such as drought, this damage made up only 3% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought (Wine Australia, 2020).
+%  From the code comments~ We might want to show the relationship between revenue and operational cost! then we have an idea of how these two are related.
+% # These are artefacts of.
+As profit can be negative it is harder to predict!
+We know after some further mapping that both operational costs and gross margin have a cubic root relationships to the logarithm of other variables.
+An interesting relationship to have!
+% # 
+% # Importantly this cubic root relationship exists between operational costs and gross margin, which is worth noting.
+However we want to know what tips it either side of them being equal as that shows that a vineyard is profitable!!
 
@@
 
+It maybe better to pursue more causal models and create decision support systems than to model factors directly.
+It is difficult to untangle the predictive and correlative nature of the variable compared to the causal reasons.
+
 This study has provided valuable insights into the multifaceted dynamics governing operational costs, different yearly effects and vineyard regions highlighting the complex interrelatedness of variables within a vineyard.

### 2023-10-10 article.tex
backing up before leaving

@@
 \citep{lukemanciniUnderstandingAustralianWine2020}
-
+\newline
 An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
-We use this data to study the economic outcomes of Australian vineyards and their statistical relationships to vineyards' utilisation of the resources.
+This dataset allowed insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue within the data.
+We use this data to study these economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
 Answering what the driving factors are behind vineyard economic outcomes, and linking these outcomes to predictor importance.
@@
 
-%%%%%%%%%%%%%%%
-% Future work %
-%%%%%%%%%%%%%%%
-
-% What remains unresolved?
-% 
-% You may or may not have a lot to say here.  Some of it may be useful in your discussion.
-The causal reasoning behind these variables behaviours. The untangling the multicollinearity of these variables.
-
+\section{Methods}
 % TODO:
@@
 
-% \if{false}
-
-% We classify regions to find the most prominent and well-defined relationships to regions.
-% We examine the variables that are the most decisive in determining region.
-% We compare the similarities and differences of these regions
-% - within the analysis and
-% - within the literature (also in their definition and history)
-
-% We examine the similarity between these classifications and the variables that are significant in classifying profitable vs not.
-% How do regional changes affect that. What is similar and what is different.
-
-% We can link these contributing factors back to the baselines - at least anecdotally, or in an unpublished section. This will be good for the thesis.
-
-% \fi
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
-Each region describing unique reputations, qualities and varietals of wine produced there.
-While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
-In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
-% \newline
-
-%
-%This analysis addresses the knowledge gap regarding the effectiveness of regional level strategies employed in the wine industry and their relation to grape quality.
-% What the hell is a regional strategy
-% And how do you define, or know what trategy each region is using?
-%
-% Through the use of classification trees this study aims to highlight the key differences in sustainable practices at a regional level and how these practices relate to the different grades of grape quality.
-
-% This does not really address what the methods are - aside from classisifcation trees in the most vague sense.
-% What are the results briefly? why should someone read this article?
-% What is interesting - what will be discussed in the article with regard to these results? What makes me want to read the article?
-% And what is the take away point of the article? If I cant be bothered to read this article, what piece of knowledge can you leave mewiwth that I could recommend the article to someone else with?
-
-\section{Methods}
-
 % FIXME:
@@
 % As with the previous paper, put the models up front it is very unclear what you have created - and less clear to understand it if there is no way to reference to it.
-
-% FIXME:
-% In this methods section, what is novel/innovative? Not currently clear, In particular - if I gave this data & these packages to a 3rd year class, which part here couldn't they do?
 
@@
 
-% FIXME: Adde a section for CV.
-
-% FIXME: Another section that needs to be added is one that describes what the sankey and chord diagrams are and what they represent.
-The reader needs to be prepped for this before the results.
-They need to be able to properly interpret them.
-
-% FIXME: Explain how some variables are one-hot-encoded and how this relates to variable importance. This is a significant resulti n the xgboost tree for the huntervalley
-
 % FIXME: the motivation is missing from methods.
-% FIXME: What is interrelated importance?
 
 \subsection{Data}
-
 % 6091 samples - total
@@
 % 10 years
-
+% 
 % 853 with reported profits
@@
 The program aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by the SWA is entered manually by winegrowers using a web based interface tool.
+Data recorded by SWA is entered manually by winegrowers using a web based interface tool.
 A total of 6091 observations were collected from 2012/2013 to 2021/2022 financial years.
-23 variables were used for each observation reflecting a vineyards account for the given year (see Table \ref{tab:vars}).
-The profit variable was transformed % FIXME: What was it before?
-and included as another separate variable, profitable; depicting whether a vineyard was profitable or not.
+23 variables were used for each observation reflecting a vineyards state for the given year (see Table \ref{tab:vars}).
 \par
-
 \begin{table}[] 
@@
   Water Type & Class & 6091 & 39 & \\
-  Profit & AUD & 853 & & \\
+  Grape Revenue & AUD & 853 & & \\
   Operating Costs & AUD & 853 & & \\ \bottomrule
@@
 The data originally contained only two multiclass variables: year and region.
-Variables that measured the same metric from different sources (such as water collected from rivers versus water from dams) were converted into multiclass variables representing the source.
-The total amount used from these variables was retained as a separate variable.
-Occurrences of multiple sources were defined as separate classes.
+Variables that measured the same metric from different sources (such as water collected from rivers versus water from dams) were converted into multiclass variables representing the source through one-hot-encoding.
+Changing each variable class into a binary value, with one indicating the presence of the class and zero indicating its absence.
+Occurrences of multiple sources were defined as their own separate classes.
+Where a class variable had a recorded amount the total amount used from these variables was retained as a separate variable; for example water used (in Mega Litres ) was also included alongside water source.
 \par
@@
 \par
-%
-% Mention one-hot-encoding.
-%
 
-%
-% How deep into the classes do we go?
-%
 \subsection{XGBoosted Trees}
-% FIXME: What do boosted trees do?
 XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-They were chosen for this analysis as they provide both a
-high predictive performance % FIXME: For what?
-and ability to effectively 
-capture complex relationships % FIXME: How?
-.
+XGBoosted trees are a boosted tree ensemble method that can be used to classify classes, or predict continuous response variables.
+They were chosen for this analysis as the data contained a mixture of class and continuous variables.
+And, XGBoosted trees are unaffected by multicollinearity, as well as offer high predictive performance for a wide variety of purposes \citep{chenXGBoostScalableTree2016}.
 An XGBoosted tree was created for each variable to show how they interacted.
-Each tree included all but the economic variables (profit and operating cost), which were only included once as response variables.
-% FIXME: what does this mean in terms of the method and outputs?
+Each tree included all but the economic variables (operating cost and revenue), which were only included once as response variables.
+This meant that every variable would have a measure of its importance to other variables, which was used to show the highly interrelated nature of variables within vineyards (see \ref{variableimp}).
 \par
@@
 The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
-% FIXME: All equations must be in the middle or end of sentances, not at the beginning.
+Each of $f_k$ functions is a classification or regression tree, such that all functions are in the set of all decision trees, given by $\mathcal{F}$, is defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
+Where each function corresponds to an independent tree structure $q$ of $\omega$ weights.
+Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf.
+The final prediction is determined by the sum of the score of the corresponding leaves, given by:
 \begin{equation}
- \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{K}(x_{i})}, \space f_{K}\in \mathcal{F},
+ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{k}(x_{i})}, \space f_{K}\in \mathcal{F},
 \end{equation}
@@
 % FIXME: The methods would benefit from some subsections (signposts)
-% FIXME: What does the equation do? Give intuition.
 
-where each function $f_K$ is a classification or regression tree, such that all functions are in the set of all decision trees $\mathcal{F}$, defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
-Where, $f_K$ corresponds to an independent tree structure $q$ of $\omega$ weights. % FIXME: Sentance fragment.
-Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf. 
-The final prediction % FIXME: What is this?
-is determined by the sum of the score of the corresponding leaves, given by $\omega$. 
-The set of functions % FIXME: You haven't described the tree as having a set of functions?
-%  I believe when you refer to the set of functions F you might benefit from putting the F in brackets here or being more explicit where the functions are.
-used by the tree is determined by minimising the 
-regularised % What is the regularised objective?
-objective function, given by:
+The set of functions, $\mathcal{F}$, used by the tree is determined by minimising a regularised objective function, given by:
 
@@
 
-The difference between the 
-prediction and actual variable % FIXME: what are the predition and actual variables?
-% When you are talking broadly about elements of xgboost I believe it will help to refer to the equation, an example or something more specific.
-This would help to address the other comment regarding giving an intuition to the method.
-is a convex loss function $l$. To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration. The function 
-$f_t$ % FIXME: Kate asked if this was $f_K$.
+The difference between a prediction, $\phi(x_i)$, and actual variable, $f_{k}(x_i)$, is a differentiable convex loss function $l$.
+To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
+% \par
+% A function $f_t$ is selected according to which value minimises (\ref{eqn:objective}).
+The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
+% FIXME: Kate asked if this was $f_K$.
 This is not the case as $f_K$ is part of the penalty function, and I know this at first glance due to working with the equation regularly for this paper.
@@
 You are clearly not
-is selected according to which value minimises (\ref{eqn:objective}). The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
 \par
@@
 Due to the mixture of continuous, binary and multiclass variables in this analysis,
-both classification and regression trees were created. % FIXME: What were these created for?
+both classification and regression trees were created.
 The difference between the trees created for this analysis was the objective function used.
@@
 % FIXME: What is the output?
+
+% FIXME: What is interrelated importance? add a section for this!
+
+% FIXME: Explain how some variables are one-hot-encoded and how this relates to variable importance. This is a significant resulti n the xgboost tree for the huntervalley
+
 \subsection{Classification and Regression Trees} % FIXME: perhaps this should be renamed surrogate models
@@
 
+% FIXME: Another section that needs to be added is one that describes what the sankey and chord diagrams are and what they represent.
+The reader needs to be prepped for this before the results.
+They need to be able to properly interpret them.
+
+% FIXME: Adde a section for CV.
+
 \section{Results}
 
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
+Each region describing unique reputations, qualities and varietals of wine produced there.
+While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
+% \newline
 % FIXME: The first result should be in the langauge of the question.
@@
 
+%%%%%%%%%%%%%%%
+% Future work %
+%%%%%%%%%%%%%%%
+
+% What remains unresolved?
+% 
+% You may or may not have a lot to say here.  Some of it may be useful in your discussion.
+The causal reasoning behind these variables behaviours. The untangling the multicollinearity of these variables.
+
 It maybe better to pursue more causal models and create decision support systems than to model factors directly.

### 2023-10-10 analysis.py

@@ -255,7 +255,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
     
     # We need at least 3 entries into a region for it to be able to be classified.
 
-    X = df[df.groupby(y_name)[y_name].transform('count')>10].copy() 
+    X = df[df.groupby(y_name)[y_name].transform('count')>10].copy()
 
     y = X[y_name]
     y = y.cat.remove_categories(list(set(y.unique().categories) - set(y.unique())))

### 2023-10-11 article.tex
Methods edits

@@
 An XGBoosted tree was created for each variable to show how they interacted.
-Each tree included all but the economic variables (operating cost and revenue), which were only included once as response variables.
-This meant that every variable would have a measure of its importance to other variables, which was used to show the highly interrelated nature of variables within vineyards (see \ref{variableimp}).
+Each tree included all but the economic variables (operating cost and revenue), which were only included within their own trees as response variables.
+This meant that every variable would have a measure of its importance to other variables (see Section \ref{sec:importance}), which was used to show the highly interrelated nature of variables within vineyards.
+The complicated interaction between bariables was illustrated using Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between any two variables (see section \ref{sec:importance}).
 \par
@@
 
-% FIXME: The methods would benefit from some subsections (signposts)
-
-The set of functions, $\mathcal{F}$, used by the tree is determined by minimising a regularised objective function, given by:
+The set of functions, $\mathcal{F}$, used by the tree is determined by minimising a regularised objective function, $\mathcal{L}$ given by:
 
@@
  \mathcal{L}(\phi)=\sum_i l(\hat{y_i}, y_i^{t-1} + f_t(x_i)) + \sum_k \Omega (f_K).
+\end{equation}, where
+\begin{equation}
+  \Omega(f) = \gamma T + {{1}\over{2}}\lambda ||\omega||^2
 \end{equation}
 
+As predictions are made using additive tree functions, XGboosted trees can be used for classification or regression.
 The difference between a prediction, $\phi(x_i)$, and actual variable, $f_{k}(x_i)$, is a differentiable convex loss function $l$.
+These properties of $l$ allow the function to be versatile in which objective we choose to optimise for, which is also important in being able to process both continuous and categorical variables.
 To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
-% \par
-% A function $f_t$ is selected according to which value minimises (\ref{eqn:objective}).
-The model complexity is penalised by the function $\Omega$, this acts to smooth weights in an attempt to prevent over fitting.
-% FIXME: Kate asked if this was $f_K$.
-This is not the case as $f_K$ is part of the penalty function, and I know this at first glance due to working with the equation regularly for this paper.
-If Kate is confused about this then your explanation sucks.
-It is to obtuse, like Kate has said you are hitting people with too many facts and not padding it enough to the context of this paper and its application or even its purpose.
-The function $f_t$ is incredibly important, especially because for the purposes of this study it changes depending on the variable and its nature.
-You are clearly not
-\par
-As predictions are made using additive tree functions, XGboosted trees can be used for classification and regression.
-Due to the mixture of continuous, binary and multiclass variables in this analysis,
-both classification and regression trees were created.
-The difference between the trees created for this analysis was the objective function used.
-% FIXME: Kate marked this with a "?" clearly highlighting the issue with the previous comment regarding f_t (at some point was on line 321)
-XGBoosted regression trees were created for continuous variables, using the
+\subsection{Loss functions}
 
-% FIXME: The below section talks about the objective functions.
-Kate askes what they are and what they related to.
-You will need to go into the objective functions, how they are different and why different functions are required to be used.
-% Perhaps a reference that talks about the importance of objective functions.
-root-mean-square as the objective function.
-Binary class variables utilised the logistic loss function as the objective.
-And, Multiclass variable used the soft max function.
-All objective functions are defined within the SKlearn library \citep{sklearn_api}, linked via an API to the XGBoost library \citep{chenXGBoostScalableTree2016}.
-\par
-% FIXME: Why are you doing this? give me reasons related to this work to keep reading.
+The functions included as parameters in equation \ref{eqn:objective} mean that traditional opimisation methods for Euclidean space cannot be used.
 Chen and Guestrin \citep{chenXGBoostScalableTree2016} illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
-Furthermore, they show the loss reduction after
-the split % FIXME: What is the split?
-is given by the function:
+Importantly a loss function can be used to fit a model iteratively to data.
+For this analysis several loss functions were used, as variables took the form of continuous, binary and multi-call data.
+The loss function for making a split within the tree structure is given by:
 
@@
  {{(\sum_{i\in I_R}g_i)^{2}}\over{\sum_{i\in I_R} h_i + \lambda}} -
- {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma,
+ {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma.
 \end{equation}
 
-with the tree structure defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
+The tree structure being defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
 Instead of enumerating all possible tree structures, a greedy algorithm iteratively adds branches to the tree minimising $ \mathcal{L}_{split} $ in (\ref{eqn:enumerate}).
-The frequency of a variable's occurrence within a tree is directly attributed to the minimisation of the objective function (or loss) through the minimisation of $ \mathcal{L}_{split} $.
+The frequency of a variable's occurrence within a tree is directly attributed to the minimisation of the loss function through the minimisation of $ \mathcal{L}_{split} $.
 \par
-% FIXME: The below section needs to come sooner
-The frequency of a variable appearing as a node within the ensemble was used as a measure of importance.
-This measure was chosen as it connected a variable to the minimisation of its associated objective function, translating the value into a
-simple count metric. % FIXME: Variable importance might need to be in its own section to fully explain it and its motivation.
-% FIXME: The creation of xgboost trees for each variable needs to be earlier and more clear as well contextualised by its motivation.
+The loss functions used for this analysis were the root-mean-square function for continuous variables.
+The logistic loss function for binary class variables.
+And, the soft max function for Multiclass variables.
+All objective functions are defined within the SKlearn library \citep{sklearn_api}, which was utilised via an API to the XGBoost library \citep{chenXGBoostScalableTree2016}.
+\par
+\subsection{Variable Importance}\label{sec:importance}
+
+Due to XGBoost creating a large amount of decision trees, the interpretability of these models is obfuscated by the intricate relationships within complicated ensembles.
+A measure of variable importance was the technique used to highlight a variables influence within the ensemble.
+Variable importance can be measured in multiple ways; we used the frequency of a variable appearing as a node within the ensemble as a measure of its importance.
+This measure was chosen as it connected a variable to the minimisation of its associated objective function.
+The measure of a variable's importance within this study can then be interpreted as how often a variable was the optimal choice in reducing the loss function of the ensemble.
+Importantly, multiclass variables being one-hot-encoded are given an importance score for each individual class; for example, each specific region will have its own importance score.
+\par
 Creating XGBoosted trees for each variable allowed the use of importance to show how strongly variables were associated with each other.
-The importance of predictor variables to economic variables was illustrated through the use of
-Sankey diagrams constructed % FIXME: More needs to be said on this.
-More needs to be talked about regarding the chord diagrams as well.
-There is no thorough explanation of how the diagrams work and what you have done with them to fully explain them!
-using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
-Other variable's interconnectedness was demonstrated through the use of a chord diagram also created using Holoviews.
-% FIXME: The above comment about sankey diagrams applies to chord diagrams.
-These need fuller descriptions.
-\par
-% FIXME: More info is needed regarding the validation and how it was performed.
-This section can be expaned on as different validation and scoring is used due to the differencein objective functions.
-This is a rather critical section and very little is said about it.
-% 
-Each variable utilised 80\% of the data to train the XGBoost ensemble, with 20\% reserved for testing and validation.
-Testing was done through the iterative minimisation of the respective objective function for the variables type.
+The importance of variables to one another was illustrated through the use of Sankey and Chord diagrams.
+These diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+Both Chord and Sankey diagrams illustrated variable importance through the size of the bands between two variables.
+The number at the end of a connection in a Sankey diagram indicated a variable's importance, or the number of times it appeared within the ensemble.
+Sankey and Chord  diagrams were presented together; with Sankey diagrams showing the connection of a variable to its 10 most important predictor variables.
+Chord diagrams were used alongside a Sankey diagram to show the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
+Chord diagrams formed circles, with variables being connected through their relative importance.
+The importance values for the Chord diagrams were taking from the models of those individual variables, with the diagram being simplified to just the ten variables in the associated Sankey diagram, for readability's sake.
+\subsection{Validation}
+As there were multiple different loss functions, multiple different forms of validation were used.
+In each case the data was split into training data, which constituted 80\% of the original data.
+The remaining 20\% was used in testing and validation.
+Data was stratified when splitting the data into these subsets to conserve the same proportion of class occurrences between training, testing and validation data.
 For continuous variables 20\% was used as testing data, minimising the root-mean-square function.
@@
 $R^2$ scores were used to determine the best regression models during validation.
-For binary and multiclass variables data was split into 80\% training, 10\% testing and 10\% validation data. 
-Due to class disparity in multiclass variables (most prominently in region) data was stratified % FIXME: What does it means to stratify the data and why is this done?
-into each subset at the same ratio of class occurrence. Validation was summarised through the accuracy, the proportion of true negatives and positives.
+For binary and multiclass variables, data was split into 80\% training, 10\% testing and 10\% validation data.
+For class variables, validation was summarised through the accuracy, the proportion of true negatives and positives.
 \par
-% FIXME: rewrite the below paragraph in the active voice.
-The use of the XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
-The further use of cross validated grid search functions allowed for the selection of better performing hyperparameters when selecting the final model.
+The XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
+This allowed us to utilise cross validated grid search functions when selecting for better performing hyperparameters.
 The performance measure for model selection was root-mean-square error for continuous variables.
@@
 Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
-% FIXME: What is the output?
 
-% FIXME: What is interrelated importance? add a section for this!
+\subsection{Surrogate Models}
 
-% FIXME: Explain how some variables are one-hot-encoded and how this relates to variable importance. This is a significant resulti n the xgboost tree for the huntervalley
-
-\subsection{Classification and Regression Trees} % FIXME: perhaps this should be renamed surrogate models
-% FIXME: It is not clear what you are trying to achieve with these models. No where do you give a definition of what a surrogate model is and why you are using it!
-% Be specific about the purpose of what these models are being used for!
-Justify why you have not simply shown one of the ensembles trees as well!
-The use of these algorithms means that the tree is more general than any one of the ensembles manifestations.
-Classification and Regression Trees were created for region, year, profit and operating cost.
+The creation of more interpretable models such as linear regression in parallel to AI systems has been used to explain variable's relationships within black box algorithms \citep{molnarInterpretableMachineLearning2022}.
+As XGBoost create an ensemble of decision trees, here we use classification and regression trees to gain insight into intricacies of the ensembles derived through XGBoost.
+Decision Trees were created for operating costs, revenue and region.
 These models describe the partitions that are useful in predicting these variables; giving insight into the trees that make up the ensembles created by XGBoost.
@@
 The same measure of accuracy as the XGBoosted trees was used for comparison.
-
-% FIXME: Another section that needs to be added is one that describes what the sankey and chord diagrams are and what they represent.
-The reader needs to be prepped for this before the results.
-They need to be able to properly interpret them.
-
-% FIXME: Adde a section for CV.
-
 \section{Results}
@@
 Have a google of this to overcome it as the figure placement is all over the shop.
-
-\subsection{Region}
-% FIXME: there are artifacts from the previous paper.
-Although the previous paper is pertinant to some of this (quite critically so) it is no longer the primary focus.
-On top of this some of these terms such as 'Region Classification' are not well defined.
-This is because previously a great deal of effort was put into defining them and describing the potential regions.
-However this has been removed.
-% FIXME: This section is much more of a: 'isn't it weird?' That region is not as significant in some areas but not others.
-That is can be used to predict things like yield but no profit etc.
-This phrasing shows why it goes after other sections.
-if you are writting by significance then it is not a significant find but it is something that is worth noting - unlike year which was just oh dear we dont know what is going on.
-Region classification performed % FIXME: what does this mean - see the above comment.
-% Note that it is not about the implications but getting in the definition of these terms.
-These things should be prepared for the reader well before the results.
-There should be no new terms here, we should just be refering to the outcomes that were established in the methods, and they should already have their motivations and contexts estabished (there is more that Kate has written if you are confused this is my translation after our meeting).
-at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
-The most prominent feature used to classify regions with the classification tree was water source (see Figure \ref{fig:region_tree}).
-This differed from the variables that illustrated the greatest importance for the XGBoosted ensemble (see Figure (\ref{fig:region_sankey}), with predictor variables being highly interrelated in importance.
-Area, water, fuel and yield were more determining factors when predicting region using XGBoost.
-% FIXME: Is this perhaps your primary test?
-The fact that Kate was asking this means that the tools you are using do not have a well defined purpose in this analysis.
-I believe this stems from the lack of contextual motivation behind these tools.
-What is it that you are testing?
-Why are these important?
-This problem is throughout the paper and cannot simply be fixed here.
-Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
-\par
-
-Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
-The regions with the most samples performed the best.
-Bordering regions were routinely grouped together and misclassified as the same region.
-Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
-The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
-%, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
-The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
-These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
-Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
-
-\begin{figure}
- \resizebox{\textwidth}{!}{
- \includegraphics{region.pdf}}
- \caption{Decision tree predicting Region.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:region_tree}
-\end{figure}
-
-\begin{figure}
- \resizebox{\textwidth}{!}{
- \includegraphics{region.png}}
-% FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
-% FIXME: Mark one Panel A and the other B
- \caption{The left-hand side depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:region_sankey}
-\end{figure}
 
@@
 
-\subsection{Profit}
+\subsection{Revenue}
 
@@
 %
+
+\subsection{Region}
+% FIXME: there are artifacts from the previous paper.
+Although the previous paper is pertinant to some of this (quite critically so) it is no longer the primary focus.
+On top of this some of these terms such as 'Region Classification' are not well defined.
+This is because previously a great deal of effort was put into defining them and describing the potential regions.
+However this has been removed.
+% FIXME: This section is much more of a: 'isn't it weird?' That region is not as significant in some areas but not others.
+That is can be used to predict things like yield but no profit etc.
+This phrasing shows why it goes after other sections.
+if you are writting by significance then it is not a significant find but it is something that is worth noting - unlike year which was just oh dear we dont know what is going on.
+Region classification performed % FIXME: what does this mean - see the above comment.
+% Note that it is not about the implications but getting in the definition of these terms.
+These things should be prepared for the reader well before the results.
+There should be no new terms here, we should just be refering to the outcomes that were established in the methods, and they should already have their motivations and contexts estabished (there is more that Kate has written if you are confused this is my translation after our meeting).
+at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
+The most prominent feature used to classify regions with the classification tree was water source (see Figure \ref{fig:region_tree}).
+This differed from the variables that illustrated the greatest importance for the XGBoosted ensemble (see Figure (\ref{fig:region_sankey}), with predictor variables being highly interrelated in importance.
+Area, water, fuel and yield were more determining factors when predicting region using XGBoost.
+% FIXME: Is this perhaps your primary test?
+The fact that Kate was asking this means that the tools you are using do not have a well defined purpose in this analysis.
+I believe this stems from the lack of contextual motivation behind these tools.
+What is it that you are testing?
+Why are these important?
+This problem is throughout the paper and cannot simply be fixed here.
+Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
+\par
+
+Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
+The regions with the most samples performed the best.
+Bordering regions were routinely grouped together and misclassified as the same region.
+Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
+%, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
+The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
+These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
+Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
+
+\begin{figure}
+ \resizebox{\textwidth}{!}{
+ \includegraphics{region.pdf}}
+ \caption{Decision tree predicting Region.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:region_tree}
+\end{figure}
+
+\begin{figure}
+ \resizebox{\textwidth}{!}{
+ \includegraphics{region.png}}
+% FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
+% FIXME: Mark one Panel A and the other B
+ \caption{The left-hand side depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:region_sankey}
+\end{figure}
+
 \section{Discussion}
-
-% FIXME: Mardi again asked about whether yield was per hectare.
-This is areally interesting consideration - we do not look at predicting yield specifically in this though.
-perhaps we should be very wary about these analysis with this respect.
 

### 2023-10-15 article.tex
backing up prior changes

@@
 
+Operating cost
+
 There was a pronounced difference in accuracy between the regression tree and the XGBoost model when predicting Operating costs.
@@
  \includegraphics{operating_costs.png}}
- % FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
- % FIXME: Mark one Panel A and the other B
- \caption{The left-hand side depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:operating_costs_sankey}
+ % FIXME: Mark one Panel A and the other B This still needs to be done in the actual pictures
+ \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The number at the end of each band in the diagram is that variable's importance.
+The right-hand side, B, depicts the importance of the 10 variables in Sankey diagram relative to one another.}\label{fig:operating_costs_sankey}
 \end{figure}

### 2023-10-15 analysis.r
added revenue surrogate model

@@ -169,3 +169,35 @@ sink("operating_costs.txt")
 print(model_tree)
 sink()
 ########################
+
+######################
+#       Revenue      #
+######################
+
+model_tree <- train(Revenue ~ .
+    , data = df[, colnames(df)[
+        !(colnames(df) %in% c(
+            "Profitable",
+            "Profit",
+            "Operating.Costs"))]]
+    , method = "rpart"
+    , trControl = trControl
+    , na.action = na.omit
+    , tuneLength = 10
+)
+
+# Each node shows
+# - the predicted value (the average value of the response).
+# Note that each terminating node is a regression model
+# - the percentage of observations in the node
+
+pdf("revenue.pdf")
+rpart.plot(model_tree$finalModel
+    , legend.x = -100
+    , box.palette = "auto"
+)
+dev.off()
+
+sink("revenue.txt")
+print(model_tree)
+sink()

### 2023-10-15 article.tex
added correct importance scores for region
backup

@@
 An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
-This dataset allowed insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue within the data.
+This dataset allowed insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue from grape sales within the data.
 We use this data to study these economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
@@
 An XGBoosted tree was created for each variable to show how they interacted.
-Each tree included all but the economic variables (operating cost and revenue), which were only included within their own trees as response variables.
+Each tree included all but the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
 This meant that every variable would have a measure of its importance to other variables (see Section \ref{sec:importance}), which was used to show the highly interrelated nature of variables within vineyards.
@@
 The importance values for the Chord diagrams were taking from the models of those individual variables, with the diagram being simplified to just the ten variables in the associated Sankey diagram, for readability's sake.
+
 \subsection{Validation}
+
 As there were multiple different loss functions, multiple different forms of validation were used.
@@
 The same measure of accuracy as the XGBoosted trees was used for comparison.
+
 \section{Results}
@@
 
+\subsection{Revenue}
+
+\begin{figure} 
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.pdf}}
+  \caption{Decision tree predicting Profit.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:revenue_tree}
+ \end{figure}
+
+ \begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.png}}
+  % FIXME: Mark one Panel A and the other B
+  \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+ \end{figure}
+
+We investigated the link between revenue to other variables in the SWA data by predicting it, and then linking each variable to revenue through variable importance.
+The prediction of revenue performed similarly to operating cost achieving an $R^2$ of 0.7716 (with a standard deviation of 0.1525).
+A regression tree was used as a surrogate model to present an example of the typical type of decision tree present within the XGBoost Ensemble, however the surrogate model only achieved an $R^2$ of 0.0961 (with a standard deviation of 0.0181) and the XGBoosted ensemble.
+\par
+The important variables when attempting to determine revenue were size, yield, fuel and water (see \ref{fig:revenue_sankey}).
+Due to regions being recorded separately for importance none appeared as the most important variables, overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
+Although performing poorly, the surrogate model highlights the importance of size in determining profits.
+Area also appearing as a variable of higher importance is show to be highly interrelated with other variables.
+The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
+Area harvested is likely also an indicator of other variables such as slashing passes its strongest connection presented.
+
 \subsection{Operating Costs}
 
-% FIXME: Try to follow the same format for each of the section in the results. It will help to include a leading sentance about why the metrics are different everytime.
-
-Operating cost
-
-There was a pronounced difference in accuracy between the regression tree and the XGBoost model when predicting Operating costs.
-With the regression tree achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) in its cross validation.
-The XGBoosted regression ensemble achieved an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
-\par
-Within the XGBoost ensemble's nodes for operating costs (see figure \ref{fig:operating_costs_sankey}) fuel, water, area and yield occurred the most, similarly to region.
-Both diesel and petrol were of more relative importance (being ranked higher) in operating costs than water was compared with region.
-It is surprising that electricity, slashing and spraying passes was not more prominent in operating costs.
-% FIXME: Me and Kate talked about this.
-The diagrams, their context and justification go into the methods.
-How to interpret them needs to be better outlined here.
-You need to affirm peoples intition of them and the significance of more ribbons going out than in as well as why we are only using 10 variables.
-However, Figure \ref{fig:year_sankey} shows that electricity, slashing and spraying are important variables in determining area and yield.
-
-Electricity in particular is used predominantly for irrigation and so is related largely to the size of a vineyard.
-However, slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
-This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
-% FIXME: May need to maintain bigger vineyards are more likely to have more staff verse small owner operators.
-Those extra staff may be trying to keep employers busy be sending them slashing (this comment was from Mardi and very hard to validate in the literature.)
-
-%
-% Is there any point in including this if its so bad?
-%
-% \begin{figure}
-%  \resizebox{\textwidth}{!}{
-%  \includegraphics{operating_costs.pdf}}
-%  \caption{Decision tree predicting Operating Costs.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_costs_tree}
-% \end{figure}
+\begin{figure} 
+  \resizebox{\textwidth}{!}{
+  \includegraphics{operating_costs.pdf}}
+  \caption{A surrogate model decision tree predicting operating costs.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_tree}
+ \end{figure}
 
@@
 
-\subsection{Revenue}
-
-Predictions of profit performed poorly compared to operating costs with the regression tree having an $R^2$ of 0.1873 (with a standard deviation of 0.0522) and the XGBoosted ensemble achieving an $R^2$ of 0.2535 (with a standard deviation of 0.3126).
-The high standard deviation in the XGBoosted tree was a bias in more accurately predicting vineyards that made profit compared to those that lost money.
-% FIXME: Kate pointed out that it is unclear what this implies (the below sentance).
-% The following paragraph goes into it though, it would be better to meld these together. Talking more about the fact that it was easier to predict those who made a profit than those who did not.
-% FIXME: It is wise to run a model purely on gross margin as well. It might have better insights than just profit! And it might over come this disparity.
-With much higher $R^2$ values being achieved in k-folds containing only those that made profits (recording a maximum of 0.7634).
-% profit summary stats
-% max 3345000
-% min -3937878
-% std 396027.633119089
-% range 7282878
+The relationships to operating cost through variable importance were found to be similar to that of revenue.
+With fuel, water, area and yield occurring the most (see figure \ref{fig:operating_costs_sankey}).
+A surprising difference is that the most important operational consideration for operating cost is the use of fungicide, compared to revenue where slashing is the most important (comparing Figure \ref{fig:region_sankey} and \ref{fig:operating_costs_sankey}).
+The variables that feed into these decisions are also very differing with diesel being the most informative to slashing and area being the most informative to the need for fungicide.
 \par
-There was a disparity of 66.63\% of vineyards recording a profit than those that did not.
-When predicting if a vineyard would be profitable or not the classification tree and XGBoosted ensemble did not perform considerably differently from this proportion.
-With the regression tree achieving an accuracy of 68.66\% (and a standard deviation of 0.01\%) and the XGBoost ensemble achieving 71.97\% accuracy (with a validation accuracy of 70.59\%).
+Again, region played a determining factor overall but not as much individually with region contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
+It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
 \par
-It was surprising that operating costs performed substantially better in $R^2$ compared to profit.
-% Is there a better way to phrase this?
-Look at the papers that use the same analysis techniques for phrases to use when soething is good, bad predictor.
-
-Interestingly the important variables when attempting to determine profit were similar to those used to classify region (see Figure \ref{fig:profit_sankey}), with the exception of water used.
-Both the regression tree and the XGBoosted ensemble used region, specifically the Hunter Valley.
-The regression tree also used Tasmania when determining profit.
-Both the Hunter valley and Tasmania are known for the production of high quality grapes used in export wines \citep{wineaustraliaNationalVintageReport2022}.
-A major difference between region and profit was the importance given to water use, with water use being a more important variable in predicting region than profit.
-
-\begin{figure} % FIXME: Interesting these 2 regions probably have the highest price of grapes/tonne.
-They are also 2 extreme climates (Hunter = lots of grazing, season rainfall = of tractor passes.
-Tasmania = V.Cool weather = lots of tractor passes.)
- \resizebox{\textwidth}{!}{
- \includegraphics{profit.pdf}}
- \caption{Decision tree predicting Profit.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:profit_tree}
-\end{figure}
-
-\begin{figure}
- \resizebox{\textwidth}{!}{
- \includegraphics{profit.png}}
- % FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
- % FIXME: Mark one Panel A and the other B
- \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:profit_sankey}
-\end{figure}
-
-% \begin{table}[]
-%  \caption{Validation and training accuracies of each multiclass variable.}
-%  \begin{tabular}{@{}ccc@{}}
-%  \toprule
-%  \textbf{Variable} & \textbf{Validation} & \textbf{Training} \\ \midrule
-%  \textbf{cover crops} & 0.364086 & 0.396418 \\
-%  \textbf{water type} & 0.742097 & 0.928905 \\
-%  \textbf{profitable} & 0.705882 & 0.719737 \\
-%  \textbf{irrigation type} & 0.841845 & 0.847554 \\
-%  \textbf{giregion} & 0.505824 & 0.568242 \\
-%  \textbf{irrigation energy} & 0.746293 & 0.836405 \\
-%  \textbf{data year id} & 0.422003 & 0.518059 \\ \bottomrule
-%  \end{tabular}
-%  \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%
-% Points to address
-%
+Comparatively to revenue, operating cost performed better The XGBoosted regression ensemble achieved an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
 
 \subsection{Region}
-% FIXME: there are artifacts from the previous paper.
-Although the previous paper is pertinant to some of this (quite critically so) it is no longer the primary focus.
-On top of this some of these terms such as 'Region Classification' are not well defined.
-This is because previously a great deal of effort was put into defining them and describing the potential regions.
-However this has been removed.
-% FIXME: This section is much more of a: 'isn't it weird?' That region is not as significant in some areas but not others.
-That is can be used to predict things like yield but no profit etc.
-This phrasing shows why it goes after other sections.
-if you are writting by significance then it is not a significant find but it is something that is worth noting - unlike year which was just oh dear we dont know what is going on.
-Region classification performed % FIXME: what does this mean - see the above comment.
-% Note that it is not about the implications but getting in the definition of these terms.
-These things should be prepared for the reader well before the results.
-There should be no new terms here, we should just be refering to the outcomes that were established in the methods, and they should already have their motivations and contexts estabished (there is more that Kate has written if you are confused this is my translation after our meeting).
-at 32.34\% (3.67\% standard deviation) and 56.82\% accuracy (50.58\% validation accuracy), for the classification tree and XGBoosted ensemble respectively.
-The most prominent feature used to classify regions with the classification tree was water source (see Figure \ref{fig:region_tree}).
-This differed from the variables that illustrated the greatest importance for the XGBoosted ensemble (see Figure (\ref{fig:region_sankey}), with predictor variables being highly interrelated in importance.
-Area, water, fuel and yield were more determining factors when predicting region using XGBoost.
-% FIXME: Is this perhaps your primary test?
-The fact that Kate was asking this means that the tools you are using do not have a well defined purpose in this analysis.
-I believe this stems from the lack of contextual motivation behind these tools.
-What is it that you are testing?
-Why are these important?
-This problem is throughout the paper and cannot simply be fixed here.
-Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
-\par
-
-Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
-The regions with the most samples performed the best.
-Bordering regions were routinely grouped together and misclassified as the same region.
-Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
-The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
-%, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
-The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
-These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
-Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
 
@@
   \includegraphics{region.png}}
-% FIXME: What are the #'s? What do the sizes mean? What are the colours? What is interrelated importance?
  % FIXME: Mark one Panel A and the other B
- \caption{The left-hand side depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:region_sankey}
+  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The number at the end of each band in the diagram is that variable's importance.
+The right-hand side, B, depicts the importance of the 10 variables in Sankey diagram relative to one another.}\label{fig:region_sankey}
  \end{figure}
+
+When considered overall, Region was a highly informative variable through measure of importance to both operating cost and revenue.
+Notably the overall importance of region to revenue was 234 (making it the third most important variable when considerng all regions together).
+The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
+These two regions are also relative opposites in winegrowing climates with the Barossa being warm and dry climate focussing on Shiraz grapes and Tasmania being a cool wet climate that grows Pinot.
+\par
+When considering all regions together, it had the most node contributions to determining operating costs with an importance of 334.
+Of all the regions, again Tasmania was the most important, followed by the Adelaide Hills.
+In contrast to revenue, both climates are considered cool and wet, and warmer drier regions such as the Barossa and Hunter Valley only contributed roughly half the same number of nodes to the ensemble.
+When looking at \ref{fig:region_sankey} the inclusion of slashing and fungicide spraying is the likely reason; with fungal and weed pressure being greater in cooler wetter regions.
+\par
+Both diesel and petrol were of more relative importance in operating costs than water was compared with region.
+similar to those used to classify region (see Figure \ref{fig:region_sankey}, except water used.
+The surrogate model relied heavily on the use of water source to classify regions, which is reflective of regional access to resources being a deciding factor in vineyard management (see Figure \ref{fig:region_tree}).
+A major difference between region and profit was the importance given to water use, with water use being a relatively more important variable in predicting region than profit (considering its rank in importance to other variables).
+\par
+The surrogate model for region performed better than other surrogate models with 32.34\% (3.67\% standard deviation).
+The prominence of types and use of water resources was in classifying region is reflective of difference of availability of water resources is when comparing different regions (see Figure \ref{fig:region_tree}).
+The XGBoost ensemble, did not perform as well as operating costs or revenue with 56.82\% accuracy (50.58\% validation accuracy).
+The difference in accuracy is in part due to the large number of classes (being 58).
+The ensemble did not differ greately from the surrogate model, with a continuing emphasis on Area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
+% Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
+\par
+Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
+The regions with the most samples performed the best.
+Bordering regions were routinely grouped together and misclassified as the same region.
+Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
 
@@
 
-% FIXME: Moved here from results. Find a place for this lost puppy!
-%
-% It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
-The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
-Due to the wide variety of uses of petrol and diesel, it is likely that they are representative of
-% other activities within the vineyard, such as pruning and harvesting. FIXME: This is new infor tht doesnt have any context (e.g no discussion yet about what fuel may be used for)
-% With predictors such as yield and area being highly interconnected as they likely operate as proxy variables to other factors, possibly other present variables.
+Regional pressures such as disease and resource access have an incredible influence in determining a vineyard's revenue and operating cost.
+
+The explored relationships between vineyard resource use, operations and geographical properties to revenue and operating costs highlight how discerning regional pressures such as disease,
 
 Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The association between yield and region is demonstrated by its position as fourth most occurring variable within the nodes of the XGBoosted ensemble which determined region (see Figure \ref{fig:region_sankey}).
-% FIXME: Clunky phrasing
-
-The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % FIXME: Maybe but not sure this assumption holds.
-It may be an artefact of the members being corporate in larger holdings.
-The larger percentage of blocks in the riverland are small.
+The association between yield and region is demonstrated by its rank of fourth-highest variable importance when  determining region (see Figure \ref{fig:region_sankey}). 
+% The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % TODO: This section is commented out due to Mardi's take:
+%  Maybe but not sure this assumption holds. It may be an artefact of the members being corporate in larger holdings. The larger percentage of blocks in the riverland are small.
 \par
-Regions with lower land costs are also warmer \citep{willchancellorMeasuringAustralianBroadacre2019}, which is known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
+Warmer regions 
+% generally have lower land costs \citep{willchancellorMeasuringAustralianBroadacre2019}, 
+are known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
 Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
@@
 An example is the optimisation of fungicide sprays using Bayesian models that forecast disease risk \citep{luDiseaseRiskForecasting2020}.
+
+Electricity in particular is used predominantly for irrigation and so is related largely to the size of a vineyard.
+However, slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
+This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
+% FIXME: May need to maintain bigger vineyards are more likely to have more staff verse small owner operators.
+Those extra staff may be trying to keep employers busy be sending them slashing (this comment was from Mardi and very hard to validate in the literature.)
+
+It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
+The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
+Due to the wide variety of uses of petrol and diesel.
+It is likely that fuel use is representative of other activities within the vineyard, such as pruning and harvesting.
+Predictors such as yield and area being highly interconnected to fuel are likely reflective of fuel required to  as they likely operate as proxy variables to other factors, possibly other present variables.
 
@@
 
-%%%%%%%%%%%%%%%
-% Future work %
-%%%%%%%%%%%%%%%
-
-% What remains unresolved?
-% 
-% You may or may not have a lot to say here.  Some of it may be useful in your discussion.
-The causal reasoning behind these variables behaviours. The untangling the multicollinearity of these variables.
+The causal reasoning behind Revenue and operational costs. The untangling the multicollinearity of these variables.
 
@@
 
-This study has provided valuable insights into the multifaceted dynamics governing operational costs, different yearly effects and vineyard regions highlighting the complex interrelatedness of variables within a vineyard.
+This study has provided valuable insights into the multifaceted dynamics governing operational costs and revenue. 
+The impact of different regions highlighted the complex interrelatedness of variables within a vineyard.
 The paper underscores how factors such as water and fuel intersect to impact operational costs and How different seasonal events affect these operations and the significance of context-specific decision-making.

### 2023-10-15 article.tex
done!

@@
 These systems, although predictable through some variables are not causal.
-We see this in the evidence of some years notaffecting outcomes even though they are known to contain the aforementioned issues.
+We see this in the evidence of some years not affecting outcomes even though they are known to contain the aforementioned issues.
 
@@
 Each tree included all but the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
+Separately profit (the difference between revenue and operational costs) was looked at in prior analyses (see appendix) but the results were not included due to low average loss values and model stability.
 This meant that every variable would have a measure of its importance to other variables (see Section \ref{sec:importance}), which was used to show the highly interrelated nature of variables within vineyards.
@@
 
-Regional pressures such as disease and resource access have an incredible influence in determining a vineyard's revenue and operating cost.
-
-The explored relationships between vineyard resource use, operations and geographical properties to revenue and operating costs highlight how discerning regional pressures such as disease,
-
+The explored relationships between vineyard resource use, operations and geographical properties to revenue and operating costs highlight how decisive regional influences can be determining a vineyard's economic outcomes.
 Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
@@
 \par
-Warmer regions 
-% generally have lower land costs \citep{willchancellorMeasuringAustralianBroadacre2019}, 
-are known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
+Warmer regions are known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
 Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
-In general if irrigated warmer regions have been associated with lower yields due to their generally lower rainfall, which can be mitigated through applying more water \citep{campsGrapeHarvestYield2012}.
-% FIXME: Rework this sentance
-It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of lower quality grapes; reflected through the variables' importance of water use in the XGBoost ensemble.
-% FIXME: Need to link lower quality to profitability.
-% FIXME: Warm inland regions are close to a major source of irrigation water which means that they have always taken advantage of this to grow higher crops.
-Which are also extremely important to our sector.)
-% FIXME: May need to explain that growers are paid @ a $ value / tonne and that the $ value varies in region.
-The practice of utilising larger quantities of water for inland Australian wine crops is partly reflected in the prior use of flood style irrigation to saturate soil \citep{bgcoombeGrapeBerryDevelopment2004}.
-% FIXME: More likely is that they have access to large water licences (Prev used for flood irrigation) That ythey now use on vineyards.
-% FIXME: Examples of regions??
+In general warmer regions are not associated with higher yields, but if a vineyard in a warmer region is sufficiently irrigated much higher yields can be achieved than in cooler regions \citep{campsGrapeHarvestYield2012}.
+It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of grapes; reflected through region using water use so prominently in the XGBoost ensemble.
+The link to water resources in defining regions is also an important consideration, as vineyards can leverage higher irrigation rates given more accessible water resources.
+A further consideration in the link between revenue and region is that grape prices are set at a regional level by buyers \citep{wineaustraliaNationalVintageReport2022}.
+It is also important to consider that some regions carry particular fame regarding the quality of their produce such as Tasmania, the Hunter Valley and Barossa Valley \citep{hallidayAustralianWineEncyclopedia2009}.
 This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
-This is possibly the connection between the presence of the Hunter Valley within the XGBoost ensemble that determined profit (see Figure \ref{fig:profit_sankey}).
+% This is possibly the connection between the presence of the Hunter Valley within the XGBoost ensemble that determined profit (see Figure \ref{fig:profit_sankey}).
 With this connection reflecting the restriction of possible strategies employable by winegrowers between different regions.
@@
 In part some winegrowing strategies are restricted simply through access to water resources, being reflected through the region classification tree (see Figure \ref{fig:region_tree}).
-Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surgace or underground water.
+Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
 Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region.
 Where, the need to pressurise irrigation systems from river water or to generate power would require larger amounts of diesel and petrol.
-% FIXME: Yes but this requires some explanation
 \par
@@
 An example of how interrelated operational costs can be, is
-the optimisation of tractor passes % FIXME: What does this mean? Multiple activities with a single pass?
-being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}. 
+the optimisation of tractor passes to achieve multiple goals in a pass, being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
+\par
+When determining revenue, similar variables were used to operational cost; with region also being of high variable importance relative to other variables (when considering all regions together in importance).
+It is difficult to extrapolate the specific influence of location on a vineyard's outcomes due to the broad and varying definition of a region.
+Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
+However, as buyers set prices at regional levels, it is still important to consider a vineyards Geographical Indicator region.
 \par
@@
 Without fully capturing more granular activities such as the specifics of what fuel was used for, it is hard to determine what decisions specifically influence the operational costs.
+Electricity in particular is used predominantly for irrigation.
+Size is also a further consideration where slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
+This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
 \par
-Although less important in the XGBoost ensembles for profit, the variables: cover crops, and number of fungicide spraying and slashing passes are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
+The reasoning for any particular decision can be widely varying. 
+% Where, a cover crop may be employed to help increase soil water retention, reduce erosion, increase biodiversity and reduce weeds \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020,delpuechAdaptingCoverCrop2018}.
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
+A more granular definition of region may help to better discern the differences in practices, and the reason for employing them.
+More sophisticated models, specifically those that utilise expert opinion, may also help to capture and address the decision-making process.
+An example is the optimisation of fungicide sprays using Bayesian models that forecast disease risk \citep{luDiseaseRiskForecasting2020}.
+\par
+% Although less important in the XGBoost ensembles for profit, the variables: cover crops, and number of fungicide spraying and slashing passes are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
 Rainfall being related to vine fungal growth, disease and weed pressure.
@@
 % FIXME: Re-structure this sentance.
-It is difficult to extrapolate findings to these methods and the reason for their use due to the broad and varying definition of the regions.
-Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation, as it is too broad to fully capture a vineyards location and its influence on more granular variables.
-The reasoning for using approaches such as cover crops can be widely varying.
-Where, a cover crop may be employed to help increase soil water retention, reduce erosion, increase biodiversity and reduce weeds \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020,delpuechAdaptingCoverCrop2018}.
-However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
-A more granular definition of region may help to better discern the differences in practices, and the reason for employing them.
-More sophisticated models, specifically those that utilise expert opinion, may also help to capture and address the decision making process.
-An example is the optimisation of fungicide sprays using Bayesian models that forecast disease risk \citep{luDiseaseRiskForecasting2020}.
 
-Electricity in particular is used predominantly for irrigation and so is related largely to the size of a vineyard.
-However, slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
-This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
-% FIXME: May need to maintain bigger vineyards are more likely to have more staff verse small owner operators.
-Those extra staff may be trying to keep employers busy be sending them slashing (this comment was from Mardi and very hard to validate in the literature.)
-
-It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
+% It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
 The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
@@
 
-The disparity in accuracy between profit and operational costs is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
+Separately revenue and operating cost did have a greater predictability than their counterpart profit.
+The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
 The difference between turning a profit or loss is dependent on decisions made and chance.
@@
 
-The causal reasoning behind Revenue and operational costs. The untangling the multicollinearity of these variables.
-
-It maybe better to pursue more causal models and create decision support systems than to model factors directly.
-It is difficult to untangle the predictive and correlative nature of the variable compared to the causal reasons.
-
 This study has provided valuable insights into the multifaceted dynamics governing operational costs and revenue.
 The impact of different regions highlighted the complex interrelatedness of variables within a vineyard.
-The paper underscores how factors such as water and fuel intersect to impact operational costs and How different seasonal events affect these operations and the significance of context-specific decision-making.
+We relate how factors such as water and fuel intersect to impact operational costs and how different seasonal events affect these operations; as well as the significance of context-specific decision-making.
 While this investigation utilised a broad regional classification, the potential benefits of adopting a more nuanced approach and incorporating expert knowledge have been highlighted.
+Further work could pursue causal models and the creation of decision support systems.
+It is difficult to untangle the predictive and correlative nature of a variable compared to the causal reasons.
 By delving deeper into the complex interplay of variables, further advancements can be made in optimising vineyard management strategies for lowering operational costs and enhancing sustainability.

### 2023-10-15 article.tex
quick reference add

@@
 This is done through analysing a new comprehensive nationwide data set using XGBoosted models.
-We further compare the relationships between different resources to address the extensive collinearity found within the data FIXME: This needs references||.
+We further compare the relationships between different resources to address the extensive collinearity found within the data \citep{chenXGBoostScalableTree2016}.
 XGBoosted models were used because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables; with importance being able to be statistically defined through multiple methods.
@@
   \includegraphics{profit.png}}
-  % FIXME: Mark one Panel A and the other B
   \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
@@
   \includegraphics{operating_costs.png}}
-  % FIXME: Mark one Panel A and the other B This still needs to be done in the actual pictures
   \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
@@
   \includegraphics{region.png}}
- % FIXME: Mark one Panel A and the other B
   \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.

### 2023-10-15 article.tex
fixed references to revenue

@@
   \resizebox{\textwidth}{!}{
-  \includegraphics{profit.pdf}}
-  \caption{Decision tree predicting Profit.
+  \includegraphics{revenue.pdf}}
+  \caption{Decision tree predicting revenue.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:revenue_tree}
@@
   \resizebox{\textwidth}{!}{
-  \includegraphics{profit.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting Profit using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+  \includegraphics{revenue.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
 The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
@@
 Due to regions being recorded separately for importance none appeared as the most important variables, overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
-Although performing poorly, the surrogate model highlights the importance of size in determining profits.
+Although performing poorly, the surrogate model highlights the importance of size in determining revenue.
 Area also appearing as a variable of higher importance is show to be highly interrelated with other variables.
@@
 The surrogate model relied heavily on the use of water source to classify regions, which is reflective of regional access to resources being a deciding factor in vineyard management (see Figure \ref{fig:region_tree}).
-A major difference between region and profit was the importance given to water use, with water use being a relatively more important variable in predicting region than profit (considering its rank in importance to other variables).
+A major difference between region and revenue was the importance given to water use, with water use being a relatively more important variable in predicting region than revenue (considering its rank in importance to other variables).
 \par

### 2023-10-15 article.tex
added profit to the appendix
added bibliography

@@
 
-\section{Abstract}
-
-Region is a driving force in determining profits.
-Fuel and water are the primary driving operational costs
-
-The finding themselves are not that significant, it is the evidence of a complex system governing these relationships that is the significant finding.
-
-They were gathered by Sustainable Winegrowing Australia
-
-What we add is that agricultural economics maybe complex systems that are obfuscated by the expertise of growers to navigate problems such as drought, pests and disease.
-These systems, although predictable through some variables are not causal.
-We see this in the evidence of some years not affecting outcomes even though they are known to contain the aforementioned issues.
-
 \section{Introduction}
@@
 Specifically within the Australian Wine and vine industry there is a need to further understand the driving relationships between resource use and economic output.
-Where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers.
-\citep{lukemanciniUnderstandingAustralianWine2020}
-\newline
+Where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
+\par
 An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
@@
 
+ \subsection{Profit}
+
+\begin{figure} 
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.pdf}}
+  \caption{Decision tree predicting revenue.
+Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:revenue_tree}
+ \end{figure}
+
+ \begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{profit.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+ \end{figure}
+
+Predictions of profit perfomed poorly compared to operating cost and revenue with an average $R^2$ of 0.2535 and standard deviation of 0.3126.
+With the large standard deviation being indicative of how unstable the models created were.
+
 \end{linenumbers}

### 2023-10-30 article.tex
title change
introduction km edits

@@
 %% \title{Title\tnoteref{label1}}
-\title{An analysis of interrelations between economic and environmental variables in Australian Winegrowing.}
+\title{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards.}
 
@@
 
-Historically strong demands for Australian wine have helped to create a thriving industry, however recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation has negatively effected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
+Historically strong demands for Australian wine have helped to create a thriving industry.
+However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation has negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
 The 2021-2022 financial year alone saw a decline of 19\% in exports solely due to tariffs \citep{wineaustraliaNationalVintageReport2022}.
-A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale are key to  introducing stronger policy and information to aid in increasing a nations agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
-Specifically within the Australian Wine and vine industry there is a need to further understand the driving relationships between resource use and economic output.
-Where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
+A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale are key to making data-informed decisions to aid in increasing a nations agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
+Specifically within the Australian Wine and vine industry there is a need to further understand the driving relationships between resource use and economic output, where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
 \par
 An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
-This dataset allowed insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue from grape sales within the data.
+This dataset allows insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue from grape sales within the data.
 We use this data to study these economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
-Answering what the driving factors are behind vineyard economic outcomes, and linking these outcomes to predictor importance.
-This is done through analysing a new comprehensive nationwide data set using XGBoosted models.
 We further compare the relationships between different resources to address the extensive collinearity found within the data \citep{chenXGBoostScalableTree2016}.
-XGBoosted models were used because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables; with importance being able to be statistically defined through multiple methods.
+We adopt XGBoosted models for this analysis because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables.
 

### 2023-10-30 article.tex
saving progress

@@
 
-% FIXME:
-% While overall this methods section seems technically correct, more work is needed on justification, motivation, and context.
-% Through the methods, motivate each decision youve made with context from the particular problem/data/question.
-Why is the basic standard procedure not adequate?
-This helps to slow it down and help the reader understand complex methods.
-%
-% As kate put it in person you are currently writting like ||||| fact fact fact. You need to be more |...||...|| with context etc inbetween each fact so that people can digest it.
-
-% As with the previous paper, put the models up front it is very unclear what you have created - and less clear to understand it if there is no way to reference to it.
-
-% FIXME: The methods needs to be restructured. Try the following order:
-%     - Why do it?
-%     - What does it do? (at a high level)
-%     - How did you do it?
-% 
-% You should also include subsections. I think subsections are a good crutch because they can always be removed if the flow is done well.
-% 
-% Avoid following a narrative. Do not do the who/what/when/where/how/why and the big reveal.
-% An aside comment might be that you can attempt to follow the narrative writing paradigm across the paper as a whole but limit your methods section to the why, what then how.
-And use the results as your reveal.
-% Or maybe dont do that - it might just be that you need to write it like a narrative as a first draft then butcher it into the necessary sections.
-That writing style is incredibly ingrained at this point.
-
-% FIXME: the motivation is missing from methods.
-
 \subsection{Data}
@@
 Australia's national wine industry sustainability program.
-The program aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by SWA is entered manually by winegrowers using a web based interface tool.
-A total of 6091 observations were collected from 2012/2013 to 2021/2022 financial years.
-23 variables were used for each observation reflecting a vineyards state for the given year (see Table \ref{tab:vars}).
+The program aims to support grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data recorded by SWA are entered manually by winegrowers using a web based interface tool.
+A total of 6091 observations were collected from 2012/2013 to 2021/2022 financial years, with each observation comprising 23 variables reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
 \par
@@
 \end{table}
-
+% FIXME: Add supplementary tables of all classes and another table of the summary statistics of continuous variables.
+Refer to this table to help with the paragraphs about one-hot encoding.
+A quick definition of one-hot encoding will also help!
 The data originally contained only two multiclass variables: year and region.
@@
 They were chosen for this analysis as the data contained a mixture of class and continuous variables.
-And, XGBoosted trees are unaffected by multicollinearity, as well as offer high predictive performance for a wide variety of purposes \citep{chenXGBoostScalableTree2016}.
+Moreover, XGBoosted trees are unaffected by multicollinearity, and offer high predictive performance for a wide variety of purposes \citep{chenXGBoostScalableTree2016}.
+\par
+% FIXME: Make it clearer why you did this "Because the purpose of the analysis was to identify relationships between variables and since there was no single outcome variable of interest an XGBoosted tree was created for each variable using the other variables as predictors
+% FIXME: forward reference here "XGBoosted trees, described in more detail below...
+% FIXME: I found this a bit confusing as a 'new reader'. How about describing this process as a set of analyses:
+% Step 1: Separate analyses of operating cost and revenue: XGBoosted models were constructed for each of these variables, with the variable as the response and all other variables (with the exception of operating cost and revenue) as predictors.
+% Step 2: For each of the operating cost and revenue models, an analysis of the inter-relationships between the ten most important predictors based on the XGBoosted model.
+% Set 3: Analysis of profit (difference between...): etc
+% Then say that the results of the analyses in Set 3 were not considered further due to how crap they were.
 An XGBoosted tree was created for each variable to show how they interacted.
-Each tree included all but the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
+Each tree included all variables except the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
 Separately profit (the difference between revenue and operational costs) was looked at in prior analyses (see appendix) but the results were not included due to low average loss values and model stability.
 This meant that every variable would have a measure of its importance to other variables (see Section \ref{sec:importance}), which was used to show the highly interrelated nature of variables within vineyards.
-The complicated interaction between bariables was illustrated using Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between any two variables (see section \ref{sec:importance}).
+The interaction between variables was illustrated using Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variable and the response (see section \ref{sec:importance}).
 \par
-Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees predict a value $y_i$ from the input $x_i$.
-The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
-Each of $f_k$ functions is a classification or regression tree, such that all functions are in the set of all decision trees, given by $\mathcal{F}$, is defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
-Where each function corresponds to an independent tree structure $q$ of $\omega$ weights.
-Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf.
-The final prediction is determined by the sum of the score of the corresponding leaves, given by:
-\begin{equation}
- \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{k}(x_{i})}, \space f_{K}\in \mathcal{F},
-\end{equation}
-
-The set of functions, $\mathcal{F}$, used by the tree is determined by minimising a regularised objective function, $\mathcal{L}$ given by:
-
-\begin{equation}\label{eqn:objective}
- \mathcal{L}(\phi)=\sum_i l(\hat{y_i}, y_i^{t-1} + f_t(x_i)) + \sum_k \Omega (f_K).
-\end{equation}, where
-\begin{equation}
-  \Omega(f) = \gamma T + {{1}\over{2}}\lambda ||\omega||^2
-\end{equation}
-
-As predictions are made using additive tree functions, XGboosted trees can be used for classification or regression.
-The difference between a prediction, $\phi(x_i)$, and actual variable, $f_{k}(x_i)$, is a differentiable convex loss function $l$.
-These properties of $l$ allow the function to be versatile in which objective we choose to optimise for, which is also important in being able to process both continuous and categorical variables.
-To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
-\subsection{Loss functions}
-
-The functions included as parameters in equation \ref{eqn:objective} mean that traditional opimisation methods for Euclidean space cannot be used.
-Chen and Guestrin \citep{chenXGBoostScalableTree2016} illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
-Importantly a loss function can be used to fit a model iteratively to data.
-For this analysis several loss functions were used, as variables took the form of continuous, binary and multi-call data.
-The loss function for making a split within the tree structure is given by:
-
-\begin{equation} \label{eqn:enumerate}
-\mathcal{L}_{split} = {{1}\over{2}} \left[ 
- {{(\sum_{i\in I_L}g_i)^{2}}\over{\sum_{i\in I_L} h_i + \lambda}} +
- {{(\sum_{i\in I_R}g_i)^{2}}\over{\sum_{i\in I_R} h_i + \lambda}} -
- {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma.
-\end{equation}
-
-The tree structure being defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
-Instead of enumerating all possible tree structures, a greedy algorithm iteratively adds branches to the tree minimising $ \mathcal{L}_{split} $ in (\ref{eqn:enumerate}).
-The frequency of a variable's occurrence within a tree is directly attributed to the minimisation of the loss function through the minimisation of $ \mathcal{L}_{split} $.
-\par
-The loss functions used for this analysis were the root-mean-square function for continuous variables.
-The logistic loss function for binary class variables.
-And, the soft max function for Multiclass variables.
-All objective functions are defined within the SKlearn library \citep{sklearn_api}, which was utilised via an API to the XGBoost library \citep{chenXGBoostScalableTree2016}.
-\par
+% FIXME: We need a digestable description of XGBoost for readers of this article.
 \subsection{Variable Importance}\label{sec:importance}
-
+% FIXME: this section can be tightende/shortened a lot - moving some of the material into the supplementary section
 Due to XGBoost creating a large amount of decision trees, the interpretability of these models is obfuscated by the intricate relationships within complicated ensembles.
@@
 The measure of a variable's importance within this study can then be interpreted as how often a variable was the optimal choice in reducing the loss function of the ensemble.
-Importantly, multiclass variables being one-hot-encoded are given an importance score for each individual class; for example, each specific region will have its own importance score.
+Importantly, multiclass variables being one-hot-encoded
+% FIXME: make sure one-hot-encoding is well defined and rephrase this so it makes more sense (possibly define it earlier as suggested in the data section.)
+are given an importance score for each individual class; for example, each specific region will have its own importance score.
 \par
-Creating XGBoosted trees for each variable allowed the use of importance to show how strongly variables were associated with each other.
-The importance of variables to one another was illustrated through the use of Sankey and Chord diagrams.
-These diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+The Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
 Both Chord and Sankey diagrams illustrated variable importance through the size of the bands between two variables.
-The number at the end of a connection in a Sankey diagram indicated a variable's importance, or the number of times it appeared within the ensemble.
-Sankey and Chord  diagrams were presented together; with Sankey diagrams showing the connection of a variable to its 10 most important predictor variables.
+The number at the end of a connection in a Sankey diagram indicates a variable's importance, or the number of times it appeared within the ensemble.
+Sankey and Chord  diagrams are presented together; with Sankey diagrams showing the connection of a variable to its ten most important predictor variables.
 Chord diagrams were used alongside a Sankey diagram to show the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
@@
 
+\section{XGBoost}
+Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees predict a value $y_i$ from the input $x_i$.
+The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
+Each of $f_k$ functions is a classification or regression tree, such that all functions are in the set of all decision trees, given by $\mathcal{F}$, is defined by ${f(x) = \omega_{q(x)}}(q : \mathbb{R}^m \rightarrow T, \omega \in \mathbb{R}^T)$.
+Where each function corresponds to an independent tree structure $q$ of $\omega$ weights.
+Each tree has $T$ leaves, which contain a continuous score, represented by $\omega_i$ for the i-th leaf.
+The final prediction is determined by the sum of the score of the corresponding leaves, given by:
+\begin{equation}
+ \hat{y_i}=\phi(x_i)=\sum_{k=1}^{K}{f_{k}(x_{i})}, \space f_{K}\in \mathcal{F},
+\end{equation}
+
+The set of functions, $\mathcal{F}$, used by the tree is determined by minimising a regularised objective function, $\mathcal{L}$ given by:
+
+\begin{equation}\label{eqn:objective}
+ \mathcal{L}(\phi)=\sum_i l(\hat{y_i}, y_i^{t-1} + f_t(x_i)) + \sum_k \Omega (f_K).
+\end{equation}, where
+\begin{equation}
+  \Omega(f) = \gamma T + {{1}\over{2}}\lambda ||\omega||^2
+\end{equation}
+
+As predictions are made using additive tree functions, XGboosted trees can be used for classification or regression.
+The difference between a prediction, $\phi(x_i)$, and actual variable, $f_{k}(x_i)$, is a differentiable convex loss function $l$.
+These properties of $l$ allow the function to be versatile in which objective we choose to optimise for, which is also important in being able to process both continuous and categorical variables.
+To optimise $l$, the difference is calculated for the i-th instance at the t-th iteration.
+\subsection{Loss functions}
+
+The functions included as parameters in equation \ref{eqn:objective} mean that traditional opimisation methods for Euclidean space cannot be used.
+Chen and Guestrin \citep{chenXGBoostScalableTree2016} illustrate, using Taylor expansions, that for a fixed structure $q(x)$ the optimal weight $\omega^*_j$ for a leaf $j$ can be derived.
+Importantly a loss function can be used to fit a model iteratively to data.
+For this analysis several loss functions were used, as variables took the form of continuous, binary and multi-call data.
+The loss function for making a split within the tree structure is given by:
+
+\begin{equation} \label{eqn:enumerate}
+\mathcal{L}_{split} = {{1}\over{2}} \left[ 
+ {{(\sum_{i\in I_L}g_i)^{2}}\over{\sum_{i\in I_L} h_i + \lambda}} +
+ {{(\sum_{i\in I_R}g_i)^{2}}\over{\sum_{i\in I_R} h_i + \lambda}} -
+ {{(\sum_{i\in I}{g_i})^{2}}\over{\sum_{i \in I} h_i + \lambda}} \right] - \gamma.
+\end{equation}
+
+The tree structure being defined using left $I_L$ and right $I_R$ instance sets of nodes, with $I = I_L \cup I_R$.
+Instead of enumerating all possible tree structures, a greedy algorithm iteratively adds branches to the tree minimising $ \mathcal{L}_{split} $ in (\ref{eqn:enumerate}).
+The frequency of a variable's occurrence within a tree is directly attributed to the minimisation of the loss function through the minimisation of $ \mathcal{L}_{split} $.
+\par
+The loss functions used for this analysis were the root-mean-square function for continuous variables, the logistic loss function for binary class variables, and the soft max function for Multiclass variables.
+All objective functions are defined within the SKlearn library \citep{sklearn_api}, which was utilised via an API to the XGBoost library \citep{chenXGBoostScalableTree2016}.
+\par
  \subsection{Year}

### 2023-11-20 article.tex
backup

@@
 \subsection{XGBoosted Trees}
-XGBoosted (eXtreme Gradient Boosting) trees were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+XGBoosted (eXtreme Gradient Boosting) trees, described in more detail below,  were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
 XGBoosted trees are a boosted tree ensemble method that can be used to classify classes, or predict continuous response variables.
@@
 
-As there were multiple different loss functions, multiple different forms of validation were used.
-In each case the data was split into training data, which constituted 80\% of the original data.
+For each model the data was split into training data, which constituted 80\% of the original data.
 The remaining 20\% was used in testing and validation.
-Data was stratified when splitting the data into these subsets to conserve the same proportion of class occurrences between training, testing and validation data.
-For continuous variables 20\% was used as testing data, minimising the root-mean-square function.
-The final model was validated using repeated k-fold cross validation for 10 folds, repeated 10 times.
+Categorical data was stratified to conserve the same proportion of class occurrences between training, testing and validation data.
+For continuous variables 20\% was used as testing data and the models were validated using 10 repetitions of the sampling process (10-fold cross validation).
 $R^2$ scores were used to determine the best regression models during validation.
-For binary and multiclass variables, data was split into 80\% training, 10\% testing and 10\% validation data.
-For class variables, validation was summarised through the accuracy, the proportion of true negatives and positives.
+$R^2$ was used instead of RMSE to avoid the issue of comparing models of different units to each other when considering how well each model extrapolated to further data.
+For binary and multiclass variables, validation was summarised through the accuracy, the proportion of true negatives and positives.
 \par
@@
 
-\subsection{Surrogate Models}
+% \subsection{Surrogate Models}
 
-The creation of more interpretable models such as linear regression in parallel to AI systems has been used to explain variable's relationships within black box algorithms \citep{molnarInterpretableMachineLearning2022}.
+% The creation of more interpretable models such as linear regression in parallel to AI systems has been used to explain variable's relationships within black box algorithms \citep{molnarInterpretableMachineLearning2022}.
 As XGBoost create an ensemble of decision trees, here we use classification and regression trees to gain insight into intricacies of the ensembles derived through XGBoost.
@@
 These trees were created using the rparts and caret packages \citep{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
-\par
-Decision trees were validated using K-fold cross validation.
+% \par
+% Decision trees were validated using K-fold cross validation.
 Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the decision trees.
@@
 
-We investigated the link between revenue to other variables in the SWA data by predicting it, and then linking each variable to revenue through variable importance.
 The prediction of revenue performed similarly to operating cost achieving an $R^2$ of 0.7716 (with a standard deviation of 0.1525).
-A regression tree was used as a surrogate model to present an example of the typical type of decision tree present within the XGBoost Ensemble, however the surrogate model only achieved an $R^2$ of 0.0961 (with a standard deviation of 0.0181) and the XGBoosted ensemble.
+The value of predictors' relative importance was then calculated through the number of nodes used within the XGBoost.
+Values for relative importance were then used to construct Sankey and Chord diagrams to compare the contribution of each variable to predicting revenue.
 \par
-The important variables when attempting to determine revenue were size, yield, fuel and water (see \ref{fig:revenue_sankey}).
-Due to regions being recorded separately for importance none appeared as the most important variables, overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
-Although performing poorly, the surrogate model highlights the importance of size in determining revenue.
-Area also appearing as a variable of higher importance is show to be highly interrelated with other variables.
-The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
+In order of importance, predictors of revenue were fuel use(petrol 307 and diesel 144), yield (285), size (216) and water use (199).
+Here, the values in the brackets indicate the relative importance of each variable (see \ref{fig:revenue_sankey}).
+Overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
+The chord diagram illustrates that vineyard area is also of high relative importance to other variables especially slashing.
+The overall importance of area to other variables is evident by its larger circumference within the chord diagram (see B in Figure \ref{fig:revenue_sankey}).
+% The strong relationship of vineyard area to revenue and its predictors  but also has a high relative importance to the other predictors, The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
 Area harvested is likely also an indicator of other variables such as slashing passes its strongest connection presented.
@@
 
-The relationships to operating cost through variable importance were found to be similar to that of revenue.
-With fuel, water, area and yield occurring the most (see figure \ref{fig:operating_costs_sankey}).
-A surprising difference is that the most important operational consideration for operating cost is the use of fungicide, compared to revenue where slashing is the most important (comparing Figure \ref{fig:region_sankey} and \ref{fig:operating_costs_sankey}).
-The variables that feed into these decisions are also very differing with diesel being the most informative to slashing and area being the most informative to the need for fungicide.
+Comparatively to revenue, operating cost performed better with the XGBoosted regression ensemble achieving an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+The relationships to operating cost through variable importance were found to be similar to that of revenue, with fuel, water, area and yield having the largest number relative importance (see figure \ref{fig:operating_costs_sankey}).
+A surprising difference is that the most important operational consideration for operating cost is the use of fungicide, compared to revenue where slashing is the most important (comparing Figure \ref{fig:region_sankey}).
+The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
 \par
-Again, region played a determining factor overall but not as much individually with region contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
+Again, region played a determining factor overall, contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
 It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
 \par
-Comparatively to revenue, operating cost performed better The XGBoosted regression ensemble achieved an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+% .
 Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
@@
 
-When considered overall, Region was a highly informative variable through measure of importance to both operating cost and revenue.
-Notably the overall importance of region to revenue was 234 (making it the third most important variable when considerng all regions together).
+When considered overall, Region was a highly informative variable based on measures of importance for both operating cost and revenue.
+As noted above, Region was the third most important variable for determining revenue.
 The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
-These two regions are also relative opposites in winegrowing climates with the Barossa being warm and dry climate focussing on Shiraz grapes and Tasmania being a cool wet climate that grows Pinot.
+These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot.
 \par
-When considering all regions together, it had the most node contributions to determining operating costs with an importance of 334.
-Of all the regions, again Tasmania was the most important, followed by the Adelaide Hills.
+As also noted above Region was also a key determinant of operating costs.
+Again Tasmania was the most important, followed by the Adelaide Hills.
 In contrast to revenue, both climates are considered cool and wet, and warmer drier regions such as the Barossa and Hunter Valley only contributed roughly half the same number of nodes to the ensemble.
-When looking at \ref{fig:region_sankey} the inclusion of slashing and fungicide spraying is the likely reason; with fungal and weed pressure being greater in cooler wetter regions.
+Based on further analysis of Regions (Figure \ref{fig:region_sankey}) the inclusion of slashing and fungicide spraying is the likely reason with fungal and weed pressure being greater in cooler wetter regions.
 \par
-Both diesel and petrol were of more relative importance in operating costs than water was compared with region.
-similar to those used to classify region (see Figure \ref{fig:region_sankey}, except water used.
-The surrogate model relied heavily on the use of water source to classify regions, which is reflective of regional access to resources being a deciding factor in vineyard management (see Figure \ref{fig:region_tree}).
-A major difference between region and revenue was the importance given to water use, with water use being a relatively more important variable in predicting region than revenue (considering its rank in importance to other variables).
-\par
-The surrogate model for region performed better than other surrogate models with 32.34\% (3.67\% standard deviation).
-The prominence of types and use of water resources was in classifying region is reflective of difference of availability of water resources is when comparing different regions (see Figure \ref{fig:region_tree}).
-The XGBoost ensemble, did not perform as well as operating costs or revenue with 56.82\% accuracy (50.58\% validation accuracy).
-The difference in accuracy is in part due to the large number of classes (being 58).
-The ensemble did not differ greately from the surrogate model, with a continuing emphasis on Area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
+The XGBoost ensemble, did not perform well when predicting operating costs or revenue with 56.82\% accuracy (50.58\% validation accuracy).
+The difference in accuracy is in part due to the large number of classes (58 regions).
+The ensemble had a great emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
 % Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
@@
 Bordering regions were routinely grouped together and misclassified as the same region.
-Two areas that suffered the most from this, specifically with the classification tree were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
-The classification tree likely had more difficulty discerning vineyards closer to the river using only water sources due to the greater access to river water in these areas.
+Two areas that suffered the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
 
@@
 
-The explored relationships between vineyard resource use, operations and geographical properties to revenue and operating costs highlight how decisive regional influences can be determining a vineyard's economic outcomes.
+This study has explored the relationships between vineyard resource use, operations and geographical properties to revenue and operating costs
+
+geographical regions with respect to a vineyards revenue and operating costs. The analysis was based on a large national study of XXX
+
+Three main findings were identified. First, the most important predicted of revenue were XXX
+
+With YYY showing the relationships among the corresponding set of predictor variables
+
+Second XXX for operating cost
+
+Third for regions...
+
+highlight how decisive regional influences can be determining a vineyard's economic outcomes. 
+\par
 Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
@@
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
-Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region.
-Where, the need to pressurise irrigation systems from river water or to generate power would require larger amounts of diesel and petrol.
+Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region because vineyards in regions without pressurised water will need to use more fuel to pressurise their irrigation systems.
 \par
@@
 The dominating factor of area likely played a large part in determining how costly a tractor pass would be, or in defining the ratio of water applied to the amount of vines.
-The node frequency was high for area but much lower in general across the other variables, which could indicate the need to be more circumstantial in determining operational costs.
+The node frequency was high for area but much lower in general across the other variables, which could indicate the need to be specific when attempting to determine the cause of a operational cost.
 Although it was attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), it is likely yet more complicated.
@@
 Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
-However, as buyers set prices at regional levels, it is still important to consider a vineyards Geographical Indicator region.
+However, as buyers set prices at regional levels, it is still important to consider this factor.
 \par
 Decisions made on the ground have far-reaching effects and are difficult to completely capture.
-A higher number of tractor passes used as a preventative measure for occurrences such as disease, may incur higher operational costs but could be critical in preventing long term losses.
-With factors such as erosion and soil health being difficult to capture but also influenced by tractor use \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
-Although, performing well in $R^2$, the ability to predict operational costs is limited by the variables incorporated.
+A larger number of tractor passes used as a preventative measure for occurrences such as disease may incur higher operational costs but could be critical in preventing long term losses.
+Although the models demonstrated a good predictive fit (via large $R^2$ values), the ability to predict operational costs is limited by the variables incorporated in the analysis.
+Other factors such as erosion and soil health are also influenced by tractor use and would contribute to these operational costs but are difficult to measure and were not available as part of the data \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
 Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions.
-Without fully capturing more granular activities such as the specifics of what fuel was used for, it is hard to determine what decisions specifically influence the operational costs.
-Electricity in particular is used predominantly for irrigation.
-Size is also a further consideration where slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
+Without fully capturing more granular activities for example the specific reasons for fuel use, it is difficult to determine what decisions specifically influence the operational costs.
+%  Electricity in particular is used predominantly for irrigation. 
+% Size is also a further consideration where slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
 This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
@@
 However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
-A more granular definition of region may help to better discern the differences in practices, and the reason for employing them.
+% A more granular definition of region may help to better discern the differences in practices, and the reason for employing them. 
 More sophisticated models, specifically those that utilise expert opinion, may also help to capture and address the decision-making process.
@@
 The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
-The difference between turning a profit or loss is dependent on decisions made and chance.
+The difference between turning a profit or loss is dependent on predictable factors unforecasted factors, farming practice and farmers' decisions.
 The difference between vineyards that make profit and those that do not could be a multitude of factors including differences in farming practices not captured within this study.
@@
 It is difficult to untangle the predictive and correlative nature of a variable compared to the causal reasons.
-By delving deeper into the complex interplay of variables, further advancements can be made in optimising vineyard management strategies for lowering operational costs and enhancing sustainability.
+By delving deeper into the complex interplay of variables, further advancements can be made in optimising vineyard management strategies for lowering operational costs, increasing revenue and enhancing sustainability.
 

### 2023-11-20 article.tex
finished first pass of edits

@@
 
+The predictive accuracy of each tree was assessed through a validation process.
 For each model the data was split into training data, which constituted 80\% of the original data.
@@
 $R^2$ scores were used to determine the best regression models during validation.
-$R^2$ was used instead of RMSE to avoid the issue of comparing models of different units to each other when considering how well each model extrapolated to further data.
+$R^2$ was used instead of RMSE to allow the comparison of  models with different units to each other when considering how well each model extrapolated to further data.
 For binary and multiclass variables, validation was summarised through the accuracy, the proportion of true negatives and positives.

### 2023-11-20 data_prep.py
fix impossible values

@@ -4,6 +4,15 @@ import numpy as np
 
 df = pd.read_feather("df.feather")
 
+#  We disallow errors in the dataset:
+# The first is tractor passes as fractions (the field is meant to be integer only)
+df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
+df.loc[df["insecticide_spraying_number_of_times_passes_per_year"]<1, "insecticide_spraying_number_of_times_passes_per_year"] = 0
+df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
+for col in df.loc[:, df.dtypes == np.float64].columns:
+    df.loc[df[col]<0, col] = np.nan
+df[df.loc[:, df.dtypes == np.float64].columns].min()
+
 # we get the indices of those that made profit
 indices = df[
     (df["total_operating_costs"]>0) & (df["total_operating_costs"]>0)].index

### 2023-11-20 summary_tables.py
added summary tables

@@ -1,0 +1,210 @@
+"""This is a quick script to make some summary supplementary tables."""
+
+from os import listdir
+from os.path import isfile, join
+import pandas as pd
+
+import re
+
+df = pd.read_csv("dfb.csv")
+
+cols = ["tonnes_grapes_harvested"
+    , "area_harvested"
+    , "water_used"
+    # , "total_tractor_passes"
+    , "total_fertiliser"
+    # , "synthetic_nitrogen_applied"
+    # , "organic_nitrogen_applied"
+    # , "synthetic_fertiliser_applied"
+    # , "organic_fertiliser_applied"
+    , "giregion"
+    , "data_year_id"
+    , "river_water"
+    , "groundwater"
+    , "surface_water_dam"
+    , "recycled_water_from_other_source"
+    , "mains_water"
+    , "other_water"
+    , "water_applied_for_frost_control"
+    , "bare_soil"
+    , "annual_cover_crop"
+    , "permanent_cover_crop_native"
+    , "permanent_cover_crop_non_native"
+    , "permanent_cover_crop_volunteer_sward"
+    , "diesel_vineyard"
+    , "electricity_vineyard"
+    , "petrol_vineyard"
+    , "vineyard_solar"
+    # , "vineyard_wind"
+    , "lpg_vineyard"
+    , "biodiesel_vineyard"
+    , "slashing_number_of_times_passes_per_year"
+    , "fungicide_spraying_number_of_times_passes_per_year"
+    , "herbicide_spraying_number_of_times_passes_per_year"
+    , "insecticide_spraying_number_of_times_passes_per_year"
+    , "irrigation_energy_diesel"
+    , "irrigation_energy_electricity"
+    , "irrigation_energy_pressure"
+    , "irrigation_energy_solar"
+    , "irrigation_type_dripper"
+    , "irrigation_type_flood"
+    , "irrigation_type_non_irrigated"
+    , "irrigation_type_overhead_sprinkler"
+    , "irrigation_type_undervine_sprinkler"
+    , "nh_disease"
+    # , "nh_frost"
+]
+
+# These are the columns that will be classes!
+
+cat_cols = [ # These are binary
+    "bare_soil"
+    , "annual_cover_crop"
+    , "permanent_cover_crop_native"
+    , "permanent_cover_crop_non_native"
+    , "permanent_cover_crop_volunteer_sward"
+    , "irrigation_energy_diesel"
[…]

### 2023-11-21 article.tex
summary tables

@@
 \subsection{Data}
-% 6091 samples - total
+% 6049 samples - total
 % 58 regions
@@
 Data recorded by SWA are entered manually by winegrowers using a web based interface tool.
-A total of 6091 observations were collected from 2012/2013 to 2021/2022 financial years, with each observation comprising 23 variables reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
+A total of 6049 observations were collected from 2012/2013 to 2021/2022 financial years, with each observation comprising 23 variables reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
 \par
@@
   Herbicide Spray & Times per year & 2026 & & \\
-  Year & Class & 6091 & 10 & \\
-  Disease & Class & 6091 & 2 & \\
-  Region & Class & 6091 & 58 & \\
+  Year & Class & 6049 & 10 & \\
+  Disease & Class & 6049 & 2 & \\
+  Region & Class & 6049 & 58 & \\
   Solar & Kilowatt Hours & 622 & & \\
-  Irrigation Type & Class & 6091 & 20 & \\
+  Irrigation Type & Class & 6049 & 20 & \\
   Petrol & Litres & 4309 & & \\
@@
   Yield & Tonnes & 5935 & & \\
-  Irrigation Energy & Class & 6091 & 16 & \\
-  Area Harvested & Hectares & 6091 & & \\
-  Electricity & Kilowatt Hours & 1015 & & \\
+  Irrigation Energy & Class & 6049 & 16 & \\
+  Area Harvested & Hectares & 6049 & & \\
+  Electricity & Kilowatt Hours & 1014 & & \\
   Insecticide Spray & Times per year & 1092 & & \\
@@
   Fungicide Spray & Times per year & 2260 & & \\
-  Cover Crop & Class & 6091 & 32 & \\
-  Water Type & Class & 6091 & 39 & \\
+  Cover Crop & Class & 6049 & 32 & \\
+  Water Type & Class & 6049 & 39 & \\
   Grape Revenue & AUD & 853 & & \\
@@
 
+ \section{Variable summaries}
+
+The Table \ref{cont_summary} below shows the ranges of each of the continuous variables:
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}lllllllll@{}}
+  \toprule
+   & count & mean & std & min & 0.25 & 0.5 & 0.75 & max \\ \midrule
+  Vineyard Solar & 622 & 22916.89 & 104808 & 1 & 1170.75 & 5500 & 14866.25 & 2300000 \\
+  Biodiesel & 25 & 6635.932 & 11768.832104 & 1 & 200 & 500 & 10000 & 37216 \\
+  Fungicide Spray & 2260 & 7.724801 & 3.279794 & 1 & 6 & 7 & 9 & 68 \\
+  LPG & 958 & 327.831399 & 861.538804 & 1 & 40 & 95.835 & 240 & 11950 \\
+  Petrol & 4309 & 825.276809 & 1556.621119 & 1 & 135 & 306.66 & 903 & 38568 \\
+  Insecticide Spray & 1092 & 1.707189 & 1.316042 & 0 & 1 & 1 & 2 & 12 \\
+  Water Used & 5846 & 7301838 & 558206600 & 0.0007 & 13.2655 & 43 & 146.875 & 42680000000 \\
+  Fertiliser & 795 & 91149.89 & 483913.4 & 1 & 560 & 4759.5 & 45148.5 & 11358000 \\
+  Diesel & 5585 & 11677.070183 & 24380.588742 & 0.1267 & 1240 & 3850 & 12500 & 591000 \\
+  Yield & 5935 & 772.902449 & 2175.113895 & 0.03 & 68 & 192.3 & 601.8795 & 72305 \\
+  Herbicide Spray & 2026 & 2.646199 & 2.598899 & 0 & 2 & 2 & 3 & 103 \\
+  Slashing & 2290 & 3.311485 & 1.826788 & 1 & 2 & 3 & 4 & 26 \\
+  Electricity & 1014 & 58223.07 & 177626.3 & 0.019 & 2160 & 9637 & 36498.25 & 3000000 \\
+  Area Harvested & 6049 & 66.52604 & 133.4525 & 2.220446E-16 & 10.13 & 24.5 & 66.8 & 2436.15 \\
+  Grape Revenue & 875 & 377972 & 606286.8 & 1 & 76000 & 172964 & 386747 & 5700000 \\
+  Operating Costs & 853 & 314187.1 & 511522.6 & 1 & 57315 & 140000 & 327408 & 4482828 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+  \section{water types}
+  % Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  water types & frequency \\ \midrule
+  river water & 1578 \\
+  groundwater & 1433 \\
+  surface water dam & 617 \\
+  recycled water from other source & 386 \\
+  groundwater and surface water dam & 256 \\
+  not listed & 235 \\
+  mains water & 170 \\
+  river water and groundwater & 147 \\
+  groundwater and recycled water from other source & 145 \\
+  other water & 101 \\
+  river water and surface water dam & 92 \\
+  groundwater and water applied for frost control & 90 \\
+  groundwater and mains water & 76 \\
+  river water and groundwater and  surface water dam & 70 \\
+  recycled water from other source and mains water & 63 \\
+  groundwater and recycled water from other source and mains water & 60 \\
+  river water and mains water & 57 \\
+  surface water dam and mains water & 56 \\
+  groundwater and other water & 33 \\
+  river water and groundwater and mains water & 30 \\
+  groundwater and surface water dam and recycled water from other source & 27 \\
+  river water and water applied for frost control & 27 \\
+  groundwater and surface water dam and mains water & 22 \\
+  surface water dam and recycled water from other source & 21 \\
+  river water and recycled water from other source & 19 \\
+  river water and other water & 19 \\
+  river water and surface water dam and mains water & 18 \\
+  river water and groundwater and surface water dam and mains water & 18 \\
+  mains water and other water & 16 \\
+  groundwater and surface water dam and water applied for frost control & 12 \\
+  surface water dam and other water & 12 \\
+  groundwater and recycled water from other source and other water & 11 \\
+  groundwater and surface water dam and recycled water from other source and mains water & 8 \\
+  recycled water from other source and mains water and other water & 8 \\
+  river water and recycled water from other source and mains water & 8 \\
+  river water and surface water dam and recycled water from other source & 8 \\
+  surface water dam and mains water and other water & 7 \\
+  recycled water from other source and other water & 7 \\
+  river water and groundwater and recycled water from other source & 6 \\
+  groundwater and mains water and other water & 5 \\
+  groundwater and surface water dam and other water & 5 \\
+  groundwater and surface water dam and mains water and other water & 5 \\
+  river water and groundwater and recycled water from other source and mains water & 5 \\
+  river water and groundwater and water applied for frost control & 5 \\
+  river water and surface water dam and water applied for frost control & 4 \\
+  surface water dam and water applied for frost control & 4 \\
+  river water and groundwater and surface water dam and recycled water from other source and mains water and other water & 4 \\
+  river water and groundwater and recycled water from other source and other water & 3 \\
+  groundwater and surface water dam and recycled water from other source and water applied for frost control & 3 \\
+  river water and groundwater and surface water dam and recycled water from other source & 3 \\
+  river water and recycled water from other source and other water & 3 \\
+  surface water dam and recycled water from other source and mains water & 2 \\
+  river water and recycled water from other source and mains water and water applied for frost control & 2 \\
+  groundwater and surface water dam and recycled water from other source and mains water and other water & 2 \\
+  river water and groundwater and mains water and other water & 2 \\
+  river water and groundwater and surface water dam and other water & 2 \\
+  river water and surface water dam and other water & 2 \\
+  river water and mains water and water applied for frost control & 2 \\
+  river water and groundwater and surface water dam and recycled water from other source and mains water & 2 \\
+  river water and mains water and other water & 2 \\
+  river water and surface water dam and mains water and other water & 2 \\
+  river water and groundwater and mains water and water applied for frost control & 1 \\
+  surface water dam and other water and water applied for frost control & 1 \\
+  water applied for frost control & 1 \\
+  groundwater and other water and water applied for frost control & 1 \\
+  other water and water applied for frost control & 1 \\
+  groundwater and surface water dam and recycled water from other source and other water and water applied for frost control & 1 \\
+  mains water and water applied for frost control & 1 \\
+  groundwater and surface water dam and recycled water from other source and other water & 1 \\
+  groundwater and mains water and water applied for frost control & 1 \\
+  river water and groundwater and surface water dam and mains water and other water & 1 \\
+  river water and surface water dam and recycled water from other source and mains water & 1 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+\section{cover crop types}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  Cover crop types & frequency \\ \midrule
+  permanent_cover_crop_volunteer_sward & 1822 \\
+  permanent_cover_crop_non_native & 936 \\
+  permanent_cover_crop_native & 490 \\
+  annual_cover_crop & 479 \\
+  groundwater and surface water dam & 406 \\
+  annual_cover_crop and permanent_cover_crop_volunteer_sward & 309 \\
+  bare_soil & 225 \\
+  permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 214 \\
+  annual_cover_crop and permanent_cover_crop_non_native & 169 \\
+  bare_soil and permanent_cover_crop_volunteer_sward & 129 \\
+  bare_soil and permanent_cover_crop_non_native & 115 \\
+  annual_cover_crop and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 101 \\
+  bare_soil and annual_cover_crop & 93 \\
+  permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 80 \\
+  bare_soil and permanent_cover_crop_native & 78 \\
+  annual_cover_crop and permanent_cover_crop_native & 78 \\
+  permanent_cover_crop_native and permanent_cover_crop_non_native & 68 \\
+  permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 44 \\
+  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 44 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_volunteer_sward & 33 \\
+  bare_soil and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 26 \\
+  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 17 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_native & 15 \\
+  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native & 15 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_non_native & 13 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 12 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 11 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native & 8 \\
+  bare_soil and permanent_cover_crop_native and permanent_cover_crop_non_native & 7 \\
+  bare_soil and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 6 \\
+  bare_soil and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 4 \\
+  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward and & 2 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+\section{irrigation type}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  Irrigation type & frequency \\ \midrule
+  dripper & 4800 \\
+  dripper and non_irrigated & 342 \\
+  Not listed & 319 \\
+  dripper and overhead_sprinkler & 201 \\
+  dripper and undervine_sprinkler & 91 \\
+  non_irrigated & 65 \\
+  undervine_sprinkler & 53 \\
+  dripper and flood & 53 \\
+  overhead_sprinkler & 46 \\
+  dripper and overhead_sprinkler and undervine_sprinkler & 28 \\
+  overhead_sprinkler and undervine_sprinkler & 12 \\
+  dripper and non_irrigated and overhead_sprinkler & 11 \\
+  flood and undervine_sprinkler & 10 \\
+  dripper and flood and undervine_sprinkler & 7 \\
+  dripper and flood and non_irrigated and overhead_sprinkler and undervine_sprinkler & 3 \\
+  dripper and flood and overhead_sprinkler & 3 \\
+  non_irrigated and undervine_sprinkler & 2 \\
+  dripper and flood and non_irrigated & 1 \\
+  dripper and non_irrigated and overhead_sprinkler and undervine_sprinkler & 1 \\
+  flood and & 1 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+  \section{irrigation energy type}
+
+  % Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  Irrigation energy type & frequency \\ \midrule
+  electricity & 2162 \\
+  not listed & 2053 \\
+  pressure & 586 \\
+  electricity and pressure & 396 \\
+  diesel & 254 \\
+  diesel and electricity & 227 \\
+  electricity and solar & 96 \\
+  diesel and electricity and pressure & 90 \\
+  diesel and pressure & 74 \\
+  solar & 50 \\
+  electricity and pressure and solar & 23 \\
+  diesel and electricity and solar & 14 \\
+  diesel and electricity and pressure and solar & 10 \\
+  pressure and solar & 9 \\
+  diesel and solar & 4 \\
+  diesel and pressure and solar and & 1 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+  \section{year}
+  % Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  data_year_id & frequency \\ \midrule
+  2021/2022 & 954 \\
+  2020/2021 & 860 \\
+  2019/2020 & 599 \\
+  2012/2013 & 590 \\
+  2013/2014 & 549 \\
+  2015/2016 & 548 \\
+  2014/2015 & 505 \\
+  2017/2018 & 493 \\
+  2016/2017 & 485 \\
+  2018/2019 & 466 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
+\section{giregion}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}ll@{}}
+  \toprule
+  giregion & frequency \\ \midrule
+  McLaren Vale & 1195 \\
+  Barossa Valley & 584 \\
+  Murray Darling & 521 \\
+  Riverland & 472 \\
+  Adelaide Hills & 454 \\
+  Langhorne Creek & 347 \\
+  Margaret River & 344 \\
+  Coonawarra & 284 \\
+  Padthaway & 202 \\
+  Wrattonbully & 195 \\
+  Clare Valley & 149 \\
+  Yarra Valley & 122 \\
+  Eden Valley & 92 \\
+  Tasmania & 89 \\
+  Swan Hill & 83 \\
+  Grampians & 73 \\
+  Orange & 72 \\
+  Hunter Valley & 70 \\
+  Bendigo & 53 \\
+  Great Southern & 51 \\
+  Rutherglen & 41 \\
+  Robe & 36 \\
+  Tumbarumba & 35 \\
+  Mornington Peninsula & 32 \\
+  King Valley & 32 \\
+  Southern Fleurieu & 30 \\
+  Heathcote & 29 \\
+  Adelaide Plains & 25 \\
+  Currency Creek & 24 \\
+   & 23 \\
+  Henty & 22 \\
+  Canberra District & 21 \\
+  Southern Flinders Ranges & 20 \\
+  Upper Goulburn & 20 \\
+  Mudgee & 20 \\
+  Mount Benson & 20 \\
+  Other & 19 \\
+  Riverina & 18 \\
+  Alpine Valleys & 15 \\
+  Barossa Zone & 14 \\
+  Pemberton & 12 \\
+  Mount Gambier & 11 \\
+  Blackwood Valley & 10 \\
+  Kangaroo Island & 10 \\
+  Big Rivers Zone Other & 9 \\
+  Geographe & 7 \\
+  Cowra & 6 \\
+  Gundagai & 5 \\
+  Strathbogie Ranges & 5 \\
+  Glenrowan & 4 \\
+  Geelong & 4 \\
+  Swan District & 4 \\
+  Goulburn Valley & 3 \\
+  Beechworth & 3 \\
+  Southern Highlands & 3 \\
+  Macedon Ranges & 2 \\
+  Pyrenees & 2 \\
+  Sunbury & 1 \\ \bottomrule
+  \end{tabular}
+  \end{table}
+
 \section{XGBoost}

### 2023-11-21 summary_tables.py

@@ -8,6 +8,13 @@ import re
 
 df = pd.read_csv("dfb.csv")
 
+df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
+df.loc[df["insecticide_spraying_number_of_times_passes_per_year"]<1, "insecticide_spraying_number_of_times_passes_per_year"] = 0
+df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
+for col in df.loc[:, df.dtypes == np.float64].columns:
+    df.loc[df[col]<0, col] = np.nan
+df[df.loc[:, df.dtypes == np.float64].columns].min()
+
 cols = ["tonnes_grapes_harvested"
     , "area_harvested"
     , "water_used"
@@ -172,39 +179,14 @@ cat_cols = [ # These are binary
     , "giregion"
 ]
 
-# We comment out these unless we really need to redo every single variable. As they are not compared to the target variables at the end!
+df = df.replace({0: np.nan})
+for col in cols:
+    print(df[col].describe())
+    print("\n\n\n\n")
 
-
-
-# cols += cat_cols
-# print(files)
-# for y_name in files:
-#     print(y_name)
-#     if y_name in cat_cols:
-#         if y_name in ["nh_disease", "nh_frost"]:
-#             train_model_b(df[cols]
-#                 , y_name)
-#         else:
-#             train_model_multi(df[cols]
-#                 , y_name)
-#     else:
-#         train_model_reg(df[cols]
-#                 , y_name)
-
-# We also do the predicted variables!
-
-# These are artefacts. As profit can be negative it is harder to predict! We know after some further mapping that both operational costs and gross margin have a cubic root relationships to the logarithm of other variables. An interesting relationship to have!
-# 
-# Importantly this cubic root relationship exists between operational costs and gross margin, which is worth noting. However we want to know what tips it either side of them being equal as that shows that a vineyard is profitable!!
-# 
-# train_model_b(df[df["profitable"].notnull()][cols + ["profitable"]]
-#                 , "profitable")
-
-# train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
-#         , "profit")
-
-# train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
-#                 , "total_operating_costs")
-
[…]

### 2023-11-21 data_prep.py
addendum to data preprocessing

@@ -13,6 +13,9 @@ for col in df.loc[:, df.dtypes == np.float64].columns:
     df.loc[df[col]<0, col] = np.nan
 df[df.loc[:, df.dtypes == np.float64].columns].min()
 
+#  We drop vineyards with incorrect area:
+df = df.drop(df[df["area_harvested"].isnull()].index)
+
 # we get the indices of those that made profit
 indices = df[
     (df["total_operating_costs"]>0) & (df["total_operating_costs"]>0)].index

### 2023-11-21 article.tex
backing up the article

@@
 
-\subsection{Data}
+\subsection{Data}\label{Data}
 % 6049 samples - total
@@
 \end{table}
-% FIXME: Add supplementary tables of all classes and another table of the summary statistics of continuous variables.
-Refer to this table to help with the paragraphs about one-hot encoding.
-A quick definition of one-hot encoding will also help!
 The data originally contained only two multiclass variables: year and region.
-Variables that measured the same metric from different sources (such as water collected from rivers versus water from dams) were converted into multiclass variables representing the source through one-hot-encoding.
-Changing each variable class into a binary value, with one indicating the presence of the class and zero indicating its absence.
-Occurrences of multiple sources were defined as their own separate classes.
-Where a class variable had a recorded amount the total amount used from these variables was retained as a separate variable; for example water used (in Mega Litres ) was also included alongside water source.
+Related binary variables, such as the use of river water and the use of dam water, were combined to create a single multiclass variable.
+This was done by first converting each combination that occurred into a unique category (such as river and damn water used, as opposed to the two separate categories prior).
+These variables were then one-hot-encoded, changing each variable class into a binary value, with one indicating the presence of the class and zero indicating its absence.
+Further details about classes and their frequency is available in the appendices.
 \par
@@
 \subsection{XGBoosted Trees}
-XGBoosted (eXtreme Gradient Boosting) trees, described in more detail below,  were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+XGBoosted (eXtreme Gradient Boosting) trees, described in more detail below (and further in the appendix), were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
 XGBoosted trees are a boosted tree ensemble method that can be used to classify classes, or predict continuous response variables.
@@
 % Then say that the results of the analyses in Set 3 were not considered further due to how crap they were.
-An XGBoosted tree was created for each variable to show how they interacted.
+As the purpose of the analysis was to identify relationships between variables and to show how they interact, an XGBoosted tree was created for every variable.
 Each tree included all variables except the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
@@
 \subsection{Variable Importance}\label{sec:importance}
-% FIXME: this section can be tightende/shortened a lot - moving some of the material into the supplementary section
 Due to XGBoost creating a large amount of decision trees, the interpretability of these models is obfuscated by the intricate relationships within complicated ensembles.
@@
 The measure of a variable's importance within this study can then be interpreted as how often a variable was the optimal choice in reducing the loss function of the ensemble.
-Importantly, multiclass variables being one-hot-encoded
-% FIXME: make sure one-hot-encoding is well defined and rephrase this so it makes more sense (possibly define it earlier as suggested in the data section.)
-are given an importance score for each individual class; for example, each specific region will have its own importance score.
+Importantly, multiclass variables being one-hot-encoded (see Section \ref{Data}) are given an importance score for each individual class; for example, each specific region will have its own importance score.
 \par
@@
 % \newline
-% FIXME: The first result should be in the langauge of the question.
-In the original document the first result presented was region, implying that the question is: 'are regions different?' which is not the forefocus of the article.
-The results need to be rearranged to show grossmargin and operating costs as more significant variables as they are the curious factors that we do not understand.
-
-% FIXME: The irritating placement of the figures is due to the way latex uses the /ref function to determine where a figure should be.
-Have a google of this to overcome it as the figure placement is all over the shop.
 
@@
 This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
-% This is possibly the connection between the presence of the Hunter Valley within the XGBoost ensemble that determined profit (see Figure \ref{fig:profit_sankey}).
-With this connection reflecting the restriction of possible strategies employable by winegrowers between different regions.
-% FIXME:  No this is because the hunter valley/tasmania get paid more / tonne.
 \par
@@
 \par
-% Although less important in the XGBoost ensembles for profit, the variables: cover crops, and number of fungicide spraying and slashing passes are likely linked to broad environmental properties of regions (see Figure \ref{fig:region_sankey} and \ref{fig:profit_sankey}).
-Rainfall being related to vine fungal growth, disease and weed pressure.
-With cover crops being an effective and sustainable method to alleviate these issues \citep{delpuechAdaptingCoverCrop2018}.
-% FIXME: Re-structure this sentance.
-
 % It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
@@
 
- \section{Variable summaries}
-
-The Table \ref{cont_summary} below shows the ranges of each of the continuous variables:
+ \section{Continuous variables}
+
+The Table below shows the ranges of each of the continuous variables:
 % Please add the following required packages to your document preamble:
@@
 
-  \section{water types}
+  \section{Categorical variables}\ref{app: cat_var}
+  The tables below describe each possible class a multiclass variable could have taken and the frequency that it occured.
+  \subsection{water types}
   % Please add the following required packages to your document preamble:
@@
 
-\section{cover crop types}
+\subsection{cover crop types}
 % Please add the following required packages to your document preamble:
@@
 
-\section{irrigation type}
+\subsection{irrigation type}
 
@@
 
-  \section{irrigation energy type}
+  \subsection{irrigation energy type}
 
@@
 
-  \section{year}
+  \subsection{year}
   % Please add the following required packages to your document preamble:
@@
 
-\section{giregion}
+\subsection{giregion}
 % Please add the following required packages to your document preamble:

### 2023-11-21 article.tex
attempt at fixing my xgboost definitions.

@@
 \par
-% FIXME: Make it clearer why you did this "Because the purpose of the analysis was to identify relationships between variables and since there was no single outcome variable of interest an XGBoosted tree was created for each variable using the other variables as predictors
-% FIXME: forward reference here "XGBoosted trees, described in more detail below...
-% FIXME: I found this a bit confusing as a 'new reader'. How about describing this process as a set of analyses:
-% Step 1: Separate analyses of operating cost and revenue: XGBoosted models were constructed for each of these variables, with the variable as the response and all other variables (with the exception of operating cost and revenue) as predictors.
-% Step 2: For each of the operating cost and revenue models, an analysis of the inter-relationships between the ten most important predictors based on the XGBoosted model.
-% Set 3: Analysis of profit (difference between...): etc
-% Then say that the results of the analyses in Set 3 were not considered further due to how crap they were.
-As the purpose of the analysis was to identify relationships between variables and to show how they interact, an XGBoosted tree was created for every variable.
-Each tree included all variables except the economic variables (operating cost and revenue from grape sales), which were only included within their own trees as response variables.
-Separately profit (the difference between revenue and operational costs) was looked at in prior analyses (see appendix) but the results were not included due to low average loss values and model stability.
-This meant that every variable would have a measure of its importance to other variables (see Section \ref{sec:importance}), which was used to show the highly interrelated nature of variables within vineyards.
-The interaction between variables was illustrated using Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variable and the response (see section \ref{sec:importance}).
-\par
-% FIXME: We need a digestable description of XGBoost for readers of this article.
+XGBoosted models were constructed with operational cost and grape revenue as the predicted variables.
+The analyses were aimed at uncovering what factors influenced these variables and to what extent.
+As the purpose of the analysis was to identify relationships between variables and to show how they interact, an XGBoosted tree was created for each of the predictor variables as well.
+Trees for the predictor variables did not include operational cost or grape revenue as predictors.
+By creating an XGBoosted tree for each variable it meant that every variable would have a measure of its relative importance to every other variable (see Section \ref{sec:importance}).
+Together these models were used to measure the interrelationships of the ten most important variables in determining operational cost and grape revenue using variable importance.
+These measures of relative importance were used to illustrate the highly interrelated nature of variables within vineyards.
+The interaction between variables was depicted through the use of Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variable and the response (see section \ref{sec:importance}).
+\par
+Due to constraints from the XGBoost library region could only be incorporated as a one-hot-encoded variable when used as a predictor.
+To better show what variables were related to region overall, another XGBoost tree was created with Region as the predicted value.
+The difference for this model was that relative variable importance would only be measured for each region specifically, as opposed to a variables overall importance in determining region.
+Separately profit (the difference between revenue and operational costs) and year was looked at in prior analyses (see appendix) but these results were not included due to low average loss values and model stability.
+\par
+XGBoosted trees are an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
+The gradient boosting aspect of the ensemble is the use of a loss function used to create new decision trees that add to the ensemble.
+Each new tree created is done so using a loss function that is optimised iteratively to improve upon prior tree's predictive power.
+The loss function can be any convex function, allowing gradient descent to traverse the loss space until, no improvements can be made via traversal.
+Because the loss function is only required to be convex, both classifiers and regressors can be used.
+Regularisation methods can also be incorporated to help prevent over fitting.
 \subsection{Variable Importance}\label{sec:importance}

### 2023-11-21 article.tex
minor edits to styling

@@
 %%Research highlights
-\begin{highlights}
- \item Highlight 1
- \item Highlight 2
- \item Highlight 3
- \item Highlight 4
-\end{highlights}
+% \begin{highlights}
+%  \item Highlight 1
+%  \item Highlight 2
+%  \item Highlight 3
+%  \item Highlight 4
+% \end{highlights}
 \end{frontmatter}
@@
  \caption{Summary of variables used in the analysis. The recorded column indicate the number of values that were either greater than zero or that were not missing.}\label{tab:summary}
+ \small
  \begin{tabular}{@{}ccccl@{}}
   \toprule
-  \textbf{Variable} & \textbf{Units} & \textbf{Recorded} & \textbf{\begin{tabular}[c]{@{}c@{}}Number of\\ Classes\end{tabular}} & \\ \midrule
+  \textbf{Variable} & \textbf{Units} & \textbf{Recorded} & \textbf{Number of Classes} & \\ \midrule
   Water Used & Mega Litres & 5846 & & \\
@@
   Insecticide Spray & Times per year & 1092 & & \\
-  Fertiliser & \begin{tabular}[c]{@{}c@{}}Kilograms \\ of Nitrogen\end{tabular} & 795 & & \\
+  Fertiliser & KGs of Nitrogen & 795 & & \\
   Fungicide Spray & Times per year & 2260 & & \\
@@
 \end{table}
+\normalsize
 The data originally contained only two multiclass variables: year and region.
@@
   Cover crop types & frequency \\ \midrule
-  permanent_cover_crop_volunteer_sward & 1822 \\
-  permanent_cover_crop_non_native & 936 \\
-  permanent_cover_crop_native & 490 \\
-  annual_cover_crop & 479 \\
+  permanent cover crop volunteer sward & 1822 \\
+  permanent cover crop non native & 936 \\
+  permanent cover crop native & 490 \\
+  annual cover crop & 479 \\
   groundwater and surface water dam & 406 \\
-  annual_cover_crop and permanent_cover_crop_volunteer_sward & 309 \\
-  bare_soil & 225 \\
-  permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 214 \\
-  annual_cover_crop and permanent_cover_crop_non_native & 169 \\
-  bare_soil and permanent_cover_crop_volunteer_sward & 129 \\
-  bare_soil and permanent_cover_crop_non_native & 115 \\
-  annual_cover_crop and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 101 \\
-  bare_soil and annual_cover_crop & 93 \\
-  permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 80 \\
-  bare_soil and permanent_cover_crop_native & 78 \\
-  annual_cover_crop and permanent_cover_crop_native & 78 \\
-  permanent_cover_crop_native and permanent_cover_crop_non_native & 68 \\
-  permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 44 \\
-  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 44 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_volunteer_sward & 33 \\
-  bare_soil and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 26 \\
-  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 17 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_native & 15 \\
-  annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native & 15 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_non_native & 13 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 12 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 11 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_non_native & 8 \\
-  bare_soil and permanent_cover_crop_native and permanent_cover_crop_non_native & 7 \\
-  bare_soil and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward & 6 \\
-  bare_soil and permanent_cover_crop_native and permanent_cover_crop_non_native and permanent_cover_crop_volunteer_sward & 4 \\
-  bare_soil and annual_cover_crop and permanent_cover_crop_native and permanent_cover_crop_volunteer_sward and & 2 \\ \bottomrule
+  annual cover crop and permanent cover crop volunteer sward & 309 \\
+  bare soil & 225 \\
+  permanent cover crop non native and permanent cover crop volunteer sward & 214 \\
+  annual cover crop and permanent cover crop non native & 169 \\
+  bare soil and permanent cover crop volunteer sward & 129 \\
+  bare soil and permanent cover crop non native & 115 \\
+  annual cover crop and permanent cover crop non native and permanent cover crop volunteer sward & 101 \\
+  bare soil and annual cover crop & 93 \\
+  permanent cover crop native and permanent cover crop volunteer sward & 80 \\
+  bare soil and permanent cover crop native & 78 \\
+  annual cover crop and permanent cover crop native & 78 \\
+  permanent cover crop native and permanent cover crop non native & 68 \\
+  permanent cover crop native and permanent cover crop non native and permanent cover crop volunteer sward & 44 \\
+  annual cover crop and permanent cover crop native and permanent cover crop non native and permanent cover crop volunteer sward & 44 \\
+  bare soil and annual cover crop and permanent cover crop volunteer sward & 33 \\
+  bare soil and permanent cover crop non native and permanent cover crop volunteer sward & 26 \\
+  annual cover crop and permanent cover crop native and permanent cover crop volunteer sward & 17 \\
+  bare soil and annual cover crop and permanent cover crop native & 15 \\
+  annual cover crop and permanent cover crop native and permanent cover crop non native & 15 \\
+  bare soil and annual cover crop and permanent cover crop non native & 13 \\
+  bare soil and annual cover crop and permanent cover crop native and permanent cover crop non native and permanent cover crop volunteer sward & 12 \\
+  bare soil and annual cover crop and permanent cover crop non native and permanent cover crop volunteer sward & 11 \\
+  bare soil and annual cover crop and permanent cover crop native and permanent cover crop non native & 8 \\
+  bare soil and permanent cover crop native and permanent cover crop non native & 7 \\
+  bare soil and permanent cover crop native and permanent cover crop volunteer sward & 6 \\
+  bare soil and permanent cover crop native and permanent cover crop non native and permanent cover crop volunteer sward & 4 \\
+  bare soil and annual cover crop and permanent cover crop native and permanent cover crop volunteer sward and & 2 \\ \bottomrule
   \end{tabular}
@@
   dripper & 4800 \\
-  dripper and non_irrigated & 342 \\
+  dripper and non irrigated & 342 \\
   Not listed & 319 \\
-  dripper and overhead_sprinkler & 201 \\
-  dripper and undervine_sprinkler & 91 \\
-  non_irrigated & 65 \\
-  undervine_sprinkler & 53 \\
+  dripper and overhead sprinkler & 201 \\
+  dripper and undervine sprinkler & 91 \\
+  non irrigated & 65 \\
+  undervine sprinkler & 53 \\
   dripper and flood & 53 \\
-  overhead_sprinkler & 46 \\
-  dripper and overhead_sprinkler and undervine_sprinkler & 28 \\
-  overhead_sprinkler and undervine_sprinkler & 12 \\
-  dripper and non_irrigated and overhead_sprinkler & 11 \\
-  flood and undervine_sprinkler & 10 \\
-  dripper and flood and undervine_sprinkler & 7 \\
-  dripper and flood and non_irrigated and overhead_sprinkler and undervine_sprinkler & 3 \\
-  dripper and flood and overhead_sprinkler & 3 \\
-  non_irrigated and undervine_sprinkler & 2 \\
-  dripper and flood and non_irrigated & 1 \\
-  dripper and non_irrigated and overhead_sprinkler and undervine_sprinkler & 1 \\
+  overhead sprinkler & 46 \\
+  dripper and overhead sprinkler and undervine sprinkler & 28 \\
+  overhead sprinkler and undervine sprinkler & 12 \\
+  dripper and non irrigated and overhead sprinkler & 11 \\
+  flood and undervine sprinkler & 10 \\
+  dripper and flood and undervine sprinkler & 7 \\
+  dripper and flood and non irrigated and overhead sprinkler and undervine sprinkler & 3 \\
+  dripper and flood and overhead sprinkler & 3 \\
+  non irrigated and undervine sprinkler & 2 \\
+  dripper and flood and non irrigated & 1 \\
+  dripper and non irrigated and overhead sprinkler and undervine sprinkler & 1 \\
   flood and & 1 \\ \bottomrule
@@
   \toprule
-  data_year_id & frequency \\ \midrule
+  data year id & frequency \\ \midrule
   2021/2022 & 954 \\

### 2023-11-21 article.tex
final draft all is good!!

@@
 \usepackage{booktabs}
+\usepackage{longtable}
 
@@
 
-\begin{figure} 
-  \resizebox{\textwidth}{!}{
-  \includegraphics{revenue.pdf}}
-  \caption{Decision tree predicting revenue.
+% \begin{figure} 
+%   \resizebox{\textwidth}{!}{
+%   \includegraphics{revenue.pdf}}
+%   \caption{Decision tree predicting revenue.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:revenue_tree}
- \end{figure}
+%  \end{figure}
 
@@
 
-\begin{figure} 
-  \resizebox{\textwidth}{!}{
-  \includegraphics{operating_costs.pdf}}
-  \caption{A surrogate model decision tree predicting operating costs.
+% \begin{figure} 
+%   \resizebox{\textwidth}{!}{
+%   \includegraphics{operating_costs.pdf}}
+%   \caption{A surrogate model decision tree predicting operating costs.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:operating_tree}
- \end{figure}
+%  \end{figure}
 
@@
 
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{region.pdf}}
-  \caption{Decision tree predicting Region.
+% \begin{figure}
+%   \resizebox{\textwidth}{!}{
+%   \includegraphics{region.pdf}}
+%   \caption{Decision tree predicting Region.
 Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:region_tree}
- \end{figure}
+%  \end{figure}
 
@@
 
-This study has explored the relationships between vineyard resource use, operations and geographical properties to revenue and operating costs
-
-geographical regions with respect to a vineyards revenue and operating costs. The analysis was based on a large national study of XXX
-
-Three main findings were identified. First, the most important predicted of revenue were XXX
-
-With YYY showing the relationships among the corresponding set of predictor variables
-
-Second XXX for operating cost
-
-Third for regions...
-
-highlight how decisive regional influences can be determining a vineyard's economic outcomes. 
+This study explored the relationships between vineyard resource use, operations and geographical properties to revenue and operating costs.
+The analysis was based on a large national study of 6049 samples collected over ten years.
+Three main findings were identified.
+First, the most important predictors of revenue and operating costs were fuel, yield and area.
+Secondly, area and fuel were highly interrelated to other variables (see Figure \ref{fig:operating_costs_sankey} and Figure \ref{fig:revenue_sankey}).
+Finally, the relative importance of predictor variables for region, differed from Revenue and operating costs, with water use being more prominent than yield.
+Region was also more prominent than illustrated in the Sankey diagrams due to the relative importance for operating cost and revenue being calculated for individual regions and not all regions together.
+In its entirety region contributed third most prominently in relative importance to revenue, and was of the most relative importance in determining operating costs.
 \par
@@
 \par
-In part some winegrowing strategies are restricted simply through access to water resources, being reflected through the region classification tree (see Figure \ref{fig:region_tree}).
+In part some winegrowing strategies are restricted simply through access to water resources.
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
@@
 
-The Table below shows the ranges of each of the continuous variables:
+Table \ref{tab:cont_vars} below shows the ranges of each of the continuous variables:
 % Please add the following required packages to your document preamble:
@@
 \begin{table}[]
-  \begin{tabular}{@{}lllllllll@{}}
+  \caption{Summary statistics of continuous variables used in XGBoosted models.}\label{tab:cont_vars}
+  \resizebox{\textwidth}{!}{\begin{tabular}{@{}lllllllll@{}}
   \toprule
@@
   Operating Costs & 853 & 314187.1 & 511522.6 & 1 & 57315 & 140000 & 327408 & 4482828 \\ \bottomrule
-  \end{tabular}
+  \end{tabular}}
   \end{table}
-
-  \section{Categorical variables}\ref{app: cat_var}
+\clearpage
+
+  \section{Categorical Variables}
   The tables below describe each possible class a multiclass variable could have taken and the frequency that it occured.
-  \subsection{water types}
+  \subsection{Water Source Types}
+  Table \ref{tab:water_type} below shows the different class types for water sources used by vineyards and their frequency of occurrences.
   % Please add the following required packages to your document preamble:
 % \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  water types & frequency \\ \midrule
+
+\begin{center}
+  \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+    \caption{Frequency and class types of water types used by vineyards.}\label{tab:water_type} \\
+
+  \hline \textbf{Water types} & \textbf{frequency}\\ \hline 
+  \endfirsthead
+
+  \multicolumn{2}{c}%
+  {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+  \hline \textbf{Water types} & \textbf{frequency} \\ \hline 
+  \endhead
+
+  \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+  \endfoot
+
+  \hline \hline
+  \endlastfoot
     river water & 1578 \\
@@
     river water and groundwater and surface water dam and mains water and other water & 1 \\
-  river water and surface water dam and recycled water from other source and mains water & 1 \\ \bottomrule
-  \end{tabular}
-  \end{table}
-
-\subsection{cover crop types}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  Cover crop types & frequency \\ \midrule
+    river water and surface water dam and recycled water from other source and mains water & 1 \\
+
+  \end{longtable}
+  \end{center}
+
+  \clearpage
+\subsection{Cover Crop Types}
+Table \ref{tab:cover_crop} below shows the different cover crop types used together and their frequency.
+
+\begin{center}
+  \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+    \caption{Frequency and class types of cover crop types used by vineyards.}\label{tab:cover_crop} \\
+
+  \hline \textbf{Cover crop types} & \textbf{frequency}\\ \hline 
+  \endfirsthead
+
+  \multicolumn{2}{c}%
+  {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+  \hline \textbf{Cover crop types} & \textbf{frequency} \\ \hline 
+  \endhead
+
+  \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+  \endfoot
+
+  \hline \hline
+  \endlastfoot
+
+  Cover crop types & frequency \\
   permanent cover crop volunteer sward & 1822 \\
@@
   bare soil and permanent cover crop native and permanent cover crop non native and permanent cover crop volunteer sward & 4 \\
-  bare soil and annual cover crop and permanent cover crop native and permanent cover crop volunteer sward and & 2 \\ \bottomrule
-  \end{tabular}
-  \end{table}
-
-\subsection{irrigation type}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  Irrigation type & frequency \\ \midrule
+  bare soil and annual cover crop and permanent cover crop native and permanent cover crop volunteer sward and & 2 \\
+  \end{longtable}
+  \end{center}
+  \clearpage
+
+\subsection{Irrigation Types}
+Below in Table \ref{tab:irig_types} are the frequency and different irrigation types.
+
+\begin{center}
+  \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+    \caption{Frequency and class types of irrigation types used by vineyards.}\label{tab:irig_types} \\
+
+  \hline \textbf{Irrigation types} & \textbf{frequency}\\ \hline 
+  \endfirsthead
+
+  \multicolumn{2}{c}%
+  {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+  \hline \textbf{Irrigation types} & \textbf{frequency} \\ \hline 
+  \endhead
+
+  \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+  \endfoot
+
+  \hline \hline
+  \endlastfoot
+  Irrigation type & frequency \\
   dripper & 4800 \\
@@
   dripper and non irrigated and overhead sprinkler and undervine sprinkler & 1 \\
-  flood and & 1 \\ \bottomrule
-  \end{tabular}
-  \end{table}
-
-  \subsection{irrigation energy type}
-
-  % Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  Irrigation energy type & frequency \\ \midrule
+  flood and & 1 \\
+\end{longtable}
+\end{center}
+
+  \clearpage
+
+  \subsection{Irrigation Energy Type}
+Below, Table \ref{irig_energy} shows the different types of energy used to power vineyards and their frequency.
+
+\begin{center}
+  \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+    \caption{Frequency and class types of irrigation energy types used by vineyards.}\label{tab:irig_energy}  \\
+
+  \hline \textbf{Irrigation Energy types} & \textbf{frequency}\\ \hline 
+  \endfirsthead
+
+  \multicolumn{2}{c}%
+  {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+  \hline \textbf{Irrigation Energy types} & \textbf{frequency} \\ \hline 
+  \endhead
+
+  \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+  \endfoot
+
+  \hline \hline
+  \endlastfoot
+
+  Irrigation energy type & frequency \\
   electricity & 2162 \\
@@
   diesel and pressure and solar and & 1 \\ \bottomrule
-  \end{tabular}
-  \end{table}
-
-  \subsection{year}
-  % Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  data year id & frequency \\ \midrule
+\end{longtable}
+\end{center}
+
+  \clearpage
+
+  \subsection{Year}
+Below in Table \ref{tab:year} is the list of years and the number of sample collected in each.
+
+    \begin{center}
+      \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+        \caption{Frequency and class types of year}\label{tab:year}  \\
+
+      \hline \textbf{Year } & \textbf{frequency}\\ \hline 
+      \endfirsthead
+
+      \multicolumn{2}{c}%
+      {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+      \hline \textbf{Year} & \textbf{frequency} \\ \hline 
+      \endhead
+
+      \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+      \endfoot
+
+      \hline \hline
+      \endlastfoot
+
+  Year & frequency \\
   2021/2022 & 954 \\
@@
   2016/2017 & 485 \\
-  2018/2019 & 466 \\ \bottomrule
-  \end{tabular}
-  \end{table}
-
-\subsection{giregion}
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-  \begin{tabular}{@{}ll@{}}
-  \toprule
-  giregion & frequency \\ \midrule
+  2018/2019 & 466 \\
+
+\end{longtable}
+\end{center}
+
+\clearpage
+
+\subsection{Region}
+Below in Table \ref{tab:region} are the number of collected samples for each region.
+\begin{center}
+  \begin{longtable}{p{0.5\linewidth} p{.35\linewidth}}
+    \caption{Frequency and class types of regions.}\label{tab:region}  \\
+
+  \hline \textbf{Regions} & \textbf{frequency}\\ \hline 
+  \endfirsthead
+
+  \multicolumn{2}{c}%
+  {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+  \hline \textbf{Regions} & \textbf{frequency} \\ \hline 
+  \endhead
+
+  \hline \multicolumn{2}{|r|}{{Continued on next page}} \\ \hline
+  \endfoot
+
+  \hline \hline
+  \endlastfoot
+
+  giregion & frequency \\
   McLaren Vale & 1195 \\
@@
   Sunbury & 1 \\ \bottomrule
-  \end{tabular}
-  \end{table}
+\end{longtable}
+\end{center}
+\clearpage
 

### 2023-11-21 article.tex
included abstract and minor changes after proof reading

@@
 
-% \begin{abstract}
-% \end{abstract}
+\begin{abstract}
+
+  The Australian wine industry is a major part of Australia's agricultural sector.
+As global demands change and new pressures on the industry present themselves, a more sustainable approach is needed.
+Through a nationwide data set, collected over ten years we link key variables in determining vineyard operational costs and revenue through the use of XGBoosted ensembles.
+We use the measure of relative importance to show the interrelated nature of these variables and the comparative influence they have on one another.
+We present these connections through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and how highly interrelated these variables are to one another.
+Furthermore, we connect these variables to different wine regions highlighting the complex interrelatedness of how location effects the use of different resources.
+The study provides valuable insights into the multifaceted dynamics governing operational costs and revenue illustrating how factors such as water and fuel use impacts operational costs and how different seasonal events affect these operations.
+
+\end{abstract}
 %%Graphical abstract`
@@
 The data originally contained only two multiclass variables: year and region.
-Related binary variables, such as the use of river water and the use of dam water, were combined to create a single multiclass variable.
-This was done by first converting each combination that occurred into a unique category (such as river and damn water used, as opposed to the two separate categories prior).
+Related binary variables, such as the use of river water and the use of dam water, were combined to create multiclass variables such as water source for this example (see Appendix for further details).
+This was done by first converting each combination that occurred into its own unique category (such as river and damn water used, as opposed to two individual and separate categories).
 These variables were then one-hot-encoded, changing each variable class into a binary value, with one indicating the presence of the class and zero indicating its absence.
@@
 To better show what variables were related to region overall, another XGBoost tree was created with Region as the predicted value.
-The difference for this model was that relative variable importance would only be measured for each region specifically, as opposed to a variables overall importance in determining region.
-Separately profit (the difference between revenue and operational costs) and year was looked at in prior analyses (see appendix) but these results were not included due to low average loss values and model stability.
+The difference for this model was that relative variable importance for each variable would be measured for the overall importance in determining region, as opposed to a variables connection to each region specifically.
+Separately profit (the difference between revenue and operational costs) and year were looked at in prior analyses (see appendix) but these results were not included due to low average loss values and model stability.
 \par
 XGBoosted trees are an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
-The gradient boosting aspect of the ensemble is the use of a loss function used to create new decision trees that add to the ensemble.
-Each new tree created is done so using a loss function that is optimised iteratively to improve upon prior tree's predictive power.
+The gradient boosting aspect of the ensemble is the use of a loss function used to create new decision trees that add to the ensemble improving its predictive power.
+Each new tree created is done so using a loss function that is optimised iteratively to improve upon prior tree.
 The loss function can be any convex function, allowing gradient descent to traverse the loss space until, no improvements can be made via traversal.
@@
 Regularisation methods can also be incorporated to help prevent over fitting.
+This makes XGBoosted trees incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods.
 \subsection{Variable Importance}\label{sec:importance}
@@
 Sankey and Chord  diagrams are presented together; with Sankey diagrams showing the connection of a variable to its ten most important predictor variables.
-Chord diagrams were used alongside a Sankey diagram to show the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
+Chord diagrams were used alongside the Sankey diagrams to show the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
 Chord diagrams formed circles, with variables being connected through their relative importance.
@@
 
- \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{revenue.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
- \end{figure}
-
 The prediction of revenue performed similarly to operating cost achieving an $R^2$ of 0.7716 (with a standard deviation of 0.1525).
 The value of predictors' relative importance was then calculated through the number of nodes used within the XGBoost.
-Values for relative importance were then used to construct Sankey and Chord diagrams to compare the contribution of each variable to predicting revenue.
-\par
-In order of importance, predictors of revenue were fuel use(petrol 307 and diesel 144), yield (285), size (216) and water use (199).
+Values for relative importance were then used to construct Sankey and Chord diagrams to compare the contribution of each variable in predicting revenue.
+\par
+In order of importance the predictors of revenue were fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199).
 Here, the values in the brackets indicate the relative importance of each variable (see \ref{fig:revenue_sankey}).
@@
 
+\begin{figure}[htb]
+  \label{fig:revenue_sankey}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{revenue.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}
+ \end{figure}
+
 \subsection{Operating Costs}
@@
 
-\begin{figure}
+Comparatively to revenue, operating cost performed better with the XGBoosted regression ensemble achieving an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
+The relationships to operating cost through variable importance were found to be similar to that of revenue, with fuel, water, area and yield having the largest number relative importance (see figure \ref{fig:operating_costs_sankey}).
+A surprising difference was the change in relative importance of activities involving tractors passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (comparing Figure \ref{fig:region_sankey}).
+The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
+\par
+Again, region played a determining factor overall, contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
+It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
+\par
+% .
+Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
+\begin{figure}[htb]
   \resizebox{\textwidth}{!}{
@@
 
-Comparatively to revenue, operating cost performed better with the XGBoosted regression ensemble achieving an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
-The relationships to operating cost through variable importance were found to be similar to that of revenue, with fuel, water, area and yield having the largest number relative importance (see figure \ref{fig:operating_costs_sankey}).
-A surprising difference is that the most important operational consideration for operating cost is the use of fungicide, compared to revenue where slashing is the most important (comparing Figure \ref{fig:region_sankey}).
-The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
-\par
-Again, region played a determining factor overall, contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
-It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
-\par
-% .
-Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
-
 \subsection{Region}
@@
 
- \begin{figure}
+When considered overall, Region was a highly informative variable based on measures of importance for both operating cost and revenue.
+As noted above, Region was the third most important variable for determining revenue.
+The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
+These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot.
+\par
+As also noted above Region was also a key determinant of operating costs.
+Again Tasmania was the most important, followed by the Adelaide Hills.
+In contrast to revenue, both climates are considered cool and wet, and warmer drier regions such as the Barossa and Hunter Valley only contributed roughly half the same number of nodes to the ensemble.
+Based on further analysis of Regions (Figure \ref{fig:region_sankey}) the inclusion of slashing and fungicide spraying is the likely reason with fungal and weed pressure being greater in cooler wetter regions.
+\par
+The XGBoost ensemble, did not perform well when predicting operating costs or revenue with 56.82\% accuracy (50.58\% validation accuracy).
+The difference in accuracy is in part due to the large number of classes (58 regions).
+The ensemble had a great emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
+% Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
+\par
+Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
+The regions with the most samples performed the best.
+Bordering regions were routinely grouped together and misclassified as the same region.
+Two areas that suffered the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+
+\begin{figure}[htb]
   \resizebox{\textwidth}{!}{
@@
  \end{figure}
-
-When considered overall, Region was a highly informative variable based on measures of importance for both operating cost and revenue.
-As noted above, Region was the third most important variable for determining revenue.
-The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
-These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot.
-\par
-As also noted above Region was also a key determinant of operating costs.
-Again Tasmania was the most important, followed by the Adelaide Hills.
-In contrast to revenue, both climates are considered cool and wet, and warmer drier regions such as the Barossa and Hunter Valley only contributed roughly half the same number of nodes to the ensemble.
-Based on further analysis of Regions (Figure \ref{fig:region_sankey}) the inclusion of slashing and fungicide spraying is the likely reason with fungal and weed pressure being greater in cooler wetter regions.
-\par
-The XGBoost ensemble, did not perform well when predicting operating costs or revenue with 56.82\% accuracy (50.58\% validation accuracy).
-The difference in accuracy is in part due to the large number of classes (58 regions).
-The ensemble had a great emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
-% Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
-\par
-Many of the regions had significantly lower reporting rates, resulting in much poorer classification performance.
-The regions with the most samples performed the best.
-Bordering regions were routinely grouped together and misclassified as the same region.
-Two areas that suffered the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
-
 \section{Discussion}
@@
 It is also important to consider that some regions carry particular fame regarding the quality of their produce such as Tasmania, the Hunter Valley and Barossa Valley \citep{hallidayAustralianWineEncyclopedia2009}.
-This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
+This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
 \par
@@
 
-Separately revenue and operating cost did have a greater predictability than their counterpart profit.
+Separately revenue and operating cost did have a greater predictability than their counterpart profit (see appendix).
 The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
@@
   \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:profit}
  \end{figure}

### 2023-12-17 article.tex
KM edits done (except tables)

--- 2023-11-21 article.tex
@@
 
-% \begin{abstract}
-% \end{abstract}
+\begin{abstract}
+
+ The Australian wine industry is a major part of Australia's agricultural sector.
+As global demands change and new pressures on the industry present themselves, a more sustainable approach is needed.
+Through a nationwide data set, collected over ten years, we link key variables in determining vineyard operational costs and revenue through the use of XGBoost.
+We use a measure of relative importance to show the interrelated nature of these variables and the comparative influence they have on one another.
+We present these connections through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and their strong interrelatedness.
+Furthermore, we connect these variables to different wine regions, highlighting the complex influence of location on the use of different resources.
+The study provides valuable insights into the multifaceted dynamics governing operational costs and revenue, illustrating how factors such as water and fuel use impacts on operational costs and how different seasonal events affect these operations.
+
+\end{abstract}
 %%Graphical abstract`
@@
 Historically strong demands for Australian wine have helped to create a thriving industry.
-However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation has negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
+However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation have negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
 The 2021-2022 financial year alone saw a decline of 19\% in exports solely due to tariffs \citep{wineaustraliaNationalVintageReport2022}.
-A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale are key to making data-informed decisions to aid in increasing a nations agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
-Specifically within the Australian Wine and vine industry there is a need to further understand the driving relationships between resource use and economic output, where these relationships can lead to determining better and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
-\par
-An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering new insights into the driving economic forces of the Australian wine industry.
-This dataset allows insights into the economic outcome of vineyards through the incorporation of operating costs and grape revenue from grape sales within the data.
-We use this data to study these economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
+A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale is key to making data-informed decisions increase a nation's agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
+Specifically within the Australian wine and vine industry, there is a need to further understand the driving relationships between resource use and economic output, which can help to determine more cost effective and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
+\par
+An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering the potential for new insights into the driving economic forces of the Australian wine industry.
+A major part of the potential for insight within this dataset comes from the incorporation of operating costs and grape revenue from grape sales within the data.
+In this paper, we use data to study economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
 We further compare the relationships between different resources to address the extensive collinearity found within the data \citep{chenXGBoostScalableTree2016}.
-We adopt XGBoosted models for this analysis because they are able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables.
+We adopt  a popular, relatively new machine learning method, XGBoost, for this analysis because it is able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables.
 
@@
 % reported since 2018
-Data used in this analysis were obtained from Sustainable Winegrowing Australia.
-Australia's national wine industry sustainability program.
-The program aims to support grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data used in this analysis were obtained from Sustainable Winegrowing Australia (SWA), Australia's national wine industry sustainability program.
+SWA aims to support grape growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
 Data recorded by SWA are entered manually by winegrowers using a web based interface tool.
@@
 The data originally contained only two multiclass variables: year and region.
-Related binary variables, such as the use of river water and the use of dam water, were combined to create a single multiclass variable.
-This was done by first converting each combination that occurred into a unique category (such as river and damn water used, as opposed to the two separate categories prior).
-These variables were then one-hot-encoded, changing each variable class into a binary value, with one indicating the presence of the class and zero indicating its absence.
-Further details about classes and their frequency is available in the appendices.
-\par
-The variable region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+For this case study, related binary variables, such as the use of river water and the use of dam water, were combined to create multiclass variables such as water source.
+Further details about these variables, their classes and their frequency is available in the Appendix.
+\par
+The variable Region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
@@
 
-\subsection{XGBoosted Trees}
-XGBoosted (eXtreme Gradient Boosting) trees, described in more detail below (and further in the appendix), were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-XGBoosted trees are a boosted tree ensemble method that can be used to classify classes, or predict continuous response variables.
+\subsection{XGBoost}
+XGBoost (eXtreme Gradient Boosting), described in more detail below (and further in the Appendix), were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+XGBoost is a type of machine learning method that constructs and ensemble of decision trees to predict or estimate an output variable (the response) based on a number of input variables.
+The ensemble, can be used to classify classes or predict a continuous response, depending on the nature of the output variable.
 They were chosen for this analysis as the data contained a mixture of class and continuous variables.
-Moreover, XGBoosted trees are unaffected by multicollinearity, and offer high predictive performance for a wide variety of purposes \citep{chenXGBoostScalableTree2016}.
-\par
-XGBoosted models were constructed with operational cost and grape revenue as the predicted variables.
-The analyses were aimed at uncovering what factors influenced these variables and to what extent.
-As the purpose of the analysis was to identify relationships between variables and to show how they interact, an XGBoosted tree was created for each of the predictor variables as well.
-Trees for the predictor variables did not include operational cost or grape revenue as predictors.
-By creating an XGBoosted tree for each variable it meant that every variable would have a measure of its relative importance to every other variable (see Section \ref{sec:importance}).
+Moreover, XGBoost is unaffected by multicollinearity, and offer high predictive performance for a wide variety of purposes, and are capable of identifying and ranking variables and interactions in order of relative importance \citep{chenXGBoostScalableTree2016}.
+\par
+Four sets of analyses were conducted.
+In the first set, two XGBoost models were developed, with operational cost and grape revenue as the response variables.
+The analysis of operational cost and revenue included all variables in Table \ref{tab:vars}.
+The second set of analyses focused on identifying relationships between the input variables themselves, creating XGBoost models for each other variable so that every variable would have a measure of its relative importance to every other variable (see Section \ref{sec:importance}).
 Together these models were used to measure the interrelationships of the ten most important variables in determining operational cost and grape revenue using variable importance.
 These measures of relative importance were used to illustrate the highly interrelated nature of variables within vineyards.
-The interaction between variables was depicted through the use of Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variable and the response (see section \ref{sec:importance}).
-\par
-Due to constraints from the XGBoost library region could only be incorporated as a one-hot-encoded variable when used as a predictor.
-To better show what variables were related to region overall, another XGBoost tree was created with Region as the predicted value.
-The difference for this model was that relative variable importance would only be measured for each region specifically, as opposed to a variables overall importance in determining region.
-Separately profit (the difference between revenue and operational costs) and year was looked at in prior analyses (see appendix) but these results were not included due to low average loss values and model stability.
-\par
-XGBoosted trees are an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
-The gradient boosting aspect of the ensemble is the use of a loss function used to create new decision trees that add to the ensemble.
-Each new tree created is done so using a loss function that is optimised iteratively to improve upon prior tree's predictive power.
-The loss function can be any convex function, allowing gradient descent to traverse the loss space until, no improvements can be made via traversal.
+The interaction between variables was depicted through the use of Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variables and the response (see section \ref{sec:importance}).
+\par
+The third analysis was an XGBoost tree with Region as the response variable.
+The difference for this model was that relative variable importance for each variable would be measured for the overall importance in determining region, as opposed to a variable's connection to each region specifically.
+The fourth analysis focussed on profit (the difference between revenue and operational costs) and year, however these results were not included due to low average loss values and model stability (see Appendix).
+\par
+XGBoost is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
+The gradient boosting aspect of the ensemble is the use of a loss function to create new decision trees that add to the ensemble, improving its predictive power.
+The loss function is optimised iteratively to improve upon prior trees.
+The loss function can be any convex function, allowing gradient descent to traverse the loss space until no substantive improvements can be made via traversal.
 Because the loss function is only required to be convex, both classifiers and regressors can be used.
 Regularisation methods can also be incorporated to help prevent over fitting.
+This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods.
 \subsection{Variable Importance}\label{sec:importance}
-Due to XGBoost creating a large amount of decision trees, the interpretability of these models is obfuscated by the intricate relationships within complicated ensembles.
-A measure of variable importance was the technique used to highlight a variables influence within the ensemble.
-Variable importance can be measured in multiple ways; we used the frequency of a variable appearing as a node within the ensemble as a measure of its importance.
-This measure was chosen as it connected a variable to the minimisation of its associated objective function.
-The measure of a variable's importance within this study can then be interpreted as how often a variable was the optimal choice in reducing the loss function of the ensemble.
-Importantly, multiclass variables being one-hot-encoded (see Section \ref{Data}) are given an importance score for each individual class; for example, each specific region will have its own importance score.
+XGBoost creates a large number of decision trees in the ensemble, it is hard to directly interpret the model and the derived intricate relationship between the variables.
+Variable importance can be measured in multiple ways, in this paper we used the frequency of a variable appearing as a node within the ensemble as a measure of its importance.
+This measure can be interpreted as how often a variable was the optimal choice in reducing the loss function of the ensemble.
+Multiclass variables are given an importance score for each individual class;  for example, in the first set of analyses each specific region will have its own importance score, as will Year, Irrigation Type, etc (see Table \ref{tab:vars}).
 \par
@@
 Both Chord and Sankey diagrams illustrated variable importance through the size of the bands between two variables.
-The number at the end of a connection in a Sankey diagram indicates a variable's importance, or the number of times it appeared within the ensemble.
-Sankey and Chord  diagrams are presented together; with Sankey diagrams showing the connection of a variable to its ten most important predictor variables.
-Chord diagrams were used alongside a Sankey diagram to show the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
+The number at the end of a connection in a Sankey diagram indicates a variable's importance (the number of times it appeared within the ensemble).
+Sankey and Chord diagrams are presented together; with Sankey diagrams showing the connection of a variable to its ten most important predictor variables and chord diagrams showing the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
 Chord diagrams formed circles, with variables being connected through their relative importance.
-The importance values for the Chord diagrams were taking from the models of those individual variables, with the diagram being simplified to just the ten variables in the associated Sankey diagram, for readability's sake.
 
@@
 The predictive accuracy of each tree was assessed through a validation process.
-For each model the data was split into training data, which constituted 80\% of the original data.
-The remaining 20\% was used in testing and validation.
-Categorical data was stratified to conserve the same proportion of class occurrences between training, testing and validation data.
-For continuous variables 20\% was used as testing data and the models were validated using 10 repetitions of the sampling process (10-fold cross validation).
+For each model, a sample of 80\% of the data was used for training the model and the remaining 20\% was used for testing and validation.
+Categorical data were stratified to conserve the same proportion of class occurrences between the training, testing and validation data.
+The models were validated using 10 repetitions of the sampling process (10-fold cross validation).
 $R^2$ scores were used to determine the best regression models during validation.
-$R^2$ was used instead of RMSE to allow the comparison of  models with different units to each other when considering how well each model extrapolated to further data.
+For analyses with continuous responses $R^2$ was used instead of RMSE to allow the comparison of  models with different units to each other when considering how well each model extrapolated to further data.
 For binary and multiclass variables, validation was summarised through the accuracy, the proportion of true negatives and positives.
 \par
+As with most machine learning methods, a key component of the XGBoost model setup was the tuning of hyperparameters.
 The XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
@@
 
+The below sections present each of the analyses conducted within this study.
+This includes the three analyses for Revenue, Operational Cost and Region, with the fourth and final analysis on profit and yield presented in the appendix.
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
- \begin{figure}
+The predictive performance of the XGBoost model for revenue performed similarly to operating cost, for achieving an $R^2$ of 0.77 (with a standard deviation of 0.15). 
+\par
+The most important predictors of revenue were fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199).
+The values attached to each variable indicate the relative importance of the variable (number of times selected in the tree ensemble, see Section \ref{fig:revenue_sankey}).
+Overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
+The chord diagram (see Figure \ref{fig:revenue_sankey}) illustrates that vineyard area is also of high relative importance to other variables, especially slashing.
+The overall importance of Area to other variables is evident by its larger circumference within the chord diagram.
+% The strong relationship of vineyard area to revenue and its predictors  but also has a high relative importance to the other predictors, The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
+Area harvested is likely also an indicator of other variables such as slashing passes its strongest connection presented.
+
+\begin{figure}[htb]
+  \label{fig:revenue_sankey}
   \resizebox{\textwidth}{!}{
   \includegraphics{revenue.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}
  \end{figure}
-
-The prediction of revenue performed similarly to operating cost achieving an $R^2$ of 0.7716 (with a standard deviation of 0.1525).
-The value of predictors' relative importance was then calculated through the number of nodes used within the XGBoost.
-Values for relative importance were then used to construct Sankey and Chord diagrams to compare the contribution of each variable to predicting revenue.
-\par
-In order of importance, predictors of revenue were fuel use(petrol 307 and diesel 144), yield (285), size (216) and water use (199).
-Here, the values in the brackets indicate the relative importance of each variable (see \ref{fig:revenue_sankey}).
-Overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
-The chord diagram illustrates that vineyard area is also of high relative importance to other variables especially slashing.
-The overall importance of area to other variables is evident by its larger circumference within the chord diagram (see B in Figure \ref{fig:revenue_sankey}).
-% The strong relationship of vineyard area to revenue and its predictors  but also has a high relative importance to the other predictors, The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
-Area harvested is likely also an indicator of other variables such as slashing passes its strongest connection presented.
 
@@
 
-\begin{figure}
+Compared to revenue, the predictive performance of XGBoost model for operating cost was slightly better, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
+Similar to revenue, the most iomportant predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
+A surprising difference was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
+The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
+\par
+Again, Region played a determining factor overall, contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
+It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
+\par
+% .
+Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
+\begin{figure}[htb]
   \resizebox{\textwidth}{!}{
   \includegraphics{operating_costs.png}}
-  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Operating Costs using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Operating Costs using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The number at the end of each band in the diagram is that variable's importance.
@@
  \end{figure}
-
-Comparatively to revenue, operating cost performed better with the XGBoosted regression ensemble achieving an $R^2$ of 0.8025 (with a standard deviation of 0.1033).
-The relationships to operating cost through variable importance were found to be similar to that of revenue, with fuel, water, area and yield having the largest number relative importance (see figure \ref{fig:operating_costs_sankey}).
-A surprising difference is that the most important operational consideration for operating cost is the use of fungicide, compared to revenue where slashing is the most important (comparing Figure \ref{fig:region_sankey}).
-The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
-\par
-Again, region played a determining factor overall, contributing to 334 nodes within the ensemble making it the most important variable when considering all regions together.
-It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
-\par
-% .
-Again the surrogate model did not perform well achieving an $R^2$ of 0.0931 (with a standard deviation of 0.0197) but showed similarly to revenue an importance placed on fungicide spraying and size (see figure \ref{fig:operating_tree}).
 
@@
 
- \begin{figure}
+Region was a highly informative variable based on measures of importance for both operating cost and revenue.
+As noted above, Region was the third most important variable for determining revenue.
+The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
+These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot.
+\par
+As also noted, above Region was also a key determinant of operating costs.
+Again Tasmania was the most important, followed by the Adelaide Hills.
+In contrast, the regions of highest relative importance were warmer and drier, such as the Barossa and Hunter Valley.
+The higher relative importance of slashing and fungicide spraying is the likely due to fungal and weed pressure being greater in cooler wetter regions variables than in drier regions.
+\par
+The XGBoost ensemble for Region achieved an accuracy of 56.82\% (and 50.58\% validation accuracy).
+The difference in accuracy compared to the other models is in part due to the large number of classes (58 regions).
+The ensemble had a great emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
+% Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
+\par
+Substantially many of the regions had lower reporting rates, resulting in much poorer classification performance.
+The regions with the most samples performed the best likely due to the disparity in sample sizes.
+Bordering regions were routinely grouped together and misclassified as the same region.
+When scrutinising each class explicitly, the two areas that suffered the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+
+\begin{figure}[htb]
   \resizebox{\textwidth}{!}{
   \includegraphics{region.png}}
-  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Region using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Region using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The number at the end of each band in the diagram is that variable's importance.
@@
 First, the most important predictors of revenue and operating costs were fuel, yield and area.
-Secondly, area and fuel were highly interrelated to other variables (see Figure \ref{fig:operating_costs_sankey} and Figure \ref{fig:revenue_sankey}).
-Finally, the relative importance of predictor variables for region, differed from Revenue and operating costs, with water use being more prominent than yield.
+Secondly, area and fuel were highly interrelated to other variables (see Figure \ref{fig:operating_costs_sankey} and Figure \ref{fig:revenue_sankey}A).
+Finally, the relative importance of predictor variables for Region, differed from Revenue and Operating costs, with Water Use being more prominent than Yield.
 Region was also more prominent than illustrated in the Sankey diagrams due to the relative importance for operating cost and revenue being calculated for individual regions and not all regions together.
-In its entirety region contributed third most prominently in relative importance to revenue, and was of the most relative importance in determining operating costs.
-\par
-Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location; making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The association between yield and region is demonstrated by its rank of fourth-highest variable importance when determining region (see Figure \ref{fig:region_sankey}).
+In its entirety, was the third most important predictor of revenue and the most important predictor for operating costs, relative to the other variables consideration in the analyses.
+\par
+Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location, making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The association between yield and region is demonstrated by yield apparing as the fourth most important variable when determining region (see Figure \ref{fig:region_sankey}).
 % The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % TODO: This section is commented out due to Mardi's take:
@@
 It is also important to consider that some regions carry particular fame regarding the quality of their produce such as Tasmania, the Hunter Valley and Barossa Valley \citep{hallidayAustralianWineEncyclopedia2009}.
-This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile (and thus quality) \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
-\par
-In part some winegrowing strategies are restricted simply through access to water resources.
+This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
+\par
+In part, some winegrowing strategies are restricted simply through access to water resources.
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
@@
 The dominating factor of area likely played a large part in determining how costly a tractor pass would be, or in defining the ratio of water applied to the amount of vines.
-The node frequency was high for area but much lower in general across the other variables, which could indicate the need to be specific when attempting to determine the cause of a operational cost.
-Although it was attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), it is likely yet more complicated.
+The relative importance was high for area but much lower in general across the other variables, which could indicate the need to be specific when attempting to determine the cause of a operational cost.
+Although these analyses attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), in reality these relationships are likely even more complicated.
 An example of how interrelated operational costs can be, is
@@
 
-Separately revenue and operating cost did have a greater predictability than their counterpart profit.
+Separately, revenue and operating cost did have a greater predictability than their counterpart profit (see Appendix).
 The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
@@
 \begin{table}[]
-  \caption{Summary statistics of continuous variables used in XGBoosted models.}\label{tab:cont_vars}
+  \caption{Summary statistics of continuous variables used in XGBoost models.}\label{tab:cont_vars}
   \resizebox{\textwidth}{!}{\begin{tabular}{@{}lllllllll@{}}
@@
 \section{XGBoost}
-Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGboosted trees predict a value $y_i$ from the input $x_i$.
+Following Chen and Guestrin \citep{chenXGBoostScalableTree2016}, XGBoost predicted a value $y_i$ from the input $x_i$.
 The method of prediction is achieved through a tree ensemble model, using $K$ additive functions to predict the output.
@@
 
-As predictions are made using additive tree functions, XGboosted trees can be used for classification or regression.
+As predictions are made using additive tree functions, XGboost can be used for classification or regression.
 The difference between a prediction, $\phi(x_i)$, and actual variable, $f_{k}(x_i)$, is a differentiable convex loss function $l$.
@@
 
- The classification tree and XGBoosted ensemble performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
+ The classification tree and XGBoost performed similarly for classifying year with 35.20\% (6.28\% standard deviation) and 51.81\% (42.20\% validation accuracy) respectively.
 Electricity and the type of irrigation were highly influential within the classification tree.
@@
   \includegraphics{year.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:year_sankey}
@@
   \includegraphics{profit.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoosted trees as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:profit}
  \end{figure}

### 2023-12-18 article.tex
mardi edits

@@
 
- The Australian wine industry is a major part of Australia's agricultural sector.
-As global demands change and new pressures on the industry present themselves, a more sustainable approach is needed.
+ The Australian wine industry is a major contributor to Australia's agricultural sector and economy.
+As global market demands change and new pressures on the industry present themselves, a more sustainable approach is needed.
 Through a nationwide data set, collected over ten years, we link key variables in determining vineyard operational costs and revenue through the use of XGBoost.
@@
 Furthermore, we connect these variables to different wine regions, highlighting the complex influence of location on the use of different resources.
-The study provides valuable insights into the multifaceted dynamics governing operational costs and revenue, illustrating how factors such as water and fuel use impacts on operational costs and how different seasonal events affect these operations.
+The study provides valuable insights into the multifaceted dynamics governing operational costs and revenue, illustrating how factors such as water and fuel use impacts operational costs and how different seasonal events affect these operations.
 
@@
 The 2021-2022 financial year alone saw a decline of 19\% in exports solely due to tariffs \citep{wineaustraliaNationalVintageReport2022}.
-A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale is key to making data-informed decisions increase a nation's agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
+A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale is key to making data-informed decisions to increase a nation's agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
 Specifically within the Australian wine and vine industry, there is a need to further understand the driving relationships between resource use and economic output, which can help to determine more cost effective and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
 \par
-An unprecedented amount of data regarding the Australian winegrowing industry has been collected through Sustainable Winegrowing Australia, offering the potential for new insights into the driving economic forces of the Australian wine industry.
+An unprecedented amount of data regarding the Australian winegrowing industry has been collected through the Sustainable Winegrowing Australia program, offering the potential for new insights into the driving economic forces of the Australian wine industry.
 A major part of the potential for insight within this dataset comes from the incorporation of operating costs and grape revenue from grape sales within the data.
@@
 \par
-The variable Region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe different unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+The variable Region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
@@
 The Barossa Valley region and Tasmania were the two most important regions in relation to revenue; these two regions are considered to be some of the highest revenue per hectare regions in Australia \citep{wineaustraliaNationalVintageReport2022}.
-These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot.
-\par
-As also noted, above Region was also a key determinant of operating costs.
-Again Tasmania was the most important, followed by the Adelaide Hills.
-In contrast, the regions of highest relative importance were warmer and drier, such as the Barossa and Hunter Valley.
-The higher relative importance of slashing and fungicide spraying is the likely due to fungal and weed pressure being greater in cooler wetter regions variables than in drier regions.
+These two regions are also relative opposites in winegrowing climates with the Barossa having a warm and dry climate focussing on Shiraz grapes and Tasmania having a cool wet climate that favours Pinot/Chardonnay \citep{wineaustraliaNationalVintageReport2022}.
+\par
+As also noted above, Region was also a key determinant of operating costs.
+Tasmania had the highest relative importance, followed by the Adelaide Hills.
+In contrast, the regions of the highest relative importance were warmer and drier, such as the Barossa.
+The higher relative importance of fungicide spraying is the likely due to fungal pressure being greater in cooler wetter regions variables than in drier regions.
 \par
@@
 The difference in accuracy compared to the other models is in part due to the large number of classes (58 regions).
-The ensemble had a great emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
+The ensemble had an emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
 % Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
 \par
-Substantially many of the regions had lower reporting rates, resulting in much poorer classification performance.
+A number of regions had lower reporting rates, resulting in much poorer classification performance.
 The regions with the most samples performed the best likely due to the disparity in sample sizes.
 Bordering regions were routinely grouped together and misclassified as the same region.
-When scrutinising each class explicitly, the two areas that suffered the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
+When scrutinising each class explicitly, the two areas that effected the most from this were the Limestone Coast (cool coastal areas in South Australia) and the warmer inland regions along the Murray Darling.
 
@@
 Region was also more prominent than illustrated in the Sankey diagrams due to the relative importance for operating cost and revenue being calculated for individual regions and not all regions together.
-In its entirety, was the third most important predictor of revenue and the most important predictor for operating costs, relative to the other variables consideration in the analyses.
+In its entirety Region was the third most important predictor of revenue and the most important predictor for operating costs, relative to the other variables consideration in the analyses.
 \par
@@
 Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
-In general warmer regions are not associated with higher yields, but if a vineyard in a warmer region is sufficiently irrigated much higher yields can be achieved than in cooler regions \citep{campsGrapeHarvestYield2012}.
 It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of grapes; reflected through region using water use so prominently in the XGBoost ensemble.
-The link to water resources in defining regions is also an important consideration, as vineyards can leverage higher irrigation rates given more accessible water resources.
+The link to water resources in defining regions is also an important consideration, as vineyards can leverage higher irrigation rates if water resources are available.
 A further consideration in the link between revenue and region is that grape prices are set at a regional level by buyers \citep{wineaustraliaNationalVintageReport2022}.
@@
 \par
-In part, some winegrowing strategies are restricted simply through access to water resources.
+In part, yield is sometimes restricted simply through access to water resources.
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
-Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region because vineyards in regions without pressurised water will need to use more fuel to pressurise their irrigation systems.
+Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region because vineyards in regions without pressurised water will need to use more fuel or electricity to pressurise their irrigation systems.
 \par
@@
 The difference between vineyards that make profit and those that do not could be a multitude of factors including differences in farming practices not captured within this study.
-Some decisions leading to latent effects such as large scale soil deposition in extreme rain events can be caused by soil compaction due to overworking a vineyard \citep{capelloPermanentCoverSoil2020}.
-
 \iffalse
@@
 
-This study has provided valuable insights into the multifaceted dynamics governing operational costs and revenue.
+This study has provided valuable insights into the multifaceted dynamics governing operational costs and revenue in vineyards.
 The impact of different regions highlighted the complex interrelatedness of variables within a vineyard.

### 2023-12-18 article.tex
fixed table

@@
  \label{tab:vars}
- \caption{Summary of variables used in the analysis. The recorded column indicate the number of values that were either greater than zero or that were not missing.}\label{tab:summary}
+ \caption{Summary of variables used in the analysis.
+The recorded column indicate the number of values that were either greater than zero or that were not missing (see Appendix for more information).}\label{tab:summary}
  \small
- \begin{tabular}{@{}ccccl@{}}
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+  \begin{tabular}{@{}lccc@{}}
   \toprule
-  \textbf{Variable} & \textbf{Units} & \textbf{Recorded} & \textbf{Number of Classes} & \\ \midrule
-  Water Used & Mega Litres & 5846 & & \\
-  Diesel & Litres & 5585 & & \\
-  Biodiesel & Litres & 25 & & \\
-  LPG & Litres & 958 & & \\
-  Herbicide Spray & Times per year & 2026 & & \\
-  Year & Class & 6049 & 10 & \\
-  Disease & Class & 6049 & 2 & \\
-  Region & Class & 6049 & 58 & \\
-  Solar & Kilowatt Hours & 622 & & \\
-  Irrigation Type & Class & 6049 & 20 & \\
-  Petrol & Litres & 4309 & & \\
-  Slashing & Times per year & 2290 & & \\
-  Yield & Tonnes & 5935 & & \\
-  Irrigation Energy & Class & 6049 & 16 & \\
-  Area Harvested & Hectares & 6049 & & \\
-  Electricity & Kilowatt Hours & 1014 & & \\
-  Insecticide Spray & Times per year & 1092 & & \\
-  Fertiliser & KGs of Nitrogen & 795 & & \\
-  Fungicide Spray & Times per year & 2260 & & \\
-  Cover Crop & Class & 6049 & 32 & \\
-  Water Type & Class & 6049 & 39 & \\
-  Grape Revenue & AUD & 853 & & \\
-  Operating Costs & AUD & 853 & & \\ \bottomrule
+  \textbf{Variable} & \textbf{Units} & \textbf{Number of Classes} & \textbf{No. Records} \\ \midrule
+  Water Used & Mega Litres &  & 5846 \\
+  Diesel & Litres &  & 5585 \\
+  Biodiesel & Litres &  & 25 \\
+  LPG & Litres &  & 958 \\
+  Herbicide Spray & No. Times per year &  & 2026 \\
+  Year & Class & 10 & 6049 \\
+  Disease & Class & 2 & 6049 \\
+  Region & Class & 58 & 6049 \\
+  Solar & Kilowatt Hours &  & 622 \\
+  Irrigation Type & Class & 20 & 6049 \\
+  Petrol & Litres &  & 4309 \\
+  Slashing & No. Times per year &  & 2290 \\
+  Yield & Tonnes &  & 5935 \\
+  Irrigation Energy & Class & 16 & 6049 \\
+  Area Harvested & Hectares &  & 6049 \\
+  Electricity & Kilowatt Hours &  & 1014 \\
+  Insecticide Spray & No. Times per year &  & 1092 \\
+  Fertiliser & KGs of Nitrogen &  & 795 \\
+  Fungicide Spray & Times per year &  & 2260 \\
+  Cover Crop & Class & 32 & 6049 \\
+  Water Type/Source & Class & 39 & 6049 \\
+  Grape Revenue & AUD &  & 853 \\
+  Operating Costs & AUD &  & 853 \\ \bottomrule
   \end{tabular}
@@
 The data originally contained only two multiclass variables: year and region.
-For this case study, related binary variables, such as the use of river water and the use of dam water, were combined to create multiclass variables such as water source.
+For this case study, related binary variables, such as the use of river water and the use of dam water, were combined to create multiclass variables such as water source/type.
 Further details about these variables, their classes and their frequency is available in the Appendix.
@@
   \includegraphics{operating_costs.png}}
-  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Operating Costs using XGBoost as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side, A,  depicts the 10 most relative  important variables in predicting Operating Costs using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The number at the end of each band in the diagram is that variable's importance.
@@
   \includegraphics{region.png}}
-  \caption{The left-hand side, A,  depicts the 10 most important variables in predicting Region using XGBoost as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side, A,  depicts the 10 most relative important variables in predicting Region using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The number at the end of each band in the diagram is that variable's importance.
@@
   \includegraphics{year.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting Year using XGBoost as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side depicts the 10 most relative important variables in predicting Year using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:year_sankey}
@@
   \includegraphics{profit.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
+  \caption{The left-hand side depicts the 10 most relative important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
 The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:profit}

### 2023-12-18 article.tex
fixed table issue and built pdf

@@
 \begin{table}[] 
- \label{tab:vars}
 \caption{Summary of variables used in the analysis.
-The recorded column indicate the number of values that were either greater than zero or that were not missing (see Appendix for more information).}\label{tab:summary}
+The recorded column indicate the number of values that were either greater than zero or that were not missing (see Appendix for more information).} \label{tab:vars}
  \small
@@
 % \usepackage{booktabs}
-\begin{table}[]
+% \begin{table}[]
   \begin{tabular}{@{}lccc@{}}

### 2023-12-18 Cover Letter/main.tex
cover letter

\documentclass{article}

\usepackage{charter} 
\usepackage{ragged2e}
\usepackage{graphicx}
\usepackage[
	a4paper,    % Paper size
	top=1in,    % Top margin
	bottom=1in, % Bottom margin
	left=1in,   % Left margin
	right=1in,  % Right margin
	%showframe  % Uncomment to show frames around the margins for debugging purposes
]{geometry}

\setlength{\parindent}{0pt}     % Paragraph indentation 
\setlength{\parskip}{1em}       % Vertical space between paragraphs

\usepackage{graphicx}       % Required for including images

\usepackage{fancyhdr}       % Required for customizing headers and footers

\fancypagestyle{firstpage}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{0pt} % No header rule
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\fancypagestyle{subsequentpages}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{1pt} % Header rule thickness
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\AtBeginDocument{\thispagestyle{firstpage}} % Use the first page headers/footers style on the first page
\pagestyle{subsequentpages} % Use the subsequent pages headers/footers style on subsequent pages

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	FIRST PAGE HEADER
%----------------------------------------------------------------------------------------

\includegraphics[scale=0.5]{logo.jpg} % Logo

\vspace{-1em} % Pull the rule closer to the logo

\rule{\linewidth}{1pt} % Horizontal rule

\bigskip\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\hfill
\begin{tabular}{l @{}}
	$29^{th}$ October, 2023\\ % Date
	\textbf{Bryce Ronald William Polley $|$ PhD Candidate}\\
	Queensland University of Technology \\
	School of Mathematical Sciences\\~\\
	Address: Level 8 Y Block, 2 George Street, \\
	Brisbane, Queensland, Australia, 4000\\
	Phone: +61 400 310 105 \\
	Email: bryce.polley@awri.com.au
\end{tabular}

\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING
%----------------------------------------------------------------------------------------

\begin{tabular}{@{} l}
	\textbf{Editors,}\\
	Journal of Economic Geography\\
\end{tabular}

\bigskip % Vertical whitespace

Dear Editor,
\\~\\
\parskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\centering \textbf{Submission of Manuscript for Consideration by \\Journal of Economic Geography}\\
\bigskip % Vertical whitespace
\justifying
\noindent

\bigskip \noindent% Vertical whitespace
We wish to submit a new manuscript entitled  ''\textit{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards}'' for your consideration by the Journal of Economic Geography.
\newline \noindent
In this manuscript we explore the contributing factors to Australian vineyards' operational costs and revenue. We provide insights into the interrelatedness of the factors contributing to operational cost and revenue. We further explore the relationships between these factors and regional differences; showing how different regional effects impact on vineyard operations and outcomes. We use a novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia. We construct and analyse several models of the multifaceted, multi-resource nature of vineyard management.
\newline \noindent
We believe that this manuscript is appropriate for publication by the Journal of Economic  Geography as it demonstrates the regional differences in operational costs and revenue of Australian vineyards. We further link these outcomes to contributing factors to rank the relative importance of different vineyard activities and resource use; illustrating the economic outcomes of Australian vineyards and linking it to different regional factors. Thank you for your consideration of this manuscript.
\bigskip 
\newline \noindent% Vertical whitespace
\includegraphics[scale=0.1]{signature_block.png}
\newline \noindent
Bryce Polley (corresponding author) on behalf of all coauthors: Kate Helmsted, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
\newline \noindent
\textit{Corresponding Author}, \textit{Queensland University of Technology}

\end{document}

### 2024-04-16 Cover Letter/main.tex
update for JAFR submission

@@
 \begin{tabular}{l @{}}
-	$29^{th}$ October, 2023\\ % Date
-	\textbf{Bryce Ronald William Polley $|$ PhD Candidate}\\
+	$16^{th}$ April, 2024\\ % Date
+	\textbf{Bryce Ronald William Boyd $|$ PhD Candidate}\\
 	Queensland University of Technology \\
@@
 \begin{tabular}{@{} l}
-	\textbf{Editors,}\\
-	Journal of Economic Geography\\
+	\textbf{Dr. John P. Brooks and Professor Yangchao Luo,}\\ Editors-in-Chief \\ Journal of Agriculture and Food Research
 \end{tabular}
@@
 
-Dear Editor,
+Dear Editors,
 \\~\\
@@
 %----------------------------------------------------------------------------------------
-\centering \textbf{Submission of Manuscript for Consideration by \\Journal of Economic Geography}\\
+\centering \textbf{Submission of Manuscript for Consideration by \\Journal of Agriculture and Food Research}\\
 \bigskip % Vertical whitespace
@@
 \bigskip \noindent% Vertical whitespace
-We wish to submit a new manuscript entitled  ''\textit{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards}'' for your consideration by the Journal of Economic Geography.
+We wish to submit a new manuscript entitled  ''\textit{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards}'' for your consideration by the Journal of Agriculture and Food Research.
 \newline \noindent
 In this manuscript we explore the contributing factors to Australian vineyards' operational costs and revenue.
-We provide insights into the interrelatedness of the factors contributing to operational cost and revenue.
-We further explore the relationships between these factors and regional differences; showing how different regional effects impact on vineyard operations and outcomes.
+We provide insights into the interrelatedness of the factors contributing to operational cost and revenue, we further explore the relationships between these factors and regional differences; showing how different regional effects impact on vineyard operations and outcomes.
 We use a novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia.
@@
 \newline \noindent
-We believe that this manuscript is appropriate for publication by the Journal of Economic  Geography as it demonstrates the regional differences in operational costs and revenue of Australian vineyards.
-We further link these outcomes to contributing factors to rank the relative importance of different vineyard activities and resource use; illustrating the economic outcomes of Australian vineyards and linking it to different regional factors.
+We believe that this manuscript is appropriate for publication by the Journal of Agriculture and Food Research as it demonstrates the regional differences in operational costs and revenue of Australian vineyards and further link these outcomes to contributing factors to rank the relative importance of different vineyard activities and resource use.
+We illustrate the economic outcomes of Australian vineyards and link these outcomes to different regional factors.
 Thank you for your consideration of this manuscript.
@@
 \newline \noindent
-Bryce Polley (corresponding author) on behalf of all coauthors: Kate Helmsted, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
+Bryce Boyd (corresponding author) on behalf of all coauthors: Kate Helmstedt, Mardi Longbottom, Kerrie Mengersen\\
 \newline \noindent

### 2024-04-16 article.tex
fixed reference error in appendix
update for JAFR submission

@@
 \documentclass[review,12pt,authoryear]{elsarticle}
+
+\let\today\relax
+\makeatletter
+\def\ps@pprintTitle{%
+    \let\@oddhead\@empty
+    \let\@evenhead\@empty
+    \def\@oddfoot{\footnotesize\itshape
+         {Submitted preprint to Journal of Agricultural and Food Research} \hfill\today}%
+    \let\@evenfoot\@oddfoot
+    }
+\makeatother
 
@@
 \begin{frontmatter}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % country={}}
-\author[label1,label2,label3]{Author}
-\date{02/08/2023}
+% \author[label1,label2,label3]{Author}
+% \date{02/08/2023}
 
@@
 % \end{highlights}
+% \fntext[fn1]{This is the first author footnote.}
 \end{frontmatter}
@@
 % \usepackage{booktabs}
-\begin{table}[]
+\begin{table}[htb]
   \caption{Summary statistics of continuous variables used in XGBoost models.}\label{tab:cont_vars}
@@
   \subsection{Irrigation Energy Type}
-Below, Table \ref{irig_energy} shows the different types of energy used to power vineyards and their frequency.
+Below, Table \ref{tab:irig_energy} shows the different types of energy used to power vineyards and their frequency.
 
@@
 
- \begin{figure}
+ \begin{figure}[htb]
   \resizebox{\textwidth}{!}{
@@
 
- \begin{figure}
+\begin{figure}[htb]
   \resizebox{\textwidth}{!}{
@@
 
-\begin{figure} 
+ \begin{figure}[htb]
   \resizebox{\textwidth}{!}{
@@
 
-Predictions of profit perfomed poorly compared to operating cost and revenue with an average $R^2$ of 0.2535 and standard deviation of 0.3126.
+Predictions of profit performed poorly compared to operating cost and revenue with an average $R^2$ of 0.2535 and standard deviation of 0.3126.
 With the large standard deviation being indicative of how unstable the models created were.

### 2024-09-03 data_prep.py
zeroing fractional tractor passes updated to drop

@@ -6,9 +6,9 @@ df = pd.read_feather("df.feather")
 
 #  We disallow errors in the dataset:
 # The first is tractor passes as fractions (the field is meant to be integer only)
-df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
-df.loc[df["insecticide_spraying_number_of_times_passes_per_year"]<1, "insecticide_spraying_number_of_times_passes_per_year"] = 0
-df.loc[df["herbicide_spraying_number_of_times_passes_per_year"]<1, "herbicide_spraying_number_of_times_passes_per_year"] = 0
+df.drop(df[df["herbicide_spraying_number_of_times_passes_per_year"]<1].index)
+df.drop(df[df["insecticide_spraying_number_of_times_passes_per_year"]<1].index)
+df.drop(df[df["herbicide_spraying_number_of_times_passes_per_year"]<1].index)
 for col in df.loc[:, df.dtypes == np.float64].columns:
     df.loc[df[col]<0, col] = np.nan
 df[df.loc[:, df.dtypes == np.float64].columns].min()

### 2024-09-05 article.tex
refined leading question - in prep for adressing reviewers edits
first done

@@
 % Regional differences in Australian winegrowing (Quality)?
+
+% The leading question:
+% How do different regional factors and resource utilisation impact the operational costs and revenue in Australian vineyards?
+% 
+% How can machine learning models like XGBoost help in understanding these relationships to optimise vineyard management strategies?
 
@@
 For binary and multiclass variables, validation was summarised through the accuracy, the proportion of true negatives and positives.
-\par
-As with most machine learning methods, a key component of the XGBoost model setup was the tuning of hyperparameters.
+
+\subsection{Hyperparameters}
+
+As part of the utilising the XGBoost model the hyperparameters of the model were tuned.
 The XGBoost library incorporates regularisation techniques built into the software to mitigate over-fitting and enhance model generalisation.
 This allowed us to utilise cross validated grid search functions when selecting for better performing hyperparameters.
+This method required three distinct types of metrics to be used for the three types of variables incorporated into the analysis (multiclass, binary and numeric).
+For consistency the metrics utilised by the grid search were aligned with the error functions used when training the model on those variables.
 The performance measure for model selection was root-mean-square error for continuous variables.
-The receiver operator characteristic's area under the curve was used for category variables \citep{hanley1982meaning}.
+The receiver operator characteristic's area under the curve was used for binary variables \citep{hanley1982meaning}.
 Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.

### 2024-09-05 article.tex
fixed comment 7 (prior one was 6)

@@
 SWA aims to support grape growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by SWA are entered manually by winegrowers using a web based interface tool.
-A total of 6049 observations were collected from 2012/2013 to 2021/2022 financial years, with each observation comprising 23 variables reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
+Data recorded by SWA are entered voluntarily by winegrowers, manually using a web based interface tool.
+There are a total of 6049 observations were collected from 2012/2013 to 2021/2022 financial years.
+Variables recorded by winegrowers are optional.
+Each vineyard record consists of observations comprising 23 variables which reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
+The data was restricted to vineyards that at minimum recorded vineyard size.
 \par
@@
 \normalsize
+Due to the optional and manual recording of data, steps were taken to remove potentially erronous entries.
+This process first involved discussions with SWA highlighting possible entry errors.
+At the end of a season any suspect entries, such as a missing entry of fuel but a vineyard having recorded the use of tractors or the other most common being the use of incorrect units (litres instead of megalitres), would warrant calling individual vineyards to clarify values and logic within the data.
+Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through the calling growers for clarification or the removal of an observation from the dataset.
 The data originally contained only two multiclass variables: year and region.
-For this case study, related binary variables, such as the use of river water and the use of dam water, were combined to create multiclass variables such as water source/type.
+The data originally contained granular readings for similar variables that related to their source, such as water use being divided into river , dam, and pressurised water sources.
+For this study, these related numeric variables, which included water use, irrigation, and electricity were combined to create multiclass variables, which would illustrated the mixture of types utilised by the grower.
+This decision was driven by a majority of vineyards utilising one source or a second as a backup, with an overwhelming percentage of water/electricity/irrigation prevailing within a single vineyard.
 Further details about these variables, their classes and their frequency is available in the Appendix.
@@
 
-Compared to revenue, the predictive performance of XGBoost model for operating cost was slightly better, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
-Similar to revenue, the most iomportant predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
+Compared to revenue, the predictive performance of the XGBoost model for operating cost was slightly better, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
+Similar to revenue, the most important predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
 A surprising difference was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
+The connection between spraying and operational costs is intuitive in that it utilises equipment and resources to do.
+However, it is surprising that although spraying is considered important, that area not harvested due to disease is not, even though this would be a direct cause.
+The lack of importance on disease directly could be because of a low amount disease reported in the dataset (137 vineyards).
+The use of spraying is also unfortunately not part of the data, and could be in response to a variety of factors such as other vineyards within the region having disease or preventative sprays.
 The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
@@
 A larger number of tractor passes used as a preventative measure for occurrences such as disease may incur higher operational costs but could be critical in preventing long term losses.
-Although the models demonstrated a good predictive fit (via large $R^2$ values), the ability to predict operational costs is limited by the variables incorporated in the analysis.
+Although the models demonstrated a good predictive fit, the ability to predict operational costs is limited by the variables incorporated in the analysis.
 Other factors such as erosion and soil health are also influenced by tractor use and would contribute to these operational costs but are difficult to measure and were not available as part of the data \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
-Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions.
-Without fully capturing more granular activities for example the specific reasons for fuel use, it is difficult to determine what decisions specifically influence the operational costs.
+The data collection process being voluntary and part of a sustainable program also limitted the ability to compare what happened between those who had to abandon crops due to disease, pests or other catastrophes such as fire, in part due to a lack of incentive to record as part of the SWA program.
+Furthermore, no comparison can be made between those that have chosen to mothball as a response to predicted outcomes, or external presssures due to them not being part of the data.
+Although this dataset contained vineyards that suffered partial losses due to disease, these limitations offer an avenue for further study that could benefit decision processes and variable relevance regarding mothballing, crop loss and external pressures.
+Without fully capturing more granular activities, for example the specific of tractor operastions and their differing fuel consumptions, it is difficult to determine what decisions specifically influence the operational costs.
+Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions when considering external risks such as disease.
 %  Electricity in particular is used predominantly for irrigation. 

### 2024-09-05 article.tex
no. 8 edit

@@
 It was surprising that electricity, slashing and spraying passes were not more prominent in operating costs due to the intrinsic nature as an agricultural expense.
+However, a consideration for a bias within the dataset may be explanatory towards to the lack of these factors contributing to expenses, with the dataset being derived from vineyards actively participating within a sustainability program.
+\par
+Although region plays a strong role, it was found that factors that can be considered regionally important were not always rated high in relative importance.
+This was noted especially when considering oversimplified models, such as the obvious partitioning of river water as the primary source of water to identify vineyards located in the riverland (with 456 of 472 vineyards in the riverland utilising riverwater).
+The occurence of factors dominated by particular practices also washes out the ability to determine the effects of practices and their benefits that are widely used by this sample of vineyards.
+A keen example is the use of techniques such as cover crops and their prevalance within the sample, where just over 85\% (5272) of vineyards utilised some form of cover crop such as grassing and only just under 4\% (225) used only bare soil (with the remaining 552 utilising a combination).
+The high percentage of vineyards using this type of sustainable practice means that its effect is will not be prominent within the model, and can only show what practices would further improve those already implementing these techniques, and how they are connected to these operating costs.
+A strength of utilising XGBoost in this context is that, a subset of particular interest can be leveraged to focus in on the combination of factors that would contribute to the specific concieved scenario.
 \par
@@
 
-Predictions of profit performed poorly compared to operating cost and revenue with an average $R^2$ of 0.2535 and standard deviation of 0.3126.
+Predictions of profit performed poorly compared to operating cost and profit with an average $R^2$ of 0.2535 and standard deviation of 0.3126.
 With the large standard deviation being indicative of how unstable the models created were.

### 2024-09-30 article.tex
article edits backup (switch computer)

@@
 We adopt  a popular, relatively new machine learning method, XGBoost, for this analysis because it is able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables.
+
+TODO: expand the litearture review and premise of the article..
+There is growing interest from the wine industry in environmentally sustainable practices in the wine industry, and several recent reviews have called for more data-driven studies to show the economic benefits of such practices.
+While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to demonstrate these benefits across different regions and climates (Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).
+
+The success stories in implementing sustainable strategies in vineyards varies in region, climate, grape variety, and scale (SWA, 2022).
+The variation shows the amount of difference that can exist between vineyards and their unique challenges, as well as the differences in their viable options.
+Pressures such as salinity, biosecurity, and water availability change from site to site.
+The complexity of these environmental interactions has led to the bridging of different research areas such as geology, pedology, hydrology and plant sciences; where the on the ground decisions become governed by these complex physical forces as well as by external pressures such as international market demands, disease pressures and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
+Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018).
+Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
+The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
+The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
+The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65% (Capello et al., 2020).
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
+A coverage of only 30% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
+The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
+An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
+Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
+It is important that the interaction of events such as erosion have on other considerations such as soil health.
+Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
+For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
+Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
+The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
+When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
+Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
+Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
+The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
+Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
 
@@
 At the end of a season any suspect entries, such as a missing entry of fuel but a vineyard having recorded the use of tractors or the other most common being the use of incorrect units (litres instead of megalitres), would warrant calling individual vineyards to clarify values and logic within the data.
-Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through the calling growers for clarification or the removal of an observation from the dataset.
-The data originally contained only two multiclass variables: year and region.
-The data originally contained granular readings for similar variables that related to their source, such as water use being divided into river , dam, and pressurised water sources.
-For this study, these related numeric variables, which included water use, irrigation, and electricity were combined to create multiclass variables, which would illustrated the mixture of types utilised by the grower.
+Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through the calling growers for clarification or the removal of an observation from the dataset after consulting viticulturalists as to its plausability.
+Several entries were removed from the study due to impossible recordings such as no fuel of any kind being consumed when tractors were used, or no electricity being used when the irrigation system was electric (most of these were able to be clarified through contacting vineyard owners).
+\par
+Due to the nature of XGBoost the data was not needed to be transformed for the algorithms consumption with the exception of multiclass variables being converted to one-hot-encoded variables, with the data originally containing only two multiclass variables: year and region.
+Only two augmenetations were done to the data for this analysis beyond data cleaning, the first was to convert the continuous variables (fuel, electricity, water, yield, etc) into ratio's of a vineyard to reflect the consumption of resources and output yield per hectare to compare the outputs to non-transformed models.
+
+The second transfomration, was done to add clarity in interpretting the relative importance scores.
+Where, variables that related to the source of a resource, such as water use being divided into river, dam, and pressurised water sources were summed into total electricity/water variables.
+The source of these (such as river, dam, pressurised water) were converted into binary variables that reflected the presence of a source being used.
+Other potentially multi-class variables such as irrigation and cover crops were also converted to reflect whether a grower simply used these types of systems as opposed to the dict amount of area covered by them.
 This decision was driven by a majority of vineyards utilising one source or a second as a backup, with an overwhelming percentage of water/electricity/irrigation prevailing within a single vineyard.
+The use of a binarisation also meant that importance measures would be better understood as they forced ensemble to partition by presence or absence of a type of system as opposed to a potentially overly specific number.
+This further helps to relate relative importance values for these variables directly to the importance of using a system over another.
+This approach was also used after finding little difference in model accuracy between whether these variables were reported as proportion of a type used (i.e the percentage of land covered by drip irrigation), direct units of a type (i.e M\\L river water used) or as a binary presence/absence.
 Further details about these variables, their classes and their frequency is available in the Appendix.
@@
 \par
-
+TODO: talk about external climatic data
 \subsection{XGBoost}
-XGBoost (eXtreme Gradient Boosting), described in more detail below (and further in the Appendix), were created using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-XGBoost is a type of machine learning method that constructs and ensemble of decision trees to predict or estimate an output variable (the response) based on a number of input variables.
-The ensemble, can be used to classify classes or predict a continuous response, depending on the nature of the output variable.
-They were chosen for this analysis as the data contained a mixture of class and continuous variables.
-Moreover, XGBoost is unaffected by multicollinearity, and offer high predictive performance for a wide variety of purposes, and are capable of identifying and ranking variables and interactions in order of relative importance \citep{chenXGBoostScalableTree2016}.
-\par
-Four sets of analyses were conducted.
-In the first set, two XGBoost models were developed, with operational cost and grape revenue as the response variables.
-The analysis of operational cost and revenue included all variables in Table \ref{tab:vars}.
-The second set of analyses focused on identifying relationships between the input variables themselves, creating XGBoost models for each other variable so that every variable would have a measure of its relative importance to every other variable (see Section \ref{sec:importance}).
-Together these models were used to measure the interrelationships of the ten most important variables in determining operational cost and grape revenue using variable importance.
-These measures of relative importance were used to illustrate the highly interrelated nature of variables within vineyards.
-The interaction between variables was depicted through the use of Sankey and Chord diagrams; with variable importance measures being used to show the strength of connection between the respective predictor variables and the response (see section \ref{sec:importance}).
-\par
-The third analysis was an XGBoost tree with Region as the response variable.
-The difference for this model was that relative variable importance for each variable would be measured for the overall importance in determining region, as opposed to a variable's connection to each region specifically.
-The fourth analysis focussed on profit (the difference between revenue and operational costs) and year, however these results were not included due to low average loss values and model stability (see Appendix).
-\par
-XGBoost is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
+XGBoost (eXtreme Gradient Boosting) is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
 The gradient boosting aspect of the ensemble is the use of a loss function to create new decision trees that add to the ensemble, improving its predictive power.
-The loss function is optimised iteratively to improve upon prior trees.
-The loss function can be any convex function, allowing gradient descent to traverse the loss space until no substantive improvements can be made via traversal.
+The loss function is optimised iteratively to improve upon prior trees (where the loss function can be any convex function), allowing gradient descent to traverse the loss space until no substantive improvements can be made (detail pertaining to the algorithm is described further in the Appendix).
 Because the loss function is only required to be convex, both classifiers and regressors can be used.
 Regularisation methods can also be incorporated to help prevent over fitting.
-This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods.
+This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods (TODO: ref).
+XGBoost analyses were conducted using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+It is a method that is widely used within agriculture for yield prediction (TODO: ref), but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables (TODO: ref).
+We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables where it is known to perform well with mixed types of predictor domains (TODO: ref).
+Furthermore we choose XGBoost as it has a good performance in predictions whislt allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models.
+\par
+The ability to classify and predict continuous response variables and categoric variables alongside one another was also a consideration in the use of XGBoost, as both were contained in the data.
+XGBoost was also used due to its ability to handle sparse data, which was present within this dataset due to the voluntary nature of data entry, with many fields being left blank during data collection.
+Tree based methods also do not require data to be transformed prior to analyses; this consideration was taken into account so that specific partitions of values could be evaluated more easily and understood within original units of the data (TODO: ref).
+A further consideration in its use was the level of interpretability offered through measures of 'relative importance' allowing for the ability to identify and rank variables and interactions by contribution to predictions \citep{chenXGBoostScalableTree2016}.
+\par
+An XGBoost model was trained for each variable so that every variable's relative importance could be caculated.
+This process was done three time using three iterations of data (three models for each variable).
+The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region (see TODO: insert a table for this) and the final were trained on data with continuous variables  transformed to be expressed as a ratio of vineyard area.
+The final dataset that consisted of ratios also included the extra regional data (but not in ratio form).
+\par
+\subsection{Sankey and Chord Diagrams}
+Originally created by Sankey to depict different pressures in steam engines (TODO: ref) we leverage Sankey diagrams to illustrate the different impact or 'pressure' each variable has on one another through the use of measurements of variable importance.
+Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+Sankey diagrams (depicted on the left as section Asection A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular represntation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
+\par
+Both Chord and Sankey diagrams illustrate variable importance through the size of the bands between two variables.
+The number at the end of a connection in the diagrams indicates a variabless importance (the number of times it appeared within the ensemble).
+
 \subsection{Variable Importance}\label{sec:importance}
+
 XGBoost creates a large number of decision trees in the ensemble, it is hard to directly interpret the model and the derived intricate relationship between the variables.
@@
 Multiclass variables are given an importance score for each individual class;  for example, in the first set of analyses each specific region will have its own importance score, as will Year, Irrigation Type, etc (see Table \ref{tab:vars}).
-\par
-The Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
-Both Chord and Sankey diagrams illustrated variable importance through the size of the bands between two variables.
-The number at the end of a connection in a Sankey diagram indicates a variable's importance (the number of times it appeared within the ensemble).
-Sankey and Chord diagrams are presented together; with Sankey diagrams showing the connection of a variable to its ten most important predictor variables and chord diagrams showing the interconnectedness of the ten most prominent variables within its associated Sankey diagram.
-Chord diagrams formed circles, with variables being connected through their relative importance.
 
@@
 
-The below sections present each of the analyses conducted within this study.
+% The below sections present each of the analyses conducted within this study.
 This includes the three analyses for Revenue, Operational Cost and Region, with the fourth and final analysis on profit and yield presented in the appendix.
@@
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
-Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region because vineyards in regions without pressurised water will need to use more fuel or electricity to pressurise their irrigation systems.
+Similarly, the connection between region and fuel use is likely an indicator of the level of infrastructure within the region due to vineyards in regions without pressurised water needing to use fuel or electricity to pressurise their irrigation systems.
+Although infrastructure between regions, especially further from cities is likely to vary, fuel price itself has little variation across regional Australia.
+It is reported by the Australian Competition and Consumer Commission that during the period of this data, that regional fuel prices tended to be higher (+5.4c/litre) and more stable than urban prices due to their primary driver being international market trends~\citep{aipFactsPricesRegional2019}.
+The importance between fuel and other variables is a complicated interaction.
+The size, number of blocks, types and age of equipment will contribute to the efficiency of its use and the amount required across a site.
+It is likely that larger operations will generaly gain from economies of scale but also risk further incurring costs from the need to redeploy equipment.
+A further connection between region and fuel is the possible requirement of more specialty equipment, either due to regional practices differing or physical requirements such as greater inclines.
+However, the style of management will also greatly contribute to how efficient both fuel and water are used, which is difficult to account for through the use of a metric.
 \par

### 2024-09-30 analysis.py

@@ -321,7 +321,7 @@ def train_model_multi(df: pd.DataFrame, y_name: str):
 
 ################################################################################
 # Setup
-df = pd.read_csv("dfb.csv")
+df = pd.read_csv("dfa.csv")
 
 cols = ["tonnes_grapes_harvested"
     , "area_harvested"
@@ -487,6 +487,19 @@ cat_cols = [ # These are binary
     , "giregion"
 ]
 
+# We can remove the region data if we want to - this was mostly for looking at predicting region...
+region_cols = [
+    "elev_Median"
+    , "elev_Max"
+    , "elev_Min ‡"
+    , "elev_Range"
+    , "rainfall_mean"
+    , "temperature_mean"
+    , "extreme_cold"
+    , "extreme_heat"
+    , "aridity"
+]
+
 # We comment out these unless we really need to redo every single variable. As they are not compared to the target variables at the end!
 # 
 # cols += cat_cols
@@ -516,8 +529,11 @@ cat_cols = [ # These are binary
 # train_model_reg(df[df["profit"].notnull()][cols+["profit"]]
 #         , "profit")
 
-train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
-                , "total_operating_costs")
+# train_model_reg(df[df["total_operating_costs"]!=0][cols+["total_operating_costs"]]
+#                 , "total_operating_costs")
 
-train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
-                , "total_grape_revenue")
+# train_model_reg(df[df["total_grape_revenue"]!=0][cols+["total_grape_revenue"]]
+#                 , "total_grape_revenue")
+
+train_model_multi(df[cols].drop(columns=region_cols)
+                , "giregion")

### 2024-09-30 data_prep.py

@@ -16,12 +16,17 @@ df = pd.read_feather("df.feather")
 df.drop(df[df["herbicide_spraying_number_of_times_passes_per_year"]<1].index)
 df.drop(df[df["insecticide_spraying_number_of_times_passes_per_year"]<1].index)
 df.drop(df[df["herbicide_spraying_number_of_times_passes_per_year"]<1].index)
+#  We drop vineyards with no area >100m2:
+df = df.drop(df[df["area_harvested"] <= 0.01].index)
+df = df.drop(df[df["vineyard_area"] <= 0.01].index)
+
 for col in df.loc[:, df.dtypes == np.float64].columns:
     df.loc[df[col]<0, col] = np.nan
 df[df.loc[:, df.dtypes == np.float64].columns].min()
 
-#  We drop vineyards with incorrect area:
+# Some floats are so small we drop them if they are null just in case
 df = df.drop(df[df["area_harvested"].isnull()].index)
+
 
 # we get the indices of those that made profit
 indices = df[
@@ -184,7 +189,7 @@ for col in df_floats.columns:
 
 area = df_floats["area_harvested"]
 # we multiply by 1000 to stop fractions when dividing.
-df_floats = df_floats.div(df_floats["area_harvested"], axis=0)/1000
+df_floats = df_floats.div(df_floats["area_harvested"], axis=0)
 df_floats["area_harvested"] = area
 df_floats = df_floats.replace({np.inf: np.nan, -np.inf: np.nan})
 add_regional_data(

### 2024-10-01 article.tex
01/10

@@
 
- The Australian wine industry is a major contributor to Australia's agricultural sector and economy.
-As global market demands change and new pressures on the industry present themselves, a more sustainable approach is needed.
 Through a nationwide data set, collected over ten years, we link key variables in determining vineyard operational costs and revenue through the use of XGBoost.
-We use a measure of relative importance to show the interrelated nature of these variables and the comparative influence they have on one another.
-We present these connections through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and their strong interrelatedness.
+We further use a measure of relative importance to show the interrelated nature of these variables and the comparative influence they have on one another.
+Connections betwen variables is presented through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and their strong interrelatedness.
 Furthermore, we connect these variables to different wine regions, highlighting the complex influence of location on the use of different resources.
-The study provides valuable insights into the multifaceted dynamics governing operational costs and revenue, illustrating how factors such as water and fuel use impacts operational costs and how different seasonal events affect these operations.
+With the Australian wine industry being a major contributor to Australia's agricultural sector and economy, this study provides valuable insights into the multifaceted dynamics governing operational costs and revenue, illustrating how factors such as water and fuel use impact operational costs and how different seasonal events affect these operations.
 
@@
 % NB Avoid titles along the lines of: “Effects of ...”, “The role of ...”, etc.  Be specific about the effect and its significance so that your reader knows what is on offer.
-%  TODO:
+%
 
@@
 
-Historically strong demands for Australian wine have helped to create a thriving industry.
+Strong demands for Australian wine have historically helped to create a thriving industry.
 However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation have negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
@@
 A greater understanding of the different underlying conditions leading to improved performance in agricultural productivity and sustainability at scale is key to making data-informed decisions to increase a nation's agricultural sustainability \citep{oecdInnovationProductivitySustainability2019}.
-Specifically within the Australian wine and vine industry, there is a need to further understand the driving relationships between resource use and economic output, which can help to determine more cost effective and efficient methods and develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
-\par
-An unprecedented amount of data regarding the Australian winegrowing industry has been collected through the Sustainable Winegrowing Australia program, offering the potential for new insights into the driving economic forces of the Australian wine industry.
-A major part of the potential for insight within this dataset comes from the incorporation of operating costs and grape revenue from grape sales within the data.
-In this paper, we use data to study economic outcomes and their statistical relationships to vineyards' utilisation of the resources.
-We further compare the relationships between different resources to address the extensive collinearity found within the data \citep{chenXGBoostScalableTree2016}.
-We adopt  a popular, relatively new machine learning method, XGBoost, for this analysis because it is able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables.
-
-TODO: expand the litearture review and premise of the article..
-There is growing interest from the wine industry in environmentally sustainable practices in the wine industry, and several recent reviews have called for more data-driven studies to show the economic benefits of such practices.
-While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to demonstrate these benefits across different regions and climates (Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).
-
-The success stories in implementing sustainable strategies in vineyards varies in region, climate, grape variety, and scale (SWA, 2022).
-The variation shows the amount of difference that can exist between vineyards and their unique challenges, as well as the differences in their viable options.
-Pressures such as salinity, biosecurity, and water availability change from site to site.
-The complexity of these environmental interactions has led to the bridging of different research areas such as geology, pedology, hydrology and plant sciences; where the on the ground decisions become governed by these complex physical forces as well as by external pressures such as international market demands, disease pressures and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
+Specifically within the Australian wine and vine industry, there is a need to further understand the driving relationships between resource use and economic output, which can help to determine more cost effective, efficient methods, and to develop benchmarks with local growers \citep{lukemanciniUnderstandingAustralianWine2020}.
+\par
+The potential for new insights into the driving economic forces of the Australian wine industry have manifested in an unprecedented amount of data regarding Australian winegrowing, collected through the Sustainable Winegrowing Australia program.
+A major part of the insights within this dataset come from the incorporation of operating costs and grape revenue, with environmental and sustainable data.
+We seek to address both the predictability of operating costs and revenue within the australian winegrowing context and examine their major driving factors to observe linked trends in sustainable practices.
+As part of this we examine the data to study economic outcomes and their statistical relationships to vineyards' utilisation of resources.
+We adopt a popular, relatively new machine learning technique, XGBoost, for this analysis because it is able to overcome multicollinearity as well as highlight the level of importance that predictor variables have on response variables \citep{chenXGBoostScalableTree2016}.
+\par
+This study is further driven by recent reveiews calling for data-driven studies to show the economic benefits of sustainable practices within the wine industry, specifically winegrowing.
+While there is evidence to suggest that environmentally sustainable pracitces can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to numerically demonstrate these benefits across different regions and climates #TODO:(Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).
+Futhermore, many different sustainable approaches exist but are often studied in isolation or are limited in their geographical and climatic conditions, restricting their generalisability.
+We embrace the variation that exists between vineyards and their unique challenges across Australia.
+Where, vineyard decisions on-the-ground are governed by complex physical forces of a regions' resources, climate, soil and geology, as well as by external pressures such as international market demands, disease and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
 Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018).
-Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
-The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
-The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
-The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65% (Capello et al., 2020).
-However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
-A coverage of only 30% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
-The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
-An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
-Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
-It is important that the interaction of events such as erosion have on other considerations such as soil health.
-Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
-For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
-Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
-The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
-When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
-Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
-Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
-The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
-Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
-
 \section{Methods}
@@
 SWA aims to support grape growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by SWA are entered voluntarily by winegrowers, manually using a web based interface tool.
+Data recorded by SWA are entered voluntarily by winegrowers, manually using a web based interface.
 There are a total of 6049 observations were collected from 2012/2013 to 2021/2022 financial years.
 Variables recorded by winegrowers are optional.
-Each vineyard record consists of observations comprising 23 variables which reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
+Each vineyard record consists of observations comprising 23 variables reflecting a vineyard's state for the given year (see Table \ref{tab:vars}).
 The data was restricted to vineyards that at minimum recorded vineyard size.
@@
 This process first involved discussions with SWA highlighting possible entry errors.
-At the end of a season any suspect entries, such as a missing entry of fuel but a vineyard having recorded the use of tractors or the other most common being the use of incorrect units (litres instead of megalitres), would warrant calling individual vineyards to clarify values and logic within the data.
-Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through the calling growers for clarification or the removal of an observation from the dataset after consulting viticulturalists as to its plausability.
-Several entries were removed from the study due to impossible recordings such as no fuel of any kind being consumed when tractors were used, or no electricity being used when the irrigation system was electric (most of these were able to be clarified through contacting vineyard owners).
-\par
-Due to the nature of XGBoost the data was not needed to be transformed for the algorithms consumption with the exception of multiclass variables being converted to one-hot-encoded variables, with the data originally containing only two multiclass variables: year and region.
-Only two augmenetations were done to the data for this analysis beyond data cleaning, the first was to convert the continuous variables (fuel, electricity, water, yield, etc) into ratio's of a vineyard to reflect the consumption of resources and output yield per hectare to compare the outputs to non-transformed models.
-
-The second transfomration, was done to add clarity in interpretting the relative importance scores.
-Where, variables that related to the source of a resource, such as water use being divided into river, dam, and pressurised water sources were summed into total electricity/water variables.
-The source of these (such as river, dam, pressurised water) were converted into binary variables that reflected the presence of a source being used.
-Other potentially multi-class variables such as irrigation and cover crops were also converted to reflect whether a grower simply used these types of systems as opposed to the dict amount of area covered by them.
-This decision was driven by a majority of vineyards utilising one source or a second as a backup, with an overwhelming percentage of water/electricity/irrigation prevailing within a single vineyard.
-The use of a binarisation also meant that importance measures would be better understood as they forced ensemble to partition by presence or absence of a type of system as opposed to a potentially overly specific number.
-This further helps to relate relative importance values for these variables directly to the importance of using a system over another.
-This approach was also used after finding little difference in model accuracy between whether these variables were reported as proportion of a type used (i.e the percentage of land covered by drip irrigation), direct units of a type (i.e M\\L river water used) or as a binary presence/absence.
+At the end of a season any suspect entries, such as a missing fuel-use in a vineyard that recorded the use of tractors, would warrant calling individual vineyards to clarify values and logic within the data.
+Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through calling growers for clarification or the removal of an observation due to its unlikely plausability, with most cases suspected of being incorrect units (commonmly litres instead of megalitres of water used) but were not able to be verified.
+\par
+Due to the nature of XGBoost (eXtreme Gradient Boosting) data was not required to be scaled before used.
+However some transformations were done, such as multiclass variables being converted to one-hot-encoded variables (the only multiclass variables originally included were year and region).
+Variables relating to resource consumption, such as water-use were originally divided into whether it was river, dam, or pressurised water but were summed into total water/electricity/diesel/petrol.
+The source of these variables (such as river, dam, pressurised water) were then converted into binary variables that reflected the presence of a source being used.
+Other variables that reflected types of operations used such as irrigation-type and cover-crops were also converted to reflect whether a grower simply used these types of systems as opposed to the original format being the specific hectares covered by them.
+This decision due to a majority of vineyards utilising one source or a second as a backup, with an overwhelming percentage of water/electricity/irrigation prevailing within a single vineyard.
+The use of a binarisation also meant that importance measures would be better understood as they forced the ensemble to partition by presence or absence of a type of system as opposed to an overly specific number of hectares.
+This further helped to utilise relative importance for these variables directly to the act of using one system over another.
+This approach was compare to using the original variables but little difference in model accuracies was found between variables reported as proportion of a type used (i.e the percentage of land covered by drip irrigation), direct units of a type (i.e M\\L river water used) or as a binary presence/absence.
 Further details about these variables, their classes and their frequency is available in the Appendix.
-\par
+\subsection{Additional regional data}
 The variable Region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Each region is explicitly defined under the Wine Australia Corporation Act of 1980 \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-\par
-TODO: talk about external climatic data
+The regional data also expanded to include summary information regarding regions' climate and terrain in the form of minimum, maximum, median and range of elevation.
+And, temperature and rainfall means alongside extreme heat and cold days; as well as a regions' aridity index.
+This data was sourced using https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-0238.2010.00100.x
+And
+https://www.wineaustralia.com/growing-making/environment-and-climate/climate-atlas
 \subsection{XGBoost}
-XGBoost (eXtreme Gradient Boosting) is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
+XGBoost is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
 The gradient boosting aspect of the ensemble is the use of a loss function to create new decision trees that add to the ensemble, improving its predictive power.
-The loss function is optimised iteratively to improve upon prior trees (where the loss function can be any convex function), allowing gradient descent to traverse the loss space until no substantive improvements can be made (detail pertaining to the algorithm is described further in the Appendix).
+The loss function is optimised iteratively to improve upon prior trees (where the loss function can be any convex function), allowing gradient descent to traverse the loss space until no substantive improvements can be made (further detail pertaining to the algorithm is described in the Appendix).
 Because the loss function is only required to be convex, both classifiers and regressors can be used.
@@
 It is a method that is widely used within agriculture for yield prediction (TODO: ref), but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables (TODO: ref).
-We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables where it is known to perform well with mixed types of predictor domains (TODO: ref).
+We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables as XGBoost is known to perform well with mixed types of predictor domains (TODO: ref).
 Furthermore we choose XGBoost as it has a good performance in predictions whislt allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models.
@@
 An XGBoost model was trained for each variable so that every variable's relative importance could be caculated.
-This process was done three time using three iterations of data (three models for each variable).
+This process was done three times using three iterations of data (three models for each variable).
 The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region (see TODO: insert a table for this) and the final were trained on data with continuous variables transformed to be expressed as a ratio of vineyard area.
 The final dataset that consisted of ratios also included the extra regional data (but not in ratio form).
-\par
+
 \subsection{Sankey and Chord Diagrams}
@@
 Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
-Sankey diagrams (depicted on the left as section Asection A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular represntation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
-\par
+Sankey diagrams (depicted on the left as section A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular represntation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
 Both Chord and Sankey diagrams illustrate variable importance through the size of the bands between two variables.
@@
 \subsection{Variable Importance}\label{sec:importance}
-
 XGBoost creates a large number of decision trees in the ensemble, it is hard to directly interpret the model and the derived intricate relationship between the variables.
@@
 \subsection{Validation}
-
 The predictive accuracy of each tree was assessed through a validation process.
@@
 The receiver operator characteristic's area under the curve was used for binary variables \citep{hanley1982meaning}.
-Multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
+And, multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
 
@@
 
+\subsection{Yield}
+
 \subsection{Region}
@@
 \section{Discussion}
+% Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
+The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
+The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
+The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65% (Capello et al., 2020).
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
+A coverage of only 30% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
+% The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
+An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
+Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
+It is important that the interaction of events such as erosion have on other considerations such as soil health.
+% Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
+For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
+Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
+The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
+% When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
+Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
+Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
+The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
+Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
 

### 2024-10-02 article_with notes.tex
discussion

--- 2024-10-01 article.tex
@@
 Strong demands for Australian wine have historically helped to create a thriving industry.
-However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation have negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,wineaustraliaAustralianWineProduction2021}.
+However, recent pressures brought on by a loss of tourism and labour due to the COVID-19 pandemic, the global freight crisis, war in Europe, tariffs and rising inflation have negatively affected the industry's outlook \citep{wineaustraliaNationalVintageReport2021,australiaAustralianWineProduction2021}.
 The 2021-2022 financial year alone saw a decline of 19\% in exports solely due to tariffs \citep{wineaustraliaNationalVintageReport2022}.
@@
 This study is further driven by recent reveiews calling for data-driven studies to show the economic benefits of sustainable practices within the wine industry, specifically winegrowing.
-While there is evidence to suggest that environmentally sustainable pracitces can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to numerically demonstrate these benefits across different regions and climates #TODO:(Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).
+While there is evidence to suggest that environmentally sustainable pracitces can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to numerically demonstrate these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023, laurentReviewIssuesMethods2021}.
 Futhermore, many different sustainable approaches exist but are often studied in isolation or are limited in their geographical and climatic conditions, restricting their generalisability.
 We embrace the variation that exists between vineyards and their unique challenges across Australia.
-Where, vineyard decisions on-the-ground are governed by complex physical forces of a regions' resources, climate, soil and geology, as well as by external pressures such as international market demands, disease and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
-Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018).
+Where, vineyard decisions on-the-ground are governed by complex physical forces of a regions' resources, climate, soil and geology, as well as by external pressures such as international market demands, disease and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,goodwinijeriepRegulatedDeficitIrrigation1992,hallSeasonTemporalVariation2011,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNonDestructiveSensingMethods2018}.
 \section{Methods}
-% TODO:
 %  From the code comments~ We might want to show the relationship between revenue and operational cost! then we have an idea of how these two are related.
@@
 Regularisation methods can also be incorporated to help prevent over fitting.
-This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods (TODO: ref).
+This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods \citep{kistenExplainableArtificialIntelligence2024}.
+\par
 XGBoost analyses were conducted using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-It is a method that is widely used within agriculture for yield prediction (TODO: ref), but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables (TODO: ref).
-We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables as XGBoost is known to perform well with mixed types of predictor domains (TODO: ref).
-Furthermore we choose XGBoost as it has a good performance in predictions whislt allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models.
+It is a method that is widely used within agriculture for yield prediction \citep{d.mariadassExtremeGradientBoosting2022,liGlobalDetrendingSignificantly2024,raviCropYieldPrediction2020}, but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables \citep{ZHANG2023104045}.
+We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables as XGBoost is known to perform well with mixed types of predictor domains \citep{doi:10.1080/15481603.2024.2349341,ZHANG2023104045}.
+Furthermore we choose XGBoost as it has a good performance in predictions whislt allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models \citep{heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}.
 \par
@@
 XGBoost was also used due to its ability to handle sparse data, which was present within this dataset due to the voluntary nature of data entry, with many fields being left blank during data collection.
-Tree based methods also do not require data to be transformed prior to analyses; this consideration was taken into account so that specific partitions of values could be evaluated more easily and understood within original units of the data (TODO: ref).
+Tree based methods also do not require data to be transformed prior to analyses; this consideration was taken into account so that specific partitions of values could be evaluated more easily and understood within original units of the data \citep{d.mariadassExtremeGradientBoosting2022}.
 A further consideration in its use was the level of interpretability offered through measures of 'relative importance' allowing for the ability to identify and rank variables and interactions by contribution to predictions \citep{chenXGBoostScalableTree2016}.
@@
 This process was done three times using three iterations of data (three models for each variable).
-The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region (see TODO: insert a table for this) and the final were trained on data with continuous variables transformed to be expressed as a ratio of vineyard area.
+The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region and the final were trained on data with continuous variables transformed to be expressed as a ratio of vineyard area.
 The final dataset that consisted of ratios also included the extra regional data (but not in ratio form).
@@
 \subsection{Sankey and Chord Diagrams}
-Originally created by Sankey to depict different pressures in steam engines (TODO: ref) we leverage Sankey diagrams to illustrate the different impact or 'pressure' each variable has on one another through the use of measurements of variable importance.
-Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
+Originally created by Sankey to depict different pressures in steam engines \citep{7536189} we leverage Sankey diagrams to illustrate the different impact or 'pressure' each variable has on one another through the use of measurements of variable importance.
+Sankey and Chord diagrams were constructed using the Holoviews python library \citep{rudigerHolovizHoloviewsVersion2020}.
 Sankey diagrams (depicted on the left as section A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular represntation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
@@
 The performance measure for model selection was root-mean-square error for continuous variables.
-The receiver operator characteristic's area under the curve was used for binary variables \citep{hanley1982meaning}.
+The receiver operator characteristic's area under the curve was used for binary variables \citep{hanleyMeaningUseArea1982}.
 And, multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
@@
 
-The predictive performance of the XGBoost model for revenue performed similarly to operating cost, for achieving an $R^2$ of 0.77 (with a standard deviation of 0.15). 
-\par
-The most important predictors of revenue were fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199).
-The values attached to each variable indicate the relative importance of the variable (number of times selected in the tree ensemble, see Section \ref{fig:revenue_sankey}).
-Overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
-The chord diagram (see Figure \ref{fig:revenue_sankey}) illustrates that vineyard area is also of high relative importance to other variables, especially slashing.
-The overall importance of Area to other variables is evident by its larger circumference within the chord diagram.
+There was little difference in predictive power between the inclusion of climatic and regional data and the use of region alone.
+With the use of extra regional data achieving an $R^2$ of 0.76 (with a standard deviation of 0.13), and when not using the extra regional data achieving an $R^2$ of 0.77 (with a standard deviation of 0.15).
+The higher number of variables included as part of the regional data likely resulting in a lower variance in predictive power when including the extra regional data.
+\par
+The most notable difference was in the importance of the predictors, where elevation was a had a high relative importance of 8, only surpassed by diesel and yield, with 9 and 45 relative importance respectively (TODO: see figure).
+The lower variance is also reflected in the lower amount of nodes (or partitions) required when leveraging the extra regional parameters, with a reduction in hundreds of splits between the two models.
+Without the extra regional parameters it was found that fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199) held the highest relative importance (TODO: see figure).
+\par
+Even without the extra regional parameters, overall region contributed to 234 nodes in the ensemble making it collectively the third most important variable.
+This places equal importance within both models on region however the nature of regions contribution is more generalised using elevation when included as a parameter.
+This alone does not necessarily make elevation a direct contributing factor but links regions that of similar terrain together.
+The relevance of this was noted when reviewing regional missclassifications, where neighbouring regions were often misclassified as each other.
+The extra numerical parameters that can be partitioned likely gives the algorithm a greater ability to partition these regions together using fewer nodes.
 % The strong relationship of vineyard area to revenue and its predictors  but also has a high relative importance to the other predictors, The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
@@
 %  \end{figure}
-
-Compared to revenue, the predictive performance of the XGBoost model for operating cost was slightly better, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
-Similar to revenue, the most important predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
-A surprising difference was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
-The connection between spraying and operational costs is intuitive in that it utilises equipment and resources to do.
-However, it is surprising that although spraying is considered important, that area not harvested due to disease is not, even though this would be a direct cause.
-The lack of importance on disease directly could be because of a low amount disease reported in the dataset (137 vineyards).
-The use of spraying is also unfortunately not part of the data, and could be in response to a variety of factors such as other vineyards within the region having disease or preventative sprays.
+Compared to revenue, the predictive performance of the XGBoost model for operating cost was slightly better when not using the extra regional parameters, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
+Similarly, when predictin without extra regional data the most important predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
+\par
+A major difference was also in the poorer performance of the model with extra regional parameters.
+Although achieving a similar $R^2$ of 0.78 and a standard deviation of 0.12 the difference lay in an outlier model recording an $R^2$ of 0.08.
+This divergence likely being due to over generalising using elevation to match neighbouring regions.
+The removal of this radical leaves the inclusion of extra regional parameters outperforming the models without them, however it overlooks the major pitfall in the ease of misattributing factors and causality when predicting these variables.
+\par
+A surprising difference between operating cost and revenue was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
+This difference was only found when not including the extra regional parameters.
+The model including extra regional parameters reflected an identical hierarchy of importance to its revenue counter part (TODO: see figs).
+\par
+The connection between spraying and operational costs is intuitive in that it utilises both the expense of equipment and resources.
+However, it is surprising that although spraying is considered important when extra regional paramters were not included, 'area not harvested due to disease' was not, even though disease would be a direct cause.
+The lack of importance on disease directly could be due to a low amount reported in the dataset (137 vineyards).
+The reason for spraying was also unfortunately not part of the data, and could be in response to a variety of factors such as other vineyards within the region having disease or preventative sprays.
 The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
@@
 However, a consideration for a bias within the dataset may be explanatory towards to the lack of these factors contributing to expenses, with the dataset being derived from vineyards actively participating within a sustainability program.
-\par
-Although region plays a strong role, it was found that factors that can be considered regionally important were not always rated high in relative importance.
-This was noted especially when considering oversimplified models, such as the obvious partitioning of river water as the primary source of water to identify vineyards located in the riverland (with 456 of 472 vineyards in the riverland utilising riverwater).
-The occurence of factors dominated by particular practices also washes out the ability to determine the effects of practices and their benefits that are widely used by this sample of vineyards.
-A keen example is the use of techniques such as cover crops and their prevalance within the sample, where just over 85\% (5272) of vineyards utilised some form of cover crop such as grassing and only just under 4\% (225) used only bare soil (with the remaining 552 utilising a combination).
-The high percentage of vineyards using this type of sustainable practice means that its effect is will not be prominent within the model, and can only show what practices would further improve those already implementing these techniques, and how they are connected to these operating costs.
-A strength of utilising XGBoost in this context is that, a subset of particular interest can be leveraged to focus in on the combination of factors that would contribute to the specific concieved scenario.
 \par
@@
 
-\subsection{Yield}
-
 \subsection{Region}
@@
  \end{figure}
+
+\subsection{Results per hectare}
+Both operating cost and revenue were predicted as ratios of area (revenue and operating cost per vineyard hectare).
+In both cases it was found that the models performed poorly with operating costs recording an $R^2$ of 0.24 (with a standard deviation of 0.15) and revenue an $R^2$ of 0.32 (with a standard deviation of 0.14).
+
 \section{Discussion}
-% Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
-The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
-The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
-The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65% (Capello et al., 2020).
-However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
-A coverage of only 30% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
-% The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
-An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
-Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
-It is important that the interaction of events such as erosion have on other considerations such as soil health.
-% Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
-For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
+
+The significance of yield and region is demonstrated by their continual apparance as important variables when predicting operating costs and revenue.
+Several physical parameters such as climate, geology and soil are predetermined by a vineyard's location, making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The relationships between vineyard resource use, operations and geographical properties is a complex one, illustrated through the highly interelated properties demonstrated in the chord diagrams.
+The difference in resources available between regions is well illustrated when considering the relative importance that water use had in predicting region.
+The difference in readily available resources between different regions is also easily demosntrated when observing the partitioning of river water as the primary source of water to identify vineyards located in the riverland (with 456 of 472 vineyards in the riverland utilising riverwater).
+\par 
+The availability of resources and geographical features of a region are significant but are also a potentially determining factor in the types of operational decisions made, and thus the reasoning as to variation in revenue and operating costs.
+This can be seen in the addition of extra regional paramters that add greater context to understanding the why or cause of operating costs and revenue.
+The effect of different operational consideration can be reflected in higher costs likely incurred in regions of greater variations in slope requiring more specialised equipment, or reflected in the types of resources available in some regions.
+The specifics of a vineyards' site are of incredible importance when determining what causes higher operational costs or revenue.
+Although it is useful to be able to predict and compare regions and their revenue and operational costs, a greater nuace to help understand the why behind these decisions, would help in specifically guiding operational decisions.
+AN example of this can is whether the reduction of tillage operations through optimising tractor efficiency would be useful and how to optimise tractor use for a specific operation.
+This example is chosen as while this practice is undertaken to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
+The interrelatedness of this decisions is far reaching as increase in tractor use can cause soil compaction which has been shown to further increase water runoff (Capello et al., 2020).
+With runoff itself being a significant factor during extreme rain events which can lead to large scale soil deposition, creating further erosion and removing topsoil and having wide spread effects for a vineyard.
+\par
+% Differences between regions also likely extends further to the types of employed methods and the reason for using them.
+% An important and ongoing consideration for vineyards that changes with location is soil, which interacts with every other practice in various ways at different time scales.
+The inclusion of soil specifically within this model was not possible but For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
 Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
-The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
 % When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
@@
 Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
-
-This study explored the relationships between vineyard resource use, operations and geographical properties to revenue and operating costs.
-The analysis was based on a large national study of 6049 samples collected over ten years.
-Three main findings were identified.
-First, the most important predictors of revenue and operating costs were fuel, yield and area.
-Secondly, area and fuel were highly interrelated to other variables (see Figure \ref{fig:operating_costs_sankey} and Figure \ref{fig:revenue_sankey}A).
-Finally, the relative importance of predictor variables for Region, differed from Revenue and Operating costs, with Water Use being more prominent than Yield.
-Region was also more prominent than illustrated in the Sankey diagrams due to the relative importance for operating cost and revenue being calculated for individual regions and not all regions together.
-In its entirety Region was the third most important predictor of revenue and the most important predictor for operating costs, relative to the other variables consideration in the analyses.
-\par
-Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location, making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The association between yield and region is demonstrated by yield apparing as the fourth most important variable when determining region (see Figure \ref{fig:region_sankey}).
-% The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % TODO: This section is commented out due to Mardi's take:
-%  Maybe but not sure this assumption holds. It may be an artefact of the members being corporate in larger holdings. The larger percentage of blocks in the riverland are small.
+\par
+Further to the consideration of including specific operations is hindered in this model due to the sample being derived specifically from vineyards already within a sustainable program.
+Making the sample inherintly biased towards the use of sustainable practices.
+A keen example is the use of techniques such as cover crops.
+Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
+The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
+The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
+The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65\% (Capello et al., 2020).
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
+A coverage of only 30\% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
+The presence of cover crops within the sample is reflects this bias, where just over 85\% (5272) of vineyards utilised some form of cover crop such as grassing and only just under 4\% (225) used only bare soil (with the remaining 552 utilising a combination).
+The high percentage of vineyards using this type of sustainable practice means that its effect is will not be prominent within the model, and can only show what practices would further improve those already implementing these techniques, and how they are connected to these operating costs.
+A strength of utilising XGBoost in this context is that, a subset of particular interest can be leveraged to focus in on the combination of factors that would contribute to the specific concieved scenario.
 \par
 Warmer regions are known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
-Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
+Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{bottingCanopyArchitectureImplicationsShiraz1996}.
 It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of grapes; reflected through region using water use so prominently in the XGBoost ensemble.
@@
 It is difficult to extrapolate the specific influence of location on a vineyard's outcomes due to the broad and varying definition of a region.
-Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
+Utilising the Geographical Indicator regions defined by Wine Australia \citep{australiaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
 However, as buyers set prices at regional levels, it is still important to consider this factor.

### 2024-10-03 article.tex
final spell check and addition of figures

@@
 
-% Below is Elseviers requirements - they are similar to most articles and a good point of reference when writting scientific articles or analyses in general.
-%1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
-% The main text should be approximately 6,000 words in length, but it should not exceed 8,000 words (excluding abstract, references, tables, figures, and appendices).
-%A maximum of 250 words abstract and up to 10 displayed items (figures and tables) is allowed. A full-length article should include an Introduction,
-%Materials and methods, Results, Discussion, Conclusions, and References, which can be accompanied by Supplementary material.
-
 \begin{document}
@@
 \begin{frontmatter}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%       Start Matter      %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-
-%% \title{Title\tnoteref{label1}}
+
 \title{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards.}
-
-% Variable importance in determining vineyard operational costs and region using statistical machine learning.
-% How are region and operational costs related?
-
-% Regional differences in Australian winegrowing (Quality)?
-
-% The leading question:
-% How do different regional factors and resource utilisation impact the operational costs and revenue in Australian vineyards?
-% 
-% How can machine learning models like XGBoost help in understanding these relationships to optimise vineyard management strategies?
-
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%      addressline={}, 
-%%      city={},
-%%      postcode={}, 
-%%      state={},
-%%      country={}}
-%% \fntext[label3]{}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%       addressline={},
-%%       city={},
-%%       postcode={},
-%%       state={},
-%%       country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%       addressline={},
-%%       city={},
-%%       postcode={},
-%%       state={},
-%%       country={}}
-%\affiliation[label1]{organization={QUT},
-% addressline={},
-% city={},
-% postcode={},
-% state={QLD},
-% country={}}
-%\affiliation[label2]{organization={AWRI},
-% addressline={},
-% city={},
-% postcode={},
-% state={SA},
-% country={}}
-%\affiliation[label3]{organization={Food Agility CRC},
-% addressline={},
-% city={},
-% postcode={},
-% state={Vic},
-% country={}}
-% \author[label1,label2,label3]{Author}
-% \date{02/08/2023}
 
@@
 We further use a measure of relative importance to show the interrelated nature of these variables and the comparative influence they have on one another.
-Connections betwen variables is presented through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and their strong interrelatedness.
+Connections between variables is presented through the use of Sankey and Chord diagrams to show the important predictors of revenue and operating costs and their strong interrelatedness.
 Furthermore, we connect these variables to different wine regions, highlighting the complex influence of location on the use of different resources.
@@
 \end{abstract}
-%%Graphical abstract`
-%\begin{graphicalabstract}
- % \includegraphics{graphical_abstract.jpeg}
-%\end{graphicalabstract}'
-
-%\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-%Keyword one \sep{} keyword two
-%% PACS codes here, in the form: \PACS code \sep code
-%\PACS{} 0000 \sep{} 1111
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-%\MSC{} 0000 \sep{} 1111
-%\end{keyword}
-%%Research highlights
-% \begin{highlights}
-%  \item Highlight 1
-%  \item Highlight 2
-%  \item Highlight 3
-%  \item Highlight 4
-% \end{highlights}
-% \fntext[fn1]{This is the first author footnote.}
+
 \end{frontmatter}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        main text       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%
-% Working title
-%
-
-% Description
-% NB Avoid titles along the lines of: “Effects of ...”, “The role of ...”, etc.  Be specific about the effect and its significance so that your reader knows what is on offer.
-%
-
-% For example, rather than write a title like “The effect of factor X on astrophysical properties of green cheese.”,
-
-% be specific about the effect and write something more like ...
-
-% “Factor X halves the lunar thermal diffusivity of  green cheese”.
-
-% You will usually find it easier to write an effective title if you make your title a sentence.
-
-% Notes
-% N/a
-
 \section{Introduction}
@@
 A major part of the insights within this dataset come from the incorporation of operating costs and grape revenue, with environmental and sustainable data.
-We seek to address both the predictability of operating costs and revenue within the australian winegrowing context and examine their major driving factors to observe linked trends in sustainable practices.
+We seek to address both the predictability of operating costs and revenue within the Australian winegrowing context and examine their major driving factors to observe linked trends in sustainable practices.
 As part of this we examine the data to study economic outcomes and their statistical relationships to vineyards' utilisation of resources.
@@
 \par
-This study is further driven by recent reveiews calling for data-driven studies to show the economic benefits of sustainable practices within the wine industry, specifically winegrowing.
-While there is evidence to suggest that environmentally sustainable pracitces can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to numerically demonstrate these benefits across different regions and climates #TODO:(Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).
-Futhermore, many different sustainable approaches exist but are often studied in isolation or are limited in their geographical and climatic conditions, restricting their generalisability.
+This study is further driven by recent reviews calling for data-driven studies to show the economic benefits of sustainable practices within the wine industry, specifically winegrowing.
+While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, and improve the quality of grapes, more research is needed to numerically demonstrate these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023, laurentReviewIssuesMethods2021}.
+Furthermore, many different sustainable approaches exist but are often studied in isolation or are limited in their geographical and climatic conditions, restricting their generalisability.
 We embrace the variation that exists between vineyards and their unique challenges across Australia.
-Where, vineyard decisions on-the-ground are governed by complex physical forces of a regions' resources, climate, soil and geology, as well as by external pressures such as international market demands, disease and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
-Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018).
+Where, vineyard decisions on-the-ground are governed by complex physical forces of a regions' resources, climate, soil and geology, as well as by external pressures such as international market demands, disease and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,goodwinijeriepRegulatedDeficitIrrigation1992,hallSeasonTemporalVariation2011,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNonDestructiveSensingMethods2018}.
+
 \section{Methods}
-% TODO:
-%  From the code comments~ We might want to show the relationship between revenue and operational cost! then we have an idea of how these two are related.
-% # These are artefacts of.
-As profit can be negative it is harder to predict!
-We know after some further mapping that both operational costs and gross margin have a cubic root relationships to the logarithm of other variables.
-An interesting relationship to have!
-% # 
-% # Importantly this cubic root relationship exists between operational costs and gross margin, which is worth noting.
-However we want to know what tips it either side of them being equal as that shows that a vineyard is profitable!!
 
 \subsection{Data}\label{Data}
-% 6049 samples - total
-% 58 regions
-% 10 years
-% 
-% 853 with reported profits
-% 46 regions
-% reported since 2018
+
 Data used in this analysis were obtained from Sustainable Winegrowing Australia (SWA), Australia's national wine industry sustainability program.
@@
  \small
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-% \begin{table}[]
   \begin{tabular}{@{}lccc@{}}
@@
 \normalsize
-Due to the optional and manual recording of data, steps were taken to remove potentially erronous entries.
+
+Due to the optional and manual recording of data, steps were taken to remove potentially erroneous entries.
 This process first involved discussions with SWA highlighting possible entry errors.
 At the end of a season any suspect entries, such as a missing fuel-use in a vineyard that recorded the use of tractors, would warrant calling individual vineyards to clarify values and logic within the data.
-Similarly suspicious entries within the data were first described to viticulturalists for scrutiny before being addressed, either through calling growers for clarification or the removal of an observation due to its unlikely plausability, with most cases suspected of being incorrect units (commonmly litres instead of megalitres of water used) but were not able to be verified.
+Similarly suspicious entries within the data were first described to viticulturists for scrutiny before being addressed, either through calling growers for clarification or the removal of an observation due to its unlikely plausibility, with most cases suspected of being incorrect units (commonly litres instead of Megalitres of water used) but were not able to be verified.
 \par
@@
 This further helped to utilise relative importance for these variables directly to the act of using one system over another.
-This approach was compare to using the original variables but little difference in model accuracies was found between variables reported as proportion of a type used (i.e the percentage of land covered by drip irrigation), direct units of a type (i.e M\\L river water used) or as a binary presence/absence.
+This approach was compare to using the original variables but little difference in model accuracies was found between variables reported as proportion of a type used (i.e the percentage of land covered by drip irrigation), direct units of a type (i.e ML river water used) or as a binary presence/absence.
 Further details about these variables, their classes and their frequency is available in the Appendix.
+
 \subsection{Additional regional data}
+
 The variable Region represented one of the 65 Geographical Indicator Regions (GI Region) used to describe unique localised traits of vineyards across Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
@@
 And, temperature and rainfall means alongside extreme heat and cold days; as well as a regions' aridity index.
-This data was sourced using https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-0238.2010.00100.x
-And
-https://www.wineaustralia.com/growing-making/environment-and-climate/climate-atlas
+This data was sourced using
+\cite{remenyiAustraliasWineFuture2019} and \cite{hallSpatialAnalysisClimate2010}.
+
 \subsection{XGBoost}
+
 XGBoost is an ensemble method that combines multiple decision trees together to create a more accurate predictive model.
@@
 The loss function is optimised iteratively to improve upon prior trees (where the loss function can be any convex function), allowing gradient descent to traverse the loss space until no substantive improvements can be made (further detail pertaining to the algorithm is described in the Appendix).
-Because the loss function is only required to be convex, both classifiers and regressors can be used.
+Because the loss function is only required to be convex, both classification and regression can be used.
 Regularisation methods can also be incorporated to help prevent over fitting.
-This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods (TODO: ref).
+This makes XGBoost incredibly versatile and accurate, whilst still being interpretable compared to other machine learning methods \citep{kistenExplainableArtificialIntelligence2024}.
+\par
 XGBoost analyses were conducted using the XGBoost library \citep{chenXGBoostScalableTree2016} in the Python Programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-It is a method that is widely used within agriculture for yield prediction (TODO: ref), but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables (TODO: ref).
-We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables as XGBoost is known to perform well with mixed types of predictor domains (TODO: ref).
-Furthermore we choose XGBoost as it has a good performance in predictions whislt allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models.
+It is a method that is widely used within agriculture for yield prediction \citep{d.mariadassExtremeGradientBoosting2022,liGlobalDetrendingSignificantly2024,raviCropYieldPrediction2020}, but is also highly capable method for financial predictions, even when dealing with multi-domain predictor variables \citep{ZHANG2023104045}.
+We utilise XGBoot due to a combination of agricultural yield prediction, financial prediction and the use of both economic and environmental variables as XGBoost is known to perform well with mixed types of predictor domains \citep{doi:10.1080/15481603.2024.2349341,ZHANG2023104045}.
+Furthermore we choose XGBoost as it has a good performance in predictions whilst allowing the use of directly comparable metrics to sanity check models against prior research (such as yield using $R^2$), offering insight into the relative performance of models lacking prior reference points in the literature such as revenue and operating costs and making the model more interpretable to audiences familiar with regression models \citep{heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}.
 \par
@@
 XGBoost was also used due to its ability to handle sparse data, which was present within this dataset due to the voluntary nature of data entry, with many fields being left blank during data collection.
-Tree based methods also do not require data to be transformed prior to analyses; this consideration was taken into account so that specific partitions of values could be evaluated more easily and understood within original units of the data (TODO: ref).
+Tree based methods also do not require data to be transformed prior to analyses; this consideration was taken into account so that specific partitions of values could be evaluated more easily and understood within original units of the data \citep{d.mariadassExtremeGradientBoosting2022}.
 A further consideration in its use was the level of interpretability offered through measures of 'relative importance' allowing for the ability to identify and rank variables and interactions by contribution to predictions \citep{chenXGBoostScalableTree2016}.
 \par
-An XGBoost model was trained for each variable so that every variable's relative importance could be caculated.
+An XGBoost model was trained for each variable so that every variable's relative importance could be calculated.
 This process was done three times using three iterations of data (three models for each variable).
-The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region (see TODO: insert a table for this) and the final were trained on data with continuous variables transformed to be expressed as a ratio of vineyard area.
+The first models were trained on the original SWA data set, the second were trained on a dataset that incorporated external data for each region and the final were trained on data with continuous variables transformed to be expressed as a ratio of vineyard area.
 The final dataset that consisted of ratios also included the extra regional data (but not in ratio form).
@@
 \subsection{Sankey and Chord Diagrams}
-Originally created by Sankey to depict different pressures in steam engines (TODO: ref) we leverage Sankey diagrams to illustrate the different impact or 'pressure' each variable has on one another through the use of measurements of variable importance.
-Sankey and Chord diagrams were constructed using the Holoviews python library \citep{philipp_rudiger_2020_3904606}.
-Sankey diagrams (depicted on the left as section A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular represntation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
+
+Originally created by Sankey to depict different pressures in steam engines \citep{7536189} we leverage Sankey diagrams to illustrate the different impact or 'pressure' each variable has on one another through the use of measurements of variable importance.
+Sankey and Chord diagrams were constructed using the Holoviews python library \citep{rudigerHolovizHoloviewsVersion2020}.
+Sankey diagrams (depicted on the left as section A in figures) show the top 10 contributing factors to a variables prediction using XGBoost and Chord diagrams, a circular representation of Sankey diagrams (depicted on the right as section B in figures) show how each of the top 10 factors relate to one another by measures of relative importance.
 Both Chord and Sankey diagrams illustrate variable importance through the size of the bands between two variables.
-The number at the end of a connection in the diagrams indicates a variabless importance (the number of times it appeared within the ensemble).
+The number at the end of a connection in the diagrams indicates a variables importance (the number of times it appeared within the ensemble).
 
 \subsection{Variable Importance}\label{sec:importance}
+
 XGBoost creates a large number of decision trees in the ensemble, it is hard to directly interpret the model and the derived intricate relationship between the variables.
@@
 \subsection{Validation}
+
 The predictive accuracy of each tree was assessed through a validation process.
@@
 The performance measure for model selection was root-mean-square error for continuous variables.
-The receiver operator characteristic's area under the curve was used for binary variables \citep{hanley1982meaning}.
+The receiver operator characteristic's area under the curve was used for binary variables \citep{hanleyMeaningUseArea1982}.
 And, multiclass variables utilised the one verse one approach to minimise sensitivity to class disparity \citep{ferriExperimentalComparisonPerformance2009,handSimpleGeneralisationArea2001}.
 
-% \subsection{Surrogate Models}
-
-% The creation of more interpretable models such as linear regression in parallel to AI systems has been used to explain variable's relationships within black box algorithms \citep{molnarInterpretableMachineLearning2022}.
-As XGBoost create an ensemble of decision trees, here we use classification and regression trees to gain insight into intricacies of the ensembles derived through XGBoost.
-Decision Trees were created for operating costs, revenue and region.
-These models describe the partitions that are useful in predicting these variables; giving insight into the trees that make up the ensembles created by XGBoost.
-These trees were created using the rparts and caret packages \citep{kuhnBuildingPredictiveModels2008,terrytherneauRpartRecursivePartitioning2022} in the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
-% \par
-% Decision trees were validated using K-fold cross validation.
-Each model was validated using 10 folds, utilising a random selection of different samples ten separate times to validate each of the decision trees.
-The same measure of accuracy as the XGBoosted trees was used for comparison.
-
 \section{Results}
-
-% The below sections present each of the analyses conducted within this study.
-This includes the three analyses for Revenue, Operational Cost and Region, with the fourth and final analysis on profit and yield presented in the appendix.
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The Australian wine-growing industry is a rich and diverse landscape that is separated into multiple Geographical Indicator Regions.
-Each region describing unique reputations, qualities and varietals of wine produced there.
-While a great deal has been done regarding individual regional properties and traits, there has been little statistical insight into broader regional comparisons; due to a lack of cross-regional and in-depth data sources \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
-In this study we use Classification Trees to compare regional differences and how these differences relate to sustainable practices.
-% \newline
-
-\subsection{Revenue}
-
-% \begin{figure} 
-%   \resizebox{\textwidth}{!}{
-%   \includegraphics{revenue.pdf}}
-%   \caption{Decision tree predicting revenue.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:revenue_tree}
-%  \end{figure}
-
-The predictive performance of the XGBoost model for revenue performed similarly to operating cost, for achieving an $R^2$ of 0.77 (with a standard deviation of 0.15). 
-\par
-The most important predictors of revenue were fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199).
-The values attached to each variable indicate the relative importance of the variable (number of times selected in the tree ensemble, see Section \ref{fig:revenue_sankey}).
-Overall regions contributed to 234 nodes in the ensemble making them collectively the third most important variable.
-The chord diagram (see Figure \ref{fig:revenue_sankey}) illustrates that vineyard area is also of high relative importance to other variables, especially slashing.
-The overall importance of Area to other variables is evident by its larger circumference within the chord diagram.
-% The strong relationship of vineyard area to revenue and its predictors  but also has a high relative importance to the other predictors, The relation to area is likely to primarily be the effect of economies of scale, shown through its strong relations to other variables in figure \ref{fig:revenue_sankey}.
-Area harvested is likely also an indicator of other variables such as slashing passes its strongest connection presented.
-
-\begin{figure}[htb]
-  \label{fig:revenue_sankey}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{revenue.png}}
-  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
-The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}
- \end{figure}
 
@@
 %  \end{figure}
-
-Compared to revenue, the predictive performance of the XGBoost model for operating cost was slightly better, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
-Similar to revenue, the most important predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
-A surprising difference was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
-The connection between spraying and operational costs is intuitive in that it utilises equipment and resources to do.
-However, it is surprising that although spraying is considered important, that area not harvested due to disease is not, even though this would be a direct cause.
-The lack of importance on disease directly could be because of a low amount disease reported in the dataset (137 vineyards).
-The use of spraying is also unfortunately not part of the data, and could be in response to a variety of factors such as other vineyards within the region having disease or preventative sprays.
+Compared to revenue, the predictive performance of the XGBoost model for operating cost was slightly better when not using the extra regional parameters, with an $R^2$ of 0.80 (with a standard deviation of 0.10).
+Similarly, when predicting without extra regional data the most important predictors of operating cost were fuel, water, area and yield (see figure \ref{fig:operating_costs_sankey}).
+\par
+A major difference was also in the poorer performance of the model with extra regional parameters.
+Although achieving a similar $R^2$ of 0.78 and a standard deviation of 0.12 the difference lay in an outlier model recording an $R^2$ of 0.08.
+This divergence likely being due to over generalising using elevation to match neighbouring regions.
+The removal of this radical leaves the inclusion of extra regional parameters outperforming the models without them, however it overlooks the major pitfall in the ease of misattributing factors and causality when predicting these variables.
+\par
+A surprising difference between operating cost and revenue was the change in relative importance of activities involving tractor passes where the use of fungicide was more important for operational costs, compared to revenue, where slashing was more important (see Figure \ref{fig:region_sankey}).
+This difference was only found when not including the extra regional parameters.
+The model including extra regional parameters reflected an identical hierarchy of importance to its revenue counter part (see Figures \ref{fig:operating_costs_sankey} and \ref{fig:operating_costs_sankey_extra}).
+\par
+The connection between spraying and operational costs is intuitive in that it utilises both the expense of equipment and resources.
+However, it is surprising that although spraying is considered important when extra regional parameters were not included, 'area not harvested due to disease' was not, even though disease would be a direct cause.
+The lack of importance on disease directly could be due to a low amount reported in the dataset (137 vineyards).
+The reason for spraying was also unfortunately not part of the data, and could be in response to a variety of factors such as other vineyards within the region having disease or preventative sprays.
 The variables that feed into these decisions are also very different with diesel having the highest relative importance to slashing, and area having the greatest relative importance to the need for fungicide.
@@
 
-\subsection{Yield}
+\begin{figure}[htb]
+  \resizebox{\textwidth}{!}{
+  \includegraphics{operating_costs_extra.png}}
+  \caption{Operational costs and variable importance with the inclusion of extra regional parameters for terrain and climate.
+The left-hand side, A,  depicts the 10 most relative  important variables in predicting Operating Costs using XGBoost as a measure of node occurrence, using a Sankey diagram.
+The number at the end of each band in the diagram is that variable's importance.
+The right-hand side, B, depicts the importance of the 10 variables in Sankey diagram relative to one another.}\label{fig:operating_costs_sankey_extra}
+\end{figure}
+
+\subsection{Revenue}
+
+There was little difference in predictive power between the inclusion of climatic and regional data and the use of region alone.
+With the use of extra regional data achieving an $R^2$ of 0.76 (with a standard deviation of 0.13), and when not using the extra regional data achieving an $R^2$ of 0.77 (with a standard deviation of 0.15).
+The higher number of variables included as part of the regional data likely resulting in a lower variance in predictive power when including the extra regional data.
+\par
+The most notable difference was in the importance of the predictors, where elevation was a had a high relative importance of 8, only surpassed by diesel and yield, with 9 and 45 relative importance respectively (see Figure~\ref{fig:revenue_sankey_extra}).
+The lower variance is also reflected in the lower amount of nodes (or partitions) required when leveraging the extra regional parameters, with a reduction in hundreds of splits between the two models.
+Without the extra regional parameters it was found that fuel use (petrol 307 and diesel 144), yield (285), size (216) and water use (199) held the highest relative importance (see figure see Figure~\ref{fig:revenue_sankey}).
+\par
+Even without the extra regional parameters, overall region contributed to 234 nodes in the ensemble making it collectively the third most important variable.
+This places equal importance within both models on region however the nature of regions contribution is more generalised using elevation when included as a parameter.
+This alone does not necessarily make elevation a direct contributing factor but links regions that of similar terrain together.
+The relevance of this was noted when reviewing regional missclassifications, where neighbouring regions were often misclassified as each other.
+The extra numerical parameters that can be partitioned likely gives the algorithm a greater ability to partition these regions together using fewer nodes.
+
+\begin{figure}[htb]
+  \resizebox{\textwidth}{!}{
+  \includegraphics{revenue.png}}
+  \caption{The left-hand side depicts the 10 most important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
+The right-hand side depicts the interrelated importance of the ten predictor variables using a chord diagram.}\label{fig:revenue_sankey}
+\end{figure}
+
+\begin{figure}[htb]
+   \resizebox{\textwidth}{!}{
+   \includegraphics{revenue_extra.png}}
+   \caption{Revenue's variable importance with the inclusion of extra regional paramters for terrain and climate.
+The left-hand side, A,  depicts the 10 most relative  important variables in predicting revenue using XGBoost as a measure of node occurrence, using a Sankey diagram.
+The number at the end of each band in the diagram is that variable's importance.
+The right-hand side, B, depicts the importance of the 10 variables in Sankey diagram relative to one another.}\label{fig:revenue_sankey_extra}
+\end{figure}
 
 \subsection{Region}
-
-% \begin{figure}
-%   \resizebox{\textwidth}{!}{
-%   \includegraphics{region.pdf}}
-%   \caption{Decision tree predicting Region.
-Each node indicates the class predicted, and the proportion of elements agreeing with nodes partitioning, with the left direction indicating a yes to the nodes rule.}\label{fig:region_tree}
-%  \end{figure}
 
@@
 The ensemble had an emphasis on area, water, fuel and yield as determining factors (see Figure (\ref{fig:region_sankey}).
-% Although water and diesel were two of the three most frequently occurring variables in predicting region, they were not as connected to the other predictor variables as Yield and area harvested were.
 \par
@@
  \end{figure}
+
+\subsection{Results per hectare}
+Both operating cost and revenue were predicted as ratios of area (revenue and operating cost per vineyard hectare).
+In both cases it was found that the models performed poorly with operating costs recording an $R^2$ of 0.24 (with a standard deviation of 0.15) and revenue an $R^2$ of 0.32 (with a standard deviation of 0.14).
+
 \section{Discussion}
-% Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
-The primary reason for employing cover crops is to reduce the presence of disease and weeds (Delpuech and Metay, 2018).
-The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding (Capello et al., 2019).
-The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65% (Capello et al., 2020).
-However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
-A coverage of only 30% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield (Delpuech and Metay, 2018).
-% The reduction of tillage operations through optimising tractor efficiency is another practice that reduces energy use in vineyards, decreasing running costs, as well as reducing soil compaction (Capello et al., 2019).
-An increase in soil compaction has been shown to increase water runoff (Capello et al., 2020).
-Runoff is a significant factor as extreme rain events can cause large scale soil deposition, creating further erosion and removing topsoil.
-It is important that the interaction of events such as erosion have on other considerations such as soil health.
-% Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
-For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
-Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
-The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
-% When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
-Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
-Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
-The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
-Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
-
-This study explored the relationships between vineyard resource use, operations and geographical properties to revenue and operating costs.
-The analysis was based on a large national study of 6049 samples collected over ten years.
-Three main findings were identified.
-First, the most important predictors of revenue and operating costs were fuel, yield and area.
-Secondly, area and fuel were highly interrelated to other variables (see Figure \ref{fig:operating_costs_sankey} and Figure \ref{fig:revenue_sankey}A).
-Finally, the relative importance of predictor variables for Region, differed from Revenue and Operating costs, with Water Use being more prominent than Yield.
-Region was also more prominent than illustrated in the Sankey diagrams due to the relative importance for operating cost and revenue being calculated for individual regions and not all regions together.
-In its entirety Region was the third most important predictor of revenue and the most important predictor for operating costs, relative to the other variables consideration in the analyses.
-\par
-Several physical parameters such as climate, geography and soil are predetermined by a vineyard's location, making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The association between yield and region is demonstrated by yield apparing as the fourth most important variable when determining region (see Figure \ref{fig:region_sankey}).
-% The association with area and region is likely a connection to the change in land costs, with inland Australian areas (particularly of lower rainfall) being substantially cheaper to buy than coastal regions, allowing larger areas to be purchased \citep{willchancellorMeasuringAustralianBroadacre2019} % TODO: This section is commented out due to Mardi's take:
-%  Maybe but not sure this assumption holds. It may be an artefact of the members being corporate in larger holdings. The larger percentage of blocks in the riverland are small.
-\par
-Warmer regions are known to be beneficial in hastening the ripening process of winegrapes \citep{webbObservedTrendsWinegrape2011}.
-Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{botting1996canopy}.
+
+\subsection{Region and Climate}
+
+The relationships between vineyard resource use, operations and geographical properties is a complex one, illustrated through the highly interelated properties demonstrated in the chord diagrams.
+Many of the contributing physical parameters such as climate, geology and soil are predetermined by a vineyard's location, making it a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The contribution of geographical properties is reflected in region's continual appearance as important variable when predicting operating costs and revenue.
+The significance of region to operational costs and revenue is linked to the difference in resources available between regions, well illustrated through considering the relative importance that water use had when predicting region.
+The difference in readily available resources between different regions is also easily demonstrated when observing the partitioning of river water as the primary source of water to identify vineyards located in the Riverland (with 456 of 472 vineyards in the Riverland utilising river water).
+\par
+The availability of resources and geographical features of a region are significant but are also a potentially determining factor in the types of operational decisions made, and thus the reasoning as to variation in revenue and operating costs.
+This can be seen in the addition of extra regional parameters that add greater context to understanding the why or cause of operating costs and revenue.
+The effect of different operational consideration can be reflected in higher costs incurred in regions of greater variations in slope requiring more specialised equipment, or reflected in the types of resources available in some regions.
+The specifics of a vineyards' site are of incredible importance when determining what causes higher operational costs or revenue, as reflected through regions significance in predicting both revenue and operational cost.
+Although it is useful to be able to predict and compare regions, their revenue and operational costs, a greater nuance to help understand the why behind these decisions, would help in specifically guiding operational decisions.
+An example of this is whether the reduction of tillage operations through optimising tractor efficiency would be useful and how to optimise tractor use for a specific operation.
+This example is chosen because, while this practice is undertaken to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
+The interrelatedness of this decisions is far reaching as increase in tractor use can cause soil compaction which has been shown to further increase water runoff \citep{capelloPermanentCoverSoil2020}.
+With runoff itself being a significant factor during extreme rain events which can lead to large scale soil deposition, creating further erosion and removing topsoil and having wide spread effects for a vineyard.
+\par
+The climatic properties of regions are also a great determiner of different practices.
+For instance, warmer regions are known to be beneficial in hastening the ripening process of wine grapes \citep{webbObservedTrendsWinegrape2011}.
+Warmer regions are also associated with lower quality grapes, caused largely due to this hastened ripening \citep{bottingCanopyArchitectureImplicationsShiraz1996}.
 It is likely that the combination of larger vineyards with higher water use is a determining factor in classifying regions which favour larger production of grapes; reflected through region using water use so prominently in the XGBoost ensemble.
@@
 This classification can be contrasted with other warmer regions of higher rainfall that use the warmer climate to concentrate their grapes, increasing the flavour profile \citep{goodwinijeriepRegulatedDeficitIrrigation1992,mgmccarthyEffectCropLoad1986}.
-\par
-In part, yield is sometimes restricted simply through access to water resources.
+
+\subsection{Resource Use}
+
+The link between arid regions and yield is also complicated through the potential restricted access to water resources.
 Regions are likely to have varying access to different water sources, such as those along the River Murray being able to utilise river water for crops, unlike most coastal regions which may be drawing from surface or underground water sources.
@@
 The size, number of blocks, types and age of equipment will contribute to the efficiency of its use and the amount required across a site.
-It is likely that larger operations will generaly gain from economies of scale but also risk further incurring costs from the need to redeploy equipment.
-A further connection between region and fuel is the possible requirement of more specialty equipment, either due to regional practices differing or physical requirements such as greater inclines.
+It is likely that larger operations will generally gain from economies of scale but also risk further incurring costs from the need to redeploy equipment.
+A further connection between region and fuel is the possible requirement of more specialist equipment, either due to regional practices differing or physical requirements such as greater inclines.
 However, the style of management will also greatly contribute to how efficient both fuel and water are used, which is difficult to account for through the use of a metric.
-\par
-Operational costs showed similar importance across fuel, water and tractor use.
+
+\subsection{Sustainable Practices}
+
+Further to the consideration of including specific operations is hindered in this model due to the sample being derived specifically from vineyards already within a sustainable program.
+Making the sample inherently biased towards the use of sustainable practices.
+A keen example is the use of techniques such as cover crops.
+Cover crops are an example of a sustainable practice in viticulture in which the area between vine rows is seeded with a crop such as grasses or native vegetation.
+The primary reason for employing cover crops is to reduce the presence of disease and weeds \citep{delpuechAdaptingCoverCrop2018}.
+The benefit of reducing diseases and weeds is especially notable, as there is less cause to utilise heavy machinery for spraying herbicides and fungicides, or for mechanical weeding \citep{capelloEffectsTractorPasses2019}.
+The presence of a cover crop can also help to increase soil water retention, reducing erosion and water runoff in shallow soils, having been shown to mitigate runoff during rain events by over 65\% \citep{capelloPermanentCoverSoil2020}.
+However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{capelloEffectsTractorPasses2019,delpuechAdaptingCoverCrop2018,goslingLongTermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
+A coverage of only 30\% is required to provide protection against erosion, yet increased cover provides the benefits of greater biodiversity at the risk of yield \citep{delpuechAdaptingCoverCrop2018}.
+The presence of cover crops within the sample reflects this bias, where just over 85\% (5272) of vineyards utilised some form of cover crop such as grassing and only just under 4\% (225) used only bare soil (with the remaining 552 utilising a combination).
+The high percentage of vineyards using this type of sustainable practice means that its effect will not be prominent within the model, and can only show what practices would further improve those already implementing these techniques, and how they are connected to these operating costs.
+A strength of utilising XGBoost in this context is that, a subset of particular interest can be leveraged to focus in on the combination of factors that would contribute to the specific conceived scenario.
+Predicting operational costs reflected this through similar importance across fuel, water and tractor use.
 The dominating factor of area likely played a large part in determining how costly a tractor pass would be, or in defining the ratio of water applied to the amount of vines.
@@
 Although these analyses attempted to capture the complexity between how variables interacted when determining operational costs (see Figure \ref{fig:operating_costs_sankey}), in reality these relationships are likely even more complicated.
-An example of how interrelated operational costs can be, is
-the optimisation of tractor passes to achieve multiple goals in a pass, being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
-\par
+An example of how interrelated operational costs can be, is the optimisation of tractor passes to achieve multiple goals in a pass, being shown to reduce energy use in vineyards, decreasing running costs, as well as reducing soil compaction \citep{capelloEffectsTractorPasses2019}.
+
+\subsection{Revenue and Operational Costs}
+
 When determining revenue, similar variables were used to operational cost; with region also being of high variable importance relative to other variables (when considering all regions together in importance).
 It is difficult to extrapolate the specific influence of location on a vineyard's outcomes due to the broad and varying definition of a region.
-Utilising the Geographical Indicator regions defined by Wine Australia \citep{wineaustraliaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
+Utilising the Geographical Indicator regions defined by Wine Australia \citep{australiaWineAustraliaOpenData2021} is a limitation in one way, as it is too broad to fully capture a vineyards location and how that influences variables at a more granular level.
 However, as buyers set prices at regional levels, it is still important to consider this factor.
@@
 Other factors such as erosion and soil health are also influenced by tractor use and would contribute to these operational costs but are difficult to measure and were not available as part of the data \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020}.
-The data collection process being voluntary and part of a sustainable program also limitted the ability to compare what happened between those who had to abandon crops due to disease, pests or other catastrophes such as fire, in part due to a lack of incentive to record as part of the SWA program.
-Furthermore, no comparison can be made between those that have chosen to mothball as a response to predicted outcomes, or external presssures due to them not being part of the data.
+The data collection process being voluntary and part of a sustainable program also limited the ability to compare what happened between those who had to abandon crops due to disease, pests or other catastrophes such as fire, in part due to a lack of incentive to record as part of the SWA program.
+Furthermore, no comparison can be made between those that have chosen to mothball as a response to predicted outcomes, or external pressures due to them not being part of the data.
 Although this dataset contained vineyards that suffered partial losses due to disease, these limitations offer an avenue for further study that could benefit decision processes and variable relevance regarding mothballing, crop loss and external pressures.
-Without fully capturing more granular activities, for example the specific of tractor operastions and their differing fuel consumptions, it is difficult to determine what decisions specifically influence the operational costs.
+Without fully capturing more granular activities, for example the specific of tractor operations and their differing fuel consumptions, it is difficult to determine what decisions specifically influence the operational costs.
 Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions when considering external risks such as disease.
-%  Electricity in particular is used predominantly for irrigation. 
-% Size is also a further consideration where slashing and spraying are measured in discrete tractor passes and show a surprising connection to the overall size of a vineyard, despite not being scaled to any measure of size.
-This would mean that, although measured as the same increment, a slashing or spraying pass in a larger vineyard would consume more fuel and wages than in a smaller vineyard.
+\par
+Separately, revenue and operating cost did have a greater predictability than their counterpart profit (see Appendix).
+The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
+The difference between turning a profit or loss is dependent on predictable and unpredictable factors, farming practice and farmers' decisions.
+The difference between vineyards that make profit and those that do not could be a multitude of factors including differences in farming practices not captured within this study.
 \par
 The reasoning for any particular decision can be widely varying.
-% Where, a cover crop may be employed to help increase soil water retention, reduce erosion, increase biodiversity and reduce weeds \citep{capelloEffectsTractorPasses2019,capelloPermanentCoverSoil2020,delpuechAdaptingCoverCrop2018}.
-However, cover crops can introduce competition with grapevines and may reduce yield depending upon the plants used and the density of the cover crop \citep{goslingLongtermChangesSoil2005,monteiroInfluenceCoverCrop2007}.
-% A more granular definition of region may help to better discern the differences in practices, and the reason for employing them. 
 More sophisticated models, specifically those that utilise expert opinion, may also help to capture and address the decision-making process.
 An example is the optimisation of fungicide sprays using Bayesian models that forecast disease risk \citep{luDiseaseRiskForecasting2020}.
-\par
-% It is reasonable that regions, being subjected to different rainfalls and temperatures, would require different amounts of water, and would have access to different water sources.
-The relation of area harvested and fuel (particularly petrol) is prominent with other predictors.
-Due to the wide variety of uses of petrol and diesel.
-It is likely that fuel use is representative of other activities within the vineyard, such as pruning and harvesting.
-Predictors such as yield and area being highly interconnected to fuel are likely reflective of fuel required to  as they likely operate as proxy variables to other factors, possibly other present variables.
-
-% In particular There is the potential that the types of herbicides used also have a long term effect on crops, reducing the presence of microorganisms and soil health, making the area more prone to disease in the long run, becoming dependent on these sprays (Coll et al., 2011; Gosling and Shepherd, 2005).
-
-% The classification tree shows a likely change in avaialble infrastructure across the years, with mroe current years being associated with pressurised water and more slashing and herbicide passes.
-
-% The availability of solar panels for electrical irrigation and electrical irrigation in general
-
-% Electricity is predominantly used for pumping and irrigation in vineyards \citep{longbottomRoleVineyardPractices2015}
-
-% Fungicides are likely associated with higher preventive sprays when disease breaks out.
-
-% Slashing passes are likely weed outbreaks or yearly cycles?
-
-% %% Climate change
-
-% %% Pests Insects and weeds
-% %% Disease
-
-% %% Electricity use and reasons for (connect to year)
-% %% Fuel use and what it is used for
-
-% Operational costs saw the only occurrence of renewable resources being included in the ten most important variables for its XGBoost ensemble.
-
-% %% Address yield and how predictable it is
-% %% different sales markets and performance?
-% %% Reasons for profit?
-
-% The winegrowing industry holds significant importance for Australia and its economy. There exists many challenges that the industry has to contend against, 
-% with disease from sources such as Mildew and Botrytis being a considerable one (Cole, 2010; Magarey et al., 1994).
-
-Separately, revenue and operating cost did have a greater predictability than their counterpart profit (see Appendix).
-The disparity in accuracy between profit and other economic outcomes is reflective of the complexity in trying to address challenges such as climate change, disease and changing market demands \citep{wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022}.
-The difference between turning a profit or loss is dependent on predictable factors unforecasted factors, farming practice and farmers' decisions.
-The difference between vineyards that make profit and those that do not could be a multitude of factors including differences in farming practices not captured within this study.
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%  Prior studies have shown that decisions can also have long-lasting positive effects at the cost of short term negatives.
-In particular the use of cover crops and remedial soil techniques offer such as trade.
-
-% with well established a cover crops requiring greater water resources to maintain but offering more protection in return (Capello et al., 2019; Delpuech and Metay, 2018; Gosling and Shepherd, 2005; Monteiro and Lopes, 2007).
-
-% Prior studies have placed an emphasis on optimisations of fungicide sprays through the development of Bayesian models to forecast disease risk (Lu et al., 2020).
-This analysis investigates at the synergy between spray strategies as a proxy of fuel use to different types of water sources and sustainable strategies with an emphasis on the use of cover crops; allowing for the modelling of interactions between the use of multiple strategies.
-With the need to balance yield, quality, and combat adversities such as disease; the creation of tools to inform decisions and assist growers with warnings of disease risks is becoming crucial in ensuring sustainable and profitable wine production (Abbal et al., 2016).
-
-% Soil is an important and ongoing consideration for vineyards and interacts with every other practice in various ways at different time scales.
-For example, cover crops have been shown to be detrimental for soil health in the short term, giving an initial reduction in soil potassium and phosphorus concentrations and no change in nitrogen levels (Gosling and Shepherd, 2005).
-Conversely, in longer time frames the presence of a cover crop can induce an increase in microorganisms, which can excrete phosphates and potassium, regenerating the soils chemical balance and helping to introduce further organic nitrogen (Coll et al., 2011).
-The studies that showed this were based in two different countries of similar climate but could have been subject to other underlying conditions not measured.
-
-% When implementing practices such as cover crops the extent of the practice, the compounding effects and potential alternatives need to also be considered.
-Alternative options are always available such the use of mulch and wood chips to increase soil health and water retention in place of more involved processes such as crop rotation (Rössert et al., 2022).
-Crop rotation offers greater benefits than other soil management plans (Brock et al., 2011), however this is often not a viable option for many vineyards although is common practice in some places (Russo et al., 2021).
-The need to look at the holistic outcomes and interactions of these practices is paramount, however with the existence of many different practices, the outcomes due to interactions are not always known; one such consideration is the use of fungicide and its potential to build up copper, causing a reduction of microorganisms in the soil.
-Linking copper build up to any particular cause is a difficult endeavour due to the need for multiple reliable soil samples which are equally effected by the same conditions, within soil (Wightwick et al., 2010).
-
-% Reasons for operational costs (connect to disease and pests)
-
-% There are several environmental concerns that affect viticulture, including loss of soil quality, lack of rain, hail, disease, fire, and frost; with climate change exacerbating these issues.
-In 2020, 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase (Canadell et al., 2021).
-In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought (Wine Australia, 2020).
-
-% In rainfed areas due to the lack of irrigation possibilities The result would be a shortening of the ripening period, with harvest occurring during the period with high temperatures, which could have a negative impact on wine quality (Salazar Parra et al.
-2010; Duchne and Schneider 2005; Jones and Davis 2000) and yield (Mira de Ordua 2010; Iglesias et al.
-2010).
-
-% Climate change in the future might move the north and south latitude boundaries of areas suitable for good quality wines (Schultz and Jones 2010), and could even lead to improvements in fruit production and quality in some areas (Olesen and Bindi 2002).
-However, other areas may be negatively affected by high temperatures and water stress due to a reduction in the amount of water available.
-
-% %% Grape quality
-
-% The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
-Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to complex forces such as international market pressure, fire, pests and disease (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012).
-
-% The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
-The most prominent being what readily available resources of each region were, particular the types of water available.
-Heavy water consumption is often linked to the mass production of grapes, where lower quality grapes are targeted in a quantity over quality strategy.
-These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
-Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield ((Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I.
-Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018)).
-There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues, where on-the-ground choices are influenced by other wine industry’s decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
-It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
-It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season (Leilei He et al., 2022).
-
-% %% connection to watrer resource availability
-% What are the water resources in the hunter valley and tasmania?
-
-% economic pressures
-
-% Historically strong demands for Australian wine have helped to create a thriving industry, however recently sharp reductions in exports to mainland China due to significant deposit tariffs have caused a decline of 19% in Australian export value during the 2021-2022 financial year (see Figure 1) (Wine Australia, 2022).
-The pressure brought on by the drop in export value has been exacerbated by loss of tourism and labour due to the COVID-19 pandemic, global freight crisis, war in Europe and rising inflation (Wine Australia, 2021a, 2020).
-These pressures within the wine industry have trickled down to vineyards, where winemakers retain unwanted wine, creating an oversupply of grapes.
-Currently the proposed strategy to soften pressures proposed by Wine Australia includes initiatives focused on market diversification, and sustainability to build resilience in the coming years.
-
-% Figure 1: The exports of Australian wine over time in Australian Dollars Free On Board, comparing exports between China and the rest of the world.
-This graphic is taken from the Wine Australia Annual Report of 2020-21(Wine Australia, 2022).
-
-%%%%%%%%%%%%%%%
-
-% %Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74% and 96.97% respectively.
-
-% These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
-However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
-The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
-
-% The difference between grape quality is most notable between warm inland regions and coastal regions such as the Riverland and Coonawarra, respectively.
-Grape quality is only described by a singular variable within this study, however in reality it is driven by market demand and subject to complex forces such as international market pressure, fire, pests and disease \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} % The original citation included years prior to this dataset just be casreful with this.
-It is a bit unclear what you are really trying to cite and its relevance.
-I think it important to chop this up.
-
-% The decision trees were able to offer some insights into the factors that influence grape quality and regional contrasts that contribute to different qualities.
-The most prominent being what readily available resources of each region were, particular the types of water available.
-% A citation or reference to a study looking at the disparity in wine regions and their resources - or use the data more to describe these regional differences.
-Is water availability apparent?
-Do some regions have access to more sohpisticated water sources such as pressurised water etc.
-
-%  Heavy water consumption is often linked to the mass production of grapes,
-
- % Really?
-proce it?
-Where in the results is this really found??
-This sounds like a link to your previous paper but you cannot really show this and you do not really produce evidence to support this claim.
-
-%  where lower quality grapes are targeted in a quantity over quality strategy.
-These types of business decisions are unfortunately obfuscated by lack of in-depth data regarding vineyard business plans.
-
- % This point is incredibly important but it, itself, is obfuscated by you odd claims that are not supported by facts.
-Better to talk about how this could improve and the weaknesses than to start spouting unsupported claims.
-
-%  Notably the literature shows that there are many complex decisions to be made on the ground depending on many compounding factors that influence both quality and yield \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}
- % Woah woah woah
-%
- % what are these decisions??? Slow down speed gonzales!! This is really interseteing and important. Discuss it, its the discussion! God damn I want to know and I wrote this!.
- %
-%
-% This is not all one paragraph surely! There are so many ideas here that should be discussed!
-%
-% This leads heavily into the next paper and shows that natural progression.
-Especially in the sense of: if regions have so much influences, what determines ones use of water, fertiliser, etc.
-And, what determines the success financially, in quality or quantity?
-How are thes evariables interacting!
-The next paper is less about predicting and much more about what interacts with what?
-Where as this paper is about how things change over landscapes and what localised events helpe to determine decisions and their outcomes.
-%
-%  .
-There are also further differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues, where on-the-ground choices are influenced by other wine industry's decisions, such as the use of sustainable practices in vineyards to sell in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
-
-% This is an interesting point.
-However it would be better to include other industries, such as the ones you touch on in your previous chapter.
-If industries such as Sake are similar - how do their decisions effect the outcomes of rice quality and Sake?
-A point of comparison and difference would be great here!
-
-%  It is incredibly difficult to attribute external business decisions to produced grape quality but it is important to acknowledge that some growers are contracted to produce grapes of a particular grade; it is difficult to know whether another consumer may have graded the grape quality differently paying more or less for the same grapes given the opportunity to purchase them.
-
-% What are these business decisions? Can you give some examples?
-
-%  It is difficult to untangle the contributing factors to the success of winegrowers and the quality of grapes produced without further specifics of choices made through out a season \citep{leileiheFruitYieldPrediction2022}.
-
-% It is difficult yes, but please at least try!
-Highlight some of these affects, some of these different contributing factors!
-What did we really learn by doing this?
-What can someone take away from this and say thank you for?
-(like Josh said)
-
-% \subsection{Model 1 GI Regions}
-
-% The first Model was used to classify GI regions
- % Interestingly the firs I am hearing of this - each model needs to be more well summarised. A table would be good.
-
-% and resulted in an accuracy of 36.48\% across 52 classes.
-The most prominent features used to classify regions were the types of water resources available (see Figure 1).
-Two regions, the Riverland and Coonawarra, were the most accurate classes being 92.74\% and 96.97\% respectively.
-
-% it might be worth only talking about the regions that were well classifified - and then talk about their determining traits.
-
-% These regions differ greatly in practice and geophysical properties, with the Riverland being a dry warm inland region and Coonawarra being a cooler, wet coastal region.
-However, they are both similar in operational scales, with vineyards being relatively large compared with other regions.
-
-%Summary statistics regarding these two regions would be a better point of comparison, for example:
-%
-% How much in size do these operational scales differ?
-% How much yield do they differ by?
-% What are the differences in rainfall etc
-% What makes them unique, different and similar?
-%
-
-% The differences in resources and practices between these regions are also significant, such as the Riverland utilising the river Murray as a water source.
-% What are the specific differences? This is a really critical part of this paper
-
-% Many of the regions had significantly lower reporting rates, resulting much poorer classification performance.
- % This is almost its own paragraph. Why was this case? is it important? what are the similarities and differences of regions that were poorly classified?
-
-%  The regions with the most samples performed the best (see Table 1).
-Notably bordering regions were routinely grouped together and misclassified as the same region, for example the two closest regions to Coonawarra, Padthaway and Wrattonbulley, were misclassified as Coonawarra even though they had 147 and 137 samples respectively.
-The same case was found for the Murray Darling, with 143 samples, it was misclassified as the Riverland.
- % This speaks highly to why regions were poorly classified.
-
-% These misclassifications are likely due to the incredibly similar regional properties and close proximity these regions have with one another.
-Other misclassifications were most likely due to lower reporting rates with many regions being under represented.
-
-% \begin{table}[]
-%   \caption{Classification accuracy of the most prominent GI Regions.}
-%   \label{tab:accuracy}
-%   \resizebox{\textwidth}{!}{
-%     \begin{tabular}{@{}llll@{}}
-%       \toprule
-%       \textbf{} & \textbf{Accuracy} & \textbf{Predicted} & \textbf{Actual} \\ \midrule
-%       \textbf{Adelaide Hills} & 30.45\% & 95 & 312 \\
-%       \textbf{Barossa Valley} & 51.00\% & 205 & 402 \\
-%       \textbf{Coonawarra} & 96.97\% & 192 & 198 \\
-%       \textbf{Langhorne Creek} & 22.84\% & 53 & 232 \\
-%       \textbf{Margaret River} & 78.82\% & 201 & 255 \\
-%       \textbf{McLaren Vale} & 52.89\% & 128 & 242 \\
-%       \textbf{Riverland} & 92.74\% & 345 &
-%     \end{tabular}
-%   }
-% \end{table}
-
-% \subsection{Climate}
-% Classifying the SWA climatic categorisation of the given regions had better performance than the GI Regions, with ~41.66\% being classified correctly.
-These categories were divided into 12 climatic classifications with 3 and 4 separate subsets for rainfall and temperature respectively.
-The decision tree behaved similarly and over classified climates with higher response rates.
-The results posed an interesting similarity with grape quality classifier, being influenced predominantly by water and area.
-The use of fungicide to separate regions that were 'Very dry' and 'Damp' can be considered as indicative of the different practices required due to climatic pressure; fungicides being more prominent in cooler regions with greater rainfall due to the higher risk of disease pressure \citep{reynoldsManagingWineQuality2010}.
-% interesting although not really appropriate for the results section.
-%  This could also potentially explain the use of contractor tractor use to discern differences in grape quality, where the lack of contractor use to prevent disease could have led to lowered quality of grapes.
-
-% You really need to show every decision tree.
-
-% \subsubsection{Rainfall}
-% The rainfall decision tree showed a greater use of fungicides sprays to discern between damp and very Dry as shown in Figure 4; with the accuracy improving to 62\% but was unable to effectively discern between dry and very dry regions (see Table 3).
-
-% was again used to differentiate between warm and cool regions, likely being due to disease pressure. The temperature classification tree
-
-% \subsection{Model 3 Grape Quality}
-% The classification of grape quality through its grade % It might even be worth considering this as pseudo grade - or better define it in some of the australian wine journals rose suggested.
-% had an accuracy of 55.72\% across 5 separate grades.
-There was a notable issue with the classification of B grade grapes when compared to A and C (see Table 2).
-% This is actually really interesting.
-%  The classification tree itself shows similarities to that of classifying regions in Model 1, with the type of water resource used being a prominent determiner.
-Although not surprising the number of contractor tractor passes is new deciding factor due disease and pests reducing the potential quality of a crop.
-The prevalence of contractor use is greater in regions such as the Barossa Valley and the McLaren Vale, this could be due to the difference in operational scales, with larger sites being more likely to have ownership of their own equipment for weeding and spraying due to the cost benefit.
-
-\fi
+Further to this the use of models such as Bayesian Networks and multi criteria decision analysis would be a useful direction to proceed in to uncover the nuanced reasons and context as to why different operational decisions are made and their direct outcomes.
+Further research in this direction could aid in the creation of effective decision support systems to help the Australian winegrowing industry.
 
@@
 
-% The intricate relationships between water resources, and management practices have been examined through advanced techniques like XGBoost ensembles, shedding light on their roles in shaping production outcomes.
-
-% The type and availability of water resources were a major contributing factor when classifying grape quality and region.
-This was seen in the two most accurately classified regions, Coonawarra and the Riverland, with the Riverland predominantly utilising river water.
-Furthermore, the study highlighted the influence of water use, fungicide application, and contractor use in differentiating grape quality, climate and region respectively.
-These models provide insight into the complex dynamics between regional characteristics, sustainable practices, and grape quality in the Australian winegrowing industry.
-It is important to acknowledge that grape quality is subject to external influences such as market demands and prior established business arrangements.
-Further in-depth data and understanding are necessary to fully grasp the nuances of decision-making and the interplay of factors impacting grape quality.
-
-% Interesting - the conclusion is actually reasonable, especially considering the disconnection of ideas and models in this paper.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%       End Matter       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliography{references} % This points to the references.bib file - the file extention is automatically added.
+\bibliography{references}
 \bibliographystyle{elsarticle-harv}
 
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
 \appendix

### 2024-10-07 article.tex
added the varietals sentance

@@
 Although this dataset contained vineyards that suffered partial losses due to disease, these limitations offer an avenue for further study that could benefit decision processes and variable relevance regarding mothballing, crop loss and external pressures.
-Without fully capturing more granular activities, for example the specific of tractor operations and their differing fuel consumptions, it is difficult to determine what decisions specifically influence the operational costs.
+Without fully capturing more granular activities or, for example the specific tractor operations and their differing fuel consumptions, it is difficult to determine what decisions specifically influence the operational costs.
+Comparisons between which varietals were planted and the differences the techniques used between them would have also greatly benefited the data.
 Reductions in fuel, water and tractor use are obvious methods to reduce operational costs but not necessarily achievable decisions when considering external risks such as disease.
